### 2022-07-23 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

# Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

<!-- badges: start -->
<!-- badges: end -->

### `r emoji_glue(":snow_capped_mountain::fish:")`: Note on "mountainous mackerel" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [**codename**](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.4.0

codename(seed = "covid and ngos", type = "ubuntu")
#> [1] "mountainous mackerel"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

### 2022-07-23 _targets.R
targetsification

@@ -1,0 +1,26 @@
+library(targets)
+library(tarchetypes)
+
+options(tidyverse.quiet = TRUE,
+        dplyr.summarise.inform = FALSE)
+
+set.seed(8588)  # From random.org
+
+tar_option_set(
+  packages = c("tidyverse"),
+  format = "qs",
+)
+
+# here::here() returns an absolute path, which then gets stored in tar_meta and
+# becomes computer-specific (i.e. /Users/andrew/Research/blah/thing.Rmd).
+# There's no way to get a relative path directly out of here::here(), but
+# fs::path_rel() works fine with it (see
+# https://github.com/r-lib/here/issues/36#issuecomment-530894167)
+here_rel <- function(...) {fs::path_rel(here::here(...))}
+
+# Load R scripts with functions to use in the pipeline
+lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
+
+# Pipeline ----------------------------------------------------------------
+list(
+)

### 2022-07-23 R/funs_data-cleaning.R
Clean Oxford data

@@ -1,0 +1,28 @@
+library(readxl)
+library(lubridate)
+
+clean_oxford <- function(df) {
+  oxford_clean <- tibble(
+    # Get a list of all the sheets in the Excel file
+    index_name = excel_sheets(df)
+  ) %>% 
+    # Read each sheet
+    mutate(data = map(index_name, ~read_excel(df, sheet = .x))) %>% 
+    # Standardize the index name based on the sheet name
+    mutate(index_name = janitor::make_clean_names(index_name)) %>% 
+    # Make each data frame cell in the list column long and a little cleaner
+    mutate(clean = map(data, ~{
+      .x %>% 
+        pivot_longer(cols = -c(country_code, country_name),
+                     names_to = "day", values_to = "value") %>% 
+        mutate(day = dmy(day))
+    })) %>% 
+    # Get rid of the original wide data frame and unnest the long clean data
+    select(-data) %>% 
+    unnest(clean) %>% 
+    # Make data wide so that there's a column for each index and row for each
+    # country-day
+    pivot_wider(names_from = "index_name", values_from = "value")
+  
+  return(oxford_clean)
+}

### 2022-07-23 _targets.R

@@ -9,6 +9,8 @@ set.seed(8588)  # From random.org
 tar_option_set(
   packages = c("tidyverse"),
   format = "qs",
+  workspace_on_error = TRUE,
+  workspaces = c()
 )
 
 # here::here() returns an absolute path, which then gets stored in tar_meta and
@@ -23,4 +25,14 @@ lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
 
 # Pipeline ----------------------------------------------------------------
 list(
+  ## Raw data files ----
+  tar_target(iccpr_who_raw,
+             here_rel("data", "raw_data", "ICCPR Derogation and WHO case data 1 3 2020 to 6 30 2021.xlsx"),
+             format = "file"),
+  tar_target(oxford_raw,
+             here_rel("data", "raw_data", "Oxford Covid Response Data 1 3 2020 to 6 20 2021.xlsx"),
+             format = "file"),
+  
+  ## Process and clean data ----
+  tar_target(oxford_clean, clean_oxford(oxford_raw))
 )

### 2022-07-23 R/funs_data-cleaning.R
Deal with country names in Oxford data

@@ -2,7 +2,7 @@ library(readxl)
 library(lubridate)
 
 clean_oxford <- function(df) {
-  oxford_clean <- tibble(
+  x <- tibble(
     # Get a list of all the sheets in the Excel file
     index_name = excel_sheets(df)
   ) %>% 
@@ -22,7 +22,16 @@ clean_oxford <- function(df) {
     unnest(clean) %>% 
     # Make data wide so that there's a column for each index and row for each
     # country-day
-    pivot_wider(names_from = "index_name", values_from = "value")
+    pivot_wider(names_from = "index_name", values_from = "value") %>% 
+    # Country names and codes fun times
+    mutate(iso3 = recode(country_code, "RKS" = "XKK")) %>% 
+    mutate(country_name = countrycode(
+      iso3, origin = "iso3c", destination = "country.name",
+      custom_match = c("XKK" = "Kosovo", "TUR" = "Türkiye")
+    )) %>% 
+    # Final column order
+    select(-country_code) %>% 
+    select(country_name, iso3, day, everything())
   
-  return(oxford_clean)
+  return(x)
 }

### 2022-07-23 _targets.R

@@ -29,10 +29,10 @@ list(
   tar_target(iccpr_who_raw,
              here_rel("data", "raw_data", "ICCPR Derogation and WHO case data 1 3 2020 to 6 30 2021.xlsx"),
              format = "file"),
-  tar_target(oxford_raw,
+  tar_target(oxford_raw_file,
              here_rel("data", "raw_data", "Oxford Covid Response Data 1 3 2020 to 6 20 2021.xlsx"),
              format = "file"),
   
   ## Process and clean data ----
-  tar_target(oxford_clean, clean_oxford(oxford_raw))
+  tar_target(oxford_clean, clean_oxford(oxford_raw_file))
 )

### 2022-07-23 R/funs_data-cleaning.R
Clean ICCPR and WHO data

@@ -1,5 +1,33 @@
 library(readxl)
 library(lubridate)
+library(countrycode)
+
+clean_iccpr_who <- function(df) {
+  x <- read_excel(df) %>% 
+    janitor::clean_names() %>% 
+    # Make this a date instead of PosixCT
+    mutate(date_reported = as.Date(date_reported)) %>% 
+    # All NAs here are actually 0s
+    replace_na(list(iccpr_derogation_filed = 0,
+                    derogation_start = 0,
+                    derogation_ineffect = 0,
+                    derogation_end = 0)) %>% 
+    # Country names and codes fun times
+    mutate(
+      country_name = countrycode(
+        country_code, origin = "iso2c", destination = "country.name",
+        custom_match = c("XK" = "Kosovo", "TR" = "Türkiye")),
+      iso3 = countrycode(
+        country_code, origin = "iso2c", destination = "iso3c",
+        custom_match = c("XK" = "XKK")
+      )
+    ) %>% 
+    # Final column order
+    select(-c(country_code, country, cow_code, who_region)) %>% 
+    select(country_name, iso3, day = date_reported, everything())
+  
+  return(x)
+}
 
 clean_oxford <- function(df) {
   x <- tibble(

### 2022-07-23 _targets.R

@@ -26,7 +26,7 @@ lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
 # Pipeline ----------------------------------------------------------------
 list(
   ## Raw data files ----
-  tar_target(iccpr_who_raw,
+  tar_target(iccpr_who_raw_file,
              here_rel("data", "raw_data", "ICCPR Derogation and WHO case data 1 3 2020 to 6 30 2021.xlsx"),
              format = "file"),
   tar_target(oxford_raw_file,
@@ -34,5 +34,6 @@ list(
              format = "file"),
   
   ## Process and clean data ----
+  tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file))
 )

### 2022-07-23 R/funs_data-cleaning.R
Use path instead of x for the sake of consistency

@@ -2,8 +2,8 @@ library(readxl)
 library(lubridate)
 library(countrycode)
 
-clean_iccpr_who <- function(df) {
-  x <- read_excel(df) %>% 
+clean_iccpr_who <- function(path) {
+  x <- read_excel(path) %>% 
     janitor::clean_names() %>% 
     # Make this a date instead of PosixCT
     mutate(date_reported = as.Date(date_reported)) %>% 
@@ -29,13 +29,13 @@ clean_iccpr_who <- function(df) {
   return(x)
 }
 
-clean_oxford <- function(df) {
+clean_oxford <- function(path) {
   x <- tibble(
     # Get a list of all the sheets in the Excel file
-    index_name = excel_sheets(df)
+    index_name = excel_sheets(path)
   ) %>% 
     # Read each sheet
-    mutate(data = map(index_name, ~read_excel(df, sheet = .x))) %>% 
+    mutate(data = map(index_name, ~read_excel(path, sheet = .x))) %>% 
     # Standardize the index name based on the sheet name
     mutate(index_name = janitor::make_clean_names(index_name)) %>% 
     # Make each data frame cell in the list column long and a little cleaner

### 2022-07-23 R/funs_data-cleaning.R
Oops, Kosovo's fake ISO3 code is actually XKX

@@ -19,7 +19,7 @@ clean_iccpr_who <- function(path) {
         custom_match = c("XK" = "Kosovo", "TR" = "Türkiye")),
       iso3 = countrycode(
         country_code, origin = "iso2c", destination = "iso3c",
-        custom_match = c("XK" = "XKK")
+        custom_match = c("XK" = "XKX")
       )
     ) %>% 
     # Final column order
@@ -52,10 +52,10 @@ clean_oxford <- function(path) {
     # country-day
     pivot_wider(names_from = "index_name", values_from = "value") %>% 
     # Country names and codes fun times
-    mutate(iso3 = recode(country_code, "RKS" = "XKK")) %>% 
+    mutate(iso3 = recode(country_code, "RKS" = "XKX")) %>% 
     mutate(country_name = countrycode(
       iso3, origin = "iso3c", destination = "country.name",
-      custom_match = c("XKK" = "Kosovo", "TUR" = "Türkiye")
+      custom_match = c("XKX" = "Kosovo", "TUR" = "Türkiye")
     )) %>% 
     # Final column order
     select(-country_code) %>% 

### 2022-07-23 R/funs_data-cleaning.R
Initial V-Dem cleaning

@@ -63,3 +63,39 @@ clean_oxford <- function(path) {
   
   return(x)
 }
+
+clean_vdem <- function(path) {
+  vdem_raw <- read_rds(path) %>% as_tibble()
+  
+  vdem_clean <- vdem_raw %>% 
+    filter(year >= 2020) %>% 
+    mutate(country_name = countrycode(
+      country_text_id, origin = "iso3c", destination = "country.name",
+      custom_match = c("XKX" = "Kosovo", "ZZB" = "Zanzibar", 
+                       "PSG" = "Palestine (Gaza)", "SML" = "Somaliland", 
+                       "TUR" = "Türkiye")
+    )) %>% 
+    select(country_name, iso3 = country_text_id, year,
+           
+           # Civil society stuff
+           v2csreprss,  # CSO repression
+           v2xcs_ccsi,  # Core civil society index (entry/exit, repression, participatory env)
+           
+           # Human rights and politics
+           # Political corruption index (less to more, 0-1) (public sector +
+           # executive + legislative + judicial corruption)
+           v2x_corr,
+           v2x_rule,  # Rule of law index
+           
+           # Rights indexes
+           v2x_civlib,  # Civil liberties index
+           v2x_clphy,  # Physical violence index
+           v2x_clpriv,  # Private civil liberties index
+           v2x_clpol,  # Political civil liberties index
+           
+           # Democracy
+           v2x_polyarchy, v2x_regime_amb
+    )
+  
+  return(vdem_clean)
+}

### 2022-07-23 _targets.R

@@ -32,8 +32,13 @@ list(
   tar_target(oxford_raw_file,
              here_rel("data", "raw_data", "Oxford Covid Response Data 1 3 2020 to 6 20 2021.xlsx"),
              format = "file"),
+  tar_target(vdem_raw_file,
+             here_rel("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v12",
+                      "V-Dem-CY-Full+Others-v12.rds"),
+             format = "file"),
   
   ## Process and clean data ----
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
-  tar_target(oxford_clean, clean_oxford(oxford_raw_file))
+  tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
+  tar_target(vdem_clean, clean_vdem(vdem_raw_file))
 )

### 2022-07-23 R/funs_data-cleaning.R
Create daily skeleton

@@ -1,5 +1,5 @@
 library(readxl)
-library(lubridate)
+suppressPackageStartupMessages(library(lubridate))
 library(countrycode)
 
 clean_iccpr_who <- function(path) {
@@ -98,4 +98,23 @@ clean_vdem <- function(path) {
     )
   
   return(vdem_clean)
+}
+
+create_daily_skeleton <- function(iccpr_who, oxford, vdem) {
+  all_countries <- list(unique(iccpr_who$iso3), 
+                        unique(oxford$iso3), 
+                        unique(vdem$iso3))
+  
+  countries_in_all_data <- reduce(all_countries, intersect)
+  
+  first_day <- min(oxford$day)
+  last_day <- max(oxford$day)
+  
+  daily_skeleton <- expand_grid(
+    iso3 = countries_in_all_data,
+    day = seq(first_day, last_day, by = "1 day")
+  ) %>% 
+    mutate(year = year(day))
+  
+  return(daily_skeleton)
 }

### 2022-07-23 _targets.R

@@ -40,5 +40,7 @@ list(
   ## Process and clean data ----
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
-  tar_target(vdem_clean, clean_vdem(vdem_raw_file))
+  tar_target(vdem_clean, clean_vdem(vdem_raw_file)),
+  
+  tar_target(skeleton, create_daily_skeleton(iccpr_who_clean, oxford_clean, vdem_clean))
 )

### 2022-07-23 R/funs_data-cleaning.R
Remove missing Oxford data

@@ -57,6 +57,10 @@ clean_oxford <- function(path) {
       iso3, origin = "iso3c", destination = "country.name",
       custom_match = c("XKX" = "Kosovo", "TUR" = "Türkiye")
     )) %>% 
+    # Get rid of countries with all missing data
+    group_by(country_name) %>% 
+    filter(!all(is.na(stringency_index))) %>% 
+    ungroup() %>% 
     # Final column order
     select(-country_code) %>% 
     select(country_name, iso3, day, everything())

### 2022-07-23 R/funs_data-cleaning.R
Add country name to skeleton

@@ -118,7 +118,12 @@ create_daily_skeleton <- function(iccpr_who, oxford, vdem) {
     iso3 = countries_in_all_data,
     day = seq(first_day, last_day, by = "1 day")
   ) %>% 
-    mutate(year = year(day))
+    mutate(year = year(day)) %>% 
+    mutate(country_name = countrycode(
+      iso3, origin = "iso3c", destination = "country.name",
+      custom_match = c("XKX" = "Kosovo", "TUR" = "Türkiye")
+    )) %>% 
+    select(country_name, iso3, day, year)
   
   return(daily_skeleton)
 }

### 2022-07-23 R/funs_data-cleaning.R
Create final daily panel data

@@ -127,3 +127,12 @@ create_daily_skeleton <- function(iccpr_who, oxford, vdem) {
   
   return(daily_skeleton)
 }
+
+make_final_data <- function(skeleton, iccpr_who, oxford, vdem) {
+  daily_final <- skeleton %>% 
+    left_join(select(iccpr_who, -country_name), by = c("iso3", "day")) %>% 
+    left_join(select(oxford, -country_name), by = c("iso3", "day")) %>% 
+    left_join(select(vdem, -country_name), by = c("iso3", "year"))
+  
+  return(daily_final)
+}

### 2022-07-23 _targets.R

@@ -42,5 +42,7 @@ list(
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
   tar_target(vdem_clean, clean_vdem(vdem_raw_file)),
   
-  tar_target(skeleton, create_daily_skeleton(iccpr_who_clean, oxford_clean, vdem_clean))
+  tar_target(skeleton, create_daily_skeleton(iccpr_who_clean, oxford_clean, vdem_clean)),
+  
+  tar_target(daily_panel, make_final_data(skeleton, iccpr_who_clean, oxford_clean, vdem_clean))
 )

### 2022-07-23 _targets.R
Save final data as standalone files

@@ -44,5 +44,13 @@ list(
   
   tar_target(skeleton, create_daily_skeleton(iccpr_who_clean, oxford_clean, vdem_clean)),
   
-  tar_target(daily_panel, make_final_data(skeleton, iccpr_who_clean, oxford_clean, vdem_clean))
+  tar_target(daily_panel, make_final_data(skeleton, iccpr_who_clean, oxford_clean, vdem_clean)),
+  
+  ## Save data ----
+  tar_target(data_stata, 
+             haven::write_dta(daily_panel, here_rel("data", "derived_data", "daily_panel.dta"))),
+  tar_target(data_csv, 
+             write_csv(daily_panel, here_rel("data", "derived_data", "daily_panel.csv"))),
+  tar_target(data_rds, 
+             saveRDS(daily_panel, here_rel("data", "derived_data", "daily_panel.rds")))
 )

### 2022-07-24 analysis/index.qmd
Use a Quarto website

---
title: "Analysis notebook"
---

This is a Quarto website. Neato.

### 2022-07-24 _targets.R
Add website target

@@ -52,5 +52,8 @@ list(
   tar_target(data_csv, 
              write_csv(daily_panel, here_rel("data", "derived_data", "daily_panel.csv"))),
   tar_target(data_rds, 
-             saveRDS(daily_panel, here_rel("data", "derived_data", "daily_panel.rds")))
+             saveRDS(daily_panel, here_rel("data", "derived_data", "daily_panel.rds"))),
+  
+  ## Analysis notebook ----
+  tar_quarto(analysis_notebook, path = "analysis")
 )

### 2022-07-24 R/funs_data-cleaning.R
Include WHO regions in data

@@ -3,6 +3,16 @@ suppressPackageStartupMessages(library(lubridate))
 library(countrycode)
 
 clean_iccpr_who <- function(path) {
+  who_regions <- tribble(
+    ~who_region, ~who_region_long,
+    "AFRO", "Regional Office for Africa",
+    "AMRO", "Regional Office for the Americas",
+    "SEARO", "Regional Office for South-East Asia",
+    "EURO", "Regional Office for Europe",
+    "EMRO", "Regional Office for the Eastern Mediterranean",
+    "WPRO", "Regional Office for the Western Pacific"
+  )
+  
   x <- read_excel(path) %>% 
     janitor::clean_names() %>% 
     # Make this a date instead of PosixCT
@@ -22,9 +32,11 @@ clean_iccpr_who <- function(path) {
         custom_match = c("XK" = "XKX")
       )
     ) %>% 
+    left_join(who_regions, by = "who_region") %>% 
     # Final column order
-    select(-c(country_code, country, cow_code, who_region)) %>% 
-    select(country_name, iso3, day = date_reported, everything())
+    select(-c(country_code, country, cow_code)) %>% 
+    select(country_name, iso3, who_region, who_region_long,
+           day = date_reported, everything())
   
   return(x)
 }

### 2022-07-24 analysis/data-final.qmd
Add cool links to final data

---
title: "Download final data"
format:
  html:
    code-tools: false
---

<table class="table">
  <thead>
    <tr>
      <th scope="col">Program</th>
      <th scope="col">File</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>CSV file for any program</td>
      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa table >}} &ensp;<span class="font-monospace">daily_panel.csv</span></a></td>
    </tr>
    <tr>
      <td><span class="font-monospace">.rds</span> file for R<br><span class="small">(load with `df <- readRDS("daily_panel.rds")`)</small></td>
      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa brands r-project >}} &ensp;<span class="font-monospace">daily_panel.rds</span></a></td>
    </tr>
    <tr>
      <td><span class="font-monospace">.dta</span> file for Stata<br><span class="small">(load with `use "daily_panel.dta"`)</small></td>
      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa table-cells >}} &ensp;<span class="font-monospace">daily_panel.dta</span></a></td>
    </tr>
  </tbody>
</table>

### 2022-07-24 analysis/data-details.qmd
Initial details of data

---
title: "Data details"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(DT)
library(kableExtra)
library(lubridate)

matrix_from_vector <- function(x, ncol) {
  n_balanced <- ceiling(length(x) / ncol) * ncol
  matrix(c(x, rep(NA, n_balanced - length(x))), ncol = ncol)
}

# Generated via random.org
set.seed(2385)

# Load data 
withr::with_dir(here::here(), {
  # Final data
  tar_load(daily_panel)
})
```

```{r extract-details}
info <- list(
  n_days = length(unique(daily_panel$day)),
  n_countries = length(unique(daily_panel$iso3)),
  n_rows = scales::label_comma()(nrow(daily_panel)),
  dates = map(list(
    start_date = min(daily_panel$day),
    end_date = max(daily_panel$day)
  ), ~list(date = ., 
           nice = format(., "%B %e, %Y"),
           year = year(.)))
)
```

## Overview

We combine data from the [International Covenant on Civil and Political Rights](https://www.ohchr.org/en/instruments-mechanisms/instruments/international-covenant-civil-and-political-rights) (ICCPR), the [World Health Organization](https://covid19.who.int/data) (WHO), [Oxford's COVID-19 Government Response Tracker](https://www.bsg.ox.ac.uk/research/research-projects/covid-19-government-response-tracker), and the [Varieties of Democracy](https://www.v-dem.net/) (V-Dem) project to track government responses, COVID case counts, and derogation status at a daily level.

### Countries and dates included

We include data from `r info$n_countries` countries for `r info$n_days` days from `r info$dates$start_date$nice` to `r info$dates$end_date$nice`, resulting in a panel with `r info$n_rows` country-day observations.

```{r show-country-table}
countries <- daily_panel %>% 
  distinct(who_region, who_region_long, country_name) %>% 
  arrange(who_region) %>% 
  mutate(who = paste0(who_region_long, " (", who_region, ")")) %>% 
  group_by(who) %>% 
  nest() %>% 
  mutate(nice_matrix = map(data, ~matrix_from_vector(.$country_name, ncol = 6))) %>% 
  mutate(matrix_rows = map_int(nice_matrix, ~nrow(.)))

row_groups <- countries$matrix_rows %>% 
  set_names(countries$who)

reduce(countries$nice_matrix, rbind) %>% 
  kbl(table.attr = "class=\'table-sm\'") %>% 
  kable_styling(bootstrap_options = c("basic"), 
                full_width = FALSE) %>% 
  pack_rows(index = row_groups, indent = FALSE, 
            label_row_css = 'background-color: rgb(2, 184, 117); color: #ffffff; text-align: center;')
```

### General overview of full data

Here's an overview of the complete panel data:

```{r glimpse-everything}
#| code-fold: false
glimpse(daily_panel)
```

## ICCPR derogations

Our data is based on reading the texts of ICCPR derogations at the UN Treaty Collection site at <https://treaties.un.org/>. We create these variables:

- **ICCPR derogation filed** (`iccpr_derogation_filed`): Date that derogation notice was filed/received by the UN. This is a dichotomous variable that is `TRUE` on the day the country files the notice and `FALSE` otherwise.
- **Derogation start** (`derogation_start`): This is the date noted within the derogation text of when the emergency provisions and suspension of rights starts. This is a dichotomous variable that is `TRUE` on the day the country starts its derogation and `FALSE` otherwise.
- **Derogation in effect** (`derogation_ineffect`): This is a dichotomous variable that is `TRUE` if the country is currently derogating and `FALSE` if not.
- **Derogation end** (`derogation_end`): This is the date noted within the derogation text of when emergency provisions stop and rights observances resume. This is a dichotomous variable that is `TRUE` on the day the country finishes its derogation and `FALSE` otherwise.

(*Note: In many cases states filed derogations after the derogations actually started, so filed dates are often after the derogation start dates. Also note that states often updated the timing of derogations to extend, so there may be multiple filing dates within one derogation period.*)

Here's an overview of these variables:

```{r glimpse-iccpr}
#| code-fold: false
daily_panel %>%
  select(country_name, iso3, day, iccpr_derogation_filed,
         derogation_start, derogation_ineffect, derogation_end) %>% 
  glimpse()
```

And here's an extract of the data from June 11, 2020, the day that Argentina filed a derogation notice:

```{r show-iccpr}
daily_panel %>%
  filter(day == ymd("2020-06-11")) %>%
  select(country_name, iso3, day, iccpr_derogation_filed,
         derogation_start, derogation_ineffect, derogation_end) %>% 
  datatable()
```

## WHO

We collect daily case counts from the [WHO's COVID-19 Dashboard](https://covid19.who.int/data). We use these variables:

- **WHO region** (`who_region` and `who_region_long`): One of six WHO regional offices:
  - Regional Office for Africa (AFRO)
  - Regional Office for the Americas (AMRO)
  - Regional Office for the Eastern Mediterranean (EMRO)
  - Regional Office for Europe (EURO)
  - Regional Office for South-East Asia (SEARO)
  - Regional Office for the Western Pacific (WPRO)
- **New cases** (`new_cases`): New confirmed cases of COVID-19. Calculated by subtracting previous cumulative case count from current cumulative cases count.
- **Cumulative cases** (`cumulative_cases`): Cumulative confirmed cases of COVID-19 reported to the WHO to date.
- **New deaths** (`new_deaths`): New confirmed COVID-19 deaths. Calculated by subtracting previous cumulative deaths from current cumulative deaths.
- **Cumulative deaths** (`cumulative_deaths`): Cumulative confirmed COVID-19 deaths reported to the WHO to date.

We omit all "other" observations not assigned to any country (like cruise ships), and remove all non-states that don't appear in the UN's ICCPR list or in V-Dem, including:

```{r show-nonstate-table}
nonstates <- c("Falkland Islands", "Fore Islands", "Guernsey", "Guam", 
               "Isle of Man ", "Jersey", "Saint Pierre and Miquelon", 
               "Pitcairn Islands", "Puerto Rico", "Reunion Island", "Tokelau", 
               "Holy See", "Wallis and Futuna", "Bonaire", "Sint Eustatius", 
               "Saba", "Mayotte", "Gibraltar ", "Greenland", "Guadeloupe")

matrix_from_vector(nonstates, ncol = 5) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("basic"), 
                full_width = FALSE) %>% 
  pack_rows("Omitted non-states", 1, 4, indent = FALSE, 
            label_row_css = 'background-color: rgb(184, 71, 67); color: #ffffff; text-align: center;')
```

Here's an overview of these variables:

```{r glimpse-who}
#| code-fold: false
daily_panel %>%
  select(country_name, iso3, day, who_region, who_region_long,
         new_cases, cumulative_cases, new_deaths, cumulative_deaths) %>% 
  glimpse()
```

## Oxford

TODO

## V-Dem

TODO

### 2022-07-24 analysis/targets.qmd
Show overview of targets workflow

---
title: "Targets workflow"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

# targets pipeline

We use [the magical **`targets`** package](https://docs.ropensci.org/targets/) to run our analysis and keep track of all dependencies automatically.

To build our entire project, run `targets::tar_make()` at the R console.

Here's our complete pipeline:

```{r show-targets-pipeline, echo=FALSE}
withr::with_dir(here::here(), {
  targets::tar_glimpse()
})
```

# Actual code

All the data processing is handled with dataset-specific functions that live in `R/funs_data-cleaning.R`, which **`targets`** then runs as needed. For the sake of transparency, here's that code:

```{r generate-code-chunks, echo=FALSE}
# MAGIC: https://gist.github.com/StevenMMortimer/e54ec050d97d79996189
generate_chunk <- function(filename) {
  paste0(c(paste0("#### `R/", filename, "`"),
           paste0('```{r, code=xfun::read_utf8(here::here("R", "', filename, '")), eval=FALSE}'), 
           "#| code-fold: true", 
           "```", "", ""), 
         sep = "\n")
}

out <- NULL
for (thing in list.files(here::here("R"))) {
  out <- c(out, generate_chunk(thing))
}
```

`r paste(knitr::knit(text = out), collapse = "\n")`

### 2022-07-24 analysis/index.qmd
Add abstract

@@
 ---
-title: "Analysis notebook"
+title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
+format:
+  html:
+    code-tools: false
 ---
 
-This is a Quarto website. Neato.
+- Suparna Chaudhry (Lewis & Clark College)
+- Audrey L. Comstock (Arizona State University)
+- Andrew Heiss (Georgia State University)
+
+In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
+However, the abuse of emergency derogations by many governments raises concerns about the long-lasting impact of these sweeping measures on human rights.
+Does a robust civil society mitigate the adverse effects arising from international treaty derogations?
+Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run.
+Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker and the CoronaNet COVID-19 Government Response Event, we first assess the variation in human rights responses among governments during the pandemic.
+We then assess the role of civil society in keeping governments accountable and mitigating the effects of policies that violate human rights.
+Our results hold important implications for understanding government compliance with international treaties and shedding light on how states of emergency may directly and indirectly lead to democratic backsliding.

### 2022-07-24 analysis/data-details.qmd
Use neat callout box

@@
 
-(*Note: In many cases states filed derogations after the derogations actually started, so filed dates are often after the derogation start dates.
-Also note that states often updated the timing of derogations to extend, so there may be multiple filing dates within one derogation period.*)
+:::{.callout-note}
+In many cases states filed derogations after the derogations actually started, so filed dates are often after the derogation start dates.
+Also note that states often updated the timing of derogations to extend, so there may be multiple filing dates within one derogation period.
+:::
 

### 2022-07-24 analysis/data-details.qmd
Add Oxford section

@@
 
-## WHO
+## WHO COVID case counts
 
@@
 
-## Oxford
+## Oxford government responses
 
-TODO
+To measure the severity of government responses to COVID-19, we use data from the [Oxford COVID-19 Government Response Tracker](https://www.bsg.ox.ac.uk/research/research-projects/covid-19-government-response-tracker).
+We use the project's overall stringency index, as well as several measures of economic policies and containment and closure policies:
+
+- **Stringency index** (`stringency_index`): TODO
+- **Cancel public events** (`c3_cancel_public_events`): Indicator of whether the country cancelled public events.
+  - **0**: no measures; **1**: recommend cancelling; **2**: require cancelling; **Blank**: no data
+- **Limits on gatherings** (`c4_restrictions_on_gaterings`): Indicator of whether the country limited public gatherings.
+  - **0**: no restrictions; **1**: restrictions on very large gatherings (the limit is above 1000 people); **2**: restrictions on gatherings between 101-1000 people; **3**: restrictions on gatherings between 11-100 people; **4**: restrictions on gatherings of 10 people or less; **Blank**: no data
+- **Close public transport** (`c5_close_public_transport`): Indicator of whether the country closed public transport.
+  - **0**: no measures; **1**: recommend closing (or significantly reduce volume/route/means of transport available); **2**: require closing (or prohibit most citizens from using it); **Blank**: no data
+- **Stay at home requirements** (`c6_stay_at_home_requirements`): Indicator of whether the country imposed shelter-in-place requirements.
+  - **0**: no measures; **1**: recommend not leaving house; **2**: require not leaving house with exceptions for daily exercise, grocery shopping, and 'essential' trips; **3**: require not leaving house with minimal exceptions (eg allowed to leave once a week, or only one person can leave at a time, etc); **Blank**: no data
+- **Restrictions on internal movement** (`c7_movement_restrictions_intern`): Indicator of whether the country imposed restrictions on internal movement between cities or regions.
+  - **0**: no measures; **1**: recommend not to travel between regions/cities; **2**: internal movement restrictions in place; **Blank**: no data
+- **International travel controls** (`c8_international_travel_control`): Indicator of whether the country imposed restrictions on international travel (for foreign travellers, not citizens)
+  - **0**: no restrictions; **1**: screening arrivals; **2**: quarantine arrivals from some or all regions; **3**: ban arrivals from some regions; **4**: ban on all regions or total border closure; **Blank**: no data
+- **Income support** (`e1_income_support`): Indicator of whether the country provided direct cash payments for those unable to work.
+  - **0**: no income support; **1**: government is replacing less than 50% of lost salary (or if a flat sum, it is less than 50% median salary); **2**: government is replacing 50% or more of lost salary (or if a flat sum, it is greater than 50% median salary); **Blank**: no data
+- **Debt relief** (`e2_debt_relief`): Indicator of whether the country froze financial obligations for households (e.g. loan freezes, eviction bans, etc.)
+  - **0**: no debt/contract relief; **1**: narrow relief, specific to one kind of contract; **2**: broad debt/contract relief; **Blank**: no data
+
+Here's an overview of these variables:
+
+```{r glimpse-oxford}
+#| code-fold: false
+daily_panel %>% 
+  select(country_name, iso3, day, stringency_index, c3_cancel_public_events,
+         c4_restrictions_on_gaterings, c5_close_public_transport,
+         c6_stay_at_home_requirements, c7_movement_restrictions_intern,
+         c8_international_travel_control, e1_income_support, e2_debt_relief) %>% 
+  glimpse()
+```
+
+And here's an extract of the data from June 11, 2020:
+
+```{r show-oxford}
+daily_panel %>%
+  filter(day == ymd("2020-06-11")) %>%
+  select(country_name, iso3, day, stringency_index, c3_cancel_public_events,
+         c4_restrictions_on_gaterings, c5_close_public_transport,
+         c6_stay_at_home_requirements, c7_movement_restrictions_intern,
+         c8_international_travel_control, e1_income_support, e2_debt_relief) %>% 
+  datatable()
+```
 

### 2022-07-24 analysis/data-details.qmd
Add link to index methodology

@@
 
-- **Stringency index** (`stringency_index`): TODO
+- **Stringency index** (`stringency_index`): Composite measure of multiple indicators (C1–C* and H1), ranging from 0 to 100.
+[See the complete methodology](https://github.com/OxCGRT/covid-policy-tracker/blob/master/documentation/index_methodology.md) for more details.
 - **Cancel public events** (`c3_cancel_public_events`): Indicator of whether the country cancelled public events.

### 2022-07-24 analysis/data-details.qmd
Add V-Dem stuff

@@
 
+## Pandem
+
+TODO
+
 ## V-Dem
 
-TODO
+Finally, to measure additional country-level characteristics, we use several variables from the Varieties of Democracy project.
+
+:::{.callout-warning}
+Importantly, since these are all measured annually and our main data is daily, these variables act more like country-level fixed effects in our models, capturing the overall level of civil society, human rights, and other characteristics on a given day.
+:::
+
+Here's an overview of these variables:
+
+- **CSO repression** (`v2csreprss`)
+- **Core civil society index** (`v2xcs_ccsi`): Combination of entry/exit requirements, repression, and participatory environment
+- **Political corruption index** (`v2x_corr`): Combination of public sector, executive, legislative, and judicial corruption (less to more, 0–1)
+- **Rule of law** (`v2x_rule`)
+- **Civil liberties** (`v2x_civlib`)
+- **Physical violence** (`v2x_clphy`)
+- **Private civil liberties** (`v2x_clpriv`)
+- **Political civil liberties index** (`v2x_clpol`)
+- **Electoral democracy index** (`v2x_polyarchy`)
+- **Regimes of the world scores** (`v2x_regime_amb`)
+
+```{r glimpse-videm}
+#| code-fold: false
+daily_panel %>% 
+  select(country_name, iso3, day, starts_with("v2")) %>%
+  glimpse()
+```

### 2022-07-24 analysis/analysis_h1.qmd
Add hypotheses and other placeholders

---
title: "1: Derogations and human rights"
format:
  html:
    code-fold: true
---

### 2022-07-24 analysis/analysis_h2.qmd
Add hypotheses and other placeholders

---
title: "2: Civil society and human rights"
format:
  html:
    code-fold: true
---

### 2022-07-24 analysis/analysis_h3.qmd
Add hypotheses and other placeholders

---
title: "3: Civil society, derogations, and economic policies"
format:
  html:
    code-fold: true
---

### 2022-07-24 analysis/hypotheses.qmd
Add hypotheses and other placeholders

---
title: "Hypotheses"
format:
  html:
    code-tools: false
---

We test three sets of hypotheses:

## 1: Formal derogations and pandemic-related human rights violations

- **H~1a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
- **H~1b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights
- **H~1c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.

## 2: Civil society repression and pandemic-related human rights violations

- **H~2a~**: If states are repressive towards civil society organizations, then they are more likely to violate non-derogable rights.
- **H~2b~**: If states are repressive towards civil society organizations, then they are more likely to use discriminatory policy measures during the pandemic.
- **H~2c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.
- **H~2d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom.

## 3: Civil society, derogations, and pandemic-related economic policies

- **H~3a~**: If civil society organizations are involved in the policy making process and consulted by the government), then states are more likely to have supportive economic policies during the pandemic.
- **H~3b~**: If states formally derogate from their international treaty provisions, then states are more likely to have supportive economic policies during the pandemic.

### 2022-07-25 analysis/data-details.qmd
Add pandem note
Add PanDem

@@
 
-We combine data from the [International Covenant on Civil and Political Rights](https://www.ohchr.org/en/instruments-mechanisms/instruments/international-covenant-civil-and-political-rights) (ICCPR), the [World Health Organization](https://covid19.who.int/data) (WHO), [Oxford's COVID-19 Government Response Tracker](https://www.bsg.ox.ac.uk/research/research-projects/covid-19-government-response-tracker), and the [Varieties of Democracy](https://www.v-dem.net/) (V-Dem) project to track government responses, COVID case counts, and derogation status at a daily level.
+We combine data from the [International Covenant on Civil and Political Rights](https://www.ohchr.org/en/instruments-mechanisms/instruments/international-covenant-civil-and-political-rights) (ICCPR), the [World Health Organization](https://covid19.who.int/data) (WHO), [Oxford's COVID-19 Government Response Tracker](https://www.bsg.ox.ac.uk/research/research-projects/covid-19-government-response-tracker), the [Pandemic Backsliding Project](https://github.com/vdeminstitute/pandem) (PanDem), and the [Varieties of Democracy](https://www.v-dem.net/) (V-Dem) project to track government responses, COVID case counts, and derogation status at a daily level.
 
@@
 
-- **Stringency index** (`stringency_index`): Composite measure of multiple indicators (C1–C* and H1), ranging from 0 to 100.
+- **Stringency index** (`stringency_index`): Composite measure of multiple indicators (C1–C8 and H1), ranging from 0 to 100.
 [See the complete methodology](https://github.com/OxCGRT/covid-policy-tracker/blob/master/documentation/index_methodology.md) for more details.
+
 - **Cancel public events** (`c3_cancel_public_events`): Indicator of whether the country cancelled public events.
   - **0**: no measures; **1**: recommend cancelling; **2**: require cancelling; **Blank**: no data
+
 - **Limits on gatherings** (`c4_restrictions_on_gaterings`): Indicator of whether the country limited public gatherings.
   - **0**: no restrictions; **1**: restrictions on very large gatherings (the limit is above 1000 people); **2**: restrictions on gatherings between 101-1000 people; **3**: restrictions on gatherings between 11-100 people; **4**: restrictions on gatherings of 10 people or less; **Blank**: no data
+
 - **Close public transport** (`c5_close_public_transport`): Indicator of whether the country closed public transport.
   - **0**: no measures; **1**: recommend closing (or significantly reduce volume/route/means of transport available); **2**: require closing (or prohibit most citizens from using it); **Blank**: no data
+
 - **Stay at home requirements** (`c6_stay_at_home_requirements`): Indicator of whether the country imposed shelter-in-place requirements.
   - **0**: no measures; **1**: recommend not leaving house; **2**: require not leaving house with exceptions for daily exercise, grocery shopping, and 'essential' trips; **3**: require not leaving house with minimal exceptions (eg allowed to leave once a week, or only one person can leave at a time, etc); **Blank**: no data
+
 - **Restrictions on internal movement** (`c7_movement_restrictions_intern`): Indicator of whether the country imposed restrictions on internal movement between cities or regions.
   - **0**: no measures; **1**: recommend not to travel between regions/cities; **2**: internal movement restrictions in place; **Blank**: no data
-- **International travel controls** (`c8_international_travel_control`): Indicator of whether the country imposed restrictions on international travel (for foreign travellers, not citizens)
+
+- **International travel controls** (`c8_international_travel_control`): Indicator of whether the country imposed restrictions on international travel (for foreign travellers, not citizens).
   - **0**: no restrictions; **1**: screening arrivals; **2**: quarantine arrivals from some or all regions; **3**: ban arrivals from some regions; **4**: ban on all regions or total border closure; **Blank**: no data
+
 - **Income support** (`e1_income_support`): Indicator of whether the country provided direct cash payments for those unable to work.
   - **0**: no income support; **1**: government is replacing less than 50% of lost salary (or if a flat sum, it is less than 50% median salary); **2**: government is replacing 50% or more of lost salary (or if a flat sum, it is greater than 50% median salary); **Blank**: no data
+
 - **Debt relief** (`e2_debt_relief`): Indicator of whether the country froze financial obligations for households (e.g. loan freezes, eviction bans, etc.)
@@
 
-## Pandem
+## Pandemic Backsliding Project (PanDem)
 
-TODO
+We also measure the severity of government responses with data from the [Pandemic Backsliding Project](https://github.com/vdeminstitute/pandem) (PanDem).
 
-## V-Dem
+:::{.callout-warning}
+Importantly, since these are all measured quarterly and our main data is daily, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response on a given day.
+:::
 
-Finally, to measure additional country-level characteristics, we use several variables from the Varieties of Democracy project.
+We use the following variables:
+
+- **Year + quarter** (`year_quarter`): Numeric representation of the year and quarter, with quarter ranging from 0.1 to 0.4 (so 2020.4 is Q4 of 2020).
+PanDem treats March 11–March 31, 2020 as Q2 instead of Q1, so we recode our data accordingly (i.e.
+March 13, 2020 is 2020.2)
+
+- **Pandemic Violations of Democratic Standards Index** (`pandem`): Index measuring the extent to which governments violated democratic standards in their COVID-19 emergency response.
+Ranges from 0–1 and incorporates details about 7 different types of violations (`type1`, `type2`, `type3`, `type4`, `type5`, `type6`, and `type7` from PanDem).
+
+- **Pandemic Backsliding Index** (`panback`): Index measuring the extent to which democracy has receded because of violations of government standards during COVID-19.
+Ranges from 0–1 and incorporates both `pandem` and V-Dem's liberal democracy index (`v2x_libdem`) in 2019.
+
+- **Discriminatory measures** (`pandem_discrim`; originally `type1`): Ordinal measure of whether emergency measure violated the principle of non-discrimination from the ICCPR.
+  - **None**: no discrimination; **Minor**: de facto minor discrimination; **Moderate**: de jure minor discrimination; **Major**: major either de factor or de jure discrimination
+
+- **Derogations from non-derogable rights** (`pandem_ndrights`; originally `type2`): Ordinal measure of whether countries derogated from non-derogable ICCPR rights.
+  - **None**: no on all items of ICCPR provisions; **Major**: yes on any item of ICCPR provisions
+
+- **Abusive enforcement** (`pandem_abusive`; originally `type3`): Ordinal measure of whether state security violated physical integrity rights while enforcing emergency measures
+  - **None**: no to very few reports; **Minor**: sometimes; **Moderate**: often; **Major**: widespread
+
+- **No time limit** (`pandem_nolimit`; originally `type4`): Ordinal measure of whether emergency measures failed to state a time limit
+  - **None**: there is a declared end date for all measures; **Minor**: there is no end date for some or all measures, but measured ended within the quarter; **Moderate**: there is no end date for some or all measures and the measures have not ended during the quarter
+
+- **Restrictions on media freedom** (`pandem_media`; originally `type7`): Ordinal measure of whether limits were placed on media freedom as part of pandemic emergency measures
+  - **None**: no de jure or de facto restrictions; **Minor**: some de jure or de facto restrictions; **Moderate**: widespread limitations on access to information but few de jure or de facto restrictions on how media can report about the pandemic; **Major**: severe restrictions, limits, and/or harassment *(the PanDem codebook has a lot more detail about how all these are coded; this index is the most complex of the 7 types)*
+
+And here's an overview of these variables:
+
+```{r glimpse-pandem}
+#| code-fold: false
+daily_panel %>% 
+  select(country_name, iso3, day, year_quarter, 
+         pandem, panback, starts_with("pandem_")) %>% 
+  glimpse()
+```
+
+## Varieties of Democracy project (V-Dem)
+
+Finally, to measure additional country-level characteristics, we use several variables from the Varieties of Democracy (V-Dem) project.
 

### 2022-07-25 R/funs_data-cleaning.R

@@ -80,6 +80,37 @@ clean_oxford <- function(path) {
   return(x)
 }
 
+clean_pandem <- function(path) {
+  pandem_raw <- read_excel(path)
+  
+  pandem_levels <- c("None" = "0", "Minor" = "1", "Moderate" = "2", "Major" = "3")
+  
+  pandem_clean <- pandem_raw %>% 
+    mutate(quarter_numeric = parse_number(quarter) / 10) %>% 
+    mutate(year_quarter = year + quarter_numeric) %>% 
+    mutate(iso3 = countrycode(country_name, 
+                              origin = "country.name", 
+                              destination = "iso3c"),
+           country_name = countrycode(iso3, origin = "iso3c", 
+                                      destination = "country.name",
+                                      custom_match = c("TUR" = "Türkiye"))) %>% 
+    select(country_name, iso3, year, year_quarter, 
+           pandem, panback, 
+           pandem_discrim = type1, 
+           pandem_ndrights = type2,
+           pandem_abusive = type3,
+           pandem_nolimit = type4,
+           pandem_media = type7) %>% 
+    # Make these 0-3 columns factors
+    mutate(across(starts_with("pandem_"), ~factor(., levels = pandem_levels, ordered = TRUE))) %>% 
+    # Add labels
+    mutate(across(starts_with("pandem_"), ~fct_recode(., !!!pandem_levels))) %>% 
+    # Drop unused levels
+    mutate(across(starts_with("pandem_"), ~fct_drop(.)))
+  
+  return(pandem_clean)
+}
+
 clean_vdem <- function(path) {
   vdem_raw <- read_rds(path) %>% as_tibble()
   
@@ -110,40 +141,46 @@ clean_vdem <- function(path) {
            v2x_clpol,  # Political civil liberties index
            
            # Democracy
-           v2x_polyarchy, v2x_regime_amb
+           v2x_polyarchy, v2x_libdem, v2x_regime_amb
     )
   
   return(vdem_clean)
 }
 
-create_daily_skeleton <- function(iccpr_who, oxford, vdem) {
+create_daily_skeleton <- function(iccpr_who, oxford, pandem, vdem) {
   all_countries <- list(unique(iccpr_who$iso3), 
                         unique(oxford$iso3), 
+                        unique(pandem$iso3),
                         unique(vdem$iso3))
   
   countries_in_all_data <- reduce(all_countries, intersect)
[…]

### 2022-07-25 _targets.R

@@ -32,6 +32,9 @@ list(
   tar_target(oxford_raw_file,
              here_rel("data", "raw_data", "Oxford Covid Response Data 1 3 2020 to 6 20 2021.xlsx"),
              format = "file"),
+  tar_target(pandem_raw_file,
+             here_rel("data", "raw_data", "pandem", "datasets", "pandem_TS_v6.xlsx"),
+             format = "file"),
   tar_target(vdem_raw_file,
              here_rel("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v12",
                       "V-Dem-CY-Full+Others-v12.rds"),
@@ -40,11 +43,16 @@ list(
   ## Process and clean data ----
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
+  tar_target(pandem_clean, clean_pandem(pandem_raw_file)),
   tar_target(vdem_clean, clean_vdem(vdem_raw_file)),
   
-  tar_target(skeleton, create_daily_skeleton(iccpr_who_clean, oxford_clean, vdem_clean)),
+  tar_target(skeleton, 
+             create_daily_skeleton(iccpr_who_clean, oxford_clean, 
+                                   pandem_clean, vdem_clean)),
   
-  tar_target(daily_panel, make_final_data(skeleton, iccpr_who_clean, oxford_clean, vdem_clean)),
+  tar_target(daily_panel, 
+             make_final_data(skeleton, iccpr_who_clean, oxford_clean, 
+                             pandem_clean, vdem_clean)),
   
   ## Save data ----
   tar_target(data_stata, 

### 2022-07-25 analysis/data-final.qmd
Use file-based targets + add data to website

@@
       <td>CSV file for any program</td>
-      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa table >}} &ensp;<span class="font-monospace">daily_panel.csv</span></a></td>
+      <td><a class="btn btn-primary" target="_blank" href="data/daily_panel.csv">{{< fa table >}} &ensp;<span class="font-monospace">daily_panel.csv</span></a></td>
     </tr>
@@
       <td><span class="font-monospace">.rds</span> file for R<br><span class="small">(load with `df <- readRDS("daily_panel.rds")`)</small></td>
-      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa brands r-project >}} &ensp;<span class="font-monospace">daily_panel.rds</span></a></td>
+      <td><a class="btn btn-primary" target="_blank" href="data/daily_panel.rds">{{< fa brands r-project >}} &ensp;<span class="font-monospace">daily_panel.rds</span></a></td>
     </tr>
@@
       <td><span class="font-monospace">.dta</span> file for Stata<br><span class="small">(load with `use "daily_panel.dta"`)</small></td>
-      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa table-cells >}} &ensp;<span class="font-monospace">daily_panel.dta</span></a></td>
+      <td><a class="btn btn-primary" target="_blank" href="data/daily_panel.dta">{{< fa table-cells >}} &ensp;<span class="font-monospace">daily_panel.dta</span></a></td>
     </tr>

### 2022-07-25 R/funs_data-cleaning.R

@@ -185,3 +185,22 @@ make_final_data <- function(skeleton, iccpr_who, oxford, pandem, vdem) {
   
   return(daily_final)
 }
+
+# When using a file-based target, {targets} requires that the function that
+# saves the file returns a path to the file. write_csv() and write_dta() both
+# invisibly return the data frame being written, and saveRDS() returns NULL, so
+# we need some wrapper functions to save the files and return the paths.
+save_csv <- function(df, path) {
+  readr::write_csv(df, path)
+  return(path)
+}
+
+save_r <- function(df, path) {
+  saveRDS(df, path)
+  return(path)
+}
+
+save_dta <- function(df, path) {
+  haven::write_dta(df, path)
+  return(path)
+}

### 2022-07-25 _targets.R

@@ -56,11 +56,23 @@ list(
   
   ## Save data ----
   tar_target(data_stata, 
-             haven::write_dta(daily_panel, here_rel("data", "derived_data", "daily_panel.dta"))),
+             save_dta(daily_panel, here_rel("data", "derived_data", "daily_panel.dta")),
+             format = "file"),
+  tar_target(data_stata_website, 
+             save_dta(daily_panel, here_rel("analysis", "data", "daily_panel.dta")),
+             format = "file"),
   tar_target(data_csv, 
-             write_csv(daily_panel, here_rel("data", "derived_data", "daily_panel.csv"))),
+             save_csv(daily_panel, here_rel("data", "derived_data", "daily_panel.csv")),
+             format = "file"),
+  tar_target(data_csv_website, 
+             save_csv(daily_panel, here_rel("analysis", "data", "daily_panel.csv")),
+             format = "file"),
   tar_target(data_rds, 
-             saveRDS(daily_panel, here_rel("data", "derived_data", "daily_panel.rds"))),
+             save_r(daily_panel, here_rel("data", "derived_data", "daily_panel.rds")),
+             format = "file"),
+  tar_target(data_rds_website, 
+             save_r(daily_panel, here_rel("analysis", "data", "daily_panel.rds")),
+             format = "file"),
   
   ## Analysis notebook ----
   tar_quarto(analysis_notebook, path = "analysis")

### 2022-07-25 _targets.R
Let targets handle the uploading instead of make

@@ -75,5 +75,10 @@ list(
              format = "file"),
   
   ## Analysis notebook ----
-  tar_quarto(analysis_notebook, path = "analysis")
+  tar_quarto(analysis_notebook, path = "analysis"),
+  tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
+  tar_target(deploy_notebook, {
+    analysis_notebook  # Force a dependency
+    processx::run(paste0("./", deploy_script))  # Run the deploy script
+  })
 )

### 2022-07-25 deploy.sh

@@ -1,0 +1,7 @@
+REMOTE_HOST="cloud"
+REMOTE_DIR="~/sites/stats/public_html/mountainous-mackerel"
+REMOTE_DEST=$REMOTE_HOST:$REMOTE_DIR
+
+echo "Uploading new changes to remote server..."
+echo
+rsync -crvP --delete analysis/_site/ $REMOTE_DEST

### 2022-07-25 _targets.R
Only upload if environment variable is set (!)

@@ -78,7 +78,9 @@ list(
   tar_quarto(analysis_notebook, path = "analysis"),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
   tar_target(deploy_notebook, {
-    analysis_notebook  # Force a dependency
-    processx::run(paste0("./", deploy_script))  # Run the deploy script
+    # Force a dependency
+    analysis_notebook
+    # Run the deploy script
+    if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
   })
 )

### 2022-10-18 analysis/analysis_h1.qmd
Run models for Peace Science presentation + targetify models
lol lazy mega commit

@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
+                      fig.width = 6, fig.height = (6 * 0.618),
+                      out.width = "80%", collapse = TRUE,
+                      dev = "png", dev.args = list(type = "cairo-png"))
+
+options(digits = 3, width = 120,
+        dplyr.summarise.inform = FALSE,
+        knitr.kable.NA = "")
+```
+
+```{r load-libraries, warning=FALSE, message=FALSE}
+library(tidyverse)
+library(targets)
+library(broom)
+library(modelsummary)
+library(kableExtra)
+library(ordinal)
+library(brms)
+library(marginaleffects)
+library(tidybayes)
+library(ggdist)
+library(bayesplot)
+library(patchwork)
+library(scales)
+
+# Bayes stuff
+BAYES_SEED <- 1234
+options(mc.cores = 4,
+        brms.backend = "cmdstanr")
+
+# Generated via random.org
+set.seed(2385)
+
+# Load targets 
+withr::with_dir(here::here(), {
+  # Final data
+  tar_load(daily_panel)
+  tar_load(weekly_panel)
+  tar_load(year_week_lookup)
+
+  # Models
+  tar_load(m_h1)
+  tar_load(m_coef_gof)
+
+  # Plotting functions
+  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+})
+
+# Tell bayesplot to use the Lakota palette for things like pp_check()
+bayesplot::color_scheme_set(clrs[c(1, 2, 4, 5, 7, 8)])
+
+weekly_panel <- weekly_panel %>% 
+  mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
+                list(z = ~as.numeric(scale(.)))))
+```
+
+# Model and priors
+
+## Official formal model
+
+$$
+\begin{aligned}
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{region} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
+&\ \beta_2\ \text{New cases}_{i}\ + \beta_3\ \text{Cumulative cases}_{i}\ + \\
+&\ \beta_4\ \text{New deaths}_{i}\ + \beta_5\ \text{Cumulative deaths}_{i}\ + \\
+&\ \beta_6\ \text{Rule of law index}_{i}\ + \beta_7\ \text{Civil liberties index}_{i}\ + \\
+&\ \beta_8\ \text{Core civil society index}_{i}\ + \beta_9\ \text{Week number}_{i} \\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
+\\
+\bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{1 \dots 9} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
+\end{aligned}
+$$
+
+## Priors
+
+```{r fig.width=6, fig.height=2.5}
+p1 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
+                fill = clrs[2]) +
+  xlim(c(-20, 20)) +
+  annotate(geom = "label", x = 0, y = 0.02, label = "Student t(1, 0, 3)") +
+  labs(x = "α and βs") +
+  theme_pandem(prior = TRUE)
+
+p2 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~dcauchy(., 0, 1), 
+                fill = clrs[4]) +
+  xlim(c(0, 10)) +
+  annotate(geom = "label", x = 5, y = 0.063, label = "Cauchy(0, 1)") +
+  labs(x = "σ") +
+  theme_pandem(prior = TRUE)
+
+p1 | p2
+```
+
+## Prior simulation
+
+```{r}
+#| code-fold: show
+ologit_priors <- c(prior(student_t(1, 0, 3), class = Intercept),
+                   prior(student_t(1, 0, 3), class = b),
+                   prior(cauchy(0, 1), class = sd, lb = 0))
+
+model_h1_prior_only <- brm(
+  bf(pandem_nolimit ~ derogation_ineffect + new_cases_z + cumulative_cases_z + 
+       new_deaths_z + cumulative_deaths_z + v2x_rule + v2x_civlib + v2xcs_ccsi +
+       year_week_num + (1 | country_name)), 
+  data = weekly_panel,
+  family = "cumulative",
+  prior = ologit_priors,
+  sample_prior = "only",
+  chains = 4, seed = BAYES_SEED, iter = 5000, refresh = 0,
+  threads = threading(2)  # Two CPUs per chain to speed things up
+)
+model_h1_prior_only
+```
+
+```{r}
+datagrid(model = model_h1_prior_only,
+         year_week_num = 1:69,
+         derogation_ineffect = 0:1) %>% 
+  add_epred_draws(model_h1_prior_only, ndraws = 100) %>% 
+  left_join(year_week_lookup, by = "year_week_num") %>% 
+  mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                      levels = 0:1,
+                                      labels = c("No", "Yes"),
+                                      ordered = TRUE)) %>% 
+  ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
+  geom_line(aes(group = paste(.draw, .category, derogation_ineffect)), alpha = 0.5, size = 0.5) +
+  scale_color_manual(values = c(clrs[c(2, 4, 6)])) +
+  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+  scale_y_continuous(labels = label_percent()) +
+  labs(x = NULL, y = "Predicted probability", 
+       color = "Severity",
+       title = "No Time Limit Measures (priors only)") +
+  theme_pandem()
+```
+
+# Posterior
+
+```{r tbl-h1, cache=TRUE, warning=FALSE}
+modelsummary(pull(m_h1, model, name = nice),
+             estimate = "{estimate}",
+             statistic = "conf.int",
+             fmt = 2, coef_map = m_coef_gof$coef_map, 
+             gof_map = m_coef_gof$gof_map,
+             metrics = c("none"))
+```
+
+# Model evaluation
+
+```{r}
+prelim_trace <- function(model) {
+  model %>% 
+    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2xcs_ccsi, sd_who_region__Intercept) %>% 
+    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
+    geom_line(size = 0.1) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_trank <- function(model) {
+  model %>% 
+    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2xcs_ccsi, sd_who_region__Intercept) %>% 
+    group_by(.variable) %>% 
+    mutate(draw_rank = rank(.value)) %>% 
+    ggplot(aes(x = draw_rank, color = factor(.chain))) +
+    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_pp <- function(model) {
+  pp_check(model, ndraws = 100, type = "bars")
+}
+```
+
+::: {.panel-tabset}
+
+### `r m_h1$nice[[1]]`
+
+```{r}
+prelim_trace(m_h1$model[[1]])
+```
+
+```{r}
+prelim_trank(m_h1$model[[1]])
+```
+
+```{r}
+prelim_pp(m_h1$model[[1]])
+```
+
+### `r m_h1$nice[[2]]`
+
+```{r}
+prelim_trace(m_h1$model[[2]])
+```
+
+```{r}
+prelim_trank(m_h1$model[[2]])
+```
+
+```{r}
+prelim_pp(m_h1$model[[2]])
+```
+
+### `r m_h1$nice[[3]]`
+
+```{r}
+prelim_trace(m_h1$model[[3]])
+```
+
+```{r}
+prelim_trank(m_h1$model[[3]])
+```
+
+```{r}
+prelim_pp(m_h1$model[[3]])
+```
+
+### `r m_h1$nice[[4]]`
+
+```{r}
+prelim_trace(m_h1$model[[4]])
+```
+
+```{r}
+prelim_trank(m_h1$model[[4]])
+```
+
+```{r}
+prelim_pp(m_h1$model[[4]])
+```
+
+:::
+
+# Interpretation and marginal effects
+
+## Predictions
+
+::: {.column-page-right}
+
+```{r fig.width=13, fig.height=6, out.width="100%"}
+h1_models_plots <- m_h1 %>% 
+  mutate(pred_plot = pmap(list(model, nice, family), ~{
+    if (..3 == "cumulative") {
+      datagrid(model = ..1,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(.x) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No derogation", "Derogation in effect"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
+        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = ..2) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      datagrid(model = ..1,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(.x) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No", "Yes"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
+        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_fill_manual(values = c(clrs[1], clrs[8])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = ..2) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h1_models_plots$pred_plot)
+```
+
+:::
+
+## Marginal effects
+
+```{r}
+# TODO: Figure out how to work with ordered logit marginal effects
+```

### 2022-10-18 analysis/analysis_h2.qmd
Run models for Peace Science presentation + targetify models
lol lazy mega commit

@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
+                      fig.width = 6, fig.height = (6 * 0.618),
+                      out.width = "80%", collapse = TRUE,
+                      dev = "png", dev.args = list(type = "cairo-png"))
+
+options(digits = 3, width = 120,
+        dplyr.summarise.inform = FALSE,
+        knitr.kable.NA = "")
+```
+
+```{r load-libraries, warning=FALSE, message=FALSE}
+library(tidyverse)
+library(targets)
+library(broom)
+library(modelsummary)
+library(kableExtra)
+library(ordinal)
+library(brms)
+library(marginaleffects)
+library(tidybayes)
+library(ggdist)
+library(bayesplot)
+library(patchwork)
+library(scales)
+
+# Bayes stuff
+BAYES_SEED <- 1234
+options(mc.cores = 4,
+        brms.backend = "cmdstanr")
+
+# Generated via random.org
+set.seed(2385)
+
+# Load targets
+withr::with_dir(here::here(), {
+  # Final data
+  tar_load(daily_panel)
+  tar_load(weekly_panel)
+  tar_load(year_week_lookup)
+
+  # Models
+  tar_load(m_h2)
+  tar_load(m_coef_gof)
+
+  # Plotting functions
+  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+})
+
+# Tell bayesplot to use the Lakota palette for things like pp_check()
+bayesplot::color_scheme_set(clrs[c(1, 2, 4, 5, 7, 8)])
+
+weekly_panel <- weekly_panel %>% 
+  mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
+                list(z = ~as.numeric(scale(.)))))
+```
+
+# Model and priors
+
+## Official formal model
+
+$$
+\begin{aligned}
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{region} [i]} + \\
+&\ \beta_1\ \text{Derogation in effect}_{i}\ + \beta_2\ \text{Civil society repression}_{i}\ +\\
+&\ \beta_3\ {(\text{Derogation} \times \text{Civil society repression})}_{i}\ + \\
+&\ \beta_4\ \text{New cases}_{i}\ + \beta_5\ \text{Cumulative cases}_{i}\ + \\
+&\ \beta_6\ \text{New deaths}_{i}\ + \beta_7\ \text{Cumulative deaths}_{i}\ + \\
+&\ \beta_8\ \text{Rule of law index}_{i}\ + \beta_9\ \text{Civil liberties index}_{i}\ + \\
+&\ \beta_{10}\ \text{Week number}_{i} \\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
+\\
+\bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{1 \dots 10} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
+\end{aligned}
+$$
+
+## Priors
+
+```{r fig.width=6, fig.height=2.5}
+p1 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
+                fill = clrs[2]) +
+  xlim(c(-20, 20)) +
+  annotate(geom = "label", x = 0, y = 0.02, label = "Student t(1, 0, 3)") +
+  labs(x = "α and βs") +
+  theme_pandem(prior = TRUE)
+
+p2 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~dcauchy(., 0, 1), 
+                fill = clrs[4]) +
+  xlim(c(0, 10)) +
+  annotate(geom = "label", x = 5, y = 0.063, label = "Cauchy(0, 1)") +
+  labs(x = "σ") +
+  theme_pandem(prior = TRUE)
+
+p1 | p2
+```
+
+## Prior simulation
+
+```{r}
+#| code-fold: show
+ologit_priors <- c(prior(student_t(1, 0, 3), class = Intercept),
+                   prior(student_t(1, 0, 3), class = b),
+                   prior(cauchy(0, 1), class = sd, lb = 0))
+
+model_h2_prior_only <- brm(
+  bf(pandem_abusive ~ derogation_ineffect*v2csreprss + new_cases_z + cumulative_cases_z + 
+       new_deaths_z + cumulative_deaths_z + v2x_rule + v2x_civlib +
+       year_week_num + (1 | who_region)), 
+  data = weekly_panel,
+  family = "cumulative",
+  prior = ologit_priors,
+  sample_prior = "only",
+  chains = 4, seed = BAYES_SEED, iter = 5000, refresh = 0,
+  threads = threading(2)  # Two CPUs per chain to speed things up
+)
+model_h2_prior_only
+```
+
+```{r}
+datagrid(model = model_h2_prior_only,
+         year_week_num = 1:69,
+         derogation_ineffect = 0:1) %>% 
+  add_epred_draws(model_h2_prior_only, ndraws = 100) %>% 
+  left_join(year_week_lookup, by = "year_week_num") %>% 
+  mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                      levels = 0:1,
+                                      labels = c("No", "Yes"),
+                                      ordered = TRUE)) %>% 
+  ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
+  geom_line(aes(group = paste(.draw, .category, derogation_ineffect)), alpha = 0.5, size = 0.5) +
+  scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+  scale_y_continuous(labels = label_percent()) +
+  labs(x = NULL, y = "Predicted probability", 
+       color = "Severity",
+       title = "Abusive enforcement (priors only)") +
+  theme_pandem()
+```
+
+# Posterior
+
+```{r tbl-h2, cache=TRUE, warning=FALSE}
+modelsummary(pull(m_h2, model, name = nice),
+             estimate = "{estimate}",
+             statistic = "conf.int",
+             fmt = 2, coef_map = m_coef_gof$coef_map, 
+             gof_map = m_coef_gof$gof_map,
+             metrics = c("none"))
+```
+
+# Model evaluation
+
+```{r}
+prelim_trace <- function(model) {
+  model %>% 
+    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2csreprss, sd_who_region__Intercept) %>% 
+    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
+    geom_line(size = 0.1) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_trank <- function(model) {
+  model %>% 
+    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2csreprss, sd_who_region__Intercept) %>% 
+    group_by(.variable) %>% 
+    mutate(draw_rank = rank(.value)) %>% 
+    ggplot(aes(x = draw_rank, color = factor(.chain))) +
+    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_pp <- function(model) {
+  pp_check(model, ndraws = 100, type = "bars")
+}
+```
+
+::: {.panel-tabset}
+
+### `r m_h2$nice[[1]]`
+
+```{r}
+prelim_trace(m_h2$model[[1]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[1]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[1]])
+```
+
+### `r m_h2$nice[[2]]`
+
+```{r}
+prelim_trace(m_h2$model[[2]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[2]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[2]])
+```
+
+### `r m_h2$nice[[3]]`
+
+```{r}
+prelim_trace(m_h2$model[[3]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[3]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[3]])
+```
+
+### `r m_h2$nice[[4]]`
+
+```{r}
+prelim_trace(m_h2$model[[4]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[4]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[4]])
+```
+
+:::
+
+# Interpretation and marginal effects
+
+## Predictions
+
+### Across time
+
+::: {.column-page-right}
+
+```{r fig.width=13, fig.height=6, out.width="100%"}
+h2_models_plots <- m_h2 %>% 
+  mutate(pred_plot = pmap(list(model, nice, family), ~{
+    if (..3 == "cumulative") {
+      datagrid(model = ..1,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(.x) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No derogation", "Derogation in effect"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
+        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = ..2) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      datagrid(model = ..1,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(.x) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No", "Yes"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
+        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_fill_manual(values = c(clrs[1], clrs[8])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = ..2) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h2_models_plots$pred_plot)
+```
+
+:::
+
+### Across civil society repression
+
+::: {.column-page-right}
+
+```{r fig.width=13, fig.height=6, out.width="100%"}
+h2_models_plots1 <- m_h2 %>% 
+  mutate(pred_plot = pmap(list(model, nice, family), ~{
+    if (..3 == "cumulative") {
+      datagrid(model = ..1,
+               v2csreprss = seq(-2.5, 3, by = 0.1),
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(.x) %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No derogation", "Derogation in effect"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = v2csreprss, y = .epred, color = .category)) +
+        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = "Civil society repression", y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = ..2) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      datagrid(model = ..1,
+               v2csreprss = seq(-2.5, 3, by = 0.1),
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(.x) %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No", "Yes"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = v2csreprss, y = .epred, color = derogation_ineffect)) +
+        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_fill_manual(values = c(clrs[1], clrs[8])) +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = "Civil society repression", y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = ..2) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h2_models_plots1$pred_plot)
+```
+
+:::
+
+## Marginal effects
+
+```{r}
+# TODO: Figure out how to work with ordered logit marginal effects
+```

### 2022-10-18 analysis/analysis_prelim-tables.qmd
Run models for Peace Science presentation + targetify models
lol lazy mega commit

---
title: "Preliminary tables"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(broom)
library(modelsummary)
library(kableExtra)
library(ordinal)

# Generated via random.org
set.seed(2385)

# Load data 
withr::with_dir(here::here(), {
  # Final data
  tar_load(daily_panel)
})
```

```{r}
cm <- c("derogation_ineffectTRUE" = "Derogation in effect",
        "new_cases" = "New cases",
        "cumulative_cases" = "Cumulative cases",
        "new_deaths" = "New deaths",
        "cumulative_deaths" = "Cumulative deaths",
        "year_quarter" = "Year quarter",
        "v2x_rule" = "Rule of law",
        "v2x_civlib" = "Civil liberties",
        "v2xcs_ccsi" = "Core civil society index",
        "v2csreprss" = "Civil society repression",
        "(Intercept)" = "Constant",
        "None|Minor" = "Cut: None | Minor",
        "Minor|Moderate" = "Cut: Minor | Moderate",
        "Moderate|Major" = "Cut: Moderate | Major")

gm <- tribble(
  ~raw,        ~clean,     ~fmt, ~omit,
  "nobs",      "N",        0,    FALSE,
  "nclusters", "Clusters", 0,    FALSE
)
```

# Preliminary hypothesis: Does having a derogation matter for pandemic policies?

## Table 1: ICCPR Derogations and Covid-19 Restriction Policies

```{r warning=FALSE}
model_1 <- glm(c3_cancel_events_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths,
               data = daily_panel,
               family = binomial(link = "logit"))

model_2 <- glm(c4_gatherings_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths,
               data = daily_panel,
               family = binomial(link = "logit"))

model_3 <- glm(c5_public_transport_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths,
               data = daily_panel,
               family = binomial(link = "logit"))

model_4 <- glm(c7_internal_movement_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths,
               data = daily_panel,
               family = binomial(link = "logit"))

model_5 <- glm(c8_intl_travel_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths,
               data = daily_panel,
               family = binomial(link = "logit"))

modelsummary(list("Cancel Public Event" = model_1,
                  "Gathering Restrictions" = model_2,
                  "Close Public Transit" = model_3,
                  "Movement"  = model_4,
                  "International Travel" = model_5),
             vcov = ~country_name, 
             fmt = 4, stars = c("*" = 0.1, "**" = 0.01, "***" = 0.001),
             coef_map = cm, gof_map = gm)
```

## Table 2: ICCPR Derogations and Covid-19 Restriction Policies with more controls

```{r warning=FALSE}
model_6 <- glm(c3_cancel_events_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths +
                 year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
               data = daily_panel,
               family = binomial(link = "logit"))

model_7 <- glm(c4_gatherings_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths +
                 year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
               data = daily_panel,
               family = binomial(link = "logit"))

model_8 <- glm(c5_public_transport_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths +
                 year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
               data = daily_panel,
               family = binomial(link = "logit"))

model_9 <- glm(c7_internal_movement_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths +
                 year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
               data = daily_panel,
               family = binomial(link = "logit"))

model_10 <- glm(c8_intl_travel_bin ~ derogation_ineffect + new_cases + 
                 cumulative_cases + new_deaths + cumulative_deaths +
                  year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
               data = daily_panel,
               family = binomial(link = "logit"))

modelsummary(list("Cancel Public Event" = model_6,
                  "Gathering Restrictions" = model_7,
                  "Close Public Transit" = model_8,
                  "Movement"  = model_9,
                  "International Travel" = model_10),
             vcov = ~country_name, 
             fmt = 4, stars = c("*" = 0.1, "**" = 0.01, "***" = 0.001),
             coef_map = cm, gof_map = gm)
```

# H~1a-1c~

## Table 3: ICCPR Derogations and Covid-19 Polices

```{r}
model_11 <- clm(pandem_discrim ~ derogation_ineffect + new_cases + 
                  cumulative_cases + new_deaths + cumulative_deaths +
                  year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
                data = daily_panel)

model_12 <- glm(pandem_ndrights ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
                 data = daily_panel,
                 family = binomial(link = "logit"))

model_13 <- clm(pandem_nolimit ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
                 data = daily_panel)

model_14 <- clm(pandem_abusive ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
                 data = daily_panel)

modelsummary(list("Discriminatory Policy" = model_11,
                  "Non-Derogable Rights" = model_12,
                  "No Time Limit Measures" = model_13,
                  "Abusive Enforcement"  = model_14),
             vcov = ~country_name, 
             fmt = 4, stars = c("*" = 0.1, "**" = 0.01, "***" = 0.001),
             coef_map = cm, gof_map = gm)
```

# H~2a-2d~

## Table 4: Repressive Towards Civil Society and Covid-19 Policies

```{r}
model_15 <- clm(pandem_discrim ~ derogation_ineffect + new_cases + 
                  cumulative_cases + new_deaths + cumulative_deaths +
                  year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi + v2csreprss,
                data = daily_panel)

model_16 <- glm(pandem_ndrights ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi + v2csreprss,
                 data = daily_panel,
                 family = binomial(link = "logit"))

model_17 <- clm(pandem_media ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi + v2csreprss,
                 data = daily_panel)

model_18 <- clm(pandem_abusive ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi + v2csreprss,
                 data = daily_panel)

modelsummary(list("Discriminatory Policy" = model_15,
                  "Non-Derogable Rights" = model_16,
                  "Limiting media" = model_17,
                  "Abusive Enforcement"  = model_18),
             vcov = ~country_name, 
             fmt = 4, stars = c("*" = 0.1, "**" = 0.01, "***" = 0.001),
             coef_map = cm, gof_map = gm)
```

## Table 5: ICCPR Derogations and Supportive Economic Policies

```{r}
model_19 <- glm(e1_income_support_bin ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
                 data = daily_panel,
                 family = binomial(link = "logit"))

model_20 <- glm(e2_debt_relief_bin ~ derogation_ineffect + new_cases + 
                   cumulative_cases + new_deaths + cumulative_deaths +
                   year_quarter + v2x_rule + v2x_civlib + v2xcs_ccsi,
                 data = daily_panel,
                 family = binomial(link = "logit"))

modelsummary(list("Income Support" = model_19,
                  "Debt Relief" = model_20),
             vcov = ~country_name, 
             fmt = 4, stars = c("*" = 0.1, "**" = 0.01, "***" = 0.001),
             coef_map = cm, gof_map = gm)
```

### 2022-10-18 analysis/analysis_prelim.qmd
Run models for Peace Science presentation + targetify models
lol lazy mega commit

---
title: "0: Derogations and policies"
format:
  html:
    code-fold: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(broom)
library(modelsummary)
library(kableExtra)
library(ordinal)
library(brms)
library(marginaleffects)
library(tidybayes)
library(ggdist)
library(bayesplot)
library(patchwork)
library(scales)

# Bayes stuff
BAYES_SEED <- 1234
options(mc.cores = 4,
        brms.backend = "cmdstanr")

# Generated via random.org
set.seed(2385)

# Load targets 
withr::with_dir(here::here(), {
  # Final data
  tar_load(daily_panel)
  tar_load(weekly_panel)
  tar_load(year_week_lookup)

  # Models
  tar_load(m_prelim_derog)
  tar_load(m_coef_gof)

  # Plotting functions
  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
})

# Tell bayesplot to use the Lakota palette for things like pp_check()
bayesplot::color_scheme_set(clrs[c(1, 2, 4, 5, 7, 8)])

weekly_panel <- weekly_panel %>% 
  mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
                list(z = ~as.numeric(scale(.)))))
```

# Model and priors

## Official formal model

$$
\begin{aligned}
\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{country} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
&\ \beta_2\ \text{New cases}_{i}\ + \beta_3\ \text{Cumulative cases}_{i}\ + \\
&\ \beta_4\ \text{New deaths}_{i}\ + \beta_5\ \text{Cumulative deaths}_{i}\ + \\
&\ \beta_6\ \text{Week number}_{i} \\
\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
\\
\bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
\beta_{1 \dots 6} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
\end{aligned}
$$

## Priors

```{r fig.width=6, fig.height=2.5}
p1 <- ggplot() +
  stat_function(geom = "area", 
                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
                fill = clrs[2]) +
  xlim(c(-20, 20)) +
  annotate(geom = "label", x = 0, y = 0.02, label = "Student t(1, 0, 3)") +
  labs(x = "α and βs") +
  theme_pandem(prior = TRUE)

p2 <- ggplot() +
  stat_function(geom = "area", 
                fun = ~dcauchy(., 0, 1), 
                fill = clrs[4]) +
  xlim(c(0, 10)) +
  annotate(geom = "label", x = 5, y = 0.063, label = "Cauchy(0, 1)") +
  labs(x = "σ") +
  theme_pandem(prior = TRUE)

p1 | p2
```

## Prior simulation

```{r}
#| code-fold: show
logit_priors <- c(prior(student_t(1, 0, 3), class = Intercept),
                    prior(student_t(1, 0, 3), class = b),
                    prior(cauchy(0, 1), class = sd, lb = 0))

model_prelim_prior_only <- brm(
  bf(c7_internal_movement_bin ~ derogation_ineffect + new_cases_z + cumulative_cases_z + 
       new_deaths_z + cumulative_deaths_z + year_week_num + (1 | country_name)), 
  data = filter(weekly_panel, c7_internal_movement_never == 0),
  family = "bernoulli",
  prior = logit_priors,
  sample_prior = "only",
  chains = 4, seed = BAYES_SEED, iter = 5000, refresh = 0,
  threads = threading(2)  # Two CPUs per chain to speed things up
)
model_prelim_prior_only
```

```{r}
datagrid(model = model_prelim_prior_only,
         year_week_num = 1:69,
         derogation_ineffect = 0:1) %>% 
  add_epred_draws(model_prelim_prior_only, ndraws = 100) %>% 
  left_join(year_week_lookup, by = "year_week_num") %>% 
  mutate(derogation_ineffect = factor(derogation_ineffect, 
                                      levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
  ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
  geom_line(aes(group = paste(.draw, derogation_ineffect)), alpha = 0.5, size = 0.5) +
  scale_color_manual(values = c(clrs[1], clrs[8])) +
  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
  scale_y_continuous(labels = label_percent()) +
  labs(x = NULL, y = "Predicted probability", 
       color = "Derogation in effect",
       title = "Movement (priors only)") +
  theme_pandem()
```

# Posterior

```{r tbl-prelim, cache=TRUE}
modelsummary(pull(m_prelim_derog, model, name = nice),
             estimate = "{estimate}",
             statistic = "conf.int",
             fmt = 2, coef_map = m_coef_gof$coef_map, 
             gof_map = m_coef_gof$gof_map,
             metrics = c("R2"))
```

# Model evaluation

```{r}
prelim_trace <- function(model) {
  model %>% 
    gather_draws(b_Intercept, b_derogation_ineffect, b_year_week_num, sd_country_name__Intercept) %>% 
    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
    geom_line(size = 0.1) +
    scale_color_viridis_d(option = "rocket", end = 0.85) +
    labs(color = "Chain") +
    facet_wrap(vars(.variable), scales = "free_y") +
    theme_pandem()
}

prelim_trank <- function(model) {
  model %>% 
    gather_draws(b_Intercept, b_derogation_ineffect, b_year_week_num, sd_country_name__Intercept) %>% 
    group_by(.variable) %>% 
    mutate(draw_rank = rank(.value)) %>% 
    ggplot(aes(x = draw_rank, color = factor(.chain))) +
    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
    scale_color_viridis_d(option = "rocket", end = 0.85) +
    labs(color = "Chain") +
    facet_wrap(vars(.variable), scales = "free_y") +
    theme_pandem()
}

prelim_pp <- function(model) {
  pp_check(model, ndraws = 100, type = "bars")
}
```

::: {.panel-tabset}

### `r m_prelim_derog$nice[[1]]`

```{r}
prelim_trace(m_prelim_derog$model[[1]])
```

```{r}
prelim_trank(m_prelim_derog$model[[1]])
```

```{r}
prelim_pp(m_prelim_derog$model[[1]])
```

### `r m_prelim_derog$nice[[2]]`

```{r}
prelim_trace(m_prelim_derog$model[[2]])
```

```{r}
prelim_trank(m_prelim_derog$model[[2]])
```

```{r}
prelim_pp(m_prelim_derog$model[[2]])
```

### `r m_prelim_derog$nice[[3]]`

```{r}
prelim_trace(m_prelim_derog$model[[3]])
```

```{r}
prelim_trank(m_prelim_derog$model[[3]])
```

```{r}
prelim_pp(m_prelim_derog$model[[3]])
```

### `r m_prelim_derog$nice[[4]]`

```{r}
prelim_trace(m_prelim_derog$model[[4]])
```

```{r}
prelim_trank(m_prelim_derog$model[[4]])
```

```{r}
prelim_pp(m_prelim_derog$model[[4]])
```

### `r m_prelim_derog$nice[[5]]`

```{r}
prelim_trace(m_prelim_derog$model[[5]])
```

```{r}
prelim_trank(m_prelim_derog$model[[5]])
```

```{r}
prelim_pp(m_prelim_derog$model[[5]])
```

:::

# Interpretation and marginal effects

## Predictions

::: {.column-page-right}

```{r fig.width=12, fig.height=6, out.width="100%"}
preliminary_models_plots <- m_prelim_derog %>% 
  mutate(pred_plot = map2(model, nice, ~{
    datagrid(model = .x,
             year_week_num = 1:69,
             derogation_ineffect = 0:1) %>% 
      add_epred_draws(.x) %>% 
      left_join(year_week_lookup, by = "year_week_num") %>% 
      mutate(derogation_ineffect = factor(derogation_ineffect, 
                                          levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_fill_manual(values = c(clrs[1], clrs[8])) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      scale_y_continuous(labels = label_percent()) +
      labs(x = NULL, y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  }))

wrap_plots(preliminary_models_plots$pred_plot)
```

:::

## Marginal effects

::: {.column-page-right}

```{r fig.width=12, fig.height=6, out.width="100%"}
preliminary_models_mfx <- m_prelim_derog %>% 
  mutate(mfx = map(model, ~{
    .x %>% 
      marginaleffects(newdata = datagrid(year_week_num = 1:69),
                      variables = "derogation_ineffect",
                      type = "response") |> 
      posteriordraws()
  })) %>% 
  mutate(mfx_plot = map2(mfx, nice, ~{
    .x %>% 
      left_join(year_week_lookup, by = "year_week_num") %>% 
      ggplot(aes(x = year_week_day, y = draw * 100)) +
      stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      labs(x = NULL,
           y = "Percentage point change\nin probability of policy",
           title = .y) +
      theme_pandem()
  }))

wrap_plots(preliminary_models_mfx$mfx_plot)
```

:::

### 2022-10-18 analysis/data-details.qmd
Run models for Peace Science presentation + targetify models
lol lazy mega commit

@@
 
-- **Cancel public events** (`c3_cancel_public_events`): Indicator of whether the country cancelled public events.
+- **Cancel public events** (`c3_cancel_events`): Indicator of whether the country cancelled public events.
   - **0**: no measures; **1**: recommend cancelling; **2**: require cancelling; **Blank**: no data
 
-- **Limits on gatherings** (`c4_restrictions_on_gaterings`): Indicator of whether the country limited public gatherings.
+- **Limits on gatherings** (`c4_gatherings`): Indicator of whether the country limited public gatherings.
   - **0**: no restrictions; **1**: restrictions on very large gatherings (the limit is above 1000 people); **2**: restrictions on gatherings between 101-1000 people; **3**: restrictions on gatherings between 11-100 people; **4**: restrictions on gatherings of 10 people or less; **Blank**: no data
 
-- **Close public transport** (`c5_close_public_transport`): Indicator of whether the country closed public transport.
+- **Close public transport** (`c5_public_transport`): Indicator of whether the country closed public transport.
   - **0**: no measures; **1**: recommend closing (or significantly reduce volume/route/means of transport available); **2**: require closing (or prohibit most citizens from using it); **Blank**: no data
 
-- **Stay at home requirements** (`c6_stay_at_home_requirements`): Indicator of whether the country imposed shelter-in-place requirements.
+- **Stay at home requirements** (`c6_stay_at_home`): Indicator of whether the country imposed shelter-in-place requirements.
   - **0**: no measures; **1**: recommend not leaving house; **2**: require not leaving house with exceptions for daily exercise, grocery shopping, and 'essential' trips; **3**: require not leaving house with minimal exceptions (eg allowed to leave once a week, or only one person can leave at a time, etc); **Blank**: no data
 
-- **Restrictions on internal movement** (`c7_movement_restrictions_intern`): Indicator of whether the country imposed restrictions on internal movement between cities or regions.
+- **Restrictions on internal movement** (`c7_internal_movement`): Indicator of whether the country imposed restrictions on internal movement between cities or regions.
   - **0**: no measures; **1**: recommend not to travel between regions/cities; **2**: internal movement restrictions in place; **Blank**: no data
 
-- **International travel controls** (`c8_international_travel_control`): Indicator of whether the country imposed restrictions on international travel (for foreign travellers, not citizens).
+- **International travel controls** (`c8_intl_travel`): Indicator of whether the country imposed restrictions on international travel (for foreign travellers, not citizens).
   - **0**: no restrictions; **1**: screening arrivals; **2**: quarantine arrivals from some or all regions; **3**: ban arrivals from some regions; **4**: ban on all regions or total border closure; **Blank**: no data
@@
 daily_panel %>% 
-  select(country_name, iso3, day, stringency_index, c3_cancel_public_events,
-         c4_restrictions_on_gaterings, c5_close_public_transport,
-         c6_stay_at_home_requirements, c7_movement_restrictions_intern,
-         c8_international_travel_control, e1_income_support, e2_debt_relief) %>% 
+  select(country_name, iso3, day, stringency_index, c3_cancel_events,
+         c4_gatherings, c5_public_transport,
+         c6_stay_at_home, c7_internal_movement,
+         c8_intl_travel, e1_income_support, e2_debt_relief) %>% 
   glimpse()
@@
   filter(day == ymd("2020-06-11")) %>%
-  select(country_name, iso3, day, stringency_index, c3_cancel_public_events,
-         c4_restrictions_on_gaterings, c5_close_public_transport,
-         c6_stay_at_home_requirements, c7_movement_restrictions_intern,
-         c8_international_travel_control, e1_income_support, e2_debt_relief) %>% 
+  select(country_name, iso3, day, stringency_index, c3_cancel_events,
+         c4_gatherings, c5_public_transport,
+         c6_stay_at_home, c7_internal_movement,
+         c8_intl_travel, e1_income_support, e2_debt_relief) %>% 
   datatable()

### 2022-10-18 R/funs_data-cleaning.R

@@ -1,5 +1,6 @@
 library(readxl)
 suppressPackageStartupMessages(library(lubridate))
+suppressPackageStartupMessages(library(clock))
 library(countrycode)
 
 clean_iccpr_who <- function(path) {
@@ -72,6 +73,31 @@ clean_oxford <- function(path) {
     # Get rid of countries with all missing data
     group_by(country_name) %>% 
     filter(!all(is.na(stringency_index))) %>% 
+    ungroup() %>% 
+    # Shorten these long column names
+    rename(
+      c3_cancel_events = c3_cancel_public_events, 
+      c4_gatherings = c4_restrictions_on_gaterings,
+      c5_public_transport = c5_close_public_transport, 
+      c6_stay_at_home = c6_stay_at_home_requirements,
+      c7_internal_movement = c7_movement_restrictions_intern, 
+      c8_intl_travel = c8_international_travel_control
+    ) %>% 
+    # Make binary versions of these columns
+    mutate(across(c(c3_cancel_events, c4_gatherings,
+                    c5_public_transport, c6_stay_at_home,
+                    c7_internal_movement, c8_intl_travel,
+                    e1_income_support, e2_debt_relief),
+                  list(bin = ~ ifelse(. > 0, 1, 0)))) %>% 
+    # Create indicators for whether policies were added, removed, or never changed
+    group_by(country_name) %>% 
+    mutate(across(c(c3_cancel_events, c4_gatherings,
+                    c5_public_transport, c6_stay_at_home,
+                    c7_internal_movement, c8_intl_travel,
+                    e1_income_support, e2_debt_relief),
+                  list(added = ~any(c(NA, diff(.)) >= 1, na.rm = TRUE),
+                       removed = ~any(c(NA, diff(.)) <= -1, na.rm = TRUE),
+                       never = ~all(c(NA, diff(.)) == 0, na.rm = TRUE)))) %>% 
     ungroup() %>% 
     # Final column order
     select(-country_code) %>% 
@@ -164,14 +190,20 @@ create_daily_skeleton <- function(iccpr_who, oxford, pandem, vdem) {
     day = seq(first_day, last_day, by = "1 day")
   ) %>% 
     mutate(year = year(day),
-           year_quarter = quarter(day, type = "year.quarter")) %>% 
+           year_quarter = quarter(day, type = "year.quarter"),
[…]

### 2022-10-18 R/funs_graphics.R

@@ -1,0 +1,47 @@
+clrs <- MetBrewer::met.brewer("Tiepolo")
+
+set_annotation_fonts <- function() {
+  ggplot2::update_geom_defaults("label", list(family = "Noto Sans", face = "plain"))
+  ggplot2::update_geom_defaults("text", list(family = "Noto Sans", face = "plain"))
+}
+
+theme_pandem <- function(base_size = 11, base_family = "Noto Sans", prior = FALSE) {
+  ret <- theme_bw(base_size, base_family) +
+    theme(panel.background = element_rect(fill = "#ffffff", colour = NA),
+          title = element_text(size = rel(1), family = "Noto Sans", face = "bold"),
+          plot.subtitle = element_text(size = rel(0.8),
+                                       family = "Noto Sans", face = "plain"),
+          plot.caption = element_text(margin = margin(t = 10), size = rel(0.6),
+                                      family = "Noto Sans", face = "plain"),
+          panel.border = element_rect(color = "grey50", fill = NA, size = 0.15),
+          panel.spacing = unit(1, "lines"),
+          panel.grid.minor = element_blank(),
+          panel.grid.major = element_line(size = 0.25, colour = "grey90"),
+          axis.line = element_blank(),
+          axis.ticks = element_blank(),
+          axis.title = element_text(size = rel(0.8),
+                                    family = "Noto Sans", face = "bold"),
+          axis.title.x = element_text(hjust = 0, margin = margin(t = 10)),
+          axis.title.y = element_text(hjust = 1, margin = margin(r = 10)),
+          legend.position = "bottom",
+          legend.title = element_text(size = rel(0.7), vjust = 0.5,
+                                      family = "Noto Sans", face = "plain"),
+          legend.key.size = unit(0.7, "line"),
+          legend.key = element_blank(),
+          legend.spacing = unit(0.1, "lines"),
+          legend.justification = "left",
+          legend.margin = margin(t = -5, b = 0, l = 0, r = 0),
+          strip.text = element_text(size = rel(0.9), hjust = 0,
+                                    family = "Noto Sans", face = "bold"),
+          strip.background = element_rect(fill = "white", colour = NA))
+  
+  if (prior) {
+    ret <- ret +
+      theme(panel.grid.major = element_blank(),
+            axis.title.y = element_blank(),
+            axis.text.y = element_blank(),
+            panel.border = element_blank())
+  } else {
+    ret
+  }
+}

### 2022-10-18 R/funs_models.R

@@ -1,0 +1,258 @@
+coef_map <- c(
+  "b_derogation_ineffect" = "Derogation in effect",
+  "b_v2csreprss" = "Civil society repression",
+  "b_derogation_ineffect:v2csreprss" = "Derogation × Civil society represstion",
+  "b_new_cases_z" = "New cases (standardized)",
+  "b_cumulative_cases_z" = "Cumulative cases (standardized)",
+  "b_new_deaths_z" = "New deaths (standardized)",
+  "b_cumulative_deaths_z" = "Cumulative deaths (standardized)",
+  "b_year_week_num" = "Year-week",
+  "b_v2x_rule" = "Rule of law",
+  "b_v2x_civlib" = "Civil liberties",
+  "b_v2xcs_ccsi" = "Core civil society index",
+  "b_Intercept" = "Constant",
+  "b_Intercept[1]" = "Cut 1",
+  "b_Intercept[2]" = "Cut 2",
+  "b_Intercept[3]" = "Cut 3",
+  "sd_country_name__Intercept" = "Country random effects σ",
+  "sd_who_region__Intercept" = "Region random effects σ"
+)
+
+gof_map <- tribble(
+  ~raw,        ~clean,     ~fmt, ~omit,
+  "nobs",      "N",        0,    TRUE
+)
+
+# This is how to get other stats like ELPD (using metrics = "LOOIC" is required
+# for that; metrics = "R2" provides nobs). I would do this for all the summary
+# tables, but for whatever reason, it takes forever to calculate ELPD/LOO stuff
+# on ordered logit models, so we just show N instead
+#
+# gm <- tribble(
+#   ~raw,        ~clean,      ~fmt, ~omit,
+#   "nobs",      "N",         0,    FALSE,
+#   "r.squared", "R2",        3,    FALSE,
+#   "elpd",      "ELPD",      1,    FALSE,
+#   "elpd.se",   "ELPD (SE)", 1,    FALSE
+# )
+# 
+# modelsummary(models_prelim,
+#              estimate = "{estimate}",
+#              statistic = "conf.int",
+#              gof_map = gm,
+#              metrics = c("R2", "LOOIC"))
+
+
+f_prelim_derog <- function(panel) {
+  BAYES_SEED <- 1757  # From random.org
+  
+  panel <- panel %>% 
+    mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
[…]

### 2022-10-18 _targets.R

@@ -1,8 +1,14 @@
 library(targets)
 library(tarchetypes)
+suppressPackageStartupMessages(library(dplyr))
 
 options(tidyverse.quiet = TRUE,
         dplyr.summarise.inform = FALSE)
+
+# Bayesian stuff
+suppressPackageStartupMessages(library(brms))
+options(mc.cores = 4,
+        brms.backend = "cmdstanr")
 
 set.seed(8588)  # From random.org
 
@@ -54,25 +60,61 @@ list(
              make_final_data(skeleton, iccpr_who_clean, oxford_clean, 
                              pandem_clean, vdem_clean)),
   
+  tar_target(weekly_panel, make_weekly_data(daily_panel)),
+  
+  tar_target(year_week_lookup, make_year_week_lookup(weekly_panel)),
+  
   ## Save data ----
-  tar_target(data_stata, 
+  ### Daily ----
+  tar_target(data_daily_stata, 
              save_dta(daily_panel, here_rel("data", "derived_data", "daily_panel.dta")),
              format = "file"),
-  tar_target(data_stata_website, 
+  tar_target(data_daily_stata_website, 
              save_dta(daily_panel, here_rel("analysis", "data", "daily_panel.dta")),
              format = "file"),
-  tar_target(data_csv, 
+  tar_target(data_daily_csv, 
              save_csv(daily_panel, here_rel("data", "derived_data", "daily_panel.csv")),
              format = "file"),
-  tar_target(data_csv_website, 
+  tar_target(data_daily_csv_website, 
              save_csv(daily_panel, here_rel("analysis", "data", "daily_panel.csv")),
              format = "file"),
-  tar_target(data_rds, 
+  tar_target(data_daily_rds, 
              save_r(daily_panel, here_rel("data", "derived_data", "daily_panel.rds")),
              format = "file"),
-  tar_target(data_rds_website, 
+  tar_target(data_daily_rds_website, 
              save_r(daily_panel, here_rel("analysis", "data", "daily_panel.rds")),
              format = "file"),
+  
+  ### Weekly ----
+  tar_target(data_weekly_stata, 
+             save_dta(weekly_panel, here_rel("data", "derived_data", "weekly_panel.dta")),
+             format = "file"),
+  tar_target(data_weekly_stata_website, 
+             save_dta(weekly_panel, here_rel("analysis", "data", "weekly_panel.dta")),
[…]

### 2023-02-14 analysis/analysis_h3.qmd
Add peace science presentation (super belatedly)

@@
 ---
-title: "3: Civil society, derogations, and economic policies"
+title: "H<sub>3</sub>: Civil society, derogations, and economic policies"
 format:
@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
+                      fig.width = 6, fig.height = (6 * 0.618),
+                      out.width = "80%", collapse = TRUE,
+                      dev = "png", dev.args = list(type = "cairo-png"))
+
+options(digits = 3, width = 120,
+        dplyr.summarise.inform = FALSE,
+        knitr.kable.NA = "")
+```
+
+```{r load-libraries, warning=FALSE, message=FALSE}
+library(tidyverse)
+library(targets)
+library(broom)
+library(modelsummary)
+library(kableExtra)
+library(ordinal)
+library(brms)
+library(marginaleffects)
+library(tidybayes)
+library(ggdist)
+library(bayesplot)
+library(patchwork)
+library(scales)
+
+# Bayes stuff
+BAYES_SEED <- 1234
+options(mc.cores = 4,
+        brms.backend = "cmdstanr")
+
+# Generated via random.org
+set.seed(2385)
+
+# Load targets 
+withr::with_dir(here::here(), {
+  # Final data
+  tar_load(daily_panel)
+  tar_load(weekly_panel)
+  tar_load(year_week_lookup)
+
+  # Models
+  tar_load(m_h3)
+  tar_load(m_coef_gof)
+
+  # Plotting functions
+  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+})
+
+# Tell bayesplot to use the Lakota palette for things like pp_check()
+bayesplot::color_scheme_set(clrs[c(1, 2, 4, 5, 7, 8)])
+
+weekly_panel <- weekly_panel %>% 
+  mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
+                list(z = ~as.numeric(scale(.)))))
+```
+
+# Model and priors
+
+## Official formal model
+
+$$
+\begin{aligned}
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{country} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
+&\ \beta_2\ \text{New cases}_{i}\ + \beta_3\ \text{Cumulative cases}_{i}\ + \\
+&\ \beta_4\ \text{New deaths}_{i}\ + \beta_5\ \text{Cumulative deaths}_{i}\ + \\
+&\ \beta_6\ \text{Rule of law index}_{i}\ + \beta_7\ \text{Civil liberties index}_{i}\ + \\
+&\ \beta_8\ \text{Core civil society index}_{i}\ + \beta_9\ \text{Week number}_{i} \\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
+\\
+\bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{1 \dots 9} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
+\end{aligned}
+$$
+
+## Priors
+
+```{r fig.width=6, fig.height=2.5}
+p1 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
+                fill = clrs[2]) +
+  xlim(c(-20, 20)) +
+  annotate(geom = "label", x = 0, y = 0.02, label = "Student t(1, 0, 3)") +
+  labs(x = "α and βs") +
+  theme_pandem(prior = TRUE)
+
+p2 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~dcauchy(., 0, 1), 
+                fill = clrs[4]) +
+  xlim(c(0, 10)) +
+  annotate(geom = "label", x = 5, y = 0.063, label = "Cauchy(0, 1)") +
+  labs(x = "σ") +
+  theme_pandem(prior = TRUE)
+
+p1 | p2
+```
+
+## Prior simulation
+
+```{r}
+#| code-fold: show
+logit_priors <- c(prior(student_t(1, 0, 3), class = Intercept),
+                    prior(student_t(1, 0, 3), class = b),
+                    prior(cauchy(0, 1), class = sd, lb = 0))
+
+model_h3_prior_only <- brm(
+  bf(e1_income_support_bin ~ derogation_ineffect + 
+       new_cases_z + cumulative_cases_z + new_deaths_z + cumulative_deaths_z + 
+       v2x_rule + v2x_civlib + v2xcs_ccsi + year_week_num + (1 | country_name)), 
+  data = weekly_panel,
+  family = "bernoulli",
+  prior = logit_priors,
+  sample_prior = "only",
+  chains = 4, seed = BAYES_SEED, iter = 5000, refresh = 0,
+  threads = threading(2)  # Two CPUs per chain to speed things up
+)
+model_h3_prior_only
+```
+
+```{r}
+datagrid(model = model_h3_prior_only,
+         year_week_num = 1:69,
+         derogation_ineffect = 0:1) %>% 
+  add_epred_draws(model_h3_prior_only, ndraws = 100) %>% 
+  left_join(year_week_lookup, by = "year_week_num") %>% 
+  mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                      levels = 0:1,
+                                      labels = c("No", "Yes"),
+                                      ordered = TRUE)) %>% 
+  ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
+  geom_line(aes(group = paste(.draw, derogation_ineffect)), alpha = 0.5, size = 0.5) +
+  scale_color_manual(values = c(clrs[1], clrs[8])) +
+  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+  scale_y_continuous(labels = label_percent()) +
+  labs(x = NULL, y = "Predicted probability", 
+       color = "Derogation in effect",
+       title = "Income support (priors only)") +
+  theme_pandem()
+```
+
+# Posterior
+
+```{r tbl-h1, cache=TRUE, warning=FALSE}
+modelsummary(pull(m_h3, model, name = nice),
+             estimate = "{estimate}",
+             statistic = "conf.int",
+             fmt = 2, coef_map = m_coef_gof$coef_map, 
+             gof_map = m_coef_gof$gof_map,
+             metrics = c("none"))
+```
+
+# Model evaluation
+
+```{r}
+prelim_trace <- function(model) {
+  model %>% 
+    gather_draws(b_Intercept, b_derogation_ineffect, b_year_week_num, sd_country_name__Intercept) %>% 
+    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
+    geom_line(size = 0.1) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_trank <- function(model) {
+  model %>% 
+    gather_draws(b_Intercept, b_derogation_ineffect, b_year_week_num, sd_country_name__Intercept) %>% 
+    group_by(.variable) %>% 
+    mutate(draw_rank = rank(.value)) %>% 
+    ggplot(aes(x = draw_rank, color = factor(.chain))) +
+    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_pp <- function(model) {
+  pp_check(model, ndraws = 100, type = "bars")
+}
+```
+
+::: {.panel-tabset}
+
+### `r m_h3$nice[[1]]`
+
+```{r}
+prelim_trace(m_h3$model[[1]])
+```
+
+```{r}
+prelim_trank(m_h3$model[[1]])
+```
+
+```{r}
+prelim_pp(m_h3$model[[1]])
+```
+
+### `r m_h3$nice[[2]]`
+
+```{r}
+prelim_trace(m_h3$model[[2]])
+```
+
+```{r}
+prelim_trank(m_h3$model[[2]])
+```
+
+```{r}
+prelim_pp(m_h3$model[[2]])
+```
+
+:::
+
+# Interpretation and marginal effects
+
+## Predictions
+
+### Derogations and time
+
+```{r fig.width=8, fig.height=4, out.width="100%"}
+h3_models_plots <- m_h3 %>% 
+  mutate(pred_plot = map2(model, nice, ~{
+    datagrid(model = .x,
+             year_week_num = 1:69,
+             derogation_ineffect = 0:1) %>% 
+      add_epred_draws(.x) %>% 
+      left_join(year_week_lookup, by = "year_week_num") %>% 
+      mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                          levels = 0:1,
+                                      labels = c("No", "Yes"),
+                                      ordered = TRUE)) %>% 
+      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
+      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+      scale_color_manual(values = c(clrs[1], clrs[8])) +
+      scale_fill_manual(values = c(clrs[1], clrs[8])) +
+      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+      scale_y_continuous(labels = label_percent()) +
+      labs(x = NULL, y = "Predicted probability", 
+           color = "Derogation in effect", fill = "Derogation in effect",
+           title = .y) +
+      theme_pandem()
+  }))
+
+wrap_plots(h3_models_plots$pred_plot)
+```
+
+### Derogations and core civil society index
+
+```{r fig.width=8, fig.height=4, out.width="100%"}
+h3_models_plots_cs <- m_h3 %>% 
+  mutate(pred_plot = map2(model, nice, ~{
+    datagrid(model = .x,
+             v2xcs_ccsi = seq(0, 1, by = 0.01),
+             derogation_ineffect = 0:1) %>% 
+      add_epred_draws(.x) %>% 
+      mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                          levels = 0:1,
+                                      labels = c("No", "Yes"),
+                                      ordered = TRUE)) %>% 
+      ggplot(aes(x = v2xcs_ccsi, y = .epred, color = derogation_ineffect)) +
+      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+      scale_color_manual(values = c(clrs[1], clrs[8])) +
+      scale_fill_manual(values = c(clrs[1], clrs[8])) +
+      scale_y_continuous(labels = label_percent()) +
+      labs(x = "Core civil society index", y = "Predicted probability", 
+           color = "Derogation in effect", fill = "Derogation in effect",
+           title = .y) +
+      theme_pandem()
+  }))
+
+wrap_plots(h3_models_plots_cs$pred_plot)
+```
+
+## Marginal effects
+
+### Derogations and time
+
+```{r fig.width=8, fig.height=4, out.width="100%"}
+h3_models_mfx <- m_h3 %>% 
+  mutate(mfx = map(model, ~{
+    .x %>% 
+      marginaleffects(newdata = datagrid(year_week_num = 1:69),
+                      variables = "derogation_ineffect",
+                      type = "response") |> 
+      posteriordraws()
+  })) %>% 
+  mutate(mfx_plot = map2(mfx, nice, ~{
+    .x %>% 
+      left_join(year_week_lookup, by = "year_week_num") %>% 
+      ggplot(aes(x = year_week_day, y = draw * 100)) +
+      stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
+      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+      labs(x = NULL,
+           y = "Percentage point change\nin probability of policy",
+           title = .y) +
+      theme_pandem()
+  }))
+
+wrap_plots(h3_models_mfx$mfx_plot)
+```
+
+### Derogations and core civil society index
+
+```{r fig.width=8, fig.height=4, out.width="100%"}
+h3_models_mfx_cs <- m_h3 %>% 
+  mutate(mfx = map(model, ~{
+    .x %>% 
+      marginaleffects(newdata = datagrid(v2xcs_ccsi = seq(0, 1, by = 0.01)),
+                      variables = "derogation_ineffect",
+                      type = "response") |> 
+      posteriordraws()
+  })) %>% 
+  mutate(mfx_plot = map2(mfx, nice, ~{
+    .x %>% 
+      ggplot(aes(x = v2xcs_ccsi, y = draw * 100)) +
+      stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
+      labs(x = "Core civil society index",
+           y = "Percentage point change\nin probability of policy",
+           title = .y) +
+      theme_pandem()
+  }))
+
+wrap_plots(h3_models_mfx_cs$mfx_plot)
+```

### 2023-02-14 analysis/hypotheses.qmd
Add peace science presentation (super belatedly)

@@
 
-## 1: Formal derogations and pandemic-related human rights violations
+## H~1~: Formal derogations and pandemic-related human rights violations
 
@@
 
-## 2: Civil society repression and pandemic-related human rights violations
+## H~2~: Civil society repression and pandemic-related human rights violations
 
@@
 
-## 3: Civil society, derogations, and pandemic-related economic policies
+## H~3~: Civil society, derogations, and pandemic-related economic policies
 

### 2023-02-14 analysis/presentations/2022-11-04_peace-science.qmd
Add peace science presentation (super belatedly)

---
title: "Pandemic Pass"
subtitle: "Treaty Derogations and<br>Human Rights Practices During COVID-19"
author:
  - name: Suparna Chaudhry
    affiliation: Lewis and Clark College
  - name: Audrey Comstock
    affiliation: Arizona State University
  - name: Andrew Heiss
    affiliation: Georgia State University
date: "November 4, 2022"
date-format: long
output: revealjs
format:
  revealjs: 
    slide-number: true
    chalkboard:
      buttons: true
    preview-links: auto
    footer: "Chaudhry, Comstock, and Heiss&emsp;|&emsp;Pandemic Pass"
    # include-before-body: testing.html
    # revealjs-plugins:
    #   - simplemenu
    theme: [simple, custom.scss]
    title-slide-attributes:
      data-background-color: "#c06636"
# resources:
  # - slides.pdf
---

```{r include=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(scales)
library(marginaleffects)
library(patchwork)
library(sf)

knitr::opts_chunk$set(fig.retina = 3, collapse = TRUE, echo = FALSE)
options(digits = 3, width = 75)

# Load targets 
withr::with_dir(here::here(), {
  tar_load(year_week_lookup)
  tar_load(civicus_map_data)

  # Models
  tar_load(m_prelim_derog)
  tar_load(m_h1)
  tar_load(m_h2)
  tar_load(m_h3)

  # Plotting functions
  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
})

clrs_lt <- colorspace::lighten(clrs, 0.9)
```

```{r make-plots, include=FALSE, cache=TRUE}
preliminary_models_plots <- m_prelim_derog %>% 
  filter(y != "c8_intl_travel_bin") %>% 
  mutate(pred_plot = map2(model, nice, ~{
    datagrid(model = .x,
             year_week_num = 1:69,
             derogation_ineffect = 0:1) %>% 
      add_epred_draws(.x) %>% 
      left_join(year_week_lookup, by = "year_week_num") %>% 
      mutate(derogation_ineffect = factor(derogation_ineffect, 
                                          levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_fill_manual(values = c(clrs[1], clrs[8])) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      scale_y_continuous(labels = label_percent()) +
      labs(x = NULL, y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  }))

h1_models_plots <- m_h1 %>% 
  mutate(pred_plot = pmap(list(model, nice, family), ~{
    if (..3 == "cumulative") {
      datagrid(model = ..1,
               year_week_num = 1:69,
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        left_join(year_week_lookup, by = "year_week_num") %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No derogation", "Derogation in effect"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        scale_y_continuous(labels = label_percent()) +
        labs(x = NULL, y = "Predicted probability", 
             color = "Severity", fill = "Severity",
             title = ..2) +
        facet_wrap(vars(derogation_ineffect)) +
        theme_pandem()
    } else {
      datagrid(model = ..1,
               year_week_num = 1:69,
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        left_join(year_week_lookup, by = "year_week_num") %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No", "Yes"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
        scale_color_manual(values = c(clrs[1], clrs[8])) +
        scale_fill_manual(values = c(clrs[1], clrs[8])) +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        scale_y_continuous(labels = label_percent()) +
        labs(x = NULL, y = "Predicted probability", 
             color = "Derogation in effect", fill = "Derogation in effect",
             title = ..2) +
        theme_pandem()
    }
  }))

h2_models_plots <- m_h2 %>% 
  mutate(pred_plot = pmap(list(model, nice, family), ~{
    if (..3 == "cumulative") {
      datagrid(model = ..1,
               v2csreprss = seq(-2.5, 3, by = 0.1),
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No derogation", "Derogation in effect"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = v2csreprss, y = .epred, color = .category)) +
        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_y_continuous(labels = label_percent()) +
        labs(x = "Civil society repression", y = "Predicted probability", 
             color = "Severity", fill = "Severity",
             title = ..2) +
        facet_wrap(vars(derogation_ineffect)) +
        theme_pandem()
    } else {
      datagrid(model = ..1,
               v2csreprss = seq(-2.5, 3, by = 0.1),
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No", "Yes"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = v2csreprss, y = .epred, color = derogation_ineffect)) +
        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
        scale_color_manual(values = c(clrs[1], clrs[8])) +
        scale_fill_manual(values = c(clrs[1], clrs[8])) +
        scale_y_continuous(labels = label_percent()) +
        labs(x = "Civil society repression", y = "Predicted probability", 
             color = "Derogation in effect", fill = "Derogation in effect",
             title = ..2) +
        theme_pandem()
    }
  }))

h3_models_plots_cs <- m_h3 %>% 
  mutate(pred_plot = map2(model, nice, ~{
    datagrid(model = .x,
             v2xcs_ccsi = seq(0, 1, by = 0.01),
             derogation_ineffect = 0:1) %>% 
      add_epred_draws(.x) %>% 
      mutate(derogation_ineffect = factor(derogation_ineffect, 
                                          levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
      ggplot(aes(x = v2xcs_ccsi, y = .epred, color = derogation_ineffect)) +
      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_fill_manual(values = c(clrs[1], clrs[8])) +
      scale_y_continuous(labels = label_percent()) +
      labs(x = "Core civil society index", y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  }))
```

# Question {background-color="`r clrs[3]`" .white}

##

![](img/covid-headlines@3x.png){fig-align="center" width="100%"}

## Emergencies and international law {.clrs-3}

### Derogations

Provisions in international law that authorize states to temporarily suspend treaty obligations

. . .

Many states do not formally derogate and just implement emergency measures

## ICCPR derogations during COVID-19 {.clrs-3}

::: {.smaller}

| Type of response                                              |   N |
|:--------------------------------------------------------------|----:|
| Derogated as intended                                         |  13 |
| Derogated; measures not temporary, proportional, or necessary |   5 |
| Emergency declared; no formal derogation                      |  74 |
| No emergency; no formal derogation                            |  28 |
| Has not ratified ICCPR                                        |   8 |

:::

## Research questions {.clrs-3}

::: {.incremental}
- Were derogating states less likely to abuse human rights as they implemented emergency measures for COVID-19?
- Can civil society organizations (CSOs) mitigate the adverse effects of these emergency measures?
:::

## Preview of findings {.clrs-3}

Derogating states are less likely to abuse emergency measures, have discriminatory policies, violate non-derogable rights, and impose measures without time limits

. . .

Permissive civil society environment increases likelihood of debt relief policies

. . .

Civil society repression increases the probability of abusive enforcement and discriminatory policies; derogations mitigate that probability

## Significance {.clrs-3}

**International law and human rights**

International treaty provisions designed to safeguard rights may end up worsening conditions when there is a lack of review from the international community

. . .

**Democratic erosion and human rights**

Emergencies are important for understanding the processes of autocratization and democratic backsliding

# Theory and hypotheses {background-color="`r clrs[5]`" .white}

## UN requirements for public health emergency measures {.clrs-5}

Suspensions of rights must be:

- Legal
- Necessary
- Proportional
- Non-discriminatory

## Derogations during COVID-19 {.clrs-5}

```{r prelim-plots}
#| echo: false
#| fig-width: 9
#| fig-height: 5
#| out-width: 100%
#| fig-align: center
wrap_plots(preliminary_models_plots$pred_plot)
```

## Pandemic as excuse {.clrs-5}

If states implement emergency measures without regard for international law, the implementation may be problematic and the process to restore rights may not be followed

- Restrictions on travel or movement by minority or opposition groups
- Limits on public assembly and protests
- Deployment of surveillance systems that impede privacy

## H~1~: Derogations & human rights {.clrs-5}

If states formally derogate from international treaty provisions, then they are less likely to:

a. Use discriminatory policy measures
b. Violate non-derogable rights
c. Impose measures with no time limits

## Civil society and human rights {.clrs-5}

Civil society helps mitigate the impact of crises, emergencies, and natural disasters

:::: {.columns}

::: {.column width="50%" .smallest}
During COVID-19, civil society:

- Broadened range of voices in policymaking process
- Pursued civic initiatives to deal with impact of economic shutdown, product shortages, and spikes in unemployment
- Assisted with public health initiatives
:::

::: {.column width="50%" .smaller}
![](img/image5.png){fig-align="center" width="100%"}
:::

::::

## H~2~: Civil society & supportive policies {.clrs-5}

If civil society organizations are involved in the policy making process, then states are more likely to have supportive economic policies

## Pandemic as excuse {.clrs-5}

Some states used emergency measures as an excuse to limit democratic participation, dissent, and challenges to economic and political rule

![](img/covid-excuse@3x.png){fig-align="center" width="100%"}

## Civil society repression {.clrs-5}

```{r civicus-map}
#| echo: false
#| fig-width: 6.5
#| fig-height: 3.1
#| out-width: 100%
#| fig-align: center
#| fig-cap: "2020 CIVICUS Monitor civic space ratings"
ggplot() +
  geom_sf(data = civicus_map_data, aes(fill = fct_rev(category)), size = 0.15, color = "black") +
  coord_sf(crs = st_crs("ESRI:54030"), datum = NA) +  # Robinson
  scale_fill_manual(values = clrs[c(8, 7, 4, 2, 1)],
                    na.translate = FALSE, name = "Civic space") +
  theme_void(base_family = "Noto Sans") +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.7, "lines"))
```

## H~3~: Civil society & human rights {.clrs-5}

If states are repressive towards civil society organizations, then they are more likely to:

a. Violate non-derogable rights
b. Use discriminatory policy measures
c. Engage in abusive enforcement of emergency measures
d. Impose limits on media freedom

# Data and results {background-color="`r clrs[1]`" .white}

## Data {.clrs-1}

- Hand-coded derogations from the ICCPR
- Oxford's COVID-19 Government Response Tracker
- V-Dem's Pandemic Backsliding Project
- WHO's case and death counts

. . .

Daily data for 139 countries from March 11, 2020 to June 30, 2021

## Modeling approach {.clrs-1}

:::: {.columns}

::: {.column width="50%" .smallest}

**Outcomes**

- Discriminatory policies (H~1~, H~3~)
- Violation of non-derogable rights (H~1~, H~3~)
- Abusive enforcement (H~1~, H~3~)
- Measures without time limits (H~1~)
- Limiting media (H~3~)
- Income support (H~2~)
- Debt relief (H~2~)

:::

::: {.column width="50%" .smallest}
**Explanatory variables**

- Derogation in effect
- Civil society repression
- Core civil society index

**Other controls**

- COVID case counts
- COVID death counts
- Rule of law index
- Civil liberties index

:::

::::

## Modeling approach {.clrs-1}

Bayesian multilevel models with country random effects and weekly time trend

. . .

*(Logistic models for binary outcomes; ordered logistic models for outcomes with multiple categories)*

. . .

**Open to other ideas too!**

## H~1~: Derogations and human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smaller}
Substantial 2–3 percentage point difference in probability of violating non-derogable rights (P[∆ < 0] = 0.964)
:::

::: {.column width="70%"}
```{r include=FALSE, eval=FALSE}
qwer <- m_h1 %>% 
  filter(y == "pandem_ndrights") %>% 
  pull(model) %>% pluck(1) %>% 
  marginaleffects(newdata = datagrid(year_week_num = c(1, 69)),
                  variables = "derogation_ineffect",
                  type = "response") %>% 
  posteriordraws() %>% 
  left_join(year_week_lookup, by = "year_week_num")

qwer %>% 
  group_by(year_week_day) %>% 
  summarize(avg_effect = mean(draw), 
            prop_less_0 = sum(draw < 0) / n())

ggplot(qwer, aes(x = year_week_day, y = draw * 100)) +
  stat_lineribbon()
```

```{r h1-nonderog}
#| fig-width: 4.5
#| fig-height: 3
#| out-width: 100%
#| fig-align: center
h1_models_plots %>% 
  filter(y == "pandem_ndrights") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Violation of non-derogable rights")
```
:::

::::

## H~1~: Derogations & human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smaller}
Substantial ≈22 percentage point difference in probability of having time-limited measures (P[∆ ≠ 0] = 1)
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h1 %>% 
  filter(y == "pandem_nolimit") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(year_week_num = c(1)),
                  variables = "derogation_ineffect",
                  type = "response")
```

```{r h1-timelimit, message=FALSE}
#| fig-width: 5
#| fig-height: 3.65
#| out-width: 100%
#| fig-align: center
h1_models_plots %>% 
  filter(y == "pandem_nolimit") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Emergency measures with no time limits") +
  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y",
               guide = guide_axis(n.dodge = 2))
```
:::

::::

## H~1~: Derogations & human rights {.clrs-1}

### Null results

No derogation-based difference in predicted probability of having discriminatory policies or abusive enforcement.

## H~2~: Civil society & supportive policies {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smaller}
Both derogation status and civil society strength substantially increase the probability of a state providing COVID-related debt relief
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h3 %>% 
  filter(y == "e2_debt_relief_bin") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(v2xcs_ccsi = c(0, 0.5, 0.75, 1)),
              variables = "derogation_ineffect",
              type = "response")
```

```{r h3-incomesupport, message=FALSE}
#| fig-width: 4.5
#| fig-height: 3
#| out-width: 100%
#| fig-align: center
h3_models_plots_cs %>% 
  filter(y == "e2_debt_relief_bin") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Debt relief policies")
```
:::

::::

## H~3~: Civil society & human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smallest}
Substantial difference in probability of implementing discriminatory policies across all levels of civil society repression (P[∆&nbsp;≠&nbsp;0] = 1) for states that derogated

Better civil society associated with less discriminatory policy in derogating states
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h2 %>% 
  filter(y == "pandem_discrim") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(v2csreprss = c(-2, 0, 2)),
                  variables = "derogation_ineffect",
                  type = "response")
```

```{r h2-discrim}
#| fig-width: 5
#| fig-height: 3.65
#| out-width: 100%
#| fig-align: center
h2_models_plots %>% 
  filter(y == "pandem_discrim") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Discriminatory policies")
```
:::

::::

## H~3~: Civil society & human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smallest}
Substantial difference in probability of abusive enforcement of emergency measures for countries with worse civil society repression in derogating states
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h2 %>% 
  filter(y == "pandem_abusive") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(v2csreprss = c(-2, 0, 2)),
              variables = "derogation_ineffect",
              type = "response")
```

```{r h2-abusive}
#| fig-width: 5
#| fig-height: 3.65
#| out-width: 100%
#| fig-align: center
h2_models_plots %>% 
  filter(y == "pandem_abusive") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Abusive enforcement")
```
:::

::::

## H~3~: Civil society & human rights {.clrs-1}

### Null results

No derogation-based difference in predicted probability of violating non-derogable rights or limiting media across range of civil society repression

# Discussion {background-color="`r clrs[8]`" .white}

## Summary {.clrs-8}

Derogating states are less likely to abuse emergency measures, have discriminatory policies, violate non-derogable rights, and impose measures without time limits

. . .

Civil society repression increases the probability of abusive enforcement and discriminatory policies; derogations mitigate that probability

. . .

Permissive civil society environment increases likelihood of debt relief policies

## Broader implications {.clrs-8}

Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties

. . .

Some states use derogations as intended; others use them as an excuse to violate rights

. . .

Understanding how states treat derogations is important for understanding declining respect for human rights and democratic backsliding

### 2023-02-14 R/funs_data-cleaning.R
Add mapping stuff as targets

@@ -2,6 +2,8 @@ library(readxl)
 suppressPackageStartupMessages(library(lubridate))
 suppressPackageStartupMessages(library(clock))
 library(countrycode)
+library(jsonlite)
+suppressPackageStartupMessages(library(sf))
 
 clean_iccpr_who <- function(path) {
   who_regions <- tribble(
@@ -243,6 +245,67 @@ make_year_week_lookup <- function(weekly_final) {
   
   return(year_week_lookup)
 }
+
+load_world_map <- function(path) {
+  world_map <- read_sf(path) %>%
+    filter(ISO_A3 != "ATA")
+  
+  return(world_map)
+}
+
+# Civicus Monitor
+# We downloaded the standalone embeddable widget
+# (https://monitor.civicus.org/widgets/world/) as an HTML file with
+# `wget https://monitor.civicus.org/widgets/world/` and saved it as index_2021-03-19.html
+#
+# We then extracted the COUNTRIES_DATA variable embedded in a <script> tag
+# (xpath = /html/body/script[5]), which is JSON-ish, but not quite. jsonlite
+# can't parse it for whatever reason, but some online JSON formatter and
+# validator could, so we ran it through that and saved the resulting clean file
+load_clean_civicus <- function(path) {
+  civicus_raw <- read_json(path) %>% as_tibble() %>% slice(1)
+  
+  civicus_lookup <- tribble(
+    ~value, ~category,
+    1, "Closed",
+    2, "Repressed",
+    3, "Obstructed",
+    4, "Narrowed",
+    5, "Open"
+  ) %>%
+    mutate(category = fct_inorder(category, ordered = TRUE))
+  
+  civicus_clean <- civicus_raw %>%
+    pivot_longer(everything(), names_to = "name", values_to = "value") %>%
+    mutate(value = map_chr(value, ~.)) %>%
+    mutate(value = parse_number(value, na = c("", "NA", "None"))) %>%
+    mutate(country_name = countrycode(name, "iso3c", "country.name",
+                                      custom_match = c("KOSOVO" = "XKK",
+                                                       "SVT" = "VCT")),
+           iso3c = countrycode(country_name, "country.name", "iso3c",
+                               custom_match = c("XKK" = "Kosovo",
+                                                "VCT" = "Saint Vincent and the Grenadines"))) %>%
[…]

### 2023-02-14 _targets.R
Really add mapping stuff as targets

@@ -45,8 +45,19 @@ list(
              here_rel("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v12",
                       "V-Dem-CY-Full+Others-v12.rds"),
              format = "file"),
+  tar_target(naturalearth_raw_file,
+             here_rel("data", "raw_data", "ne_110m_admin_0_countries",
+                      "ne_110m_admin_0_countries.shp"),
+             format = "file"),
+  tar_target(civicus_raw_file,
+             here_rel("data", "raw_data", "Civicus", "civicus_2021-03-19.json"),
+             format = "file"),
   
   ## Process and clean data ----
+  tar_target(world_map, load_world_map(naturalearth_raw_file)),
+  tar_target(civicus_clean, load_clean_civicus(civicus_raw_file)),
+  tar_target(civicus_map_data, create_civicus_map_data(civicus_clean, world_map)),
+  
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
   tar_target(pandem_clean, clean_pandem(pandem_raw_file)),

### 2023-02-14 R/funs_data-cleaning.R
Add ICCPR action history to final data

@@ -44,6 +44,19 @@ clean_iccpr_who <- function(path) {
   return(x)
 }
 
+clean_iccpr_action <- function(path) {
+  x <- read_excel(path) %>% 
+    janitor::clean_names() %>% 
+    rename(prior_iccpr_other_action = other_prior_iccpr_post_commitment_treaty_actions,
+           country_name = country) %>% 
+    mutate(across(starts_with("prior_"), ~case_when(
+      . == 1 ~ TRUE,
+      is.na(.) ~ FALSE
+    )))
+  
+  return(x)
+}
+
 clean_oxford <- function(path) {
   x <- tibble(
     # Get a list of all the sheets in the Excel file
@@ -210,9 +223,10 @@ create_daily_skeleton <- function(iccpr_who, oxford, pandem, vdem) {
   return(daily_skeleton)
 }
 
-make_final_data <- function(skeleton, iccpr_who, oxford, pandem, vdem) {
+make_final_data <- function(skeleton, iccpr_who, iccpr_action, oxford, pandem, vdem) {
   daily_final <- skeleton %>% 
     left_join(select(iccpr_who, -country_name), by = c("iso3", "day")) %>% 
+    left_join(select(iccpr_action, -country_name), by = c("iso3")) %>% 
     left_join(select(oxford, -country_name), by = c("iso3", "day")) %>% 
     left_join(select(pandem, -c(country_name, year)), by = c("iso3", "year_quarter")) %>% 
     left_join(select(vdem, -country_name), by = c("iso3", "year"))
@@ -222,7 +236,8 @@ make_final_data <- function(skeleton, iccpr_who, oxford, pandem, vdem) {
 
 make_weekly_data <- function(daily_final) {
   weekly_final <- daily_final %>% 
-    group_by(year_week, year_week_day, country_name, iso3, who_region, who_region_long) %>% 
+    group_by(year_week, year_week_day, country_name, iso3, who_region, who_region_long, 
+             prior_iccpr_derogations, prior_iccpr_other_action) %>% 
     summarize(across(c(new_cases, new_deaths), ~sum(., na.rm = TRUE)),
               across(c(iccpr_derogation_filed, derogation_start, 
                        derogation_ineffect, derogation_end), ~max(., na.rm = TRUE)),

### 2023-02-14 _targets.R

@@ -35,6 +35,9 @@ list(
   tar_target(iccpr_who_raw_file,
              here_rel("data", "raw_data", "ICCPR Derogation and WHO case data 1 3 2020 to 6 30 2021.xlsx"),
              format = "file"),
+  tar_target(iccpr_treaty_action_file,
+             here_rel("data", "raw_data", "ICCPR Treaty Action Variables.xlsx"),
+             format = "file"),
   tar_target(oxford_raw_file,
              here_rel("data", "raw_data", "Oxford Covid Response Data 1 3 2020 to 6 20 2021.xlsx"),
              format = "file"),
@@ -59,6 +62,7 @@ list(
   tar_target(civicus_map_data, create_civicus_map_data(civicus_clean, world_map)),
   
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
+  tar_target(iccpr_action_clean, clean_iccpr_action(iccpr_treaty_action_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
   tar_target(pandem_clean, clean_pandem(pandem_raw_file)),
   tar_target(vdem_clean, clean_vdem(vdem_raw_file)),
@@ -68,8 +72,8 @@ list(
                                    pandem_clean, vdem_clean)),
   
   tar_target(daily_panel, 
-             make_final_data(skeleton, iccpr_who_clean, oxford_clean, 
-                             pandem_clean, vdem_clean)),
+             make_final_data(skeleton, iccpr_who_clean, iccpr_action_clean, 
+                             oxford_clean, pandem_clean, vdem_clean)),
   
   tar_target(weekly_panel, make_weekly_data(daily_panel)),
   

### 2023-02-14 analysis/presentations/2022-11-04_peace-science.qmd
Remove old references to old plugins

@@
     footer: "Chaudhry, Comstock, and Heiss&emsp;|&emsp;Pandemic Pass"
-    # include-before-body: testing.html
-    # revealjs-plugins:
-    #   - simplemenu
     theme: [simple, custom.scss]

### 2023-02-14 R/funs_models.R
Remove H3

@@ -209,50 +209,3 @@ f_h2 <- function(panel) {
   
   return(h2_models)
 }
-
-
-f_h3 <- function(panel) {
-  BAYES_SEED <- 8412  # From random.org
-  
-  panel <- panel %>% 
-    mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
-                  list(z = ~as.numeric(scale(.)))))
-
-  never_filter <- function(x) {
-    panel %>% 
-      filter(.data[[x]] == 0)
-  }
-  
-  logit_priors <- c(prior(student_t(1, 0, 3), class = Intercept),
-                    prior(student_t(1, 0, 3), class = b),
-                    prior(cauchy(0, 1), class = sd, lb = 0))
-  
-  h3_model <- function(y, data) {
-    form <- glue::glue(y, " ~ derogation_ineffect + new_cases_z + cumulative_cases_z + ",
-                       "new_deaths_z + cumulative_deaths_z + ",
-                       "v2x_rule + v2x_civlib + v2xcs_ccsi + ",
-                       "year_week_num + (1 | country_name)") %>% 
-      as.formula()
-    
-    rlang::inject(
-      brm(
-        bf(!!form), 
-        data = data,
-        family = bernoulli(),
-        prior = logit_priors,
-        chains = 4, seed = BAYES_SEED,
-        threads = threading(2)  # Two CPUs per chain to speed things up
-      )
-    )
-  }
-  
-  h3_models <- tribble(
-    ~nice, ~y, ~never,
-    "Income Support", "e1_income_support_bin", "e1_income_support_never",
-    "Debt Relief", "e2_debt_relief_bin", "e2_debt_relief_never"
-  ) %>% 
-    mutate(data = map(never, ~never_filter(.))) %>% 
-    mutate(model = map2(y, data, ~h3_model(.x, .y)))
-  
-  return(h3_models)
-}

### 2023-02-14 _targets.R

@@ -124,8 +124,7 @@ list(
   tar_target(m_prelim_derog, f_prelim_derog(weekly_panel)),
   tar_target(m_h1, f_h1(weekly_panel)),
   tar_target(m_h2, f_h2(weekly_panel)),
-  tar_target(m_h3, f_h3(weekly_panel)),
-  
+
   tar_target(m_coef_gof, lst(coef_map, gof_map)),
   
   ## Graphics ----

### 2023-02-14 analysis/analysis_prelim.qmd
Add peace science presentation (super belatedly)
New preliminary models with ICCPR history indicators

@@
 ---
-title: "0: Derogations and policies"
+title: "Derogations and emergency policies"
 format:
@@
 model_prelim_prior_only <- brm(
-  bf(c7_internal_movement_bin ~ derogation_ineffect + new_cases_z + cumulative_cases_z + 
-       new_deaths_z + cumulative_deaths_z + year_week_num + (1 | country_name)), 
+  bf(c7_internal_movement_bin ~ derogation_ineffect + 
+       new_cases_z + cumulative_cases_z + 
+       new_deaths_z + cumulative_deaths_z + 
+       prior_iccpr_derogations + prior_iccpr_other_action + 
+       year_week_num + (1 | country_name)), 
   data = filter(weekly_panel, c7_internal_movement_never == 0),
@@
   ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
-  geom_line(aes(group = paste(.draw, derogation_ineffect)), alpha = 0.5, size = 0.5) +
+  geom_line(aes(group = paste(.draw, derogation_ineffect)), alpha = 0.5, linewidth = 0.5) +
   scale_color_manual(values = c(clrs[1], clrs[8])) +
@@
     ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
-    geom_line(size = 0.1) +
+    geom_line(linewidth = 0.1) +
     scale_color_viridis_d(option = "rocket", end = 0.85) +

### 2023-02-14 R/funs_models.R

@@ -6,6 +6,8 @@ coef_map <- c(
   "b_cumulative_cases_z" = "Cumulative cases (standardized)",
   "b_new_deaths_z" = "New deaths (standardized)",
   "b_cumulative_deaths_z" = "Cumulative deaths (standardized)",
+  "b_prior_iccpr_derogationsTRUE" = "Past ICCPR derogation",
+  "b_prior_iccpr_other_actionTRUE" = "Past ICCPR action",
   "b_year_week_num" = "Year-week",
   "b_v2x_rule" = "Rule of law",
   "b_v2x_civlib" = "Civil liberties",
@@ -63,7 +65,9 @@ f_prelim_derog <- function(panel) {
   
   prelim_derog_model <- function(y, data) {
     form <- glue::glue(y, " ~ derogation_ineffect + new_cases_z + cumulative_cases_z + ",
-                       "new_deaths_z + cumulative_deaths_z + year_week_num + (1 | country_name)") %>% 
+                       "new_deaths_z + cumulative_deaths_z + ",
+                       "prior_iccpr_derogations + prior_iccpr_other_action + ",
+                       "year_week_num + (1 | country_name)") %>% 
       as.formula()
     
     # Use rlang::inject() to evaluate the formula object before running the model

### 2023-02-15 _targets.R
Changes related to reorganization

@@ -131,11 +131,11 @@ list(
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
   
   ## Analysis notebook ----
-  tar_quarto(analysis_notebook, path = "analysis"),
+  tar_quarto(website, path = "."),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
   tar_target(deploy_notebook, {
     # Force a dependency
-    analysis_notebook
+    website
     # Run the deploy script
     if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
   })

### 2023-02-15 deploy.sh

@@ -4,4 +4,4 @@ REMOTE_DEST=$REMOTE_HOST:$REMOTE_DIR
 
 echo "Uploading new changes to remote server..."
 echo
-rsync -crvP --delete analysis/_site/ $REMOTE_DEST
+rsync -crvP --exclude '*_cache' --delete _site/ $REMOTE_DEST

### 2023-02-15 analysis/analysis_h1.qmd
Add peace science presentation (super belatedly)
Control for prior history in H1 and H2

@@
 ---
-title: "1: Derogations and human rights"
+title: "H<sub>1</sub>: Derogations and human rights"
 format:
@@
 &\ \beta_4\ \text{New deaths}_{i}\ + \beta_5\ \text{Cumulative deaths}_{i}\ + \\
-&\ \beta_6\ \text{Rule of law index}_{i}\ + \beta_7\ \text{Civil liberties index}_{i}\ + \\
-&\ \beta_8\ \text{Core civil society index}_{i}\ + \beta_9\ \text{Week number}_{i} \\
+&\ \beta_6\ \text{Past ICCPR derogation}_{i}\ + \beta_7\ \text{Past ICCPR action}_{i}\ + \\
+&\ \beta_8\ \text{Rule of law index}_{i}\ + \beta_9\ \text{Civil liberties index}_{i}\ + \\
+&\ \beta_{10}\ \text{Core civil society index}_{i}\ + \beta_{11}\ \text{Week number}_{i} \\
 \alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
@@
 \bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\beta_{1 \dots 9} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{1 \dots 11} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
 \sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
@@
 model_h1_prior_only <- brm(
-  bf(pandem_nolimit ~ derogation_ineffect + new_cases_z + cumulative_cases_z + 
-       new_deaths_z + cumulative_deaths_z + v2x_rule + v2x_civlib + v2xcs_ccsi +
+  bf(pandem_nolimit ~ derogation_ineffect + 
+       new_cases_z + cumulative_cases_z + 
+       new_deaths_z + cumulative_deaths_z + 
+       prior_iccpr_derogations + prior_iccpr_other_action +
+       v2x_rule + v2x_civlib + v2xcs_ccsi +
        year_week_num + (1 | country_name)), 
@@
   ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
-  geom_line(aes(group = paste(.draw, .category, derogation_ineffect)), alpha = 0.5, size = 0.5) +
+  geom_line(aes(group = paste(.draw, .category, derogation_ineffect)), alpha = 0.5, linewidth = 0.5) +
   scale_color_manual(values = c(clrs[c(2, 4, 6)])) +
@@
                derogation_ineffect = 0:1) %>% 
-        add_epred_draws(.x) %>% 
+        add_epred_draws(..1) %>% 
         left_join(year_week_lookup, by = "year_week_num") %>% 

### 2023-02-15 analysis/analysis_h2.qmd
Add peace science presentation (super belatedly)
Control for prior history in H1 and H2

--- 2022-07-24 analysis/analysis_h2.qmd
@@
 ---
-title: "2: Civil society and human rights"
+title: "H<sub>2</sub>: Civil society and human rights"
 format:
@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
+                      fig.width = 6, fig.height = (6 * 0.618),
+                      out.width = "80%", collapse = TRUE,
+                      dev = "png", dev.args = list(type = "cairo-png"))
+
+options(digits = 3, width = 120,
+        dplyr.summarise.inform = FALSE,
+        knitr.kable.NA = "")
+```
+
+```{r load-libraries, warning=FALSE, message=FALSE}
+library(tidyverse)
+library(targets)
+library(broom)
+library(modelsummary)
+library(kableExtra)
+library(ordinal)
+library(brms)
+library(marginaleffects)
+library(tidybayes)
+library(ggdist)
+library(bayesplot)
+library(patchwork)
+library(scales)
+
+# Bayes stuff
+BAYES_SEED <- 1234
+options(mc.cores = 4,
+        brms.backend = "cmdstanr")
+
+# Generated via random.org
+set.seed(2385)
+
+# Load targets
+withr::with_dir(here::here(), {
+  # Final data
+  tar_load(daily_panel)
+  tar_load(weekly_panel)
+  tar_load(year_week_lookup)
+
+  # Models
+  tar_load(m_h2)
+  tar_load(m_coef_gof)
+
+  # Plotting functions
+  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+})
+
+# Tell bayesplot to use the Lakota palette for things like pp_check()
+bayesplot::color_scheme_set(clrs[c(1, 2, 4, 5, 7, 8)])
+
+weekly_panel <- weekly_panel %>% 
+  mutate(across(c(new_cases, new_deaths, cumulative_cases, cumulative_deaths),
+                list(z = ~as.numeric(scale(.)))))
+```
+
+# Model and priors
+
+## Official formal model
+
+$$
+\begin{aligned}
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{region} [i]} + \\
+&\ \beta_1\ \text{Derogation in effect}_{i}\ + \beta_2\ \text{Civil society repression}_{i}\ +\\
+&\ \beta_3\ {(\text{Derogation} \times \text{Civil society repression})}_{i}\ + \\
+&\ \beta_4\ \text{New cases}_{i}\ + \beta_5\ \text{Cumulative cases}_{i}\ + \\
+&\ \beta_6\ \text{New deaths}_{i}\ + \beta_7\ \text{Cumulative deaths}_{i}\ + \\
+&\ \beta_8\ \text{Past ICCPR derogation}_{i}\ + \beta_9\ \text{Past ICCPR action}_{i}\ + \\
+&\ \beta_{10}\ \text{Rule of law index}_{i}\ + \beta_{11}\ \text{Civil liberties index}_{i}\ + \\
+&\ \beta_{12}\ \text{Week number}_{i} \\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
+\\
+\bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{1 \dots 12} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
+\end{aligned}
+$$
+
+## Priors
+
+```{r fig.width=6, fig.height=2.5}
+p1 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
+                fill = clrs[2]) +
+  xlim(c(-20, 20)) +
+  annotate(geom = "label", x = 0, y = 0.02, label = "Student t(1, 0, 3)") +
+  labs(x = "α and βs") +
+  theme_pandem(prior = TRUE)
+
+p2 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~dcauchy(., 0, 1), 
+                fill = clrs[4]) +
+  xlim(c(0, 10)) +
+  annotate(geom = "label", x = 5, y = 0.063, label = "Cauchy(0, 1)") +
+  labs(x = "σ") +
+  theme_pandem(prior = TRUE)
+
+p1 | p2
+```
+
+## Prior simulation
+
+```{r}
+#| code-fold: show
+ologit_priors <- c(prior(student_t(1, 0, 3), class = Intercept),
+                   prior(student_t(1, 0, 3), class = b),
+                   prior(cauchy(0, 1), class = sd, lb = 0))
+
+model_h2_prior_only <- brm(
+  bf(pandem_abusive ~ derogation_ineffect*v2csreprss + 
+       new_cases_z + cumulative_cases_z + 
+       new_deaths_z + cumulative_deaths_z + 
+       prior_iccpr_derogations + prior_iccpr_other_action +
+       v2x_rule + v2x_civlib +
+       year_week_num + (1 | who_region)), 
+  data = weekly_panel,
+  family = "cumulative",
+  prior = ologit_priors,
+  sample_prior = "only",
+  chains = 4, seed = BAYES_SEED, iter = 5000, refresh = 0,
+  threads = threading(2)  # Two CPUs per chain to speed things up
+)
+model_h2_prior_only
+```
+
+```{r}
+datagrid(model = model_h2_prior_only,
+         year_week_num = 1:69,
+         derogation_ineffect = 0:1) %>% 
+  add_epred_draws(model_h2_prior_only, ndraws = 100) %>% 
+  left_join(year_week_lookup, by = "year_week_num") %>% 
+  mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                      levels = 0:1,
+                                      labels = c("No", "Yes"),
+                                      ordered = TRUE)) %>% 
+  ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
+  geom_line(aes(group = paste(.draw, .category, derogation_ineffect)), alpha = 0.5, linewidth = 0.5) +
+  scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+  scale_y_continuous(labels = label_percent()) +
+  labs(x = NULL, y = "Predicted probability", 
+       color = "Severity",
+       title = "Abusive enforcement (priors only)") +
+  theme_pandem()
+```
+
+# Posterior
+
+```{r tbl-h2, cache=TRUE, warning=FALSE}
+modelsummary(pull(m_h2, model, name = nice),
+             estimate = "{estimate}",
+             statistic = "conf.int",
+             fmt = 2, coef_map = m_coef_gof$coef_map, 
+             gof_map = m_coef_gof$gof_map,
+             metrics = c("none"))
+```
+
+# Model evaluation
+
+```{r}
+prelim_trace <- function(model) {
+  model %>% 
+    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2csreprss, sd_who_region__Intercept) %>% 
+    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
+    geom_line(linewidth = 0.1) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_trank <- function(model) {
+  model %>% 
+    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2csreprss, sd_who_region__Intercept) %>% 
+    group_by(.variable) %>% 
+    mutate(draw_rank = rank(.value)) %>% 
+    ggplot(aes(x = draw_rank, color = factor(.chain))) +
+    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+prelim_pp <- function(model) {
+  pp_check(model, ndraws = 100, type = "bars")
+}
+```
+
+::: {.panel-tabset}
+
+### `r m_h2$nice[[1]]`
+
+```{r}
+prelim_trace(m_h2$model[[1]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[1]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[1]])
+```
+
+### `r m_h2$nice[[2]]`
+
+```{r}
+prelim_trace(m_h2$model[[2]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[2]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[2]])
+```
+
+### `r m_h2$nice[[3]]`
+
+```{r}
+prelim_trace(m_h2$model[[3]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[3]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[3]])
+```
+
+### `r m_h2$nice[[4]]`
+
+```{r}
+prelim_trace(m_h2$model[[4]])
+```
+
+```{r}
+prelim_trank(m_h2$model[[4]])
+```
+
+```{r}
+prelim_pp(m_h2$model[[4]])
+```
+
+:::
+
+# Interpretation and marginal effects
+
+## Predictions
+
+### Across time
+
+::: {.column-page-right}
+
+```{r fig.width=13, fig.height=6, out.width="100%"}
+h2_models_plots <- m_h2 %>% 
+  mutate(pred_plot = pmap(list(model, nice, family), ~{
+    if (..3 == "cumulative") {
+      datagrid(model = ..1,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(..1) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No derogation", "Derogation in effect"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
+        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = ..2) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      datagrid(model = ..1,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(..1) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No", "Yes"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
+        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_fill_manual(values = c(clrs[1], clrs[8])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = ..2) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h2_models_plots$pred_plot)
+```
+
+:::
+
+### Across civil society repression
+
+::: {.column-page-right}
+
+```{r fig.width=13, fig.height=6, out.width="100%"}
+h2_models_plots1 <- m_h2 %>% 
+  mutate(pred_plot = pmap(list(model, nice, family), ~{
+    if (..3 == "cumulative") {
+      datagrid(model = ..1,
+               v2csreprss = seq(-2.5, 3, by = 0.1),
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(..1) %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No derogation", "Derogation in effect"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = v2csreprss, y = .epred, color = .category)) +
+        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = "Civil society repression", y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = ..2) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      datagrid(model = ..1,
+               v2csreprss = seq(-2.5, 3, by = 0.1),
+               derogation_ineffect = 0:1) %>% 
+        add_epred_draws(..1) %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No", "Yes"),
+                                            ordered = TRUE)) %>% 
+        ggplot(aes(x = v2csreprss, y = .epred, color = derogation_ineffect)) +
+        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_fill_manual(values = c(clrs[1], clrs[8])) +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = "Civil society repression", y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = ..2) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h2_models_plots1$pred_plot)
+```
+
+:::
+
+## Marginal effects
+
+```{r}
+# TODO: Figure out how to work with ordered logit marginal effects
+```

### 2023-02-15 R/funs_models.R

@@ -126,6 +126,7 @@ f_h1 <- function(panel) {
   h1_model <- function(y, family, prior) {
     form <- glue::glue(y, " ~ derogation_ineffect + new_cases_z + cumulative_cases_z + ",
                        "new_deaths_z + cumulative_deaths_z + ", 
+                       "prior_iccpr_derogations + prior_iccpr_other_action + ",
                        "v2x_rule + v2x_civlib + v2xcs_ccsi + ",
                        "year_week_num + (1 | who_region)") %>% 
       as.formula()
@@ -183,6 +184,7 @@ f_h2 <- function(panel) {
     form <- glue::glue(y, " ~ derogation_ineffect*v2csreprss + ",
                        "new_cases_z + cumulative_cases_z + ",
                        "new_deaths_z + cumulative_deaths_z + ", 
+                       "prior_iccpr_derogations + prior_iccpr_other_action + ",
                        "v2x_rule + v2x_civlib + ",
                        "year_week_num + (1 | who_region)") %>% 
       as.formula()

### 2023-02-15 analysis/hypotheses.qmd
Adjust hypothesis count

@@
 
-We test three sets of hypotheses:
+We test two sets of hypotheses:
 
@@
 - **H~2d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom.
-
-## H~3~: Civil society, derogations, and pandemic-related economic policies
-
-- **H~3a~**: If civil society organizations are involved in the policy making process and consulted by the government), then states are more likely to have supportive economic policies during the pandemic.
-- **H~3b~**: If states formally derogate from their international treaty provisions, then states are more likely to have supportive economic policies during the pandemic.

### 2023-02-15 analysis/data-details.qmd
Add details about prior actions

@@
   datatable()
+```
+
+## Prior ICCPR actions
+
+To account for whether countries have engaged with ICCPR derogation procedures prior to the COVID-19 pandemic (signalling that they perhaps care about treaty obligations), we include two binary indicators for (1) whether the country has derogated from the ICCPR in the past and (2) whether the country has completed any other type of ICCPR-related actions.
+These are time-invariant and act more like overall indicator variables, like continent or region.
+
+```{r glimpse-prior-action}
+#| code-fold: false
+daily_panel %>% 
+  group_by(country_name, iso3) %>% 
+  summarize(across(starts_with("prior"), ~mean(.))) %>% 
+  ungroup() %>% 
+  glimpse()
 ```

### 2023-02-15 slides/2022-11-04_peace-science_files/libs/revealjs/plugin/reveal-simple-menu/simplemenu.js
Include supporting files for slides

@@ -1,0 +1,291 @@
+
+/*****************************************************************
+ * @author: Martijn De Jongh (Martino), martijn.de.jongh@gmail.com
+ * https://github.com/Martinomagnifico
+ *
+ * Simplemenu.js for Reveal.js 
+ * Version 1.1.2
+ * 
+ * @license 
+ * MIT licensed
+ *
+ * Thanks to:
+ *  - Hakim El Hattab, Reveal.js 
+ ******************************************************************/
+
+
+(function (global, factory) {
+  typeof exports === 'object' && typeof module !== 'undefined' ? module.exports = factory() :
+  typeof define === 'function' && define.amd ? define(factory) :
+  (global = typeof globalThis !== 'undefined' ? globalThis : global || self, global.Simplemenu = factory());
+})(this, (function () { 'use strict';
+
+  function _toConsumableArray(arr) {
+    return _arrayWithoutHoles(arr) || _iterableToArray(arr) || _unsupportedIterableToArray(arr) || _nonIterableSpread();
+  }
+
+  function _arrayWithoutHoles(arr) {
+    if (Array.isArray(arr)) return _arrayLikeToArray(arr);
+  }
+
+  function _iterableToArray(iter) {
+    if (typeof Symbol !== "undefined" && iter[Symbol.iterator] != null || iter["@@iterator"] != null) return Array.from(iter);
+  }
+
+  function _unsupportedIterableToArray(o, minLen) {
+    if (!o) return;
+    if (typeof o === "string") return _arrayLikeToArray(o, minLen);
+    var n = Object.prototype.toString.call(o).slice(8, -1);
+    if (n === "Object" && o.constructor) n = o.constructor.name;
+    if (n === "Map" || n === "Set") return Array.from(o);
+    if (n === "Arguments" || /^(?:Ui|I)nt(?:8|16|32)(?:Clamped)?Array$/.test(n)) return _arrayLikeToArray(o, minLen);
+  }
+
+  function _arrayLikeToArray(arr, len) {
+    if (len == null || len > arr.length) len = arr.length;
+
+    for (var i = 0, arr2 = new Array(len); i < len; i++) arr2[i] = arr[i];
+
+    return arr2;
+  }
+
+  function _nonIterableSpread() {
+    throw new TypeError("Invalid attempt to spread non-iterable instance.\nIn order to be iterable, non-array objects must have a [Symbol.iterator]() method.");
+  }
+
+  var Plugin = function Plugin() {
[…]

### 2023-03-01 manuscript/manuscript.qmd
Tons of edits and additions

---
title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
subtitle: "Research note"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  email: schaudhry@lclark.edu
- name: Audrey L. Comstock
  affiliation: Arizona State University
  email: audrey.comstock@asu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
abstract: |
  In an effort to combat the Covid-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations. However, the abuse of emergency derogations by many governments raises concerns about the long-lasting impact of these sweeping measures on human rights. Does a robust civil society mitigate the adverse effects arising from international treaty derogations? Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run. Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker and the CoronaNet COVID-19 Government Response Event, we first assess the variation in human rights responses among governments during the pandemic. We then assess the role of civil society in keeping governments accountable and mitigating the effects of policies that violate human rights. Our results hold important implications for understanding government compliance with international treaties and shedding light on how states of emergency may directly and indirectly lead to democratic backsliding.
date: February 27, 2023
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
bibliography: bib.bib
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
                      warning = FALSE, message = FALSE)

knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE)

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r libraries-data}
library(tidyverse)
library(scales)
library(lubridate)
library(tidybayes)
library(marginaleffects)
library(patchwork)
library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Generated via random.org
set.seed(8511)

# Load targets 
weekly_panel <- tar_read(weekly_panel) %>% 
  mutate(year_week_day = ymd(year_week_day))
tar_load(year_week_lookup)
tar_load(m_prelim_derog)

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
```

Combating a deadly virus like SARS-CoV-2 requires extraordinary public health measures. In an effort to fight the Covid-19 pandemic, governments across the world imposed restrictions on the freedom of movement, association, assembly and privacy. While countries are ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights principles. Some human rights treaties recognize states' needs for emergency measures when confronted with public health concerns and contain escape clauses that establish safeguards to prevent permanent erosion of these rights during emergencies. This temporary suspension of state compliance with its international treaty provisions occurs through a process known as derogations. However, the abuse of emergency measures during these derogations raises concerns about the long-lasting impact of these measures on human rights. Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].

Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that in May 2023 it would be ending the emergency measures instated at the beginning of the pandemic—a move that that has occurred in many other countries as well. With the lifting of these measures, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this period. What did derogations mean for compliance? Are derogating and non-derogating states different in terms of how they implemented emergency measures during Covid-19? Were derogating states more likely to exploit the provisions and to abuse human rights during Covid-19? These questions are important for the study of international law, compliance, and times of crises.

The violations of human rights during this period also raises questions about factors that can mitigate the adverse impacts of emergency measures. Existing literature on emergencies such as wars and natural disasters shows that civil society organizations (CSOs) play a pivotal role during these periods [@pellingdill; @cash2013]. They provide services and resources, and hold governments accountable to ensure inclusive policies during these emergencies. Can these groups also prevent or reduce the likelihood of human rights violations during public health emergencies? Does a robust civil society mitigate the adverse effects arising from international treaty derogations during the pandemic?

Typically, when international relations scholars write about human rights treaty compliance, they write of positive human rights recognition. Compliance is an adherence to the expectations and written standards within the treaties, often described as a "state of conformity or identity between an actor’s behavior and a specified rule" [@RaustialaSlaughter:2002, 539]. However, there are treaty qualifications that are legal addendums to commitment that alter or even limit compliance expectations. When states comply with these legal actions, they are in fact, limiting their obligations to human rights standards. In this research note, we focus on one such legal addendum—the treaty derogation. If and when states adhere to the written expectations of treaty derogations, they will be suspending human rights recognition during asset time of domestic emergency and crisis. If states are complying the derogation expectations, we should in fact expect a reduction in compliance with the overall treaty during that time and human rights standards. We explore how the potential decline in compliance may actually support human rights during the COVID-19 period since emergency mitigation strategies were targeting a reduction in disease outbreak. Human rights compliance literature generally finds that NGOs and civil society are key actors in facilitating treaty obligation actualization [e.g., @KeckSikkink:1998; @Merry:2009]. We explore what, if any, role civil society organizations had during this period.

Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, the CoronaNet Covid-19 Government Response Event, and United Nations treaty derogation data we first assess the variation in human rights responses among governments during the pandemic. We then analyze the effect of derogating (or not) on human rights practices during the pandemic and assess the role of civil society in mitigating the effects of rights-violating policies. Our main takeaways are that derogating states did largely what they signaled they would through the derogations. States that derogated and specified time limits to the emergency provisions generally stuck to the time limits and did implement the emergency practices they signaled they would. Additionally, we find that, on average, derogating states are less likely to abuse emergency measures. On the other hand, states that did not derogate from their international human rights obligations and still impose emergency measures are more likely impose discriminatory policies, violate non-derogable rights such right to life AND WHAT ELSE, and impose emergency measures without time limits. Explain this result more - ANDREW "Civil society repression increases probability of abusive enforcement and discriminatory policies; derogations mitigate that probability."

This research note makes three contributions. First, while there is a significant literature on compliance with international law, especially human rights treaties [e.g. @Simmons2009; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations. Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of the types of states that are likely to comply with these escape clauses. The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster typical of treaty derogations, making it easier to compare the long-term human rights effects of these derogations cross nationally.

Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Simmons2009; @Merry:2009]. But we know much less about how international treaty provisions intended to safeguard rights might actually worsen human rights outcomes. While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g. @Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
 EXPAND ON RESULT In this scenario, when international factors may not be enough to constrain state actions, assessing the effect of domestic-level factors such as a robust civil society that potentially can mitigate the adverse effects of derogations becomes essential. WE FIND.

Finally, studying states' compliance with their international human rights treaty obligations as well as the (unintended) effects of derogations is important given trends in democratic backsliding and declining respect for human rights. Nearly 75 percent of the world’s population now lives in a country that faced democratic deterioration last year [@freedomhouse2021]. The pandemic led to further backsliding with many countries around the world postponing elections, arresting protestors, and cracking down on citizens to consolidate power—all under the guise of public health measures. These periods of emergencies may be an important piece to understand the processes of autocratization and democratic backsliding since leaders often use emergency decrees to  allow states to aggrandize executive power [@LuhrmannRooney2021; @LevitskyZiblatt:2018; @Bermeo2016]. The temporary suspension of states' international legal obligation is one possible pathway that could lead to long-term rights violations and democratic erosion.

Below, we first explore the purpose of derogations during the pandemic. In doing so, we focus specifically on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime. Following this, we introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic. We then assess states' behavior towards civil society and these organizations' role in mitigating the adverse effects of emergency measures. We conclude with the implications' of our findings for the study of international law, times of crises, and the role of civil society.

# Derogations and international law

Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility in responding to events. These times of crises include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies. Derogations afford states the opportunity to adjust to these changing conditions [@HBHelferFariss]. Once states derogating from their international legal obligations, it also provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions. This information allows these actors—at least in principle—to challenge measures that are excessive, vague, or outlast the intended time frame of their implementation [@helfer2021].

While legal critical assessments of the use of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially led to the violation of non-derogable rights—rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that cannot be suspended even during emergencies. Research finds results that support this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013]. Even states that officially derogate from their international legal obligations are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].

These findings suggest that states may not be discerning the significance of derogable v/s. non-derogable rights status during emergencies. This also raises the question on the extent to which this distinction is a valuable one during the pandemic—which is a public health emergency different from other states of emerency normally brought on by war or internal political unrest. Overall, there is a general expectation that following derogations, we may observe some decline in respect for human rights.

## Derogations under COVID-19

Research finds that democracies, especially those with strong rule of law, are more likely to derogate [@HBHelferFariss]. The expectation is that states with strong democratic institutions need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.

A first question of interest before we examine the potential impact of derogations is what states are more likely to make them in the first place. Though not much research has focused on derogations in political science, most of it has focused on the determinants of submitting derogations to the UN rather than implications of their application. This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss]. The argument is that democratic states understand international law and international legal mechanisms to be more legitimate and recognize their importance more so than other states. Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important. Looking to current derogation practices, there is not a perfect overlap in the states that issued derogations in the past towards the ICCPR and those that did during the COVID-19 pandemic. Of course, since derogations are a type of action submitted after treaty commitment, the only states that would be submitting them are ones that already have ratified or otherwise committed to the ICCPR. As of 2023, there are 173 states party to the ICCPR. This means that most, but not all of the UN’s 193 members had access to submitting derogations during COVID-19. Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.

Some states new to derogations introducing them during the COVID-19 pandemic [@Comstock2023]. However, past treaty action behavior may help explain entering derogations during this time. To examine this first step, we code two variables 1) past ICCPR derogation behavior and 2) past ICCPR post-commitment action behavior. States receive a 1 if they engaged in this legal behavior before 2020 and 0 if they did not. This will allow us to empirically test for the first step of engagement before analyzing any impact COVID-19 era derogations had on different rights practices. There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019.

## Derogations from the ICCPR

In this research note, we look at states' derogations from the International Covenant on Civil and Political Rights (ICCPR) during the pandemic. ICCPR is arguably one of the most extensive human rights treaties and the exceptional circumstances brought on by the global pandemic have led to numerous extensive derogations to the ICCPR. The ICCPR was adopted by the UN General Assembly on December 19, 1966 and came into force on March 23, 1976. It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights. The ICCPR is one of only a few select treaties that requires states to report on the suspension or derogation of certain civil and political rights only under specific situations of "public emergency which threatens the life of the nation" [@iccpr]. The other treaties with this requirement include the European Convention Protection of Human Rights and Fundamental Freedoms and the American Convention on Human Rights. However, the ICCPR is the only one with global, rather than regional, membership and participation to track during the pandemic.

To legally derogate, government must first officially declare a state of emergency (Article 4). These derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory. However, it is worth noting that states cannot derogate from the entirety of the ICCPR. Seven rights as enshrined in the ICCPR are non-derogable: right to life, prohibit torture or cruel, inhumane punishment, prohibit slavery, provide the right to recognition as a person before the law, and protect freedom freedom of thought, conscience and religion. On the other hand, several ICCPR Articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).

Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination. However,not every country formally derogated—a number of countries (HOW MANY) implemented emergency measures without formally derogating from ICCPR. The impact of these emergencies may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020]. Below, we first theorize how derogations can protect against rights violations and the excesses of emergency measures, followed by the role of civil society in mitigating these violations.

# The impact of derogations on human rights outcomes

During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21). If states determine that emergency measures are justified without regard for international law, the processes for restoring rights at the end of the emergency may not be followed. As such, many states used the pandemic an as excuse to centralize authority and implemented measures that didn’t meet international legal standards. Many states imposed restrictions on travel or movement by minority or opposition groups, limiting public assembly and democratic protests, and deployed surveillance systems that impeded privacy. For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021]. In Bulgaria, checkpoint controls went into effect in neighborhoods largely composed of the Roma community [@bulgaria2020]. In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country's new citizenship law seen as discriminatory towards Muslims [@bbc2020].

If states determine that emergency measures are justified without regard for international law, then we expect that states which did not file for derogations to the ICCPR were less likely to meet the criteria for derogations—necessity, proportionality, non-discrimination and time delimited, and therefore more likely to impose measures that violated these criteria.

> **Hypothesis 1~a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
>
> **Hypothesis 1~b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
>
> **Hypothesis 1~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.

## Civil society and human rights

Given the divergent state responses during the pandemic, these differing responses can have lasting impacts on human rights trends. In the absence of adequate review from the international community during the pandemic, certain domestic-level factors may be necessary to mitigate the adverse effects of emergency measures. While both the theoretical and empirical literature has established that states may violate both derogable and non-derogable rights during emergencies, we have comparatively little evidence on factors that may safeguard against these rights violations during periods of derogation.[^exceptions]

There is a robust literature on how civil society is important for improving human rights outcomes [@KeckSikkink:1998; @RisseRoppSikkink; @landman2005; @HBTsutsui; @neumayer2005; @Jetschke2011]. Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run [@fair2017].

As such, we expect that a strong and vibrant civil society may help mitigate the adverse impact arising from derogations and emergency measures imposed by governments during the pandemic. Civil society organizations broaden the range of voices in the policymaking process and help amplify the concerns and interests of citizens [@murdie2014]. Civic initiatives have not only played a major role in countering the pandemic's severity, but have also pressured authorities into implementing better policies to deal with the impact of economic shutdowns, product shortages and spikes in unemployment. For instance, in Brazil, a broad coalition of CSOs lobbied the government to implement a relief fund, successfully securing a monthly stipend three times the amount initially suggested by the government [@carnegie2021, 5].

Government attitudes towards civil society organizations—whether collaboration, accommodation or repression—may also affect their respect (or lack thereof) for human rights during the pandemic. Since March 2020, many advocacy and rights-based organizations have highlighted how the pandemic disproportionately affects vulnerable and marginalized communities. These organizations also monitor and document human rights violations by the state, which can both provide information about a state's level of compliance with domestic and international human rights obligations as well can be used as evidence in court.

However, over the past three decades, more than 130 states have repressed civil society organizations [@chaudhry2022]. These trends have been further exacerbated during the pandemic. As a result, a number of advocacy and rights-based organizations have either shut down or significantly downsized their capabilities, hampering their ability to document and monitor rights violations by the state and its agents. Without such monitoring, states may be more likely violate non-derogable rights during the pandemic, as states know that organizations will be less likely to collect evidence, write briefs, provide legal aid to victims and give expert testimony at trials. As such, we expect that states will have greater leverage to impact non-derogable rights if there is insufficient evidence that can be presented to international monitoring institutions or regional courts such as the European Court of Human Rights.

> **Hypothesis 2~a~**: If states are repressive towards civil society organizations, then they are more likely to violate non-derogable rights.

Given reduced monitoring by rights-based organizations, states may also have greater leeway to pass discriminatory policy measures and engage in more abusive enforcement of these measures. This is particularly the case for state actions towards refugees and asylum-seekers. For instance, Greece has increasingly criminalized humanitarian assistance and in 2019, introduced burdensome rules for the registration of nongovernmental organizations working on migration [@Cosse:2022]. Once the pandemic began, with decreased oversight from CSOs owing to the above rules, Greek officials intercepted and turned back boats filled with asylum-seekers and also summarily returned asylum-seekers at the land border with Turkey. In doing so, they both violated the non-derogable principle of non-refoulement, as well as Article 7 of ICCPR is the person sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].

> **Hypothesis 2~b~**: If states are repressive towards civil society organizations, then they are more likely to use discriminatory policy measures during the pandemic.
> 
> **Hypothesis 2~c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.

Finally, this trend of state repression of CSOs and NGOs may also impact state behavior towards the media. States typically repress CSOs because they make democratic participation, particularly protests, more likely, highlight rights violations, and challenge economic policies—all of which may be seen as even more threatening during the pandemic [@MurdieBhasin2011; @BBCM2014; @KeckSikkink:1998; @RisseRoppSikkink; @Boulding2014; @chaudhry2022].

With many governments trying to centralize authority and decisionmaking, a free and independent media may pose challenges similar to those posed by CSOs. As such, if states are engaging in repression of CSOs during the pandemic, they may be more likely to repress the media as well. For instance, over the last five years, Hungary has increasingly cracked down on CSOs. During the pandemic, this extended towards the media as well as the country passed a new law which sought to punish anyone who published 'false information' on the pandemic with up to five years in jail [@brechenmacheretal2020, 3]. Similarly, Zimbabwe criminalized false reporting on the government's response to the pandemic and subsequently, used this law to crack down on media freedom. On the other hand, a vibrant civil society may help protect against limitations on media freedom. For instance, in Tunisia, the parliament withdrew a draft law restricting media freedom after it faced criticism from CSOs [@brechenmacheretal2020, 3].

> **Hypothesis 2~d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom during the pandemic.

# Data and methods

```{r extract-details}
info <- list(
  n_weeks = length(unique(weekly_panel$year_week)),
  n_countries = length(unique(weekly_panel$iso3)),
  n_rows = label_comma()(nrow(weekly_panel)),
  dates = map(list(
    start_date = min(weekly_panel$year_week_day),
    end_date = max(weekly_panel$year_week_day)
  ), ~list(date = ., 
           nice = format(., "%B %e, %Y"),
           year = year(.)))
)
```

While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, civil society, and state actions during the pandemic. We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl:data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023]. We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021. While the Oxford Government Response Tracker provides daily data on a variety of government actions, submitting formal treaty responses to the ICCPR is less likely to be as immediate of an action as implementing domestic emergency actions like stay at home orders. Accordingly, we collapse this daily data into weekly totals, resulting in a panel with `r info$n_rows` country-week observations.

## Outcome variables

We examine the relationship between formal treaty derogations and different human rights and policy outcomes in three stages (see @tbl:vars). First, to explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel. Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures. As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in a given week.

Next, to test our first hypothesis, we explore the relationship between formal treaty derogations on four different human rights and four policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures. Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations. We do not collapse these variables into binary indicators and instead work with them as ordered categories. Finally, we test our second hypothesis by exploring the more complex relationship treaty derogations, civil society regulations, and four policy outcomes from the PanDem project. We look at the same four policies as the first hypothesis with one exception: since civil society is often concerned with advocacy and information provision during public emergencies, we replace whether emergency measures have a time limit with an ordinal measure of whether limits were placed on media freedom as part of a state's pandemic response. Importantly, since these PanDem variables are all measured quarterly, while our main data is recorded weekly, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response each week.

## Explanatory variables

### Key explanatory variables

We use three different primary independent variables to explain changes in these policy and human rights outcomes. For our preliminary models and in models related to our first hypothesis, we rely on a binary indicator of whether a state was under ICCPR derogation conditions during a given week. We collected and coded this data from real time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023]. During the 15 months covered in this project, there were 115 formal derogations submitted mentioning 8 specific ICCPR articles; 15 derogations made no reference to any specific article. Of the derogations that indicated specific time limits, the average derogation duration was 41 days. We mark the start and end date of each derogation so that each country-week observation indicates whether a state has derogated.

For our second hypothesis, we use a measure of civil society repression from the Varieties of Democracy (V-Dem) project [@vdem-v12], derived from a latent Bayesian item response theory model of expert opinions, with values ranging from roughly −3 to 3 (where higher values represent less repression). This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti=civil society environment.

### Confounding variables

Finally, we include several other variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions. To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID cases and the count of new COVID deaths, as well as the cumulative count of cases and deaths. Since formally derogating from treaties can serve as a signal that countries take their obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR bureaucratic actions prior to the pandemic. Finally, we include several annual measures of human rights, democracy, and civil society: V-Dem's rule of law index, civil liberties index, and core civil society index.

## Modeling strategy

Our data is structured as a balanced time series cross-sectional (TSCS) panel that has a natural hierarchical structure, with `r info$n_weeks` weekly observations nested within `r info$n_countries` countries. Accordingly, we model our outcome variables with multilevel or hierarchical models, with country-specific random offsets for each intercept and a fixed global year trend. When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, use use ordered logistic regression models. We take a Bayesian approach to better model the uncertainty inherent in measuring the effect of derogations on these broad policies.

We include complete formal specifications of all our models, likelihoods, and priors in the appendix. Here we include two simplified model specifications to help demonstrate the intuition behind these multilevel TSCS models. For binary outcomes, we use a logistic model with country-specific offsets:

::: {.column-page-inset-right}

$$
\begin{aligned}
&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}, \alpha_k) \\
\\
&\ \textbf{Distribution parameters} \\
\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} & \text{Probability of outcome} \\
&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\
\\
&\ \textbf{Priors} \\
\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-country variability,}
\end{aligned}
$$

:::

while for ordered outcomes, we use ordered logistic regression with country-specific offsets:

::: {.column-page-inset-right}

$$
\begin{aligned}
&\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\,logit}(\phi_{it_j}, \alpha_k) \\
\\
&\ \textbf{Distribution parameters} \\
\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} & \text{Cumulative probability of outcome} \\
&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\
\\
&\ \textbf{Priors} \\
\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-country variability} \\
\alpha_k \sim&\ \mathcal{N}(0, 1) & \text{Boundaries between thresholds}
\end{aligned}
$$

:::

In all model specifications, we include new and cumulative COVID cases and deaths as controls. In our first hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society. In our second hypothesis, we omit the core civil society index and instead include the interaction of civil society repression and derogations.

Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in the appendix). Instead we rely on probability-scale predictions and marginal effects whenever possible. We use median values from our models' posterior distributions as point estimates and provide credible intervals using the 95% highest posterior density. For hypothesis testing, we report the posterior probability that predictions and marginal effects are above or below zero.

# Analysis

December 2020 special issue - called for people to do more descriptive stuff on covid restrictions

## Derogations and emergency policies

Before exploring whether COVID-induced derogations are associated with specific human rights abuses, it is useful to examine how derogating and non-derogating states approached the implementation of their emergency public health measures. @fig-preds-prelim-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic. In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-prelim-2 for the percentage-point differences between the predictions for the two groups of states).

For instance, all countries were highly likely to cancel public events in March and April 2020, with greater than a 95% chance. Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021. Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability. The difference in predicted probabilities between the two types of countries ranged from 3 to 8 percentage points ($p(\Delta > 0) = 1$). The differences in policy implementation are starker for other types of emergency measures. Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between 80% and 90%, representing a 15–18 percentage point difference ($p(\Delta > 0) = 1$). Derogating states were also far more likely to close public transit, beginning with a roughly 80% probability in March 2020 and ending in June 2021 with a 60% probability. Non-derogating states followed a similar downward trend over time, but 23–25 percentage points lower than their derogating counterparts ($p(\Delta > 0) = 1$). Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to roughly 70% by June 2021. Non-derogating states, on the other hand, began the pandemic with a predicted 75% probability of limiting movement, dropping to only 25% by June 2021, representing a 20–40 percentage point difference ($p(\Delta > 0) = 1$). The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel. Both derogating and non-derogating states had a nearly 100% probability of limiting international movement across all 15 months.

```{r fig-preds-prelim, fig.width=12, fig.height=6, out.width="100%"}
#| include: true
#| echo: false
#| layout-ncol: 1
#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
#| fig-subcap:
#|   - "Predicted probabilities"
#|   - "Contrasts between predicted probabilities"

# Predicted probabilities
preliminary_models_plots <- m_prelim_derog %>% 
  mutate(pred_plot = map2(model, nice, ~{
    datagrid(model = .x,
             year_week_num = 1:69,
             derogation_ineffect = 0:1) %>% 
      add_epred_draws(.x) %>% 
      left_join(year_week_lookup, by = "year_week_num") %>% 
      mutate(derogation_ineffect = factor(derogation_ineffect, 
                                          levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_fill_manual(values = c(clrs[1], clrs[8])) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      scale_y_continuous(labels = label_percent()) +
      labs(x = NULL, y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  }))

wrap_plots(preliminary_models_plots$pred_plot)

# Marginal effects / contrasts
preliminary_models_mfx <- m_prelim_derog %>% 
  mutate(mfx = map(model, ~{
    .x %>% 
      comparisons(newdata = datagrid(year_week_num = 1:69),
                  variables = "derogation_ineffect",
                  type = "response") |> 
      posteriordraws()
  })) %>% 
  mutate(mfx_plot = map2(mfx, nice, ~{
    .x %>% 
      left_join(year_week_lookup, by = "year_week_num") %>% 
      ggplot(aes(x = year_week_day, y = draw * 100)) +
      stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      labs(x = NULL,
           y = "Percentage point difference\nin probability of policy",
           title = .y) +
      theme_pandem()
  }))

wrap_plots(preliminary_models_mfx$mfx_plot)
```

While these results do not speak directly to our main hypotheses, they do reveal important insights into the motivations of derogating states. Countries that undertook emergency safety measures seem to have taken the time to follow international law and uphold their treaty obligations. International law thus appears to not just be an empty commitment—states that took the time to formally derogate were more likely to keep these emergency measures in place for longer and take their derogations seriously.

## Derogations and human rights

Based on our model results, derogations have different ASDF ASDF

There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement. Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an 85% predicted probability of no discriminatory policy in March 2020 rising to nearly 100% by June 2021. The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.

In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status. For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability). The middle ground of only minor issues with time limits is exceptionally unlikely. Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues. The derogation process requires that emergency measures be timebound and limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.

TODO: Kind of help with non derogable

Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they don't do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies. They help with bureaucratic processes, not the implementation of those processes.

## Derogations, civil society, and human rights

Civil society repression, on the other hand, adds an additional wrinkle to the story

The derogations vs. non-derogations split matters when looking across the range of civil society repression. In countries with bad repression that don't derogate, there's a higher likelihood of discriminatory policies and abusive enforcement. Countries with similar levels of civil society repression that *do* formally derogate, however, have a nearly 100% predicted probability having no discriminatory policies. The derogation process could thus serve as a strong signal that emergency measures are indeed planned to be limited and non-discriminatory even in countries where civil society is under threat.

TODO: Abusive enforcement

TODO: Not much with limiting media

TODO: Nothing at all with nonderogable rights

CS deals with more on-the-ground stuff and monitoring abuses and shaming them - CS deals with more of the implementation of

On paper vs. in practice

# Discussion

@Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic. However, @richardson2020 see derogations as a less clear path for states during COVID-19, describing the actions, their conditions, and human rights as a "tangled morass" (p. 106).

Is this good news? Derogators aren't using this opportunistically - derogators are doing the process correctly

Enforcement is carried out by lower level officials - principal agent issues and local bureaucrats not caring about international law - the time limit and nonderogable rights things are done by high level people

# Conclusion

In conclusion mention that future research can point to bureaucrats vs. policy makers and

Contribution = this is good news

This research note examines states' responses to their international human rights treaty obligations, particularly the ICCPR, during the Covid-19 pandemic and its subsequent impact on human rights violations.

Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties.

The first few months of the pandemic raised numerous concerns that the pandemic would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide

# Appendix {.appendix}

### Formal statistical models

$$
\begin{aligned}
&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered logit}(\phi_{it_j}, \alpha_k) \\
\\
&\ \textbf{Models for distribution parameters} \\
\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\
\\
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
\alpha_k \sim&\ \mathcal{N}(0, 1)
\end{aligned}
$$

### Complete regression results

#### Derogations and emergency policies

#### Derogations and human rights

#### Derogations, civil society, and human rights

# References

### 2023-03-02 R/funs_plots.R
Move plot data processing to targets

@@ -1,0 +1,155 @@
+# Storing ggplot objects as rds files is BAD 
+#   (https://github.com/hadley/ggplot2-book/issues/344)
+#
+# But it's also necessary/recommended when working with {targets}---the basic
+# walkthrough in the manual even shows how to create a plot as a target
+# 
+# ggplot objects are strange beasts because they store a copy of the overall
+# environment when serialized into rds files. When working with regular-sized
+# datasets, this isn't really ever a problem. But when working with tidy
+# data frames of MCMC chains with millions and millions of rows, this can create
+# rds files that are hundreds or thousands of MBs, which is wild.
+#
+# This post goes into more details about how to fix it: 
+# https://ropensci.org/blog/2022/12/06/save-ggplot2-targets/
+#
+# Instead of using stat_lineribbon() on tidy MCMC draws like I normally do, it's
+# better to collapse the data down first with median_qi() and then use
+# geom_lineribbon(). This results in much tinier data frames and plots render
+# immediately in .qmd files
+
+library(marginaleffects)
+
+build_prelim_plot_data <- function(model_df, year_week_lookup) {
+  model_df_preds_mfx <- model_df %>% 
+    mutate(pred_data = map(model, ~{
+      datagrid(model = .x,
+               year_week_num = 1:69,
+               derogation_ineffect = 0:1) %>% 
+        tidybayes::add_epred_draws(.x) %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                            levels = 0:1,
+                                            labels = c("No", "Yes"),
+                                            ordered = TRUE)) %>% 
+        group_by(year_week_day, derogation_ineffect) %>% 
+        tidybayes::median_qi(.epred, .width = c(0.5, 0.8, 0.95))
+    })) %>% 
+    mutate(mfx_data = map(model, ~{
+      .x %>% 
+        comparisons(newdata = datagrid(year_week_num = 1:69),
+                    variables = "derogation_ineffect",
+                    type = "response")  %>%  
+        posteriordraws() %>% 
+        left_join(year_week_lookup, by = "year_week_num") %>% 
+        group_by(year_week_day) %>% 
+        tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
+    }))
+  
[…]

### 2023-03-02 _targets.R

@@ -130,6 +130,18 @@ list(
   ## Graphics ----
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
   
+  ## Plots ----
+  tar_target(prelim_plot_data, 
+             build_prelim_plot_data(m_prelim_derog, year_week_lookup)),
+  
+  tar_target(h1_plot_data, 
+             build_h1_plot_data(m_h1, year_week_lookup)),
+  
+  # tar_target(h1_preds_plot, 
+  #            build_h1_plots_preds(m_h1, year_week_lookup, graphic_functions)),
+  # tar_target(h1_mfx_plot, 
+  #            build_h1_plots_mfx(m_h1, year_week_lookup, graphic_functions)),
+  
   ## Analysis notebook ----
   tar_quarto(website, path = "."),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),

### 2023-03-02 R/funs_plots.R
Remove these old ggplot-storing functions

@@ -103,53 +103,3 @@ build_h1_plot_data <- function(model_df, year_week_lookup) {
   
   return(model_df_preds_mfx)
 }
-
-build_h1_plots_mfx <- function(model_df, year_week_lookup, plot_funs) {
-  list2env(plot_funs, env = environment())
-  
-  h1_models_mfx <- model_df %>% 
-    mutate(mfx = map(model, ~{
-      .x %>% 
-        comparisons(newdata = datagrid(year_week_num = 1:69),
-                    variables = "derogation_ineffect",
-                    type = "response", re_formula = NA)  %>%  
-        posteriordraws()
-    })) %>% 
-    mutate(mfx_plot = pmap(list(mfx, nice, family, model, y), ~{
-      if (..3 == "cumulative") {
-        df <- ..1 %>% 
-          left_join(year_week_lookup, by = "year_week_num") %>% 
-          mutate(group = factor(group, levels = levels(..4$data[[..5]]), ordered = TRUE)) %>% 
-          select(year_week_day, draw, group)
-        
-        df %>% 
-          ggplot(aes(x = year_week_day, y = draw * 100)) +
-          tidybayes::stat_lineribbon(aes(color = group, fill = group), alpha = 0.3) +
-          geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-          scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
-          scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") + 
-          scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-          labs(x = NULL,
-               y = "Percentage point difference\nin probability of outcome level",
-               title = ..2) +
-          facet_wrap(vars(group), ncol = 2) +
-          theme_pandem()
-      } else {
-        df <- ..1 %>% 
-          left_join(year_week_lookup, by = "year_week_num") %>% 
-          select(year_week_day, draw)
-        
-        df %>% 
-          ggplot(aes(x = year_week_day, y = draw * 100, )) +
-          tidybayes::stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
-          geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-          scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-          labs(x = NULL,
[…]

### 2023-03-03 R/funs_plots.R
Use named arguments instead of ..1, ..2, etc.

@@ -79,17 +79,17 @@ build_h1_plot_data <- function(model_df, year_week_lookup) {
           tidybayes::median_qi(.epred, .width = c(0.5, 0.8, 0.95))
       }
     })) %>% 
-    mutate(mfx_data = pmap(list(model, family, model, y), ~{
-      mfx <- ..1 %>% 
+    mutate(mfx_data = pmap(list(model, family, y), \(.model, .family, .y) {
+      mfx <- .model %>% 
         comparisons(newdata = datagrid(year_week_num = 1:69),
                     variables = "derogation_ineffect",
                     type = "response")  %>%  
         posteriordraws() %>% 
         left_join(year_week_lookup, by = "year_week_num") 
       
-      if (..2 == "cumulative") {
+      if (.family == "cumulative") {
         mfx <- mfx %>% 
-          mutate(group = factor(group, levels = levels(..3$data[[..4]]), ordered = TRUE)) %>% 
+          mutate(group = factor(group, levels = levels(.model$data[[.y]]), ordered = TRUE)) %>% 
           group_by(year_week_day, group) %>% 
           tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
       } else {

### 2023-03-03 R/funs_plots.R
Add function for processing H2 plot data

@@ -103,3 +103,55 @@ build_h1_plot_data <- function(model_df, year_week_lookup) {
   
   return(model_df_preds_mfx)
 }
+
+
+build_h2_plot_data <- function(model_df) {
+  model_df_preds_mfx <- model_df %>% 
+    mutate(pred_data = pmap(list(model, family), \(.model, .family) {
+      if (.family == "cumulative") {
+        df <- datagrid(model = .model,
+                       v2csreprss = seq(-2.5, 3, by = 0.1),
+                       derogation_ineffect = 0:1) %>% 
+          tidybayes::add_epred_draws(.model) %>% 
+          mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                              levels = 0:1,
+                                              labels = c("No derogation", "Derogation in effect"),
+                                              ordered = TRUE)) %>% 
+          group_by(v2csreprss, .category, derogation_ineffect) %>% 
+          tidybayes::median_qi(.epred, .width = c(0.5, 0.8, 0.95))
+      } else {
+        df <- datagrid(model = .model,
+                       v2csreprss = seq(-2.5, 3, by = 0.1),
+                       derogation_ineffect = 0:1) %>% 
+          tidybayes::add_epred_draws(.model) %>% 
+          mutate(derogation_ineffect = factor(derogation_ineffect, 
+                                              levels = 0:1,
+                                              labels = c("No", "Yes"),
+                                              ordered = TRUE)) %>% 
+          group_by(v2csreprss, derogation_ineffect) %>% 
+          tidybayes::median_qi(.epred, .width = c(0.5, 0.8, 0.95))
+      }
+    })) %>% 
+    mutate(mfx_data = pmap(list(model, family, y), \(.model, .family, .y) {
+      mfx <- .model %>%
+        comparisons(newdata = datagrid(v2csreprss = seq(-2.5, 3, by = 0.1)),
+                    variables = "derogation_ineffect",
+                    type = "response")  %>%
+        posteriordraws()
+
+      if (.family == "cumulative") {
+        mfx <- mfx %>%
+          mutate(group = factor(group, levels = levels(.model$data[[.y]]), ordered = TRUE)) %>%
+          group_by(v2csreprss, group) %>%
+          tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
+      } else {
+        mfx <- mfx %>%
+          group_by(v2csreprss) %>%
[…]

### 2023-03-03 _targets.R
Targets for all plot data processing stuff

@@ -131,16 +131,9 @@ list(
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
   
   ## Plots ----
-  tar_target(prelim_plot_data, 
-             build_prelim_plot_data(m_prelim_derog, year_week_lookup)),
-  
-  tar_target(h1_plot_data, 
-             build_h1_plot_data(m_h1, year_week_lookup)),
-  
-  # tar_target(h1_preds_plot, 
-  #            build_h1_plots_preds(m_h1, year_week_lookup, graphic_functions)),
-  # tar_target(h1_mfx_plot, 
-  #            build_h1_plots_mfx(m_h1, year_week_lookup, graphic_functions)),
+  tar_target(prelim_plot_data, build_prelim_plot_data(m_prelim_derog, year_week_lookup)),
+  tar_target(h1_plot_data, build_h1_plot_data(m_h1, year_week_lookup)),
+  tar_target(h2_plot_data, build_h2_plot_data(m_h2)),
   
   ## Analysis notebook ----
   tar_quarto(website, path = "."),

### 2023-03-03 manuscript/manuscript.qmd
Nicer date title

@@
 toc: true
+language: 
+  title-block-published: "Latest version"
 ---

### 2023-03-03 manuscript/manuscript.qmd
Add fast-building plots for predictions and mfx

@@
   mutate(year_week_day = ymd(year_week_day))
-tar_load(year_week_lookup)
-tar_load(m_prelim_derog)
+
+tar_load(prelim_plot_data)
+tar_load(h1_plot_data)
+tar_load(h2_plot_data)
 
@@
 
-# Predicted probabilities
-preliminary_models_plots <- m_prelim_derog %>% 
-  mutate(pred_plot = map2(model, nice, ~{
-    datagrid(model = .x,
-             year_week_num = 1:69,
-             derogation_ineffect = 0:1) %>% 
-      add_epred_draws(.x) %>% 
-      left_join(year_week_lookup, by = "year_week_num") %>% 
-      mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                          levels = 0:1,
-                                      labels = c("No", "Yes"),
-                                      ordered = TRUE)) %>% 
-      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
-      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+prelim_plots <- prelim_plot_data %>% 
+  mutate(pred_plot = map2(pred_data, nice, ~{
+    .x %>% 
+      ggplot(aes(x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+                 color = derogation_ineffect, fill = after_scale(color), 
+                 group = forcats::fct_rev(ordered(.width)))) +
+      geom_lineribbon(alpha = 0.3) +
       scale_color_manual(values = c(clrs[1], clrs[8])) +
@@
       theme_pandem()
-  }))
-
-wrap_plots(preliminary_models_plots$pred_plot)
-
-# Marginal effects / contrasts
-preliminary_models_mfx <- m_prelim_derog %>% 
-  mutate(mfx = map(model, ~{
+  })) %>% 
+  mutate(mfx_plot = map2(mfx_data, nice, ~{
     .x %>% 
-      comparisons(newdata = datagrid(year_week_num = 1:69),
-                  variables = "derogation_ineffect",
-                  type = "response") |> 
-      posteriordraws()
-  })) %>% 
-  mutate(mfx_plot = map2(mfx, nice, ~{
-    .x %>% 
-      left_join(year_week_lookup, by = "year_week_num") %>% 
-      ggplot(aes(x = year_week_day, y = draw * 100)) +
-      stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
+      mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>% 
+      ggplot(aes(x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+                 group = forcats::fct_rev(ordered(.width)))) +
+      geom_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
       scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
@@
 
-wrap_plots(preliminary_models_mfx$mfx_plot)
+wrap_plots(prelim_plots$pred_plot)
+wrap_plots(prelim_plots$mfx_plot)
 ```
@@
 
+```{r fig-preds-mfx-h1, fig.width=13, fig.height=6, out.width="100%"}
+#| include: true
+#| echo: false
+#| layout-ncol: 1
+#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
+#| fig-subcap:
+#|   - "Predicted probabilities"
+#|   - "Contrasts between predicted probabilities"
+
+h1_plots <- h1_plot_data %>% 
+  # Predicted probabilties
+  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        ggplot(aes(
+          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+          color = .category, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = .nice_name) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      .data %>% 
+        ggplot(aes(
+          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+          color = derogation_ineffect, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = NULL, y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = .nice_name) +
+        theme_pandem()
+    }
+  })) %>% 
+  # Marginal effects / contrasts
+  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+          color = group, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome level",
+             title = .nice_name) +
+        facet_wrap(vars(group), ncol = 2) +
+        theme_pandem()
+    } else {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome",
+             title = .nice_name) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h1_plots$pred_plot)
+wrap_plots(h1_plots$mfx_plot)
+```
+
 ## Derogations, civil society, and human rights
@@
 The derogation process could thus serve as a strong signal that emergency measures are indeed planned to be limited and non-discriminatory even in countries where civil society is under threat.
+
+```{r fig-preds-mfx-h2, fig.width=13, fig.height=6, out.width="100%"}
+#| include: true
+#| echo: false
+#| layout-ncol: 1
+#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
+#| fig-subcap:
+#|   - "Predicted probabilities"
+#|   - "Contrasts between predicted probabilities"
+
+h2_plots <- h2_plot_data %>% 
+  # Predicted probabilities
+  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        ggplot(aes(
+          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
+          color = .category, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = "Civil society repression", y = "Predicted probability", 
+             color = "Severity", fill = "Severity",
+             title = .nice_name) +
+        facet_wrap(vars(derogation_ineffect)) +
+        theme_pandem()
+    } else {
+      .data %>% 
+        ggplot(aes(
+          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
+          color = derogation_ineffect, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        scale_color_manual(values = c(clrs[1], clrs[8])) +
+        scale_y_continuous(labels = label_percent()) +
+        labs(x = "Civil society repression", y = "Predicted probability", 
+             color = "Derogation in effect", fill = "Derogation in effect",
+             title = .nice_name) +
+        theme_pandem()
+    }
+  })) %>% 
+  # Marginal effects / contrasts
+  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
+          color = group, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome level",
+             title = .nice_name) +
+        facet_wrap(vars(group), ncol = 2) +
+        theme_pandem()
+    } else {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome",
+             title = .nice_name) +
+        theme_pandem()
+    }
+  }))
+
+wrap_plots(h2_plots$pred_plot)
+wrap_plots(h2_plots$mfx_plot)
+```
 

### 2023-03-03 manuscript/manuscript.qmd
No need for fill scale

@@
 
-```{r fig-preds-prelim, fig.width=12, fig.height=6, out.width="100%"}
+```{r fig-preds-mfx-prelim, fig.width=12, fig.height=6, out.width="100%", cache=FALSE}
 #| include: true
@@
       scale_color_manual(values = c(clrs[1], clrs[8])) +
-      scale_fill_manual(values = c(clrs[1], clrs[8])) +
       scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +

### 2023-03-03 manuscript/manuscript.qmd
Use correct figure chunk names

@@
 Before exploring whether COVID-induced derogations are associated with specific human rights abuses, it is useful to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
-@fig-preds-prelim-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
-In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-prelim-2 for the percentage-point differences between the predictions for the two groups of states).
+@fig-preds-mfx-prelim-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
+In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-prelim-2 for the percentage-point differences between the predictions for the two groups of states).
 

### 2023-03-03 manuscript/manuscript.qmd
super minor changes

@@
 For our second hypothesis, we use a measure of civil society repression from the Varieties of Democracy (V-Dem) project [@vdem-v12], derived from a latent Bayesian item response theory model of expert opinions, with values ranging from roughly −3 to 3 (where higher values represent less repression).
-This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti=civil society environment.
+This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti-civil society environment.
 
@@
 Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues.
-The derogation process requires that emergency measures be timebound and limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
+The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
 

### 2023-03-06 R/funs_models.R
Build model tables in targets

@@ -1,3 +1,46 @@
+# Running modelsummary() on Bayesian models takes a while because of all the
+# calculations involved in creating the GOF statistics. With modelsummary 0.7+,
+# though it's now possible to build the base model with modelsummary(..., output
+# = "modelsummary_list"), save that as an intermediate object, and then feed it
+# through modelsummary() again with whatever other output you want. The
+# modelsummary_list-based object thus acts like an output-agnostic ur-model.
+
+build_modelsummary <- function(model_df) {
+  msl <- model_df %>% 
+    pull(model, name = nice) %>% 
+    modelsummary::modelsummary(output = "modelsummary_list",
+                               statistic = "[{conf.low}, {conf.high}]",
+                               ci_method = "hdi",
+                               metrics = c("R2"))
+  return(msl)
+}
+
+# This is how to get other stats like ELPD (using metrics = "LOOIC" is required
+# for that; metrics = "R2" provides nobs). I would do this for all the summary
+# tables, but for whatever reason, it takes forever to calculate ELPD/LOO stuff
+# on ordered logit models, so we just show N instead
+#
+# gm <- tribble(
+#   ~raw,        ~clean,      ~fmt, ~omit,
+#   "nobs",      "N",         0,    FALSE,
+#   "r.squared", "R2",        3,    FALSE,
+#   "elpd",      "ELPD",      1,    FALSE,
+#   "elpd.se",   "ELPD (SE)", 1,    FALSE
+# )
+# 
+# modelsummary(models_prelim,
+#              estimate = "{estimate}",
+#              statistic = "conf.int",
+#              gof_map = gm,
+#              metrics = c("R2", "LOOIC"))
+
+gof_map <- tribble(
+  ~raw,          ~clean,                 ~fmt, ~omit,
+  "nobs",        "N",                    0,    FALSE,
+  "r.squared",   "\\(R^2\\) (total)",    2,    FALSE,
+  "r2.marginal", "\\(R^2\\) (marginal)", 2,    FALSE
+)
+
 coef_map <- c(
   "b_derogation_ineffect" = "Derogation in effect",
   "b_v2csreprss" = "Civil society repression",
@@ -20,31 +63,8 @@ coef_map <- c(
   "sd_who_region__Intercept" = "Region random effects σ"
 )
 
-gof_map <- tribble(
-  ~raw,        ~clean,     ~fmt, ~omit,
[…]

### 2023-03-06 _targets.R

@@ -125,8 +125,13 @@ list(
   tar_target(m_h1, f_h1(weekly_panel)),
   tar_target(m_h2, f_h2(weekly_panel)),
 
-  tar_target(m_coef_gof, lst(coef_map, gof_map)),
-  
+  ## Model tables ----
+  # Build tables here because they take a while
+  tar_target(modelsummary_functions, lst(coef_map, gof_map)),
+  tar_target(models_tbl_prelim, build_modelsummary(m_prelim_derog)),
+  tar_target(models_tbl_h1, build_modelsummary(m_h1)),
+  tar_target(models_tbl_h2, build_modelsummary(m_h2)),
+
   ## Graphics ----
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
   

### 2023-03-06 manuscript/manuscript.qmd
New stuff in analysis

@@
 abstract: |
-  In an effort to combat the Covid-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
+  In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
 However, the abuse of emergency derogations by many governments raises concerns about the long-lasting impact of these sweeping measures on human rights.
@@
 Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run.
-Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker and the CoronaNet COVID-19 Government Response Event, we first assess the variation in human rights responses among governments during the pandemic.
+Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we first assess the variation in human rights responses among governments during the pandemic.
 We then assess the role of civil society in keeping governments accountable and mitigating the effects of policies that violate human rights.
 Our results hold important implications for understanding government compliance with international treaties and shedding light on how states of emergency may directly and indirectly lead to democratic backsliding.
-date: February 27, 2023
+date: March 5, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
 bibliography: bib.bib
+link-citations: true
+format: 
+  html:
+    appendix-style: plain
     toc: true
+  pdf:
+    mainfont: "IBM Plex Serif"
+    sansfont: "Noto Sans"
+    linestretch: 1.25
+    toc: false
 language: 
@@
 library(patchwork)
+library(colorspace)
+library(modelsummary)
+library(kableExtra)
 library(targets)
@@
 
-tar_load(prelim_plot_data)
-tar_load(h1_plot_data)
-tar_load(h2_plot_data)
+tar_load(c(prelim_plot_data, h1_plot_data, h2_plot_data))
+tar_load(c(models_tbl_prelim, models_tbl_h1, models_tbl_h2))
 
-# Plotting functions
+# Plotting and modelsummary functions
 invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
 ```
@@
 Combating a deadly virus like SARS-CoV-2 requires extraordinary public health measures.
-In an effort to fight the Covid-19 pandemic, governments across the world imposed restrictions on the freedom of movement, association, assembly and privacy.
+In an effort to fight the COVID-19 pandemic, governments across the world imposed restrictions on the freedom of movement, association, assembly and privacy.
 While countries are ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights principles.
@@
 What did derogations mean for compliance?
-Are derogating and non-derogating states different in terms of how they implemented emergency measures during Covid-19?
-Were derogating states more likely to exploit the provisions and to abuse human rights during Covid-19?
+Are derogating and non-derogating states different in terms of how they implemented emergency measures during COVID-19?
+Were derogating states more likely to exploit the provisions and to abuse human rights during COVID-19?
 These questions are important for the study of international law, compliance, and times of crises.
@@
 
-Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, the CoronaNet Covid-19 Government Response Event, and United Nations treaty derogation data we first assess the variation in human rights responses among governments during the pandemic.
+Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data we first assess the variation in human rights responses among governments during the pandemic.
 We then analyze the effect of derogating (or not) on human rights practices during the pandemic and assess the role of civil society in mitigating the effects of rights-violating policies.
@@
 Additionally, we find that, on average, derogating states are less likely to abuse emergency measures.
-On the other hand, states that did not derogate from their international human rights obligations and still impose emergency measures are more likely impose discriminatory policies, violate non-derogable rights such right to life AND WHAT ELSE, and impose emergency measures without time limits.
-Explain this result more - ANDREW "Civil society repression increases probability of abusive enforcement and discriminatory policies; derogations mitigate that probability."
+On the other hand, states that did not derogate from their international human rights obligations and still impose emergency measures are more likely impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
+We also find that civil society can play an important mitigating role in curbing discriminatory policies.
+States with low levels of civil society repression are the least likely to abuse human rights when enforcing emergency policies.
+The likelihood of discriminatory policies increase as civil society repression worsens, but derogations appear to protect against worsening abuses—even in countries with severe civil society repression, those that derogate have a nearly 100% probability of no discriminatory policies.
 
@@
 @Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
- EXPAND ON RESULT In this scenario, when international factors may not be enough to constrain state actions, assessing the effect of domestic-level factors such as a robust civil society that potentially can mitigate the adverse effects of derogations becomes essential.
-WE FIND.
+TODO: EXPAND ON RESULT In this scenario, when international factors may not be enough to constrain state actions, assessing the effect of domestic-level factors such as a robust civil society that potentially can mitigate the adverse effects of derogations becomes essential.
+TODO: WE FIND.
 
@@
 We then assess states' behavior towards civil society and these organizations' role in mitigating the adverse effects of emergency measures.
-We conclude with the implications' of our findings for the study of international law, times of crises, and the role of civil society.
+We conclude with the implications of our findings for the study of international law, times of crises, and the role of civil society.
 
@@
 
-Some states new to derogations introducing them during the COVID-19 pandemic [@Comstock2023].
+Some states new to derogations introducing them during the COVID-19 pandemic [@Comstock:2023].
 However, past treaty action behavior may help explain entering derogations during this time.
@@
 While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, civil society, and state actions during the pandemic.
-We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl:data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
+We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
 We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021.
@@
 
+```{r tbl-data-sources}
+#| tbl-cap: Sources for outcomes and explanatory variables
+#| include: true
+#| echo: false
+tribble(
+  ~`Outcome variables`,                        ~Description,                                                        ~Frequency,  ~Source,
+  "Emergency public health measures",          "Binary indicator",                                                  "Weekly",    "Oxford Covid-19 Government Response Tracker",
+  "Human rights and policy outcomes",          "Ordered categories for the severity of abuses",                     "Quarterly", "Varieties of Democracy PanDem",
+  "ICCPR derogation",                          "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
+  "Civil society repression",                  "≈ −3 to 3; higher values indicate less repression",                 "Annual",    "Varieties of Democracy",
+  "New and cumulative COVID deaths and cases", "Counts",                                                            "Weekly",    "World Health Organization",
+  "Past ICCPR derogation or action",           "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
+  "Rule of law index",                         "0–1; higher values represent greater respect for rule of law",      "Annual",    "Varieties of Democracy",
+  "Civil liberties index",                     "0–1; higher values represent greater respect for civil liberties",  "Annual",    "Varieties of Democracy",
+  "Core civil society index",                  "0–1; higher values represent better environment for civil society", "Annual",    "Varieties of Democracy"
+) %>% 
+  kbl(booktabs = TRUE) %>% 
+  kable_styling() %>% 
+  pack_rows(index = c(" " = 2, "Primary explanatory variables" = 2, "Secondary explanatory variables" = 5),
+            indent = FALSE)
+
+```
+
 ## Outcome variables
 
-We examine the relationship between formal treaty derogations and different human rights and policy outcomes in three stages (see @tbl:vars).
-First, to explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
+We examine the relationship between formal treaty derogations and different human rights and policy outcomes in three stages (see @tbl-vars).
+First, to explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
 Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.
@@
 
-Next, to test our first hypothesis, we explore the relationship between formal treaty derogations on four different human rights and four policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
+| General question                                    | Outcomes                                                                                                                                  | Main explanatory variable                   | Other explanatory variables                                                                                                                                              |
+|----------------|------------------|----------------|----------------------|
+| Derogations and emergency measures                  | Cancelling public events, Gathering restrictions, Close public transit, Restrictions on internal movement, Limits on international travel | ICCPR derogation in effect                  | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions                                                                     |
+| Derogations and human rights (H~1~)                 | Discriminatory policy, Violations of non-derogable rights, No time limit measures, Abusive enforcement                                    | ICCPR derogation in effect                  | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index, core civil society index |
+| Derogations, human rights, and civil society (H~2~) | Discriminatory policy, Violations of non-derogable rights, Limiting media, Abusive enforcement                                            | ICCPR derogation × civil society repression | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index                           |
+
+: Overview of outcomes and explanatory variables included in models {#tbl-vars}
+
+Next, to test our first hypothesis, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
 Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations.
@@
 Instead we rely on probability-scale predictions and marginal effects whenever possible.
+To do so, we calculate predicted probabilities holding all explanatory variables at their typical values (median or mode).
+For logistic models, we estimate the probability of having a specific policy or violating a specific human right; for ordered models, we estimate the probability of each possible level of severity.
+To determine the effect of derogations, we calculate the contrast in predicted probabilities when derogation is set to true vs.
+when it is set to false (e.g., in ordered logistic models we compare the probability of major discrimination under derogation with the probability of major discrimination without derogation).
 We use median values from our models' posterior distributions as point estimates and provide credible intervals using the 95% highest posterior density.
-For hypothesis testing, we report the posterior probability that predictions and marginal effects are above or below zero.
+For hypothesis testing, we report the posterior probability that the between-level predicted differences are above or below zero.
 
 # Analysis
-
-December 2020 special issue - called for people to do more descriptive stuff on covid restrictions
 
@@
 
-Based on our model results, derogations have different ASDF ASDF
-
+The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-h1).
 There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement.
@@
 The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.
+Moreover, countries that derogate were substantially less likely to engage in no abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic.
+Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.
+The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way—are unclear.
 
@@
 The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
+Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)
 
-TODO: Kind of help with non derogable
-
-Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they don't do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
-They help with bureaucratic processes, not the implementation of those processes.
+Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
+We thus find strong support only for H~1c~—that derogating states are less likely to impose measures with no time limits.
+We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~1b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~1c~).
+Derogations seem help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
 
@@
 h1_plots <- h1_plot_data %>% 
-  # Predicted probabilties
+  # Predicted probabilities
   mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
@@
         geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
+        scale_color_manual(values = darken(clrs[c(2, 4, 6, 7)], 0.5, space = "combined"), guide = "none") +
         scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
@@
 
-Civil society repression, on the other hand, adds an additional wrinkle to the story
+Civil society repression, on the other hand, adds an additional wrinkle to the story.
+Countries with severe level of repression (i.e.
+negative repression scores) that do not derogate have a 25% predicted probability of minor to major discriminatory policies, while similarly repressive countries that do derogate have nearly a 0% predicted probability of discriminatory policies (see @fig-preds-mfx-h2).
+This represents a substantial 5–10 percentage point difference between derogating and non-derogating countries with repressive civil societies.
+The derogation process could thus serve as a strong signal that emergency measures are indeed planned to be limited and non-discriminatory even in countries where civil society is under threat.
 
-The derogations vs.
-non-derogations split matters when looking across the range of civil society repression.
-In countries with bad repression that don't derogate, there's a higher likelihood of discriminatory policies and abusive enforcement.
-Countries with similar levels of civil society repression that *do* formally derogate, however, have a nearly 100% predicted probability having no discriminatory policies.
-The derogation process could thus serve as a strong signal that emergency measures are indeed planned to be limited and non-discriminatory even in countries where civil society is under threat.
+However, there are no other substantial differences in predicted probabilities across derogation status for another other outcomes.
+Derogating and non-derogating states are generally equally likely to violate non-derogable rights (H~2a~), limit media (H~2d~), and engage in abusive enforcement (H~2c~) regardless of their level of civil society repression.
+We thus only find support for H~2b~—if states are repressive toward civil society organizations, they are more likely to use discriminatory policies, while derogations strongly negate this trend.
+This may be advocacy-focused civil society organizations in particular tend to deal with on-the-ground monitoring, advocacy, and shaming, and is more likely to call out governments for unequal and unjust implementation of their emergency policies.
 
@@
 
-TODO: Abusive enforcement
-
-TODO: Not much with limiting media
-
-TODO: Nothing at all with nonderogable rights
-
-CS deals with more on-the-ground stuff and monitoring abuses and shaming them - CS deals with more of the implementation of
-
-On paper vs. in practice
-
 # Discussion
@@
 
-This research note examines states' responses to their international human rights treaty obligations, particularly the ICCPR, during the Covid-19 pandemic and its subsequent impact on human rights violations.
+This research note examines states' responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and its subsequent impact on human rights violations.
 
@@
 
+```{r tbl-results-full-prelim}
+#| tbl-cap: Complete results from models showing relationship between derogations and emergency policies
+#| include: true
+#| echo: false
+notes <- c(
+  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models; 95% credible intervals (highest density posterior interval, or HDPI) in brackets.", 
+  "Total \\(R^2\\) considers the variance of both population and group effects;", 
+  "marginal \\(R^2\\) only takes population effects into account."
+)
+
+modelsummary(models_tbl_prelim,
+             estimate = "{estimate}",
+             statistic = "[{conf.low}, {conf.high}]",
+             coef_map = coef_map,
+             gof_map = gof_map,
+             output = "kableExtra",
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+  kable_styling(htmltable_class = "table-sm light-border") %>% 
+  footnote(general = notes, footnote_as_chunk = TRUE)
+```
+
 #### Derogations and human rights
+
+```{r tbl-results-full-h1}
+#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~1~)
+#| include: true
+#| echo: false
+modelsummary(models_tbl_h1,
+             estimate = "{estimate}",
+             statistic = "[{conf.low}, {conf.high}]",
+             coef_map = coef_map,
+             gof_map = gof_map,
+             output = "kableExtra",
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+  kable_styling(htmltable_class = "table-sm light-border") %>% 
+  footnote(general = notes, footnote_as_chunk = TRUE)
+```
 
@@
 
+```{r tbl-results-full-h2}
+#| tbl-cap: Complete results from models showing relationship between derogations, civil society, and human rights (H~2~)
+#| include: true
+#| echo: false
+modelsummary(models_tbl_h2,
+             estimate = "{estimate}",
+             statistic = "[{conf.low}, {conf.high}]",
+             coef_map = coef_map,
+             gof_map = gof_map,
+             output = "kableExtra",
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+  kable_styling(htmltable_class = "table-sm light-border") %>% 
+  footnote(general = notes, footnote_as_chunk = TRUE)
+```
+
 # References

### 2023-03-06 analysis/analysis_h1.qmd
Update analysis notebook to use new targets

@@
 library(patchwork)
+library(colorspace)
 library(scales)
@@
 
+tar_config_set(store = here::here('_targets'),
+               script = here::here('_targets.R'))
+
 # Load targets 
-withr::with_dir(here::here(), {
 # Final data
@@
 tar_load(m_h1)
-  tar_load(m_coef_gof)
+tar_load(h1_plot_data)
+tar_load(models_tbl_h1)
 
@@
 invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
-})
+invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
 
@@
 ```{r tbl-h1, cache=TRUE, warning=FALSE}
-modelsummary(pull(m_h1, model, name = nice),
+notes <- c(
+  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models; 95% credible intervals (highest density posterior interval, or HDPI) in brackets.", 
+  "Total \\(R^2\\) considers the variance of both population and group effects;", 
+  "marginal \\(R^2\\) only takes population effects into account."
+)
+
+modelsummary(models_tbl_h1,
              estimate = "{estimate}",
-             statistic = "conf.int",
-             fmt = 2, coef_map = m_coef_gof$coef_map, 
-             gof_map = m_coef_gof$gof_map,
-             metrics = c("none"))
+             statistic = "[{conf.low}, {conf.high}]",
+             coef_map = coef_map,
+             gof_map = gof_map,
+             output = "kableExtra",
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+  kable_styling(htmltable_class = "table-sm light-border") %>% 
+  footnote(general = notes, footnote_as_chunk = TRUE)
 ```
@@
 ```{r fig.width=13, fig.height=6, out.width="100%"}
-h1_models_plots <- m_h1 %>% 
-  mutate(pred_plot = pmap(list(model, nice, family), ~{
-    if (..3 == "cumulative") {
-      datagrid(model = ..1,
-               year_week_num = 1:69,
-               derogation_ineffect = 0:1) %>% 
-        add_epred_draws(..1) %>% 
-        left_join(year_week_lookup, by = "year_week_num") %>% 
-        mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                            levels = 0:1,
-                                            labels = c("No derogation", "Derogation in effect"),
-                                            ordered = TRUE)) %>% 
-        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
-        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+h1_plots <- h1_plot_data %>% 
+  # Predicted probabilities
+  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        ggplot(aes(
+          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+          color = .category, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
         scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
         scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
@@
              color = "Severity", fill = "Severity",
-             title = ..2) +
+             title = .nice_name) +
         facet_wrap(vars(derogation_ineffect)) +
@@
     } else {
-      datagrid(model = ..1,
-               year_week_num = 1:69,
-               derogation_ineffect = 0:1) %>% 
-        add_epred_draws(..1) %>% 
-        left_join(year_week_lookup, by = "year_week_num") %>% 
-        mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                            levels = 0:1,
-                                            labels = c("No", "Yes"),
-                                            ordered = TRUE)) %>% 
-        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
-        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+      .data %>% 
+        ggplot(aes(
+          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+          color = derogation_ineffect, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
         scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_fill_manual(values = c(clrs[1], clrs[8])) +
         scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
@@
              color = "Derogation in effect", fill = "Derogation in effect",
-             title = ..2) +
+             title = .nice_name) +
+        theme_pandem()
+    }
+  })) %>% 
+  # Marginal effects / contrasts
+  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+          color = group, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        scale_color_manual(values = darken(clrs[c(2, 4, 6, 7)], 0.5, space = "combined"), guide = "none") +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome level",
+             title = .nice_name) +
+        facet_wrap(vars(group), ncol = 2) +
+        theme_pandem()
+    } else {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome",
+             title = .nice_name) +
         theme_pandem()
@@
 
-wrap_plots(h1_models_plots$pred_plot)
+wrap_plots(h1_plots$pred_plot)
 ```
@@
 
+::: {.column-page-right}
+
 ## Marginal effects
@@
 ```{r}
-# TODO: Figure out how to work with ordered logit marginal effects
+wrap_plots(h1_plots$mfx_plot)
 ```
+
+:::

### 2023-03-06 analysis/analysis_h2.qmd
Update analysis notebook to use new targets

@@
 # Load targets
-withr::with_dir(here::here(), {
 # Final data
@@
 tar_load(m_h2)
-  tar_load(m_coef_gof)
+tar_load(h2_plot_data)
+tar_load(models_tbl_h2)
 
@@
 ```{r tbl-h2, cache=TRUE, warning=FALSE}
-modelsummary(pull(m_h2, model, name = nice),
+notes <- c(
+  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models; 95% credible intervals (highest density posterior interval, or HDPI) in brackets.", 
+  "Total \\(R^2\\) considers the variance of both population and group effects;", 
+  "marginal \\(R^2\\) only takes population effects into account."
+)
+
+modelsummary(models_tbl_h2,
              estimate = "{estimate}",
-             statistic = "conf.int",
-             fmt = 2, coef_map = m_coef_gof$coef_map, 
-             gof_map = m_coef_gof$gof_map,
-             metrics = c("none"))
+             statistic = "[{conf.low}, {conf.high}]",
+             coef_map = coef_map,
+             gof_map = gof_map,
+             output = "kableExtra",
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+  kable_styling(htmltable_class = "table-sm light-border") %>% 
+  footnote(general = notes, footnote_as_chunk = TRUE)
 ```
@@
 
-### Across time
-
 ::: {.column-page-right}
@@
 ```{r fig.width=13, fig.height=6, out.width="100%"}
-h2_models_plots <- m_h2 %>% 
-  mutate(pred_plot = pmap(list(model, nice, family), ~{
-    if (..3 == "cumulative") {
-      datagrid(model = ..1,
-               year_week_num = 1:69,
-               derogation_ineffect = 0:1) %>% 
-        add_epred_draws(..1) %>% 
-        left_join(year_week_lookup, by = "year_week_num") %>% 
-        mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                            levels = 0:1,
-                                            labels = c("No derogation", "Derogation in effect"),
-                                            ordered = TRUE)) %>% 
-        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
-        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
+h2_plots <- h2_plot_data %>% 
+  # Predicted probabilities
+  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        ggplot(aes(
+          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
+          color = .category, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
         scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
         scale_y_continuous(labels = label_percent()) +
-        labs(x = NULL, y = "Predicted probability", 
+        labs(x = "Civil society repression", y = "Predicted probability", 
              color = "Severity", fill = "Severity",
-             title = ..2) +
+             title = .nice_name) +
         facet_wrap(vars(derogation_ineffect)) +
@@
     } else {
-      datagrid(model = ..1,
-               year_week_num = 1:69,
-               derogation_ineffect = 0:1) %>% 
-        add_epred_draws(..1) %>% 
-        left_join(year_week_lookup, by = "year_week_num") %>% 
-        mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                            levels = 0:1,
-                                            labels = c("No", "Yes"),
-                                            ordered = TRUE)) %>% 
-        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
-        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+      .data %>% 
+        ggplot(aes(
+          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
+          color = derogation_ineffect, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
         scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_fill_manual(values = c(clrs[1], clrs[8])) +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
         scale_y_continuous(labels = label_percent()) +
-        labs(x = NULL, y = "Predicted probability", 
+        labs(x = "Civil society repression", y = "Predicted probability", 
              color = "Derogation in effect", fill = "Derogation in effect",
-             title = ..2) +
+             title = .nice_name) +
+        theme_pandem()
+    }
+  })) %>% 
+  # Marginal effects / contrasts
+  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
+    if (.family == "cumulative") {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
+          color = group, fill = after_scale(color),
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome level",
+             title = .nice_name) +
+        facet_wrap(vars(group), ncol = 2) +
+        theme_pandem()
+    } else {
+      .data %>% 
+        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
+        ggplot(aes(
+          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
+          group = forcats::fct_rev(ordered(.width))
+        )) +
+        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
+        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
+        labs(x = NULL,
+             y = "Percentage point difference\nin probability of outcome",
+             title = .nice_name) +
         theme_pandem()
@@
 
-wrap_plots(h2_models_plots$pred_plot)
+wrap_plots(h2_plots$pred_plot)
 ```
@@
 
-### Across civil society repression
-
 ::: {.column-page-right}
-
-```{r fig.width=13, fig.height=6, out.width="100%"}
-h2_models_plots1 <- m_h2 %>% 
-  mutate(pred_plot = pmap(list(model, nice, family), ~{
-    if (..3 == "cumulative") {
-      datagrid(model = ..1,
-               v2csreprss = seq(-2.5, 3, by = 0.1),
-               derogation_ineffect = 0:1) %>% 
-        add_epred_draws(..1) %>% 
-        mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                            levels = 0:1,
-                                            labels = c("No derogation", "Derogation in effect"),
-                                            ordered = TRUE)) %>% 
-        ggplot(aes(x = v2csreprss, y = .epred, color = .category)) +
-        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = "Civil society repression", y = "Predicted probability", 
-             color = "Severity", fill = "Severity",
-             title = ..2) +
-        facet_wrap(vars(derogation_ineffect)) +
-        theme_pandem()
-    } else {
-      datagrid(model = ..1,
-               v2csreprss = seq(-2.5, 3, by = 0.1),
-               derogation_ineffect = 0:1) %>% 
-        add_epred_draws(..1) %>% 
-        mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                            levels = 0:1,
-                                            labels = c("No", "Yes"),
-                                            ordered = TRUE)) %>% 
-        ggplot(aes(x = v2csreprss, y = .epred, color = derogation_ineffect)) +
-        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
-        scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_fill_manual(values = c(clrs[1], clrs[8])) +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = "Civil society repression", y = "Predicted probability", 
-             color = "Derogation in effect", fill = "Derogation in effect",
-             title = ..2) +
-        theme_pandem()
-    }
-  }))
-
-wrap_plots(h2_models_plots1$pred_plot)
-```
-
-:::
 
@@
 ```{r}
-# TODO: Figure out how to work with ordered logit marginal effects
+wrap_plots(h2_plots$mfx_plot)
 ```
+
+:::

### 2023-03-06 analysis/analysis_prelim.qmd
Update analysis notebook to use new targets

@@
 tar_load(m_prelim_derog)
-  tar_load(m_coef_gof)
+tar_load(prelim_plot_data)
+tar_load(models_tbl_prelim)
 
@@
 ```{r tbl-prelim, cache=TRUE}
-modelsummary(pull(m_prelim_derog, model, name = nice),
+notes <- c(
+  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models; 95% credible intervals (highest density posterior interval, or HDPI) in brackets.", 
+  "Total \\(R^2\\) considers the variance of both population and group effects;", 
+  "marginal \\(R^2\\) only takes population effects into account."
+)
+
+modelsummary(models_tbl_prelim,
              estimate = "{estimate}",
-             statistic = "conf.int",
-             fmt = 2, coef_map = m_coef_gof$coef_map, 
-             gof_map = m_coef_gof$gof_map,
-             metrics = c("R2"))
+             statistic = "[{conf.low}, {conf.high}]",
+             coef_map = coef_map,
+             gof_map = gof_map,
+             output = "kableExtra",
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+  kable_styling(htmltable_class = "table-sm light-border") %>% 
+  footnote(general = notes, footnote_as_chunk = TRUE)
 ```
@@
 ```{r fig.width=12, fig.height=6, out.width="100%"}
-preliminary_models_plots <- m_prelim_derog %>% 
-  mutate(pred_plot = map2(model, nice, ~{
-    datagrid(model = .x,
-             year_week_num = 1:69,
-             derogation_ineffect = 0:1) %>% 
-      add_epred_draws(.x) %>% 
-      left_join(year_week_lookup, by = "year_week_num") %>% 
-      mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                          levels = 0:1,
-                                      labels = c("No", "Yes"),
-                                      ordered = TRUE)) %>% 
-      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
-      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
+prelim_plots <- prelim_plot_data %>% 
+  mutate(pred_plot = map2(pred_data, nice, ~{
+    .x %>% 
+      ggplot(aes(x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+                 color = derogation_ineffect, fill = after_scale(color), 
+                 group = forcats::fct_rev(ordered(.width)))) +
+      geom_lineribbon(alpha = 0.3) +
       scale_color_manual(values = c(clrs[1], clrs[8])) +
-      scale_fill_manual(values = c(clrs[1], clrs[8])) +
       scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
@@
       theme_pandem()
+  })) %>% 
+  mutate(mfx_plot = map2(mfx_data, nice, ~{
+    .x %>% 
+      mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>% 
+      ggplot(aes(x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+                 group = forcats::fct_rev(ordered(.width)))) +
+      geom_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
+      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
+      labs(x = NULL,
+           y = "Percentage point difference\nin probability of policy",
+           title = .y) +
+      theme_pandem()
   }))
 
-wrap_plots(preliminary_models_plots$pred_plot)
+wrap_plots(prelim_plots$pred_plot)
 ```
@@
 ```{r fig.width=12, fig.height=6, out.width="100%"}
-preliminary_models_mfx <- m_prelim_derog %>% 
-  mutate(mfx = map(model, ~{
-    .x %>% 
-      marginaleffects(newdata = datagrid(year_week_num = 1:69),
-                      variables = "derogation_ineffect",
-                      type = "response") |> 
-      posteriordraws()
-  })) %>% 
-  mutate(mfx_plot = map2(mfx, nice, ~{
-    .x %>% 
-      left_join(year_week_lookup, by = "year_week_num") %>% 
-      ggplot(aes(x = year_week_day, y = draw * 100)) +
-      stat_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
-      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-      labs(x = NULL,
-           y = "Percentage point change\nin probability of policy",
-           title = .y) +
-      theme_pandem()
-  }))
-
-wrap_plots(preliminary_models_mfx$mfx_plot)
+wrap_plots(prelim_plots$mfx_plot)
 ```

### 2023-03-07 R/funs_models.R
lol typo

@@ -44,7 +44,7 @@ gof_map <- tribble(
 coef_map <- c(
   "b_derogation_ineffect" = "Derogation in effect",
   "b_v2csreprss" = "Civil society repression",
-  "b_derogation_ineffect:v2csreprss" = "Derogation × Civil society represstion",
+  "b_derogation_ineffect:v2csreprss" = "Derogation × Civil society repression",
   "b_new_cases_z" = "New cases (standardized)",
   "b_cumulative_cases_z" = "Cumulative cases (standardized)",
   "b_new_deaths_z" = "New deaths (standardized)",

### 2023-03-08 manuscript/manuscript.qmd
Audrey and Suparna edits and additions

@@
   In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
-However, the abuse of emergency derogations by many governments raises concerns about the long-lasting impact of these sweeping measures on human rights.
-Does a robust civil society mitigate the adverse effects arising from international treaty derogations?
-Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run.
-Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we first assess the variation in human rights responses among governments during the pandemic.
-We then assess the role of civil society in keeping governments accountable and mitigating the effects of policies that violate human rights.
-Our results hold important implications for understanding government compliance with international treaties and shedding light on how states of emergency may directly and indirectly lead to democratic backsliding.
+Did states actually end up complying with their pandemic-era derogations?
+Did derogating and non-derogating states implement emergency measures differently, and if so, how?
+Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
+On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
+Our results support the role that flexibility mechanisms such as derogations play in international law.
 date: March 5, 2023
@@
 
-Combating a deadly virus like SARS-CoV-2 requires extraordinary public health measures.
+# Introduction
+
+Do emergency provisions during crises do more harm or good when it comes to human rights?
+Combating a deadly virus like SARS-CoV-2 required extraordinary public health measures.
 In an effort to fight the COVID-19 pandemic, governments across the world imposed restrictions on the freedom of movement, association, assembly and privacy.
-While countries are ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights principles.
-Some human rights treaties recognize states' needs for emergency measures when confronted with public health concerns and contain escape clauses that establish safeguards to prevent permanent erosion of these rights during emergencies.
-This temporary suspension of state compliance with its international treaty provisions occurs through a process known as derogations.
-However, the abuse of emergency measures during these derogations raises concerns about the long-lasting impact of these measures on human rights.
+While countries were ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights law.
+Some human rights treaties recognize states' needs for emergency measures during public health concerns and contain escape clauses that establish safeguards to prevent permanent erosion of these rights during emergencies.
+The temporary suspension of state compliance with its international treaty obligations occurs through a legal action known as derogation.
+However, the potential abuse of emergency measures during  derogations raises concerns about the long-lasting impact of emergency measures on human rights.
 Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].
 
-Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that in May 2023 it would be ending the emergency measures instated at the beginning of the pandemic—a move that that has occurred in many other countries as well.
-With the lifting of these measures, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this period.
+Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023, ending measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
+With the lifting of these measures, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this emergency period.
 What did derogations mean for compliance?
+Did states deviate from treaty provisions when they said they would?
 Are derogating and non-derogating states different in terms of how they implemented emergency measures during COVID-19?
@@
 
-The violations of human rights during this period also raises questions about factors that can mitigate the adverse impacts of emergency measures.
+The potential violations of human rights during this period also raises questions about factors that can mitigate the adverse impacts of emergency measures.
 Existing literature on emergencies such as wars and natural disasters shows that civil society organizations (CSOs) play a pivotal role during these periods [@pellingdill; @cash2013].
 They provide services and resources, and hold governments accountable to ensure inclusive policies during these emergencies.
+Human rights compliance literature also generally finds that NGOs and civil society are key actors in facilitating treaty obligation actualization [e.g., @KeckSikkink:1998; @Merry:2009].
 Can these groups also prevent or reduce the likelihood of human rights violations during public health emergencies?
@@
 In this research note, we focus on one such legal addendum—the treaty derogation.
-If and when states adhere to the written expectations of treaty derogations, they will be suspending human rights recognition during asset time of domestic emergency and crisis.
-If states are complying the derogation expectations, we should in fact expect a reduction in compliance with the overall treaty during that time and human rights standards.
-We explore how the potential decline in compliance may actually support human rights during the COVID-19 period since emergency mitigation strategies were targeting a reduction in disease outbreak.
-Human rights compliance literature generally finds that NGOs and civil society are key actors in facilitating treaty obligation actualization [e.g., @KeckSikkink:1998; @Merry:2009].
-We explore what, if any, role civil society organizations had during this period.
+If and when states adhere to the written expectations of treaty derogations, they will be suspending human rights recognition during a set time of domestic emergency and crisis.
+If states are complying with the derogation expectations, we should in fact expect a reduction in compliance with the overall human rights treaty during that time.
+We explore how the potential decline in compliance may actually support human rights during the COVID-19 period since emergency mitigation strategies also require certain rules and standards to prevent the abuse of rights.
 
-Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data we first assess the variation in human rights responses among governments during the pandemic.
+Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation in human rights responses among governments during the pandemic.
 We then analyze the effect of derogating (or not) on human rights practices during the pandemic and assess the role of civil society in mitigating the effects of rights-violating policies.
-Our main takeaways are that derogating states did largely what they signaled they would through the derogations.
-States that derogated and specified time limits to the emergency provisions generally stuck to the time limits and did implement the emergency practices they signaled they would.
-Additionally, we find that, on average, derogating states are less likely to abuse emergency measures.
-On the other hand, states that did not derogate from their international human rights obligations and still impose emergency measures are more likely impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
+We find that states that derogated and specified time limits to the emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies, the abusive enforcement of these policies, and not violating non-derogable rights.
+On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
+
 We also find that civil society can play an important mitigating role in curbing discriminatory policies.
-States with low levels of civil society repression are the least likely to abuse human rights when enforcing emergency policies.
-The likelihood of discriminatory policies increase as civil society repression worsens, but derogations appear to protect against worsening abuses—even in countries with severe civil society repression, those that derogate have a nearly 100% probability of no discriminatory policies.
+States with low levels of civil society repression were the least likely to abuse human rights when enforcing emergency policies.
+The likelihood of discriminatory policies increased as civil society repression worsened, but derogations appear to protect against worsening abuses—even in countries with severe civil society repression, those that derogated had a nearly 100% probability of no discriminatory policies.
 
-This research note makes three contributions.
+This research note makes two contributions.
 First, while there is a significant literature on compliance with international law, especially human rights treaties [e.g.
@@
 Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Simmons2009; @Merry:2009].
-But we know much less about how international treaty provisions intended to safeguard rights might actually worsen human rights outcomes.
 While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
 @Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
-TODO: EXPAND ON RESULT In this scenario, when international factors may not be enough to constrain state actions, assessing the effect of domestic-level factors such as a robust civil society that potentially can mitigate the adverse effects of derogations becomes essential.
-TODO: WE FIND.
+We find that overall, derogating states do not increase abuse and violations.
+This finding speaks to international legal compliance scholarship and supports the idea that states are being sincere about their intentions and not, byinlarge, using actions like derogation to cover abusive behavior.
 
-Finally, studying states' compliance with their international human rights treaty obligations as well as the (unintended) effects of derogations is important given trends in democratic backsliding and declining respect for human rights.
-Nearly 75 percent of the world’s population now lives in a country that faced democratic deterioration last year [@freedomhouse2021].
-The pandemic led to further backsliding with many countries around the world postponing elections, arresting protestors, and cracking down on citizens to consolidate power—all under the guise of public health measures.
-These periods of emergencies may be an important piece to understand the processes of autocratization and democratic backsliding since leaders often use emergency decrees to  allow states to aggrandize executive power [@LuhrmannRooney2021; @LevitskyZiblatt:2018; @Bermeo2016].
-The temporary suspension of states' international legal obligation is one possible pathway that could lead to long-term rights violations and democratic erosion.
-
-Below, we first explore the purpose of derogations during the pandemic.
+The rest of the paper proceeds as follows.
+First, we explore the purpose of derogations during the pandemic.
 In doing so, we focus specifically on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
 Following this, we introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic.
-We then assess states' behavior towards civil society and these organizations' role in mitigating the adverse effects of emergency measures.
+We then assess states' behavior towards civil society and these organizations' potential role in mitigating the adverse effects of emergency measures.
 We conclude with the implications of our findings for the study of international law, times of crises, and the role of civil society.
@@
 Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility in responding to events.
-These times of crises include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
+Times of crises can include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
 Derogations afford states the opportunity to adjust to these changing conditions [@HBHelferFariss].
-Once states derogating from their international legal obligations, it also provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
+When states derogate from their international legal obligations,the derogation provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
 This information allows these actors—at least in principle—to challenge measures that are excessive, vague, or outlast the intended time frame of their implementation [@helfer2021].
 
-While legal critical assessments of the use of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially led to the violation of non-derogable rights—rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that cannot be suspended even during emergencies.
+While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially lead to the violation of non-derogable rights—rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
 Research finds results that support this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
@@
 non-derogable rights status during emergencies.
-This also raises the question on the extent to which this distinction is a valuable one during the pandemic—which is a public health emergency different from other states of emerency normally brought on by war or internal political unrest.
+This also raises the question of the extent to which this distinction is a valuable one during the pandemic.
+Most derogations to the ICCPR have been for internal conflict or natural disaster - is a public health emergency different from other states of emergency?
 Overall, there is a general expectation that following derogations, we may observe some decline in respect for human rights.
+However, previous studies have not explored the relationship between derogations and public health emergencies.
 
@@
 
-Research finds that democracies, especially those with strong rule of law, are more likely to derogate [@HBHelferFariss].
+A first question of interest before we examine the potential impact of derogations is what states are more likely to submit them in the first place.
+Though not much research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the UN rather than implications of when they are in effect.
+This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss].
 The expectation is that states with strong democratic institutions need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
-
-A first question of interest before we examine the potential impact of derogations is what states are more likely to make them in the first place.
-Though not much research has focused on derogations in political science, most of it has focused on the determinants of submitting derogations to the UN rather than implications of their application.
-This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss].
-The argument is that democratic states understand international law and international legal mechanisms to be more legitimate and recognize their importance more so than other states.
+Additionally, democratic states understand international law and international legal mechanisms to be more legitimate and recognize their importance more so than other states.
 Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important.
-Looking to current derogation practices, there is not a perfect overlap in the states that issued derogations in the past towards the ICCPR and those that did during the COVID-19 pandemic.
-Of course, since derogations are a type of action submitted after treaty commitment, the only states that would be submitting them are ones that already have ratified or otherwise committed to the ICCPR.
-As of 2023, there are 173 states party to the ICCPR.
-This means that most, but not all of the UN’s 193 members had access to submitting derogations during COVID-19.
-Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.
-
-Some states new to derogations introducing them during the COVID-19 pandemic [@Comstock:2023].
-However, past treaty action behavior may help explain entering derogations during this time.
-To examine this first step, we code two variables 1) past ICCPR derogation behavior and 2) past ICCPR post-commitment action behavior.
-States receive a 1 if they engaged in this legal behavior before 2020 and 0 if they did not.
-This will allow us to empirically test for the first step of engagement before analyzing any impact COVID-19 era derogations had on different rights practices.
-There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019.
 
@@
 In this research note, we look at states' derogations from the International Covenant on Civil and Political Rights (ICCPR) during the pandemic.
-ICCPR is arguably one of the most extensive human rights treaties and the exceptional circumstances brought on by the global pandemic have led to numerous extensive derogations to the ICCPR.
-The ICCPR was adopted by the UN General Assembly on December 19, 1966 and came into force on March 23, 1976.
+ICCPR is a foundational human rights treaty with broad human rights coverage.
+Exceptional circumstances brought on by the global pandemic have led to numerous, extensive derogations to the ICCPR removing state obligations to important rights areas of assembly, movement, and others.
+The ICCPR was adopted by the UN General Assembly on December 19, 1966 and entered into force on March 23, 1976.
 It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights.
@@
 However, the ICCPR is the only one with global, rather than regional, membership and participation to track during the pandemic.
+As of 2023, there are 173 states party to the ICCPR.
+This means that most, but not all of the UN’s 193 members had the ability to submit derogations during COVID-19.
+Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.
 
 To legally derogate, government must first officially declare a state of emergency (Article 4).
-These derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory.
+Derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory.
 However, it is worth noting that states cannot derogate from the entirety of the ICCPR.
-Seven rights as enshrined in the ICCPR are non-derogable: right to life, prohibit torture or cruel, inhumane punishment, prohibit slavery, provide the right to recognition as a person before the law, and protect freedom freedom of thought, conscience and religion.
-On the other hand, several ICCPR Articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
+Seven rights enshrined in the ICCPR are non-derogable: the right to life, the prohibition of torture or cruel, inhumane punishment,the prohibition of slavery, the right to recognition as a person before the law, and the protection of freedom of thought, conscience and religion.
+The international legal community viewed these rights as so fundamental and important that no type of crisis could justify their suspension.
+Other types of rights are seen as able to be suspended during crises.
+Several ICCPR Articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
 
 Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination.
-However,not every country formally derogated—a number of countries (HOW MANY) implemented emergency measures without formally derogating from ICCPR.
-The impact of these emergencies may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020].
+However,not every country formally derogated—a number of countries implemented emergency measures without formally derogating from ICCPR.
+Over 100 countries issued emergency declarations while fewer than 30 submitted COVID-19 related derogations [@icnl-covid].
+These COVID-19 emergency declarations may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020].
 Below, we first theorize how derogations can protect against rights violations and the excesses of emergency measures, followed by the role of civil society in mitigating these violations.
@@
 
-If states determine that emergency measures are justified without regard for international law, then we expect that states which did not file for derogations to the ICCPR were less likely to meet the criteria for derogations—necessity, proportionality, non-discrimination and time delimited, and therefore more likely to impose measures that violated these criteria.
+We expect that states that did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the the criteria for derogations—necessity, proportionality, non-discrimination and time delimited seriously, and therefore more likely to impose measures that violated these criteria.
+If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.
 
@@
 
-Given the divergent state responses during the pandemic, these differing responses can have lasting impacts on human rights trends.
+Given the divergent state responses during the pandemic, differing responses can have lasting impacts on human rights trends.
 In the absence of adequate review from the international community during the pandemic, certain domestic-level factors may be necessary to mitigate the adverse effects of emergency measures.
-While both the theoretical and empirical literature has established that states may violate both derogable and non-derogable rights during emergencies, we have comparatively little evidence on factors that may safeguard against these rights violations during periods of derogation.[^exceptions]
+While both the theoretical and empirical literature has established that states may violate both derogable and non-derogable rights during emergencies, we have comparatively little evidence regarding factors that may safeguard against these rights violations during periods of derogation.
 
@@
 Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run [@fair2017].
-
 As such, we expect that a strong and vibrant civil society may help mitigate the adverse impact arising from derogations and emergency measures imposed by governments during the pandemic.
@@
 
-We use three different primary independent variables to explain changes in these policy and human rights outcomes.
+We use three different primary independent variables to explain changes in policy and human rights outcomes.
 For our preliminary models and in models related to our first hypothesis, we rely on a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
@@
 Finally, we include several other variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
-To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID cases and the count of new COVID deaths, as well as the cumulative count of cases and deaths.
-Since formally derogating from treaties can serve as a signal that countries take their obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR bureaucratic actions prior to the pandemic.
+To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
+
+Since formally derogating from treaties can serve as a signal that countries take their obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR post-commitment actions, such as RUDs, prior to the pandemic.
+Some states new to derogations introduced them during the COVID-19 pandemic [@Comstock:2023].
+However, past treaty action behavior may help explain entering derogations during this time.
+To examine this first step, we code two binary variables: (1) past ICCPR derogation behavior before 2020 and (2) past ICCPR post-commitment action behavior before 2020.
+This allows us to empirically test for the first step of engagement before analyzing any impact COVID-19 era derogations had on different rights practices.
+There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019.
 Finally, we include several annual measures of human rights, democracy, and civil society: V-Dem's rule of law index, civil liberties index, and core civil society index.
@@
 
-In all model specifications, we include new and cumulative COVID cases and deaths as controls.
+In all model specifications, we include new and cumulative COVID-19 cases and deaths as controls.
 In our first hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society.
@@
 Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.
-The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way—are unclear.
+
+The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems.
+While elites or diplomats are the ones submitting derogations and committing to complying with international law, they are not the actors implementing emergency measures.
+During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially when it relates to human rights.
+Lower-level officials such as those in the police, state security forces and regional bureaucrats are often the ones enforcing emergency measures.
+Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations.
+Further, as previous research shows, during high levels of stress and fear, decisionmaking is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
+As such, these differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations and claim that emergency measures will be implemented proportional to the threat at hand.
 
@@
 We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~1b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~1c~).
-Derogations seem help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
+Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
 
@@
 We thus only find support for H~2b~—if states are repressive toward civil society organizations, they are more likely to use discriminatory policies, while derogations strongly negate this trend.
-This may be advocacy-focused civil society organizations in particular tend to deal with on-the-ground monitoring, advocacy, and shaming, and is more likely to call out governments for unequal and unjust implementation of their emergency policies.
+This may be because advocacy-focused civil society organizations in particular tend to deal with on-the-ground monitoring, advocacy, and shaming, and is more likely to call out governments for unequal and unjust implementation of their emergency policies.
 
@@
 
-# Discussion
+# Discussion and conclusion
 
+The first few months of the pandemic raised numerous concerns that the pandemic would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
+
+Derogations are an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
+This research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.
+
+Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties without concern about the inability to adapt to times of emergency.
+International legal scholarship has been divided over expectations of derogations during the pandemic.
 @Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
@@
 
-Is this good news? Derogators aren't using this opportunistically - derogators are doing the process correctly
+Looking back to our driving questions, what we find is that there is good news for compliance when we look to derogations.
+Derogators did not systematically use this legal 'escape clause' opportunistically—derogators used the process correctly.
+Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
+Though not many states submitted COVID-19 related derogations, the ones that did, it seems, used derogations as intended by international law.
+Once the time of derogation ended, states readjusted to typical compliance once more.
+States, therefore complied with the intent of derogations and the purpose of the treaty.
+This is an important finding for international legal behavior and compliance studies.
+Even in a time of crisis, overall, states considered it important to follow through with the processes and expectations of international human rights law bureaucracy.
 
-Enforcement is carried out by lower level officials - principal agent issues and local bureaucrats not caring about international law - the time limit and nonderogable rights things are done by high level people
+We suggest that future research look to differences in levels within states during crisis.
+There may be key differences in enforcement between local level bureaucrats and higher level elites.
+While higher level diplomats submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.
+non-derogable rights.
+This distinction may explain some variation in policy and discrimination practices during the pandemic.
 
-# Conclusion
+We suggest future research look to other times of crisis and derogation use to compare COVID-19 area derogation compliance with that in other crises.
+This note contributed to the study of derogations by building on the analysis between derogation submission and outcome.
+States did end up taking the restriction measures they signaled they would via the derogations.
 
-In conclusion mention that future research can point to bureaucrats vs. policy makers and
-
-Contribution = this is good news
-
-This research note examines states' responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and its subsequent impact on human rights violations.
-
-Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties.
-
-The first few months of the pandemic raised numerous concerns that the pandemic would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide
+Overall, we find that states used derogations as they were intended during the COVID-19 pandemic and no large-scale misuse of the legal action took place.
+It can be an important signal for states to submit derogations, both to international and domestic communities, about intentions to mitigate the spread of disease through the temporary suspension of human rights.
 

### 2023-03-08 manuscript/manuscript.qmd
Clean up tables + minor edits

@@
 Our results support the role that flexibility mechanisms such as derogations play in international law.
-date: March 5, 2023
+date: March 8, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 link-citations: true
-format: 
-  html:
-    appendix-style: plain
-    toc: true
-  pdf:
-    mainfont: "IBM Plex Serif"
-    sansfont: "Noto Sans"
-    linestretch: 1.25
-    toc: false
-language: 
-  title-block-published: "Latest version"
 ---
@@
 ```{r setup, include=FALSE}
+fmt_out <- knitr::pandoc_to()
+
 knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
@@
 ```
-
-# Introduction
 
@@
 
-With many governments trying to centralize authority and decisionmaking, a free and independent media may pose challenges similar to those posed by CSOs.
+With many governments trying to centralize authority and decisionmaking, a free and independent media may pose challenges similar to those posed by CSOs [@Themudo:2013].
 As such, if states are engaging in repression of CSOs during the pandemic, they may be more likely to repress the media as well.
@@
 ) %>% 
-  kbl(booktabs = TRUE) %>% 
-  kable_styling() %>% 
+  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]") %>% 
+  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA)) %>% 
   pack_rows(index = c(" " = 2, "Primary explanatory variables" = 2, "Secondary explanatory variables" = 5),
-            indent = FALSE)
+            indent = FALSE) %>% 
+  column_spec(c(1, 2, 4), width = "0.25\\\\textwidth") %>% 
+  column_spec(3, width = "0.1\\\\textwidth")
+```
 
-```
+\newpage
 
@@
 
+\begingroup\footnotesize
+
 | General question                                    | Outcomes                                                                                                                                  | Main explanatory variable                   | Other explanatory variables                                                                                                                                              |
@@
 : Overview of outcomes and explanatory variables included in models {#tbl-vars}
+
+\endgroup
 
@@
 Our data is structured as a balanced time series cross-sectional (TSCS) panel that has a natural hierarchical structure, with `r info$n_weeks` weekly observations nested within `r info$n_countries` countries.
-Accordingly, we model our outcome variables with multilevel or hierarchical models, with country-specific random offsets for each intercept and a fixed global year trend.
+Accordingly, we model our outcome variables with multilevel or hierarchical models, with region-specific random offsets for each intercept and a fixed global year trend.
 When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, use use ordered logistic regression models.
@@
 &\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}, \alpha_k) \\
-\\
+\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
 &\ \textbf{Distribution parameters} \\
-\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} & \text{Probability of outcome} \\
+\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + & \text{Probability of outcome} \\
 &\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\
-\\
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random region offsets} \\[0.75em]
 &\ \textbf{Priors} \\
 \beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-country variability,}
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-region variability,}
 \end{aligned}
@@
 
-while for ordered outcomes, we use ordered logistic regression with country-specific offsets:
+\noindent while for ordered outcomes, we use ordered logistic regression with country-specific offsets:
 
@@
 &\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\,logit}(\phi_{it_j}, \alpha_k) \\
-\\
+\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
 &\ \textbf{Distribution parameters} \\
-\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} & \text{Cumulative probability of outcome} \\
+\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + & \text{Cumulative probability of outcome} \\
 &\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\
-\\
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random region offsets} \\[0.75em]
 &\ \textbf{Priors} \\
 \beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-country variability} \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-region variability} \\
 \alpha_k \sim&\ \mathcal{N}(0, 1) & \text{Boundaries between thresholds}
@@
 
-The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems.
-While elites or diplomats are the ones submitting derogations and committing to complying with international law, they are not the actors implementing emergency measures.
-During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially when it relates to human rights.
-Lower-level officials such as those in the police, state security forces and regional bureaucrats are often the ones enforcing emergency measures.
-Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations.
-Further, as previous research shows, during high levels of stress and fear, decisionmaking is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
-As such, these differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations and claim that emergency measures will be implemented proportional to the threat at hand.
-
-In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status.
-For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability).
-The middle ground of only minor issues with time limits is exceptionally unlikely.
-Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues.
-The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
-Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)
-
-Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
-We thus find strong support only for H~1c~—that derogating states are less likely to impose measures with no time limits.
-We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~1b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~1c~).
-Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
-
 ```{r fig-preds-mfx-h1, fig.width=13, fig.height=6, out.width="100%"}
@@
 ```
+
+The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems.
+While elites or diplomats are the ones submitting derogations and committing to complying with international law, they are not the actors implementing emergency measures.
+During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially when it relates to human rights.
+Lower-level officials such as those in the police, state security forces and regional bureaucrats are often the ones enforcing emergency measures.
+Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations.
+Further, as previous research shows, during high levels of stress and fear, decisionmaking is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
+As such, these differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations and claim that emergency measures will be implemented proportional to the threat at hand.
+
+In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status.
+For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability).
+The middle ground of only minor issues with time limits is exceptionally unlikely.
+Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues.
+The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
+Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)
+
+Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
+We thus find strong support only for H~1c~—that derogating states are less likely to impose measures with no time limits.
+We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~1b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~1c~).
+Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
 
@@
 Looking back to our driving questions, what we find is that there is good news for compliance when we look to derogations.
-Derogators did not systematically use this legal 'escape clause' opportunistically—derogators used the process correctly.
+Derogators did not systematically use this legal "escape clause" opportunistically—derogators used the process correctly.
 Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
@@
 
-### Formal statistical models
+### H~1~: Derogations and human rights (ordered logistic regression)
 
@@
 &\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered logit}(\phi_{it_j}, \alpha_k) \\
-\\
+\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
 &\ \textbf{Models for distribution parameters} \\
-\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} \\
+\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
 &\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
@@
 &\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\
-\\
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
 &\ \textbf{Priors} \\
-\beta_{0 \dots 11} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
 \sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
@@
 
-### Complete regression results
+### H~1~: Derogations and human rights (logistic regression)
 
-#### Derogations and emergency policies
+$$
+\begin{aligned}
+&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
+\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
+&\ \textbf{Distribution parameters} \\
+\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
+&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
+&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
+&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
+&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
+&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
+&\ \textbf{Priors} \\
+\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
+\end{aligned}
+$$
+
+### H~2~: Derogations, civil society, and human rights (ordered logistic regression)
+
+$$
+\begin{aligned}
+&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
+\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
+&\ \textbf{Models for distribution parameters} \\
+\phi_{it_j} =&\ (\beta_0 + b_{0_j})\ + \\
+&\ \beta_1 \text{Derogation in effect}_{it} + \beta_2 \text{Civil society repression}_{it} \\
+&\ \beta_3\ {(\text{Derogation} \times \text{Civil society repression})}_{i}\ + \\
+&\ \beta_4\ \text{New cases}_{it}\ + \beta_5\ \text{Cumulative cases}_{it}\ + \\
+&\ \beta_6\ \text{New deaths}_{it}\ + \beta_7\ \text{Cumulative deaths}_{it}\ + \\
+&\ \beta_8\ \text{Past ICCPR derogation}_{it}\ + \beta_9\ \text{Past ICCPR action}_{it}\ + \\
+&\ \beta_{10}\ \text{Rule of law index}_{it}\ + \beta_{11}\ \text{Civil liberties index}_{it}\ + \\
+&\ \beta_{12}\ \text{Week number}_{it} \\
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
+&\ \textbf{Priors} \\
+\beta_{0 \dots 12} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
+\alpha_k \sim&\ \mathcal{N}(0, 1)
+\end{aligned}
+$$
+
+### H~2~: Derogations, civil society, and human rights (logistic regression)
+
+$$
+\begin{aligned}
+&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
+\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
+&\ \textbf{Distribution parameters} \\
+\pi_{it_j} =&\ (\beta_0 + b_{0_j})\ + \\
+&\ \beta_1 \text{Derogation in effect}_{it} + \beta_2 \text{Civil society repression}_{it} \\
+&\ \beta_3\ {(\text{Derogation} \times \text{Civil society repression})}_{i}\ + \\
+&\ \beta_4\ \text{New cases}_{it}\ + \beta_5\ \text{Cumulative cases}_{it}\ + \\
+&\ \beta_6\ \text{New deaths}_{it}\ + \beta_7\ \text{Cumulative deaths}_{it}\ + \\
+&\ \beta_8\ \text{Past ICCPR derogation}_{it}\ + \beta_9\ \text{Past ICCPR action}_{it}\ + \\
+&\ \beta_{10}\ \text{Rule of law index}_{it}\ + \beta_{11}\ \text{Civil liberties index}_{it}\ + \\
+&\ \beta_{12}\ \text{Week number}_{it} \\
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
+&\ \textbf{Priors} \\
+\beta_{0 \dots 12} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
+\end{aligned}
+$$
+
+\newpage
 
 ```{r tbl-results-full-prelim}
-#| tbl-cap: Complete results from models showing relationship between derogations and emergency policies
+#| tbl-cap: Results from models showing relationship between derogations and emergency policies
 #| include: true
@@
 notes <- c(
-  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models; 95% credible intervals (highest density posterior interval, or HDPI) in brackets.", 
-  "Total \\(R^2\\) considers the variance of both population and group effects;", 
-  "marginal \\(R^2\\) only takes population effects into account."
+  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
+  r"(95\\% credible intervals (highest density posterior interval, or HDPI) in brackets.)",
+  "Total $R^2$ considers the variance of both population and group effects;",
+  "marginal $R^2$ only takes population effects into account."
 )
@@
              output = "kableExtra",
-             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-  kable_styling(htmltable_class = "table-sm light-border") %>% 
-  footnote(general = notes, footnote_as_chunk = TRUE)
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
+             escape = FALSE) %>% 
+  kable_styling(htmltable_class = "table-sm light-border",
+                font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
+  footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
+  column_spec(1:6, width = "0.16\\\\textwidth")
 ```
 
-#### Derogations and human rights
+\newpage
 
@@
              output = "kableExtra",
-             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-  kable_styling(htmltable_class = "table-sm light-border") %>% 
-  footnote(general = notes, footnote_as_chunk = TRUE)
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
+             escape = FALSE) %>% 
+  kable_styling(htmltable_class = "table-sm light-border",
+                font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
+  footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
+  column_spec(1:5, width = "0.18\\\\textwidth")
 ```
 
-#### Derogations, civil society, and human rights
+\newpage
 
 ```{r tbl-results-full-h2}
-#| tbl-cap: Complete results from models showing relationship between derogations, civil society, and human rights (H~2~)
+#| tbl-cap: Models showing relationship between derogations, civil society, and human rights (H~2~)
 #| include: true
@@
              output = "kableExtra",
-             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-  kable_styling(htmltable_class = "table-sm light-border") %>% 
-  footnote(general = notes, footnote_as_chunk = TRUE)
+             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
+             escape = FALSE) %>% 
+  kable_styling(htmltable_class = "table-sm light-border",
+                font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
+  footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
+  column_spec(1:5, width = "0.18\\\\textwidth")
 ```
 
+\clearpage
+
 # References

### 2023-03-08 analysis/manuscript.qmd
Add special manuscript iframe page

---
title: Manuscript
toc: false
page-layout: full
---

<p class="text-center"><a class="btn btn-success" href="/manuscript/output/manuscript.html" target="manuscript">{{< fa file-lines >}} &ensp;HTML</a> <a class="btn btn-success" href="/manuscript/output/manuscript.html" target="_blank">{{< fa arrow-up-right-from-square >}} &ensp;HTML (new window)</a> <a class="btn btn-success" href="/manuscript/output/manuscript.pdf" target="manuscript">{{< fa file-pdf >}} &ensp;PDF</a> <a class="btn btn-success" target="_blank" href="/manuscript/output/manuscript.pdf">{{< fa file-arrow-down >}} &ensp;PDF (download)</a></p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="manuscript"></iframe>

### 2023-03-08 index.qmd
Use newer abstract

--- 2023-03-06 manuscript/manuscript.qmd
@@
 title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
-subtitle: "Research note"
-author:
-- name: Suparna Chaudhry
-  affiliation: Lewis and Clark College
-  email: schaudhry@lclark.edu
-- name: Audrey L. Comstock
-  affiliation: Arizona State University
-  email: audrey.comstock@asu.edu
-- name: Andrew Heiss
-  affiliation: Georgia State University
-  email: aheiss@gsu.edu
-abstract: |
-  In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
-However, the abuse of emergency derogations by many governments raises concerns about the long-lasting impact of these sweeping measures on human rights.
-Does a robust civil society mitigate the adverse effects arising from international treaty derogations?
-Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run.
-Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we first assess the variation in human rights responses among governments during the pandemic.
-We then assess the role of civil society in keeping governments accountable and mitigating the effects of policies that violate human rights.
-Our results hold important implications for understanding government compliance with international treaties and shedding light on how states of emergency may directly and indirectly lead to democratic backsliding.
-date: March 5, 2023
-code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
-bibliography: bib.bib
-link-citations: true
 format:
   html:
-    appendix-style: plain
-    toc: true
-  pdf:
-    mainfont: "IBM Plex Serif"
-    sansfont: "Noto Sans"
-    linestretch: 1.25
-    toc: false
-language: 
-  title-block-published: "Latest version"
+    code-tools: false
+sidebar: false
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
-                      warning = FALSE, message = FALSE)
+- Suparna Chaudhry (Lewis & Clark College)
+- Audrey L. Comstock (Arizona State University)
+- Andrew Heiss (Georgia State University)
 
-knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
-                      fig.width = 6, fig.height = (6 * 0.618),
-                      out.width = "80%", collapse = TRUE)
-
-options(digits = 3, width = 120,
-        dplyr.summarise.inform = FALSE,
-        knitr.kable.NA = "")
-```
-
-```{r libraries-data}
-library(tidyverse)
-library(scales)
-library(lubridate)
-library(tidybayes)
-library(marginaleffects)
-library(patchwork)
-library(colorspace)
-library(modelsummary)
-library(kableExtra)
-library(targets)
-
-tar_config_set(store = here::here('_targets'),
-               script = here::here('_targets.R'))
-
-# Generated via random.org
-set.seed(8511)
-
-# Load targets 
-weekly_panel <- tar_read(weekly_panel) %>% 
-  mutate(year_week_day = ymd(year_week_day))
-
-tar_load(c(prelim_plot_data, h1_plot_data, h2_plot_data))
-tar_load(c(models_tbl_prelim, models_tbl_h1, models_tbl_h2))
-
-# Plotting and modelsummary functions
-invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
-invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
-```
-
-Combating a deadly virus like SARS-CoV-2 requires extraordinary public health measures.
-In an effort to fight the COVID-19 pandemic, governments across the world imposed restrictions on the freedom of movement, association, assembly and privacy.
-While countries are ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights principles.
-Some human rights treaties recognize states' needs for emergency measures when confronted with public health concerns and contain escape clauses that establish safeguards to prevent permanent erosion of these rights during emergencies.
-This temporary suspension of state compliance with its international treaty provisions occurs through a process known as derogations.
-However, the abuse of emergency measures during these derogations raises concerns about the long-lasting impact of these measures on human rights.
-Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].
-
-Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that in May 2023 it would be ending the emergency measures instated at the beginning of the pandemic—a move that that has occurred in many other countries as well.
-With the lifting of these measures, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this period.
-What did derogations mean for compliance?
-Are derogating and non-derogating states different in terms of how they implemented emergency measures during COVID-19?
-Were derogating states more likely to exploit the provisions and to abuse human rights during COVID-19?
-These questions are important for the study of international law, compliance, and times of crises.
-
-The violations of human rights during this period also raises questions about factors that can mitigate the adverse impacts of emergency measures.
-Existing literature on emergencies such as wars and natural disasters shows that civil society organizations (CSOs) play a pivotal role during these periods [@pellingdill; @cash2013].
-They provide services and resources, and hold governments accountable to ensure inclusive policies during these emergencies.
-Can these groups also prevent or reduce the likelihood of human rights violations during public health emergencies?
-Does a robust civil society mitigate the adverse effects arising from international treaty derogations during the pandemic?
-
-Typically, when international relations scholars write about human rights treaty compliance, they write of positive human rights recognition.
-Compliance is an adherence to the expectations and written standards within the treaties, often described as a "state of conformity or identity between an actor’s behavior and a specified rule" [@RaustialaSlaughter:2002, 539].
-However, there are treaty qualifications that are legal addendums to commitment that alter or even limit compliance expectations.
-When states comply with these legal actions, they are in fact, limiting their obligations to human rights standards.
-In this research note, we focus on one such legal addendum—the treaty derogation.
-If and when states adhere to the written expectations of treaty derogations, they will be suspending human rights recognition during asset time of domestic emergency and crisis.
-If states are complying the derogation expectations, we should in fact expect a reduction in compliance with the overall treaty during that time and human rights standards.
-We explore how the potential decline in compliance may actually support human rights during the COVID-19 period since emergency mitigation strategies were targeting a reduction in disease outbreak.
-Human rights compliance literature generally finds that NGOs and civil society are key actors in facilitating treaty obligation actualization [e.g., @KeckSikkink:1998; @Merry:2009].
-We explore what, if any, role civil society organizations had during this period.
-
-Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data we first assess the variation in human rights responses among governments during the pandemic.
-We then analyze the effect of derogating (or not) on human rights practices during the pandemic and assess the role of civil society in mitigating the effects of rights-violating policies.
-Our main takeaways are that derogating states did largely what they signaled they would through the derogations.
-States that derogated and specified time limits to the emergency provisions generally stuck to the time limits and did implement the emergency practices they signaled they would.
-Additionally, we find that, on average, derogating states are less likely to abuse emergency measures.
-On the other hand, states that did not derogate from their international human rights obligations and still impose emergency measures are more likely impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
-We also find that civil society can play an important mitigating role in curbing discriminatory policies.
-States with low levels of civil society repression are the least likely to abuse human rights when enforcing emergency policies.
-The likelihood of discriminatory policies increase as civil society repression worsens, but derogations appear to protect against worsening abuses—even in countries with severe civil society repression, those that derogate have a nearly 100% probability of no discriminatory policies.
-
-This research note makes three contributions.
-First, while there is a significant literature on compliance with international law, especially human rights treaties [e.g.
-@Simmons2009; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations.
-Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of the types of states that are likely to comply with these escape clauses.
-The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster typical of treaty derogations, making it easier to compare the long-term human rights effects of these derogations cross nationally.
-
-Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Simmons2009; @Merry:2009].
-But we know much less about how international treaty provisions intended to safeguard rights might actually worsen human rights outcomes.
-While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
-@Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
-TODO: EXPAND ON RESULT In this scenario, when international factors may not be enough to constrain state actions, assessing the effect of domestic-level factors such as a robust civil society that potentially can mitigate the adverse effects of derogations becomes essential.
-TODO: WE FIND.
-
-Finally, studying states' compliance with their international human rights treaty obligations as well as the (unintended) effects of derogations is important given trends in democratic backsliding and declining respect for human rights.
-Nearly 75 percent of the world’s population now lives in a country that faced democratic deterioration last year [@freedomhouse2021].
-The pandemic led to further backsliding with many countries around the world postponing elections, arresting protestors, and cracking down on citizens to consolidate power—all under the guise of public health measures.
-These periods of emergencies may be an important piece to understand the processes of autocratization and democratic backsliding since leaders often use emergency decrees to  allow states to aggrandize executive power [@LuhrmannRooney2021; @LevitskyZiblatt:2018; @Bermeo2016].
-The temporary suspension of states' international legal obligation is one possible pathway that could lead to long-term rights violations and democratic erosion.
-
-Below, we first explore the purpose of derogations during the pandemic.
-In doing so, we focus specifically on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
-Following this, we introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic.
-We then assess states' behavior towards civil society and these organizations' role in mitigating the adverse effects of emergency measures.
-We conclude with the implications of our findings for the study of international law, times of crises, and the role of civil society.
-
-# Derogations and international law
-
-Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility in responding to events.
-These times of crises include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
-Derogations afford states the opportunity to adjust to these changing conditions [@HBHelferFariss].
-Once states derogating from their international legal obligations, it also provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
-This information allows these actors—at least in principle—to challenge measures that are excessive, vague, or outlast the intended time frame of their implementation [@helfer2021].
-
-While legal critical assessments of the use of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially led to the violation of non-derogable rights—rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that cannot be suspended even during emergencies.
-Research finds results that support this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
-Even states that officially derogate from their international legal obligations are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].
-
-These findings suggest that states may not be discerning the significance of derogable v/s.
-non-derogable rights status during emergencies.
-This also raises the question on the extent to which this distinction is a valuable one during the pandemic—which is a public health emergency different from other states of emerency normally brought on by war or internal political unrest.
-Overall, there is a general expectation that following derogations, we may observe some decline in respect for human rights.
-
-## Derogations under COVID-19
-
-Research finds that democracies, especially those with strong rule of law, are more likely to derogate [@HBHelferFariss].
-The expectation is that states with strong democratic institutions need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
-
-A first question of interest before we examine the potential impact of derogations is what states are more likely to make them in the first place.
-Though not much research has focused on derogations in political science, most of it has focused on the determinants of submitting derogations to the UN rather than implications of their application.
-This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss].
-The argument is that democratic states understand international law and international legal mechanisms to be more legitimate and recognize their importance more so than other states.
-Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important.
-Looking to current derogation practices, there is not a perfect overlap in the states that issued derogations in the past towards the ICCPR and those that did during the COVID-19 pandemic.
-Of course, since derogations are a type of action submitted after treaty commitment, the only states that would be submitting them are ones that already have ratified or otherwise committed to the ICCPR.
-As of 2023, there are 173 states party to the ICCPR.
-This means that most, but not all of the UN’s 193 members had access to submitting derogations during COVID-19.
-Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.
-
-Some states new to derogations introducing them during the COVID-19 pandemic [@Comstock:2023].
-However, past treaty action behavior may help explain entering derogations during this time.
-To examine this first step, we code two variables 1) past ICCPR derogation behavior and 2) past ICCPR post-commitment action behavior.
-States receive a 1 if they engaged in this legal behavior before 2020 and 0 if they did not.
-This will allow us to empirically test for the first step of engagement before analyzing any impact COVID-19 era derogations had on different rights practices.
-There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019.
-
-## Derogations from the ICCPR
-
-In this research note, we look at states' derogations from the International Covenant on Civil and Political Rights (ICCPR) during the pandemic.
-ICCPR is arguably one of the most extensive human rights treaties and the exceptional circumstances brought on by the global pandemic have led to numerous extensive derogations to the ICCPR.
-The ICCPR was adopted by the UN General Assembly on December 19, 1966 and came into force on March 23, 1976.
-It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights.
-The ICCPR is one of only a few select treaties that requires states to report on the suspension or derogation of certain civil and political rights only under specific situations of "public emergency which threatens the life of the nation" [@iccpr].
-The other treaties with this requirement include the European Convention Protection of Human Rights and Fundamental Freedoms and the American Convention on Human Rights.
-However, the ICCPR is the only one with global, rather than regional, membership and participation to track during the pandemic.
-
-To legally derogate, government must first officially declare a state of emergency (Article 4).
-These derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory.
-However, it is worth noting that states cannot derogate from the entirety of the ICCPR.
-Seven rights as enshrined in the ICCPR are non-derogable: right to life, prohibit torture or cruel, inhumane punishment, prohibit slavery, provide the right to recognition as a person before the law, and protect freedom freedom of thought, conscience and religion.
-On the other hand, several ICCPR Articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
-
-Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination.
-However,not every country formally derogated—a number of countries (HOW MANY) implemented emergency measures without formally derogating from ICCPR.
-The impact of these emergencies may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020].
-Below, we first theorize how derogations can protect against rights violations and the excesses of emergency measures, followed by the role of civil society in mitigating these violations.
-
-# The impact of derogations on human rights outcomes
-
-During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21).
-If states determine that emergency measures are justified without regard for international law, the processes for restoring rights at the end of the emergency may not be followed.
-As such, many states used the pandemic an as excuse to centralize authority and implemented measures that didn’t meet international legal standards.
-Many states imposed restrictions on travel or movement by minority or opposition groups, limiting public assembly and democratic protests, and deployed surveillance systems that impeded privacy.
-For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
-In Bulgaria, checkpoint controls went into effect in neighborhoods largely composed of the Roma community [@bulgaria2020].
-In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country's new citizenship law seen as discriminatory towards Muslims [@bbc2020].
-
-If states determine that emergency measures are justified without regard for international law, then we expect that states which did not file for derogations to the ICCPR were less likely to meet the criteria for derogations—necessity, proportionality, non-discrimination and time delimited, and therefore more likely to impose measures that violated these criteria.
-
-> **Hypothesis 1~a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
->
-> **Hypothesis 1~b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
->
-> **Hypothesis 1~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
-
-## Civil society and human rights
-
-Given the divergent state responses during the pandemic, these differing responses can have lasting impacts on human rights trends.
-In the absence of adequate review from the international community during the pandemic, certain domestic-level factors may be necessary to mitigate the adverse effects of emergency measures.
-While both the theoretical and empirical literature has established that states may violate both derogable and non-derogable rights during emergencies, we have comparatively little evidence on factors that may safeguard against these rights violations during periods of derogation.[^exceptions]
-
-There is a robust literature on how civil society is important for improving human rights outcomes [@KeckSikkink:1998; @RisseRoppSikkink; @landman2005; @HBTsutsui; @neumayer2005; @Jetschke2011].
-Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run [@fair2017].
-
-As such, we expect that a strong and vibrant civil society may help mitigate the adverse impact arising from derogations and emergency measures imposed by governments during the pandemic.
-Civil society organizations broaden the range of voices in the policymaking process and help amplify the concerns and interests of citizens [@murdie2014].
-Civic initiatives have not only played a major role in countering the pandemic's severity, but have also pressured authorities into implementing better policies to deal with the impact of economic shutdowns, product shortages and spikes in unemployment.
-For instance, in Brazil, a broad coalition of CSOs lobbied the government to implement a relief fund, successfully securing a monthly stipend three times the amount initially suggested by the government [@carnegie2021, 5].
-
-Government attitudes towards civil society organizations—whether collaboration, accommodation or repression—may also affect their respect (or lack thereof) for human rights during the pandemic.
-Since March 2020, many advocacy and rights-based organizations have highlighted how the pandemic disproportionately affects vulnerable and marginalized communities.
-These organizations also monitor and document human rights violations by the state, which can both provide information about a state's level of compliance with domestic and international human rights obligations as well can be used as evidence in court.
-
-However, over the past three decades, more than 130 states have repressed civil society organizations [@chaudhry2022].
-These trends have been further exacerbated during the pandemic.
-As a result, a number of advocacy and rights-based organizations have either shut down or significantly downsized their capabilities, hampering their ability to document and monitor rights violations by the state and its agents.
-Without such monitoring, states may be more likely violate non-derogable rights during the pandemic, as states know that organizations will be less likely to collect evidence, write briefs, provide legal aid to victims and give expert testimony at trials.
-As such, we expect that states will have greater leverage to impact non-derogable rights if there is insufficient evidence that can be presented to international monitoring institutions or regional courts such as the European Court of Human Rights.
-
-> **Hypothesis 2~a~**: If states are repressive towards civil society organizations, then they are more likely to violate non-derogable rights.
-
-Given reduced monitoring by rights-based organizations, states may also have greater leeway to pass discriminatory policy measures and engage in more abusive enforcement of these measures.
-This is particularly the case for state actions towards refugees and asylum-seekers.
-For instance, Greece has increasingly criminalized humanitarian assistance and in 2019, introduced burdensome rules for the registration of nongovernmental organizations working on migration [@Cosse:2022].
-Once the pandemic began, with decreased oversight from CSOs owing to the above rules, Greek officials intercepted and turned back boats filled with asylum-seekers and also summarily returned asylum-seekers at the land border with Turkey.
-In doing so, they both violated the non-derogable principle of non-refoulement, as well as Article 7 of ICCPR is the person sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
-
-> **Hypothesis 2~b~**: If states are repressive towards civil society organizations, then they are more likely to use discriminatory policy measures during the pandemic.
-> 
-> **Hypothesis 2~c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.
-
-Finally, this trend of state repression of CSOs and NGOs may also impact state behavior towards the media.
-States typically repress CSOs because they make democratic participation, particularly protests, more likely, highlight rights violations, and challenge economic policies—all of which may be seen as even more threatening during the pandemic [@MurdieBhasin2011; @BBCM2014; @KeckSikkink:1998; @RisseRoppSikkink; @Boulding2014; @chaudhry2022].
-
-With many governments trying to centralize authority and decisionmaking, a free and independent media may pose challenges similar to those posed by CSOs.
-As such, if states are engaging in repression of CSOs during the pandemic, they may be more likely to repress the media as well.
-For instance, over the last five years, Hungary has increasingly cracked down on CSOs.
-During the pandemic, this extended towards the media as well as the country passed a new law which sought to punish anyone who published 'false information' on the pandemic with up to five years in jail [@brechenmacheretal2020, 3].
-Similarly, Zimbabwe criminalized false reporting on the government's response to the pandemic and subsequently, used this law to crack down on media freedom.
-On the other hand, a vibrant civil society may help protect against limitations on media freedom.
-For instance, in Tunisia, the parliament withdrew a draft law restricting media freedom after it faced criticism from CSOs [@brechenmacheretal2020, 3].
-
-> **Hypothesis 2~d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom during the pandemic.
-
-# Data and methods
-
-```{r extract-details}
-info <- list(
-  n_weeks = length(unique(weekly_panel$year_week)),
-  n_countries = length(unique(weekly_panel$iso3)),
-  n_rows = label_comma()(nrow(weekly_panel)),
-  dates = map(list(
-    start_date = min(weekly_panel$year_week_day),
-    end_date = max(weekly_panel$year_week_day)
-  ), ~list(date = ., 
-           nice = format(., "%B %e, %Y"),
-           year = year(.)))
-)
-```
-
-While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, civil society, and state actions during the pandemic.
-We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
-We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021.
-While the Oxford Government Response Tracker provides daily data on a variety of government actions, submitting formal treaty responses to the ICCPR is less likely to be as immediate of an action as implementing domestic emergency actions like stay at home orders.
-Accordingly, we collapse this daily data into weekly totals, resulting in a panel with `r info$n_rows` country-week observations.
-
-```{r tbl-data-sources}
-#| tbl-cap: Sources for outcomes and explanatory variables
-#| include: true
-#| echo: false
-tribble(
-  ~`Outcome variables`,                        ~Description,                                                        ~Frequency,  ~Source,
-  "Emergency public health measures",          "Binary indicator",                                                  "Weekly",    "Oxford Covid-19 Government Response Tracker",
-  "Human rights and policy outcomes",          "Ordered categories for the severity of abuses",                     "Quarterly", "Varieties of Democracy PanDem",
-  "ICCPR derogation",                          "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
-  "Civil society repression",                  "≈ −3 to 3; higher values indicate less repression",                 "Annual",    "Varieties of Democracy",
-  "New and cumulative COVID deaths and cases", "Counts",                                                            "Weekly",    "World Health Organization",
-  "Past ICCPR derogation or action",           "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
-  "Rule of law index",                         "0–1; higher values represent greater respect for rule of law",      "Annual",    "Varieties of Democracy",
-  "Civil liberties index",                     "0–1; higher values represent greater respect for civil liberties",  "Annual",    "Varieties of Democracy",
-  "Core civil society index",                  "0–1; higher values represent better environment for civil society", "Annual",    "Varieties of Democracy"
-) %>% 
-  kbl(booktabs = TRUE) %>% 
-  kable_styling() %>% 
-  pack_rows(index = c(" " = 2, "Primary explanatory variables" = 2, "Secondary explanatory variables" = 5),
-            indent = FALSE)
-
-```
-
-## Outcome variables
-
-We examine the relationship between formal treaty derogations and different human rights and policy outcomes in three stages (see @tbl-vars).
-First, to explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
-Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.
-As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in a given week.
-
-| General question                                    | Outcomes                                                                                                                                  | Main explanatory variable                   | Other explanatory variables                                                                                                                                              |
-|----------------|------------------|----------------|----------------------|
-| Derogations and emergency measures                  | Cancelling public events, Gathering restrictions, Close public transit, Restrictions on internal movement, Limits on international travel | ICCPR derogation in effect                  | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions                                                                     |
-| Derogations and human rights (H~1~)                 | Discriminatory policy, Violations of non-derogable rights, No time limit measures, Abusive enforcement                                    | ICCPR derogation in effect                  | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index, core civil society index |
-| Derogations, human rights, and civil society (H~2~) | Discriminatory policy, Violations of non-derogable rights, Limiting media, Abusive enforcement                                            | ICCPR derogation × civil society repression | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index                           |
-
-: Overview of outcomes and explanatory variables included in models {#tbl-vars}
-
-Next, to test our first hypothesis, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
-Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations.
-We do not collapse these variables into binary indicators and instead work with them as ordered categories.
-Finally, we test our second hypothesis by exploring the more complex relationship treaty derogations, civil society regulations, and four policy outcomes from the PanDem project.
-We look at the same four policies as the first hypothesis with one exception: since civil society is often concerned with advocacy and information provision during public emergencies, we replace whether emergency measures have a time limit with an ordinal measure of whether limits were placed on media freedom as part of a state's pandemic response.
-Importantly, since these PanDem variables are all measured quarterly, while our main data is recorded weekly, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response each week.
-
-## Explanatory variables
-
-### Key explanatory variables
-
-We use three different primary independent variables to explain changes in these policy and human rights outcomes.
-For our preliminary models and in models related to our first hypothesis, we rely on a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
-We collected and coded this data from real time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023].
-During the 15 months covered in this project, there were 115 formal derogations submitted mentioning 8 specific ICCPR articles; 15 derogations made no reference to any specific article.
-Of the derogations that indicated specific time limits, the average derogation duration was 41 days.
-We mark the start and end date of each derogation so that each country-week observation indicates whether a state has derogated.
-
-For our second hypothesis, we use a measure of civil society repression from the Varieties of Democracy (V-Dem) project [@vdem-v12], derived from a latent Bayesian item response theory model of expert opinions, with values ranging from roughly −3 to 3 (where higher values represent less repression).
-This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti-civil society environment.
-
-### Confounding variables
-
-Finally, we include several other variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
-To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID cases and the count of new COVID deaths, as well as the cumulative count of cases and deaths.
-Since formally derogating from treaties can serve as a signal that countries take their obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR bureaucratic actions prior to the pandemic.
-Finally, we include several annual measures of human rights, democracy, and civil society: V-Dem's rule of law index, civil liberties index, and core civil society index.
-
-## Modeling strategy
-
-Our data is structured as a balanced time series cross-sectional (TSCS) panel that has a natural hierarchical structure, with `r info$n_weeks` weekly observations nested within `r info$n_countries` countries.
-Accordingly, we model our outcome variables with multilevel or hierarchical models, with country-specific random offsets for each intercept and a fixed global year trend.
-When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, use use ordered logistic regression models.
-We take a Bayesian approach to better model the uncertainty inherent in measuring the effect of derogations on these broad policies.
-
-We include complete formal specifications of all our models, likelihoods, and priors in the appendix.
-Here we include two simplified model specifications to help demonstrate the intuition behind these multilevel TSCS models.
-For binary outcomes, we use a logistic model with country-specific offsets:
-
-::: {.column-page-inset-right}
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}, \alpha_k) \\
-\\
-&\ \textbf{Distribution parameters} \\
-\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} & \text{Probability of outcome} \\
-&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\
-\\
-&\ \textbf{Priors} \\
-\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-country variability,}
-\end{aligned}
-$$
-
-:::
-
-while for ordered outcomes, we use ordered logistic regression with country-specific offsets:
-
-::: {.column-page-inset-right}
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\,logit}(\phi_{it_j}, \alpha_k) \\
-\\
-&\ \textbf{Distribution parameters} \\
-\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} & \text{Cumulative probability of outcome} \\
-&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\
-\\
-&\ \textbf{Priors} \\
-\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-country variability} \\
-\alpha_k \sim&\ \mathcal{N}(0, 1) & \text{Boundaries between thresholds}
-\end{aligned}
-$$
-
-:::
-
-In all model specifications, we include new and cumulative COVID cases and deaths as controls.
-In our first hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society.
-In our second hypothesis, we omit the core civil society index and instead include the interaction of civil society repression and derogations.
-
-Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in the appendix).
-Instead we rely on probability-scale predictions and marginal effects whenever possible.
-To do so, we calculate predicted probabilities holding all explanatory variables at their typical values (median or mode).
-For logistic models, we estimate the probability of having a specific policy or violating a specific human right; for ordered models, we estimate the probability of each possible level of severity.
-To determine the effect of derogations, we calculate the contrast in predicted probabilities when derogation is set to true vs.
-when it is set to false (e.g., in ordered logistic models we compare the probability of major discrimination under derogation with the probability of major discrimination without derogation).
-We use median values from our models' posterior distributions as point estimates and provide credible intervals using the 95% highest posterior density.
-For hypothesis testing, we report the posterior probability that the between-level predicted differences are above or below zero.
-
-# Analysis
-
-## Derogations and emergency policies
-
-Before exploring whether COVID-induced derogations are associated with specific human rights abuses, it is useful to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
-@fig-preds-mfx-prelim-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
-In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-prelim-2 for the percentage-point differences between the predictions for the two groups of states).
-
-For instance, all countries were highly likely to cancel public events in March and April 2020, with greater than a 95% chance.
-Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021.
-Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability.
-The difference in predicted probabilities between the two types of countries ranged from 3 to 8 percentage points ($p(\Delta > 0) = 1$).
-The differences in policy implementation are starker for other types of emergency measures.
-Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between 80% and 90%, representing a 15–18 percentage point difference ($p(\Delta > 0) = 1$).
-Derogating states were also far more likely to close public transit, beginning with a roughly 80% probability in March 2020 and ending in June 2021 with a 60% probability.
-Non-derogating states followed a similar downward trend over time, but 23–25 percentage points lower than their derogating counterparts ($p(\Delta > 0) = 1$).
-Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to roughly 70% by June 2021.
-Non-derogating states, on the other hand, began the pandemic with a predicted 75% probability of limiting movement, dropping to only 25% by June 2021, representing a 20–40 percentage point difference ($p(\Delta > 0) = 1$).
-The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel.
-Both derogating and non-derogating states had a nearly 100% probability of limiting international movement across all 15 months.
-
-```{r fig-preds-mfx-prelim, fig.width=12, fig.height=6, out.width="100%", cache=FALSE}
-#| include: true
-#| echo: false
-#| layout-ncol: 1
-#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
-#| fig-subcap:
-#|   - "Predicted probabilities"
-#|   - "Contrasts between predicted probabilities"
-
-prelim_plots <- prelim_plot_data %>% 
-  mutate(pred_plot = map2(pred_data, nice, ~{
-    .x %>% 
-      ggplot(aes(x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
-                 color = derogation_ineffect, fill = after_scale(color), 
-                 group = forcats::fct_rev(ordered(.width)))) +
-      geom_lineribbon(alpha = 0.3) +
-      scale_color_manual(values = c(clrs[1], clrs[8])) +
-      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-      scale_y_continuous(labels = label_percent()) +
-      labs(x = NULL, y = "Predicted probability", 
-           color = "Derogation in effect", fill = "Derogation in effect",
-           title = .y) +
-      theme_pandem()
-  })) %>% 
-  mutate(mfx_plot = map2(mfx_data, nice, ~{
-    .x %>% 
-      mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>% 
-      ggplot(aes(x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
-                 group = forcats::fct_rev(ordered(.width)))) +
-      geom_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
-      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-      labs(x = NULL,
-           y = "Percentage point difference\nin probability of policy",
-           title = .y) +
-      theme_pandem()
-  }))
-
-wrap_plots(prelim_plots$pred_plot)
-wrap_plots(prelim_plots$mfx_plot)
-```
-
-While these results do not speak directly to our main hypotheses, they do reveal important insights into the motivations of derogating states.
-Countries that undertook emergency safety measures seem to have taken the time to follow international law and uphold their treaty obligations.
-International law thus appears to not just be an empty commitment—states that took the time to formally derogate were more likely to keep these emergency measures in place for longer and take their derogations seriously.
-
-## Derogations and human rights
-
-The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-h1).
-There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement.
-Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an 85% predicted probability of no discriminatory policy in March 2020 rising to nearly 100% by June 2021.
-The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.
-Moreover, countries that derogate were substantially less likely to engage in no abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic.
-Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.
-The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way—are unclear.
-
-In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status.
-For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability).
-The middle ground of only minor issues with time limits is exceptionally unlikely.
-Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues.
-The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
-Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)
-
-Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
-We thus find strong support only for H~1c~—that derogating states are less likely to impose measures with no time limits.
-We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~1b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~1c~).
-Derogations seem help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
-
-```{r fig-preds-mfx-h1, fig.width=13, fig.height=6, out.width="100%"}
-#| include: true
-#| echo: false
-#| layout-ncol: 1
-#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
-#| fig-subcap:
-#|   - "Predicted probabilities"
-#|   - "Contrasts between predicted probabilities"
-
-h1_plots <- h1_plot_data %>% 
-  # Predicted probabilities
-  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
-    if (.family == "cumulative") {
-      .data %>% 
-        ggplot(aes(
-          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
-          color = .category, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = NULL, y = "Predicted probability", 
-             color = "Severity", fill = "Severity",
-             title = .nice_name) +
-        facet_wrap(vars(derogation_ineffect)) +
-        theme_pandem()
-    } else {
-      .data %>% 
-        ggplot(aes(
-          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
-          color = derogation_ineffect, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = NULL, y = "Predicted probability", 
-             color = "Derogation in effect", fill = "Derogation in effect",
-             title = .nice_name) +
-        theme_pandem()
-    }
-  })) %>% 
-  # Marginal effects / contrasts
-  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
-    if (.family == "cumulative") {
-      .data %>% 
-        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
-        ggplot(aes(
-          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
-          color = group, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        scale_color_manual(values = darken(clrs[c(2, 4, 6, 7)], 0.5, space = "combined"), guide = "none") +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-        labs(x = NULL,
-             y = "Percentage point difference\nin probability of outcome level",
-             title = .nice_name) +
-        facet_wrap(vars(group), ncol = 2) +
-        theme_pandem()
-    } else {
-      .data %>% 
-        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
-        ggplot(aes(
-          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
-        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
-        labs(x = NULL,
-             y = "Percentage point difference\nin probability of outcome",
-             title = .nice_name) +
-        theme_pandem()
-    }
-  }))
-
-wrap_plots(h1_plots$pred_plot)
-wrap_plots(h1_plots$mfx_plot)
-```
-
-## Derogations, civil society, and human rights
-
-Civil society repression, on the other hand, adds an additional wrinkle to the story.
-Countries with severe level of repression (i.e.
-negative repression scores) that do not derogate have a 25% predicted probability of minor to major discriminatory policies, while similarly repressive countries that do derogate have nearly a 0% predicted probability of discriminatory policies (see @fig-preds-mfx-h2).
-This represents a substantial 5–10 percentage point difference between derogating and non-derogating countries with repressive civil societies.
-The derogation process could thus serve as a strong signal that emergency measures are indeed planned to be limited and non-discriminatory even in countries where civil society is under threat.
-
-However, there are no other substantial differences in predicted probabilities across derogation status for another other outcomes.
-Derogating and non-derogating states are generally equally likely to violate non-derogable rights (H~2a~), limit media (H~2d~), and engage in abusive enforcement (H~2c~) regardless of their level of civil society repression.
-We thus only find support for H~2b~—if states are repressive toward civil society organizations, they are more likely to use discriminatory policies, while derogations strongly negate this trend.
-This may be advocacy-focused civil society organizations in particular tend to deal with on-the-ground monitoring, advocacy, and shaming, and is more likely to call out governments for unequal and unjust implementation of their emergency policies.
-
-```{r fig-preds-mfx-h2, fig.width=13, fig.height=6, out.width="100%"}
-#| include: true
-#| echo: false
-#| layout-ncol: 1
-#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
-#| fig-subcap:
-#|   - "Predicted probabilities"
-#|   - "Contrasts between predicted probabilities"
-
-h2_plots <- h2_plot_data %>% 
-  # Predicted probabilities
-  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
-    if (.family == "cumulative") {
-      .data %>% 
-        ggplot(aes(
-          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
-          color = .category, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = "Civil society repression", y = "Predicted probability", 
-             color = "Severity", fill = "Severity",
-             title = .nice_name) +
-        facet_wrap(vars(derogation_ineffect)) +
-        theme_pandem()
-    } else {
-      .data %>% 
-        ggplot(aes(
-          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
-          color = derogation_ineffect, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = "Civil society repression", y = "Predicted probability", 
-             color = "Derogation in effect", fill = "Derogation in effect",
-             title = .nice_name) +
-        theme_pandem()
-    }
-  })) %>% 
-  # Marginal effects / contrasts
-  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
-    if (.family == "cumulative") {
-      .data %>% 
-        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
-        ggplot(aes(
-          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
-          color = group, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
-        labs(x = NULL,
-             y = "Percentage point difference\nin probability of outcome level",
-             title = .nice_name) +
-        facet_wrap(vars(group), ncol = 2) +
-        theme_pandem()
-    } else {
-      .data %>% 
-        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
-        ggplot(aes(
-          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
-        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        labs(x = NULL,
-             y = "Percentage point difference\nin probability of outcome",
-             title = .nice_name) +
-        theme_pandem()
-    }
-  }))
-
-wrap_plots(h2_plots$pred_plot)
-wrap_plots(h2_plots$mfx_plot)
-```
-
-# Discussion
-
-@Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
-However, @richardson2020 see derogations as a less clear path for states during COVID-19, describing the actions, their conditions, and human rights as a "tangled morass" (p.
-106).
-
-Is this good news? Derogators aren't using this opportunistically - derogators are doing the process correctly
-
-Enforcement is carried out by lower level officials - principal agent issues and local bureaucrats not caring about international law - the time limit and nonderogable rights things are done by high level people
-
-# Conclusion
-
-In conclusion mention that future research can point to bureaucrats vs. policy makers and
-
-Contribution = this is good news
-
-This research note examines states' responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and its subsequent impact on human rights violations.
-
-Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties.
-
-The first few months of the pandemic raised numerous concerns that the pandemic would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide
-
-# Appendix {.appendix}
-
-### Formal statistical models
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered logit}(\phi_{it_j}, \alpha_k) \\
-\\
-&\ \textbf{Models for distribution parameters} \\
-\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} \\
-&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
-&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
-&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
-&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
-&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\
-\\
-&\ \textbf{Priors} \\
-\beta_{0 \dots 11} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
-\alpha_k \sim&\ \mathcal{N}(0, 1)
-\end{aligned}
-$$
-
-### Complete regression results
-
-#### Derogations and emergency policies
-
-```{r tbl-results-full-prelim}
-#| tbl-cap: Complete results from models showing relationship between derogations and emergency policies
-#| include: true
-#| echo: false
-notes <- c(
-  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models; 95% credible intervals (highest density posterior interval, or HDPI) in brackets.", 
-  "Total \\(R^2\\) considers the variance of both population and group effects;", 
-  "marginal \\(R^2\\) only takes population effects into account."
-)
-
-modelsummary(models_tbl_prelim,
-             estimate = "{estimate}",
-             statistic = "[{conf.low}, {conf.high}]",
-             coef_map = coef_map,
-             gof_map = gof_map,
-             output = "kableExtra",
-             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-  kable_styling(htmltable_class = "table-sm light-border") %>% 
-  footnote(general = notes, footnote_as_chunk = TRUE)
-```
-
-#### Derogations and human rights
-
-```{r tbl-results-full-h1}
-#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~1~)
-#| include: true
-#| echo: false
-modelsummary(models_tbl_h1,
-             estimate = "{estimate}",
-             statistic = "[{conf.low}, {conf.high}]",
-             coef_map = coef_map,
-             gof_map = gof_map,
-             output = "kableExtra",
-             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-  kable_styling(htmltable_class = "table-sm light-border") %>% 
-  footnote(general = notes, footnote_as_chunk = TRUE)
-```
-
-#### Derogations, civil society, and human rights
-
-```{r tbl-results-full-h2}
-#| tbl-cap: Complete results from models showing relationship between derogations, civil society, and human rights (H~2~)
-#| include: true
-#| echo: false
-modelsummary(models_tbl_h2,
-             estimate = "{estimate}",
-             statistic = "[{conf.low}, {conf.high}]",
-             coef_map = coef_map,
-             gof_map = gof_map,
-             output = "kableExtra",
-             fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-  kable_styling(htmltable_class = "table-sm light-border") %>% 
-  footnote(general = notes, footnote_as_chunk = TRUE)
-```
-
-# References
+In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
+Did states actually end up complying with their pandemic-era derogations?
+Did derogating and non-derogating states implement emergency measures differently, and if so, how?
+Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
+On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
+Our results support the role that flexibility mechanisms such as derogations play in international law.

### 2023-03-08 _targets.R
Temporarily stop using targets for site building bc paths are messed up

@@ -141,12 +141,13 @@ list(
   tar_target(h2_plot_data, build_h2_plot_data(m_h2)),
   
   ## Analysis notebook ----
-  tar_quarto(website, path = "."),
-  tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
-  tar_target(deploy_notebook, {
-    # Force a dependency
-    website
-    # Run the deploy script
-    if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
-  })
+  # tar_quarto(website, path = "analysis", quiet = FALSE),
+  # tar_quarto(manuscript, path = "manuscript"),
+  # tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
+  # tar_target(deploy_notebook, {
+  #   # Force a dependency
+  #   website
+  #   # Run the deploy script
+  #   if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
+  # })
 )

### 2023-03-08 analysis/analysis_h1.qmd
Fix plot size for marginal effects
Better space between chunks in model math
targets-ify trace/trank/pp plots

@@
 invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+invisible(list2env(tar_read(diagnostic_functions), .GlobalEnv))
 invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
@@
 \begin{aligned}
-\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\[0.75em]
 \operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{region} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
@@
 &\ \beta_{10}\ \text{Core civil society index}_{i}\ + \beta_{11}\ \text{Week number}_{i} \\
-\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
-\\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\[0.75em]
 \bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
@@
 ```{r}
-prelim_trace <- function(model) {
-  model %>% 
-    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2xcs_ccsi, sd_who_region__Intercept) %>% 
-    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
-    geom_line(linewidth = 0.1) +
-    scale_color_viridis_d(option = "rocket", end = 0.85) +
-    labs(color = "Chain") +
-    facet_wrap(vars(.variable), scales = "free_y") +
-    theme_pandem()
-}
-
-prelim_trank <- function(model) {
-  model %>% 
-    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2xcs_ccsi, sd_who_region__Intercept) %>% 
-    group_by(.variable) %>% 
-    mutate(draw_rank = rank(.value)) %>% 
-    ggplot(aes(x = draw_rank, color = factor(.chain))) +
-    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
-    scale_color_viridis_d(option = "rocket", end = 0.85) +
-    labs(color = "Chain") +
-    facet_wrap(vars(.variable), scales = "free_y") +
-    theme_pandem()
-}
-
-prelim_pp <- function(model) {
-  pp_check(model, ndraws = 100, type = "bars")
-}
+params_to_show <- c("b_derogation_ineffect", "b_year_week_num", 
+                    "b_v2xcs_ccsi", "sd_who_region__Intercept")
 ```
@@
 ```{r}
-prelim_trace(m_h1$model[[1]])
+plot_trace(m_h1$model[[1]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h1$model[[1]])
+plot_trank(m_h1$model[[1]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h1$model[[1]])
+plot_pp(m_h1$model[[1]])
 ```
@@
 ```{r}
-prelim_trace(m_h1$model[[2]])
+plot_trace(m_h1$model[[2]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h1$model[[2]])
+plot_trank(m_h1$model[[2]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h1$model[[2]])
+plot_pp(m_h1$model[[2]])
 ```
@@
 ```{r}
-prelim_trace(m_h1$model[[3]])
+plot_trace(m_h1$model[[3]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h1$model[[3]])
+plot_trank(m_h1$model[[3]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h1$model[[3]])
+plot_pp(m_h1$model[[3]])
 ```
@@
 ```{r}
-prelim_trace(m_h1$model[[4]])
+plot_trace(m_h1$model[[4]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h1$model[[4]])
+plot_trank(m_h1$model[[4]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h1$model[[4]])
+plot_pp(m_h1$model[[4]])
 ```
@@
 
-```{r}
+```{r fig.width=13, fig.height=6, out.width="100%"}
 wrap_plots(h1_plots$mfx_plot)

### 2023-03-08 analysis/analysis_h2.qmd
Fix plot size for marginal effects
Better space between chunks in model math
targets-ify trace/trank/pp plots

@@
 \begin{aligned}
-\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\[0.75em]
 \operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{region} [i]} + \\
@@
 &\ \beta_{12}\ \text{Week number}_{i} \\
-\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
-\\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\[0.75em]
 \bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
@@
 ```{r}
-prelim_trace <- function(model) {
-  model %>% 
-    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2csreprss, sd_who_region__Intercept) %>% 
-    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
-    geom_line(linewidth = 0.1) +
-    scale_color_viridis_d(option = "rocket", end = 0.85) +
-    labs(color = "Chain") +
-    facet_wrap(vars(.variable), scales = "free_y") +
-    theme_pandem()
-}
-
-prelim_trank <- function(model) {
-  model %>% 
-    gather_draws(b_derogation_ineffect, b_year_week_num, b_v2csreprss, sd_who_region__Intercept) %>% 
-    group_by(.variable) %>% 
-    mutate(draw_rank = rank(.value)) %>% 
-    ggplot(aes(x = draw_rank, color = factor(.chain))) +
-    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
-    scale_color_viridis_d(option = "rocket", end = 0.85) +
-    labs(color = "Chain") +
-    facet_wrap(vars(.variable), scales = "free_y") +
-    theme_pandem()
-}
-
-prelim_pp <- function(model) {
-  pp_check(model, ndraws = 100, type = "bars")
-}
+params_to_show <- c("b_derogation_ineffect", "b_year_week_num", 
+                    "b_v2csreprss", "sd_who_region__Intercept")
 ```
@@
 ```{r}
-prelim_trace(m_h2$model[[1]])
+plot_trace(m_h2$model[[1]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h2$model[[1]])
+plot_trank(m_h2$model[[1]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h2$model[[1]])
+plot_pp(m_h2$model[[1]])
 ```
@@
 ```{r}
-prelim_trace(m_h2$model[[2]])
+plot_trace(m_h2$model[[2]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h2$model[[2]])
+plot_trank(m_h2$model[[2]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h2$model[[2]])
+plot_pp(m_h2$model[[2]])
 ```
@@
 ```{r}
-prelim_trace(m_h2$model[[3]])
+plot_trace(m_h2$model[[3]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h2$model[[3]])
+plot_trank(m_h2$model[[3]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h2$model[[3]])
+plot_pp(m_h2$model[[3]])
 ```
@@
 ```{r}
-prelim_trace(m_h2$model[[4]])
+plot_trace(m_h2$model[[4]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_h2$model[[4]])
+plot_trank(m_h2$model[[4]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_h2$model[[4]])
+plot_pp(m_h2$model[[4]])
 ```
@@
 
-```{r}
+```{r fig.width=13, fig.height=6, out.width="100%"}
 wrap_plots(h2_plots$mfx_plot)

### 2023-03-08 R/funs_plots.R

@@ -1,3 +1,36 @@
+# Diagnostic plots
+
+plot_trace <- function(model, params) {
+  model %>% 
+    tidybayes::gather_draws(!!!syms(params)) %>% 
+    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
+    geom_line(linewidth = 0.1) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+plot_trank <- function(model, params) {
+  model %>% 
+    tidybayes::gather_draws(!!!syms(params)) %>% 
+    group_by(.variable) %>% 
+    mutate(draw_rank = rank(.value)) %>% 
+    ggplot(aes(x = draw_rank, color = factor(.chain))) +
+    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    labs(color = "Chain") +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme_pandem()
+}
+
+plot_pp <- function(model) {
+  bayesplot::pp_check(model, ndraws = 100, type = "bars")
+}
+
+
+
+
 # Storing ggplot objects as rds files is BAD 
 #   (https://github.com/hadley/ggplot2-book/issues/344)
 #

### 2023-03-08 _targets.R

@@ -134,6 +134,7 @@ list(
 
   ## Graphics ----
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
+  tar_target(diagnostic_functions, lst(plot_trace, plot_trank, plot_pp)),
   
   ## Plots ----
   tar_target(prelim_plot_data, build_prelim_plot_data(m_prelim_derog, year_week_lookup)),

### 2023-03-08 analysis/model-details.qmd
Add table of MCMC runtimes

---
title: "Model details"
---

```{r setup, include=FALSE}
options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(brms)
library(glue)
library(gt)
library(lubridate)
library(here)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Models
tar_load(c(m_prelim_derog, m_h1, m_h2))
```

```{r calculate-model-times, include=FALSE}
model_times <- bind_rows(list("**Preliminary models: Derogations and emergency policies**" = m_prelim_derog, 
                              "**H<sub>1</sub>: Derogations and human rights**" = m_h1, 
                              "**H<sub>2</sub>: Derogations, civil society, and human rights**" = m_h2), 
                         .id = "hypothesis") %>%  
  mutate(family = case_when(
    family == "cumulative" ~ "Ordered logistic",
    family == "bernoulli" | is.na(family) ~ "Logistic"
  )) %>% 
  mutate(duration = map(model, ~{
    .$fit %>% 
      rstan::get_elapsed_time() %>% 
      as_tibble() %>% 
      summarize(total = as.duration(max(warmup + sample)))
  })) %>% 
  select(-prior, -y, -never, -data, -model) %>% 
  unnest(duration)

dur <- as.period(as.duration(sum(model_times$total)))

total_run_time <- glue("{hours} hours, {minutes} minutes, and {seconds} seconds",
                       hours = hour(dur), minutes = minute(dur), seconds = round(second(dur), 0))
```

We ran these models on a 2021 M1 MacBook Pro with 32 GB of RAM, with 4 MCMC chains spread across 8 cores, with two CPU threads per chain, using Stan through brms through cmdstanr.

In total, it took `r total_run_time` to run everything.

```{r}
model_times %>% 
  gt(rowname_col = "nice",
     groupname_col = "hypothesis",
     process_md = TRUE) %>% 
  cols_label(
    family = md("**Model family**"),
    total = md("**Total time** (i.e. longest chain)")
  ) %>% 
  tab_stubhead(label = md("**Outcome**")) %>% 
  cols_align(
    align = "left",
    columns = everything()
  ) %>%
  summary_rows(
    groups = everything(), 
    columns = c(total),
    fns = list(Total = ~as.duration(sum(.)))
  ) %>% 
  grand_summary_rows(
    columns = c(total),
    fns = list(`Overall total` = ~as.duration(sum(.)))
  ) %>% 
  tab_style(
    style = cell_text(align = "left", indent = px(15)),
    locations = cells_stub()
  ) %>% 
  opt_horizontal_padding(scale = 3) %>% 
  opt_vertical_padding(scale = 1) %>% 
  tab_options(table.width = pct(80))
```

### 2023-03-08 analysis/model-details.qmd
Include code for fun

@@
 
-```{r calculate-model-times, include=FALSE}
+```{r calculate-model-times}
 model_times <- bind_rows(list("**Preliminary models: Derogations and emergency policies**" = m_prelim_derog, 

### 2023-03-09 _targets.R
Make website project and manuscript subproject targets

@@ -142,13 +142,13 @@ list(
   tar_target(h2_plot_data, build_h2_plot_data(m_h2)),
   
   ## Analysis notebook ----
-  # tar_quarto(website, path = "analysis", quiet = FALSE),
-  # tar_quarto(manuscript, path = "manuscript"),
-  # tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
-  # tar_target(deploy_notebook, {
-  #   # Force a dependency
-  #   website
-  #   # Run the deploy script
-  #   if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
-  # })
+  tar_quarto(manuscript, path = "manuscript", quiet = FALSE),
+  tar_quarto(website, path = ".", quiet = FALSE),
+  tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
+  tar_target(deploy, {
+    # Force a dependency
+    website
+    # Run the deploy script
+    if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
+  })
 )

### 2023-03-10 analysis/manuscript.qmd
Make website project and manuscript subproject targets
Use {targets} to render fancy and manuscripty versions of the paper through Quarto profiles!
THIS IS SO COOL!

Once https://github.com/quarto-dev/quarto-r/issues/95 and/or https://github.com/ropensci/tarchetypes/issues/139 get fixed this will be a ton easier too

@@
 
-<p class="text-center"><a class="btn btn-success" href="/manuscript/output/manuscript.html" target="manuscript">{{< fa file-lines >}} &ensp;HTML</a> <a class="btn btn-success" href="/manuscript/output/manuscript.html" target="_blank">{{< fa arrow-up-right-from-square >}} &ensp;HTML (new window)</a> <a class="btn btn-success" href="/manuscript/output/manuscript.pdf" target="manuscript">{{< fa file-pdf >}} &ensp;PDF</a> <a class="btn btn-success" target="_blank" href="/manuscript/output/manuscript.pdf">{{< fa file-arrow-down >}} &ensp;PDF (download)</a></p>
+```{r include=FALSE}
+# Add the manuscript target as a dependency to the overall website
+targets::tar_load(c(manuscript_nice, manuscript_manuscripty))
+```
+
+```{=html}
+<p class="text-center">
+  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="manuscript">
+  <i class="fa-solid fa-file-lines"></i> &ensp;HTML
+  </a>
+
+  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="_blank">
+  <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(new window)</small>
+  </a>
+
+  <a class="btn btn-primary btn-sm mb-1" href="/manuscript/output/manuscript.pdf" target="manuscript">
+  <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
+  </a>
+
+  <a class="btn btn-primary btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
+  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
+  </a>
+
+  <a class="btn btn-info btn-sm mb-1" href="/manuscript/output-ms/manuscript-ms.pdf" target="manuscript">
+  <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
+  </a>
+
+  <a class="btn btn-info btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript-ms.pdf">
+  <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>
+  </a>
+</p>
 
 <iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="manuscript"></iframe>
+```

### 2023-03-10 manuscript/manuscript.qmd
Use {targets} to render fancy and manuscripty versions of the paper through Quarto profiles!
THIS IS SO COOL!

Once https://github.com/quarto-dev/quarto-r/issues/95 and/or https://github.com/ropensci/tarchetypes/issues/139 get fixed this will be a ton easier too

@@
 title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
+short-title: "Pandemic Pass?"
 subtitle: "Research note"
@@
 ```{r setup, include=FALSE}
+if (is.null(knitr::pandoc_to())) {
+  fmt_out <- "interactive"
+} else {
   fmt_out <- knitr::pandoc_to()
+}
 
@@
 library(kableExtra)
+library(gt)
 library(targets)
@@
 ```
+
+\begingroup\small\sffamily
+
+::: {.callout-note}
+## Scope of this paper
+
+This paper is not a traditional research article with statistical tests of novel theories.
+Instead it is a research note aimed at examining unexplored relationships between derogations, human rights, and civil society.
+We look forward to feedback and ideas to make it a strong research note!
+
+:::
+
+\endgroup
 
@@
 We then analyze the effect of derogating (or not) on human rights practices during the pandemic and assess the role of civil society in mitigating the effects of rights-violating policies.
-We find that states that derogated and specified time limits to the emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies, the abusive enforcement of these policies, and not violating non-derogable rights.
+We find that states that derogated and specified time limits to the emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights.
 On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
@@
 Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination.
-However,not every country formally derogated—a number of countries implemented emergency measures without formally derogating from ICCPR.
+However, not every country formally derogated—a number of countries implemented emergency measures without formally derogating from ICCPR.
 Over 100 countries issued emergency declarations while fewer than 30 submitted COVID-19 related derogations [@icnl-covid].
@@
 # The impact of derogations on human rights outcomes
+
+## Derogations and human rights
 
@@
 
-We expect that states that did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the the criteria for derogations—necessity, proportionality, non-discrimination and time delimited seriously, and therefore more likely to impose measures that violated these criteria.
+We expect that states that did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited seriously, and therefore more likely to impose measures that violated these criteria.
 If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.
@@
 
-\newpage
-
 ## Outcome variables
@@
 Importantly, since these PanDem variables are all measured quarterly, while our main data is recorded weekly, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response each week.
+
+\newpage
 
@@
 The first few months of the pandemic raised numerous concerns that the pandemic would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
-
 Derogations are an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
@@
 
-Looking back to our driving questions, what we find is that there is good news for compliance when we look to derogations.
+With the caveat that our results are by no means causal—we have not attempted to use any formal strategy to identify an unconfounded causal effect, and our findings are merely descriptive—we find that we find that there is good news for compliance and derogations.
 Derogators did not systematically use this legal "escape clause" opportunistically—derogators used the process correctly.
 Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
-Though not many states submitted COVID-19 related derogations, the ones that did, it seems, used derogations as intended by international law.
+Though not many states submitted COVID-19 related derogations, it seems that the ones that did used derogations as intended by international law.
 Once the time of derogation ended, states readjusted to typical compliance once more.
@@
 This distinction may explain some variation in policy and discrimination practices during the pandemic.
-
-We suggest future research look to other times of crisis and derogation use to compare COVID-19 area derogation compliance with that in other crises.
+We also recommend that future research compare COVID-19 area derogation compliance with that in other crises.
 This note contributed to the study of derogations by building on the analysis between derogation submission and outcome.
 States did end up taking the restriction measures they signaled they would via the derogations.
+Finally, the relationship and interaction between civil society, human rights, and international law remains an important and ripe field for future research.
 
@@
   "Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
-  r"(95\\% credible intervals (highest density posterior interval, or HDPI) in brackets.)",
-  "Total $R^2$ considers the variance of both population and group effects;",
-  "marginal $R^2$ only takes population effects into account."
+  "95% credible intervals (highest density posterior interval, or HDPI) in brackets.",
+  "Total \\(R^2\\) considers the variance of both population and group effects;",
+  "marginal \\(R^2\\) only takes population effects into account."
 )
 
+if (fmt_out == "latex") {
+  notes <- notes %>% 
+    str_replace_all(r"[\\]", r"[\\\\]") %>% 
+    str_replace_all("%", r"(\\\\%)")
+} else {
+  notes <- paste(notes, collapse = " ")
+}
+
+if (fmt_out == "latex") { 
   modelsummary(models_tbl_prelim,
@@
     column_spec(1:6, width = "0.16\\\\textwidth")
+} else {
+  modelsummary(models_tbl_prelim,
+               estimate = "{estimate}",
+               statistic = "[{conf.low}, {conf.high}]",
+               coef_map = coef_map,
+               gof_map = gof_map,
+               output = "gt",
+               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
+    tab_footnote(footnote = notes) %>% 
+    tab_style(style = cell_text(weight = "bold"),
+              locations = cells_column_labels())
+}
 ```
@@
 #| echo: false
+if (fmt_out == "latex") {
   modelsummary(models_tbl_h1,
@@
     column_spec(1:5, width = "0.18\\\\textwidth")
+} else {
+  modelsummary(models_tbl_h1,
+               estimate = "{estimate}",
+               statistic = "[{conf.low}, {conf.high}]",
+               coef_map = coef_map,
+               gof_map = gof_map,
+               output = "gt",
+               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
+               escape = FALSE) %>% 
+    tab_footnote(footnote = notes) %>% 
+    tab_style(style = cell_text(weight = "bold"),
+              locations = cells_column_labels())
+}
 ```
@@
 #| echo: false
+if (fmt_out == "latex") {
   modelsummary(models_tbl_h2,
@@
     column_spec(1:5, width = "0.18\\\\textwidth")
+} else {
+  modelsummary(models_tbl_h2,
+               estimate = "{estimate}",
+               statistic = "[{conf.low}, {conf.high}]",
+               coef_map = coef_map,
+               gof_map = gof_map,
+               output = "gt",
+               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
+               escape = FALSE) %>% 
+    tab_footnote(footnote = notes) %>% 
+    tab_style(style = cell_text(weight = "bold"),
+              locations = cells_column_labels())
+}
 ```

### 2023-03-10 _targets.R

@@ -142,7 +142,23 @@ list(
   tar_target(h2_plot_data, build_h2_plot_data(m_h2)),
   
   ## Analysis notebook ----
-  tar_quarto(manuscript, path = "manuscript", quiet = FALSE),
+  tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE),
+  
+  # Temporary target here to switch profiles. Eventually this can be removed
+  # once quarto::quarto_render() adds support for profiles
+  # (https://github.com/quarto-dev/quarto-r/issues/95) and
+  # tarchetypes::tar_quarto() gets expanded to use that support
+  # (https://github.com/ropensci/tarchetypes/issues/139)
+  tar_target(manuscript_manuscripty, {
+    manuscript_nice  # Force a dependency
+    
+    # Render manuscript-y output with the ms profile
+    withr::with_envvar(new = c("QUARTO_PROFILE" = "ms"), {
+      quarto::quarto_render("manuscript", quiet = FALSE)
+      return(unlist(tarchetypes::tar_quarto_files("manuscript")))
+    })
+  }, format = "file"),
+  
   tar_quarto(website, path = ".", quiet = FALSE),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
   tar_target(deploy, {

### 2023-03-11 analysis/presentations/2022-11-04_peace-science/index.qmd
Move presentations to blog-esque listing section

---
title: "Peace Science Society"
date: 2022-11-04
toc: false
resources: slides
abstract: |
  Annual conference of the Peace Science Society; Denver, Colorado
abstract-title: "Location"
published-title: "Presented"
image: first-slide.png
page-layout: full
---

<a class="btn btn-success btn-lg" target="_blank" href="slides/2022-11-04_peace-science.html">{{< fa arrow-up-right-from-square >}} &ensp;View all slides in new window</a>

<div class="ratio ratio-16x9">
<iframe class="slide-deck" src="slides/2022-11-04_peace-science.html"></iframe>
</div>

### 2023-03-11 slides/2022-11-04_peace-science_files/libs/revealjs/plugin/reveal-simple-menu/simplemenu.esm.js

@@ -1,283 +1,0 @@
-
-/*****************************************************************
- * @author: Martijn De Jongh (Martino), martijn.de.jongh@gmail.com
- * https://github.com/Martinomagnifico
- *
- * Simplemenu.js for Reveal.js 
- * Version 1.1.2
- * 
- * @license 
- * MIT licensed
- *
- * Thanks to:
- *  - Hakim El Hattab, Reveal.js 
- ******************************************************************/
-
-
-function _toConsumableArray(arr) {
-  return _arrayWithoutHoles(arr) || _iterableToArray(arr) || _unsupportedIterableToArray(arr) || _nonIterableSpread();
-}
-
-function _arrayWithoutHoles(arr) {
-  if (Array.isArray(arr)) return _arrayLikeToArray(arr);
-}
-
-function _iterableToArray(iter) {
-  if (typeof Symbol !== "undefined" && iter[Symbol.iterator] != null || iter["@@iterator"] != null) return Array.from(iter);
-}
-
-function _unsupportedIterableToArray(o, minLen) {
-  if (!o) return;
-  if (typeof o === "string") return _arrayLikeToArray(o, minLen);
-  var n = Object.prototype.toString.call(o).slice(8, -1);
-  if (n === "Object" && o.constructor) n = o.constructor.name;
-  if (n === "Map" || n === "Set") return Array.from(o);
-  if (n === "Arguments" || /^(?:Ui|I)nt(?:8|16|32)(?:Clamped)?Array$/.test(n)) return _arrayLikeToArray(o, minLen);
-}
-
-function _arrayLikeToArray(arr, len) {
-  if (len == null || len > arr.length) len = arr.length;
-
-  for (var i = 0, arr2 = new Array(len); i < len; i++) arr2[i] = arr[i];
-
-  return arr2;
-}
-
-function _nonIterableSpread() {
-  throw new TypeError("Invalid attempt to spread non-iterable instance.\nIn order to be iterable, non-array objects must have a [Symbol.iterator]() method.");
-}
-
-var Plugin = function Plugin() {
-  var selectionArray = function selectionArray(container, selectors) {
-    var selections = container.querySelectorAll(selectors);
-    var selectionarray = Array.prototype.slice.call(selections);
-    return selectionarray;
-  };
-
-  var isBefore = function isBefore(a, b) {
-    var all = document.getElementsByTagName('*');
-
-    for (var i = 0; i < all.length; ++i) {
[…]

### 2023-03-11 slides/2022-11-04_peace-science_files/libs/revealjs/plugin/simple-menu/plugin-src.js

@@ -1,264 +1,0 @@
-const Plugin = () => {
-
-	const selectionArray = function (container, selectors) {
-		let selections = container.querySelectorAll(selectors);
-		let selectionarray = Array.prototype.slice.call(selections);
-		return selectionarray
-	};
-
-	const isBefore = function( a, b ) {
-		var all = document.getElementsByTagName('*');
-	
-		for( var i = 0; i < all.length; ++i ) {
-			if( all[i] === a )
-				return true;  
-			else if( all[i] === b )
-				return false;
-		}
-	}
-
-	const isStack = function (section) {
-		let isStack = false;
-		for (let i = 0; i < section.childNodes.length; i++) {
-			if (section.childNodes[i].tagName == "SECTION") {
-				isStack = true
-				break;
-			}
-		}
-		return isStack;
-	};
-
-	const simpleMenu = function (deck, options) {
-
-		let viewport = (deck.getRevealElement()).tagName == "BODY" ? document : deck.getRevealElement();
-		let menus = selectionArray(viewport, `.${options.menuclass}`);
-		let menubars = selectionArray(viewport, `.${options.menubarclass}`);
-		let slides = deck.getSlidesElement();
-		let sections = slides.querySelectorAll("section");
-		let langattribute = deck.getConfig().internation ? deck.getConfig().internation.langattribute : false;
-
-		sections.forEach(section => {
-			if (!isStack(section) && section.parentNode.tagName == "SECTION") {
-
-				const parentAttributes = [...section.parentNode.attributes];
-				parentAttributes.reduce((attrs, attribute) => {
-					if (attribute.name == "data-name" ) {
-						section.setAttribute(`data-simplemenuname`, attribute.value)
-					} else if (attribute.name == "id" || attribute.name == "name" ) {
-						section.setAttribute(`data-simplemenu${attribute.name}`, attribute.value)
-					}
-				}, {});
-			}
-		})
-		
-
-
-		const compare = function (eventSelector, element) {
-			let compareThis = '';
-
-			if (options.selectby == 'name' || options.selectby == 'data-name') {
-				compareThis = element.textContent || (element.querySelector('a').textContent);
-
-				if (deck.hasPlugin( 'internation' ) && element.hasAttribute(langattribute)) {
-					compareThis = element.getAttribute(langattribute);
-				}
-			}
-			else if (options.selectby == 'id') {
[…]

### 2023-03-11 slides/2022-11-04_peace-science_files/libs/revealjs/plugin/simple-menu/simplemenu.js

@@ -1,291 +1,0 @@
-
-/*****************************************************************
- * @author: Martijn De Jongh (Martino), martijn.de.jongh@gmail.com
- * https://github.com/Martinomagnifico
- *
- * Simplemenu.js for Reveal.js 
- * Version 1.1.2
- * 
- * @license 
- * MIT licensed
- *
- * Thanks to:
- *  - Hakim El Hattab, Reveal.js 
- ******************************************************************/
-
-
-(function (global, factory) {
-  typeof exports === 'object' && typeof module !== 'undefined' ? module.exports = factory() :
-  typeof define === 'function' && define.amd ? define(factory) :
-  (global = typeof globalThis !== 'undefined' ? globalThis : global || self, global.Simplemenu = factory());
-})(this, (function () { 'use strict';
-
-  function _toConsumableArray(arr) {
-    return _arrayWithoutHoles(arr) || _iterableToArray(arr) || _unsupportedIterableToArray(arr) || _nonIterableSpread();
-  }
-
-  function _arrayWithoutHoles(arr) {
-    if (Array.isArray(arr)) return _arrayLikeToArray(arr);
-  }
-
-  function _iterableToArray(iter) {
-    if (typeof Symbol !== "undefined" && iter[Symbol.iterator] != null || iter["@@iterator"] != null) return Array.from(iter);
-  }
-
-  function _unsupportedIterableToArray(o, minLen) {
-    if (!o) return;
-    if (typeof o === "string") return _arrayLikeToArray(o, minLen);
-    var n = Object.prototype.toString.call(o).slice(8, -1);
-    if (n === "Object" && o.constructor) n = o.constructor.name;
-    if (n === "Map" || n === "Set") return Array.from(o);
-    if (n === "Arguments" || /^(?:Ui|I)nt(?:8|16|32)(?:Clamped)?Array$/.test(n)) return _arrayLikeToArray(o, minLen);
-  }
-
-  function _arrayLikeToArray(arr, len) {
-    if (len == null || len > arr.length) len = arr.length;
-
-    for (var i = 0, arr2 = new Array(len); i < len; i++) arr2[i] = arr[i];
-
-    return arr2;
-  }
-
-  function _nonIterableSpread() {
-    throw new TypeError("Invalid attempt to spread non-iterable instance.\nIn order to be iterable, non-array objects must have a [Symbol.iterator]() method.");
-  }
-
-  var Plugin = function Plugin() {
-    var selectionArray = function selectionArray(container, selectors) {
[…]

### 2023-03-11 presentations/2022-11-04_peace-science/slides/2022-11-04_peace-science_files/libs/revealjs/plugin/simplemenu/plugin-src.js
lol move back a level

@@ -1,0 +1,264 @@
+const Plugin = () => {
+
+	const selectionArray = function (container, selectors) {
+		let selections = container.querySelectorAll(selectors);
+		let selectionarray = Array.prototype.slice.call(selections);
+		return selectionarray
+	};
+
+	const isBefore = function( a, b ) {
+		var all = document.getElementsByTagName('*');
+	
+		for( var i = 0; i < all.length; ++i ) {
+			if( all[i] === a )
+				return true;  
+			else if( all[i] === b )
+				return false;
+		}
+	}
+
+	const isStack = function (section) {
+		let isStack = false;
+		for (let i = 0; i < section.childNodes.length; i++) {
+			if (section.childNodes[i].tagName == "SECTION") {
+				isStack = true
+				break;
+			}
+		}
+		return isStack;
+	};
+
+	const simpleMenu = function (deck, options) {
+
+		let viewport = (deck.getRevealElement()).tagName == "BODY" ? document : deck.getRevealElement();
+		let menus = selectionArray(viewport, `.${options.menuclass}`);
+		let menubars = selectionArray(viewport, `.${options.menubarclass}`);
+		let slides = deck.getSlidesElement();
+		let sections = slides.querySelectorAll("section");
+		let langattribute = deck.getConfig().internation ? deck.getConfig().internation.langattribute : false;
+
+		sections.forEach(section => {
+			if (!isStack(section) && section.parentNode.tagName == "SECTION") {
+
+				const parentAttributes = [...section.parentNode.attributes];
+				parentAttributes.reduce((attrs, attribute) => {
+					if (attribute.name == "data-name" ) {
+						section.setAttribute(`data-simplemenuname`, attribute.value)
+					} else if (attribute.name == "id" || attribute.name == "name" ) {
+						section.setAttribute(`data-simplemenu${attribute.name}`, attribute.value)
+					}
+				}, {});
+			}
+		})
+		
+
+
+		const compare = function (eventSelector, element) {
+			let compareThis = '';
+
+			if (options.selectby == 'name' || options.selectby == 'data-name') {
+				compareThis = element.textContent || (element.querySelector('a').textContent);
+
+				if (deck.hasPlugin( 'internation' ) && element.hasAttribute(langattribute)) {
+					compareThis = element.getAttribute(langattribute);
+				}
+			}
+			else if (options.selectby == 'id') {
[…]

### 2023-03-11 presentations/2022-11-04_peace-science/slides/2022-11-04_peace-science_files/libs/revealjs/plugin/simplemenu/simplemenu.esm.js

@@ -1,0 +1,283 @@
+
+/*****************************************************************
+ * @author: Martijn De Jongh (Martino), martijn.de.jongh@gmail.com
+ * https://github.com/Martinomagnifico
+ *
+ * Simplemenu.js for Reveal.js 
+ * Version 1.1.2
+ * 
+ * @license 
+ * MIT licensed
+ *
+ * Thanks to:
+ *  - Hakim El Hattab, Reveal.js 
+ ******************************************************************/
+
+
+function _toConsumableArray(arr) {
+  return _arrayWithoutHoles(arr) || _iterableToArray(arr) || _unsupportedIterableToArray(arr) || _nonIterableSpread();
+}
+
+function _arrayWithoutHoles(arr) {
+  if (Array.isArray(arr)) return _arrayLikeToArray(arr);
+}
+
+function _iterableToArray(iter) {
+  if (typeof Symbol !== "undefined" && iter[Symbol.iterator] != null || iter["@@iterator"] != null) return Array.from(iter);
+}
+
+function _unsupportedIterableToArray(o, minLen) {
+  if (!o) return;
+  if (typeof o === "string") return _arrayLikeToArray(o, minLen);
+  var n = Object.prototype.toString.call(o).slice(8, -1);
+  if (n === "Object" && o.constructor) n = o.constructor.name;
+  if (n === "Map" || n === "Set") return Array.from(o);
+  if (n === "Arguments" || /^(?:Ui|I)nt(?:8|16|32)(?:Clamped)?Array$/.test(n)) return _arrayLikeToArray(o, minLen);
+}
+
+function _arrayLikeToArray(arr, len) {
+  if (len == null || len > arr.length) len = arr.length;
+
+  for (var i = 0, arr2 = new Array(len); i < len; i++) arr2[i] = arr[i];
+
+  return arr2;
+}
+
+function _nonIterableSpread() {
+  throw new TypeError("Invalid attempt to spread non-iterable instance.\nIn order to be iterable, non-array objects must have a [Symbol.iterator]() method.");
+}
+
+var Plugin = function Plugin() {
+  var selectionArray = function selectionArray(container, selectors) {
+    var selections = container.querySelectorAll(selectors);
+    var selectionarray = Array.prototype.slice.call(selections);
+    return selectionarray;
+  };
+
+  var isBefore = function isBefore(a, b) {
+    var all = document.getElementsByTagName('*');
+
+    for (var i = 0; i < all.length; ++i) {
[…]

### 2023-03-11 presentations/2023-03-17_isa/index.qmd
Add placeholder page for ISA

---
title: "International Studies Association"
date: 2023-03-17
toc: false
resources: slides
abstract: |
  Annual conference of the International Studies Association (ISA); Montréal, Canada
abstract-title: "Location"
published-title: "Presented"
# image: first-slide.png
page-layout: full
---

Slides here.

### 2023-03-11 presentations/2023-03-17_isa/index.qmd
Embed Google slides

@@
 published-title: "Presented"
-# image: first-slide.png
+image: first-slide.png
 page-layout: full
@@
 
-Slides here.
+<div class="ratio ratio-16x9">
+<iframe class="slide-deck" src="https://docs.google.com/presentation/d/e/2PACX-1vQl8N_xTO7QyJNFuEvJXrUDV77_DLbgnjZ4X6WFx0q3ScXPEa6Y-uvUanzHs2ds5Gt8E-5Px0Biqa3l/embed?start=false&loop=false&delayms=3000" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
+</div>

### 2023-03-11 manuscript/manuscript.qmd
Add ORCIDs and URLs

@@
   affiliation: Lewis and Clark College
+  orcid: 0000-0003-3506-7590
+  url: https://www.suparnachaudhry.com/
   email: schaudhry@lclark.edu
@@
   affiliation: Arizona State University
+  orcid: 0000-0003-1772-5350
+  url: https://audreylcomstock.weebly.com/
   email: audrey.comstock@asu.edu
@@
   affiliation: Georgia State University
+  orcid: 0000-0002-3948-3914
+  url: https://www.andrewheiss.com
   email: aheiss@gsu.edu

### 2023-03-11 analysis/hypotheses.qmd
Remove questions and methods section

@@
 - **H~1a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
-- **H~1b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights
+- **H~1b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
 - **H~1c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
@@
 - **H~2c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.
-- **H~2d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom.
+- **H~2d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom during the pandemic.

### 2023-03-11 index.qmd
Remove questions and methods section

@@
 title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
+author:
+- name: Suparna Chaudhry
+  affiliation: Lewis and Clark College
+  orcid: 0000-0003-3506-7590
+  url: https://www.suparnachaudhry.com/
+  email: schaudhry@lclark.edu
+- name: Audrey L. Comstock
+  affiliation: Arizona State University
+  orcid: 0000-0003-1772-5350
+  url: https://audreylcomstock.weebly.com/
+  email: audrey.comstock@asu.edu
+- name: Andrew Heiss
+  affiliation: Georgia State University
+  orcid: 0000-0002-3948-3914
+  url: https://www.andrewheiss.com
+  email: aheiss@gsu.edu
 format:
@@
 ---
-
-- Suparna Chaudhry (Lewis & Clark College)
-- Audrey L. Comstock (Arizona State University)
-- Andrew Heiss (Georgia State University)
 

### 2023-03-11 analysis/manuscript.qmd
Nicer button colors

@@
 
-  <a class="btn btn-primary btn-sm mb-1" href="/manuscript/output/manuscript.pdf" target="manuscript">
+  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/manuscript.pdf" target="manuscript">
   <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
@@
 
-  <a class="btn btn-primary btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
+  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
   <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
@@
 
-  <a class="btn btn-info btn-sm mb-1" href="/manuscript/output-ms/manuscript-ms.pdf" target="manuscript">
+  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/manuscript-ms.pdf" target="manuscript">
   <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
@@
 
-  <a class="btn btn-info btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript-ms.pdf">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript-ms.pdf">
   <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>

### 2023-04-11 manuscript/manuscript.qmd
Suparna edits

@@
 
-Do emergency provisions during crises do more harm or good when it comes to human rights?
+What is the impact of escape clauses or emergency provisions in international law on human rights?
 Combating a deadly virus like SARS-CoV-2 required extraordinary public health measures.
@@
 While countries were ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights law.
-Some human rights treaties recognize states' needs for emergency measures during public health concerns and contain escape clauses that establish safeguards to prevent permanent erosion of these rights during emergencies.
+Some human rights treaties recognize states' needs for emergency measures during public health concerns and contain escape clauses that establish safeguards to prevent the permanent erosion of these rights during emergencies.
 The temporary suspension of state compliance with its international treaty obligations occurs through a legal action known as derogation.
-However, the potential abuse of emergency measures during derogations raises concerns about the long-lasting impact of emergency measures on human rights.
+However, the potential abuse of emergency measures during derogations raises concerns about their impact on human rights.
 Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].
 
-Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023, ending measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
-With the lifting of these measures, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this emergency period.
-What did derogations mean for compliance?
-Did states deviate from treaty provisions when they said they would?
-Are derogating and non-derogating states different in terms of how they implemented emergency measures during COVID-19?
-Were derogating states more likely to exploit the provisions and to abuse human rights during COVID-19?
-These questions are important for the study of international law, compliance, and times of crises.
-
-The potential violations of human rights during this period also raises questions about factors that can mitigate the adverse impacts of emergency measures.
-Existing literature on emergencies such as wars and natural disasters shows that civil society organizations (CSOs) play a pivotal role during these periods [@pellingdill; @cash2013].
-They provide services and resources, and hold governments accountable to ensure inclusive policies during these emergencies.
-Human rights compliance literature also generally finds that NGOs and civil society are key actors in facilitating treaty obligation actualization [e.g., @KeckSikkink:1998; @Merry:2009].
-Can these groups also prevent or reduce the likelihood of human rights violations during public health emergencies?
-Does a robust civil society mitigate the adverse effects arising from international treaty derogations during the pandemic?
+Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023, lifting measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
+With this move, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this emergency period.
+What did derogations mean for compliance with international law?
+Did derogating and non-derogating states differ in terms of how they implemented emergency measures during COVID-19?
+Were derogating states more likely to exploit emergency provisions and to abuse human rights during COVID-19?
+These questions are important for the study of international law, compliance, and human rights during times of crises.
 
@@
 
-Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation in human rights responses among governments during the pandemic.
-We then analyze the effect of derogating (or not) on human rights practices during the pandemic and assess the role of civil society in mitigating the effects of rights-violating policies.
+Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation among derogating and non-derogating states on the implementation of emergency measures.
+We then analyze the effect of derogating (or not) on human rights practices during the pandemic.
 We find that states that derogated and specified time limits to the emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights.
-On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
-
-We also find that civil society can play an important mitigating role in curbing discriminatory policies.
-States with low levels of civil society repression were the least likely to abuse human rights when enforcing emergency policies.
-The likelihood of discriminatory policies increased as civil society repression worsened, but derogations appear to protect against worsening abuses—even in countries with severe civil society repression, those that derogated had a nearly 100% probability of no discriminatory policies.
+On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely to impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
 
@@
 Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of the types of states that are likely to comply with these escape clauses.
-The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster typical of treaty derogations, making it easier to compare the long-term human rights effects of these derogations cross nationally.
+The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally.
+In making this comparison, this research note makes an important descriptive contribution showing how the implementation of policy measures differed across derogating and non-derogating states during the pandemic.
 
@@
 @Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
-We find that overall, derogating states do not increase abuse and violations.
+We find that overall, derogating states during the pandemic do not increase abuse and violations.
 This finding speaks to international legal compliance scholarship and supports the idea that states are being sincere about their intentions and not, byinlarge, using actions like derogation to cover abusive behavior.
@@
 The rest of the paper proceeds as follows.
-First, we explore the purpose of derogations during the pandemic.
-In doing so, we focus specifically on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
+First, we explore the purpose of derogations in international law, and then specifically during the pandemic.
+In doing so, we focus on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
 Following this, we introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic.
-We then assess states' behavior towards civil society and these organizations' potential role in mitigating the adverse effects of emergency measures.
-We conclude with the implications of our findings for the study of international law, times of crises, and the role of civil society.
+We conclude with the implications of our research note for the study of international law during times of crises.
 
@@
 Derogations afford states the opportunity to adjust to these changing conditions [@HBHelferFariss].
-When states derogate from their international legal obligations,the derogation provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
+When states derogate from their international legal obligations, the derogation provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
 This information allows these actors—at least in principle—to challenge measures that are excessive, vague, or outlast the intended time frame of their implementation [@helfer2021].
 
-While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially lead to the violation of non-derogable rights—rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
+While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency (Greene 2020, 36).
+These include rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
 Research finds results that support this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
@@
 This also raises the question of the extent to which this distinction is a valuable one during the pandemic.
-Most derogations to the ICCPR have been for internal conflict or natural disaster - is a public health emergency different from other states of emergency?
+Most derogations to the ICCPR have been for internal conflict or natural disaster.
+Is a public health emergency different from other states of emergency?
 Overall, there is a general expectation that following derogations, we may observe some decline in respect for human rights.
@@
 
-A first question of interest before we examine the potential impact of derogations is what states are more likely to submit them in the first place.
-Though not much research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the UN rather than implications of when they are in effect.
+A first question of interest before we examine the potential impact of derogations is the nature of states that are more likely to submit them in the first place.
+Little research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the United Nations (UN) rather than the effects of these derogations.
 This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss].
@@
 However, it is worth noting that states cannot derogate from the entirety of the ICCPR.
-Seven rights enshrined in the ICCPR are non-derogable: the right to life, the prohibition of torture or cruel, inhumane punishment,the prohibition of slavery, the right to recognition as a person before the law, and the protection of freedom of thought, conscience and religion.
+Seven rights enshrined in the ICCPR are non-derogable: the right to life, the prohibition of torture or cruel, inhumane punishment, the prohibition of slavery, the right to recognition as a person before the law, and the protection of freedom of thought, conscience and religion.
 The international legal community viewed these rights as so fundamental and important that no type of crisis could justify their suspension.
@@
 These COVID-19 emergency declarations may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020].
-Below, we first theorize how derogations can protect against rights violations and the excesses of emergency measures, followed by the role of civil society in mitigating these violations.
+Below, we first theorize how derogations led to variation in the implementation of emergency measures during the pandemic, followed by their impact on human rights.
 
-# The impact of derogations on human rights outcomes
+# The impact of derogations during Covid-19
+
+## Derogations and emergency policies
+
+During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21).
+
+AUDREY - write expectations about what how derogating and non-derogating states differed in their likelihood of imposing emergency policies - canceling public events, gathering restrictions, closing public transit, movement and international travel
 
@@
 
-During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21).
-If states determine that emergency measures are justified without regard for international law, the processes for restoring rights at the end of the emergency may not be followed.
-As such, many states used the pandemic an as excuse to centralize authority and implemented measures that didn’t meet international legal standards.
-Many states imposed restrictions on travel or movement by minority or opposition groups, limiting public assembly and democratic protests, and deployed surveillance systems that impeded privacy.
+If states determine that emergency measures are justified without regard for international law, the processes for respecting rights during crises and restoring rights at the end of the emergency may not be followed.
+As such, many states used the pandemic an as excuse to centralize authority and implemented measures that did not meet international legal standards.
+Many states imposed in discriminatory and abusive policy enforcement--they imposed restrictions on travel or movement by minority or opposition groups.
 For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
@@
 
-We expect that states that did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited seriously, and therefore more likely to impose measures that violated these criteria.
+The pandemic also led to the violation of non-derogable rights, or rights that cannot be suspendended in any circumstances, in many states.
+For instance, Greek officials intercepted and turned back boats filled with asylum-seekers and also summarily returned asylum-seekers at the land border with Turkey.
+In doing so, they both violated the non-derogable principle of non-refoulement, as well as Article 7 of ICCPR is the person sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
+Many other states  changed rules of confinement affecting detainees, which also potentially violated non-derogable rights.
+For instance, in March 2020, a French government order extended the duration of pre-trial detention, raising the issue of prison overcrowding and leading to inhumane and degrading treatment [@Lebret:2020].
+As such, this was not just a violation of a non-derogable right, but also raised questions regarding the proportionality of emergency measures.
+
+We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited seriously, and therefore more likely to impose measures that violated these criteria.
 If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.
@@
 > **Hypothesis 1~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
-
-## Civil society and human rights
-
-Given the divergent state responses during the pandemic, differing responses can have lasting impacts on human rights trends.
-In the absence of adequate review from the international community during the pandemic, certain domestic-level factors may be necessary to mitigate the adverse effects of emergency measures.
-While both the theoretical and empirical literature has established that states may violate both derogable and non-derogable rights during emergencies, we have comparatively little evidence regarding factors that may safeguard against these rights violations during periods of derogation.
-
-There is a robust literature on how civil society is important for improving human rights outcomes [@KeckSikkink:1998; @RisseRoppSikkink; @landman2005; @HBTsutsui; @neumayer2005; @Jetschke2011].
-Previous literature finds that an effective civil society response can mitigate the impact of natural disasters and crises and even increase political engagement in the long-run [@fair2017].
-As such, we expect that a strong and vibrant civil society may help mitigate the adverse impact arising from derogations and emergency measures imposed by governments during the pandemic.
-Civil society organizations broaden the range of voices in the policymaking process and help amplify the concerns and interests of citizens [@murdie2014].
-Civic initiatives have not only played a major role in countering the pandemic's severity, but have also pressured authorities into implementing better policies to deal with the impact of economic shutdowns, product shortages and spikes in unemployment.
-For instance, in Brazil, a broad coalition of CSOs lobbied the government to implement a relief fund, successfully securing a monthly stipend three times the amount initially suggested by the government [@carnegie2021, 5].
-
-Government attitudes towards civil society organizations—whether collaboration, accommodation or repression—may also affect their respect (or lack thereof) for human rights during the pandemic.
-Since March 2020, many advocacy and rights-based organizations have highlighted how the pandemic disproportionately affects vulnerable and marginalized communities.
-These organizations also monitor and document human rights violations by the state, which can both provide information about a state's level of compliance with domestic and international human rights obligations as well can be used as evidence in court.
-
-However, over the past three decades, more than 130 states have repressed civil society organizations [@chaudhry2022].
-These trends have been further exacerbated during the pandemic.
-As a result, a number of advocacy and rights-based organizations have either shut down or significantly downsized their capabilities, hampering their ability to document and monitor rights violations by the state and its agents.
-Without such monitoring, states may be more likely violate non-derogable rights during the pandemic, as states know that organizations will be less likely to collect evidence, write briefs, provide legal aid to victims and give expert testimony at trials.
-As such, we expect that states will have greater leverage to impact non-derogable rights if there is insufficient evidence that can be presented to international monitoring institutions or regional courts such as the European Court of Human Rights.
-
-> **Hypothesis 2~a~**: If states are repressive towards civil society organizations, then they are more likely to violate non-derogable rights.
-
-Given reduced monitoring by rights-based organizations, states may also have greater leeway to pass discriminatory policy measures and engage in more abusive enforcement of these measures.
-This is particularly the case for state actions towards refugees and asylum-seekers.
-For instance, Greece has increasingly criminalized humanitarian assistance and in 2019, introduced burdensome rules for the registration of nongovernmental organizations working on migration [@Cosse:2022].
-Once the pandemic began, with decreased oversight from CSOs owing to the above rules, Greek officials intercepted and turned back boats filled with asylum-seekers and also summarily returned asylum-seekers at the land border with Turkey.
-In doing so, they both violated the non-derogable principle of non-refoulement, as well as Article 7 of ICCPR is the person sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
-
-> **Hypothesis 2~b~**: If states are repressive towards civil society organizations, then they are more likely to use discriminatory policy measures during the pandemic.
-> 
-> **Hypothesis 2~c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.
-
-Finally, this trend of state repression of CSOs and NGOs may also impact state behavior towards the media.
-States typically repress CSOs because they make democratic participation, particularly protests, more likely, highlight rights violations, and challenge economic policies—all of which may be seen as even more threatening during the pandemic [@MurdieBhasin2011; @BBCM2014; @KeckSikkink:1998; @RisseRoppSikkink; @Boulding2014; @chaudhry2022].
-
-With many governments trying to centralize authority and decisionmaking, a free and independent media may pose challenges similar to those posed by CSOs [@Themudo:2013].
-As such, if states are engaging in repression of CSOs during the pandemic, they may be more likely to repress the media as well.
-For instance, over the last five years, Hungary has increasingly cracked down on CSOs.
-During the pandemic, this extended towards the media as well as the country passed a new law which sought to punish anyone who published 'false information' on the pandemic with up to five years in jail [@brechenmacheretal2020, 3].
-Similarly, Zimbabwe criminalized false reporting on the government's response to the pandemic and subsequently, used this law to crack down on media freedom.
-On the other hand, a vibrant civil society may help protect against limitations on media freedom.
-For instance, in Tunisia, the parliament withdrew a draft law restricting media freedom after it faced criticism from CSOs [@brechenmacheretal2020, 3].
-
-> **Hypothesis 2~d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom during the pandemic.
 
@@
 
-While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, civil society, and state actions during the pandemic.
+While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, human rights, and state actions during the pandemic.
 We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
@@
 
-# Discussion and conclusion
+# Conclusion
 
-The first few months of the pandemic raised numerous concerns that the pandemic would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
-Derogations are an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
-This research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.
-
+The Covid-19 pandemic raised numerous concerns that the cris would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
+Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.
+We focused specifically on the impact of derogations--an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
 Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties without concern about the inability to adapt to times of emergency.
@@
 Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
-Though not many states submitted COVID-19 related derogations, it seems that the ones that did used derogations as intended by international law.
-Once the time of derogation ended, states readjusted to typical compliance once more.
+Though a majority of states did not submit COVID-19 related derogations, the ones that did used derogations as intended by international law.
+Once the derogation ended, states readjusted to typical compliance once more.
 States, therefore complied with the intent of derogations and the purpose of the treaty.
@@
 
-We suggest that future research look to differences in levels within states during crisis.
+We suggest that future research look to different incentives facing state officials during crisis.
 There may be key differences in enforcement between local level bureaucrats and higher level elites.
@@
 This distinction may explain some variation in policy and discrimination practices during the pandemic.
-We also recommend that future research compare COVID-19 area derogation compliance with that in other crises.
-This note contributed to the study of derogations by building on the analysis between derogation submission and outcome.
-States did end up taking the restriction measures they signaled they would via the derogations.
-Finally, the relationship and interaction between civil society, human rights, and international law remains an important and ripe field for future research.
+We also recommend that future research analyze the impact of other factors that can mitigate the adverse impacts of emergency measures.
+In the absence of adequate review from the international community and treaty monitoring bodies during crises, certain domestic-level factors, such as a robust civil society and media, may be necessary to mitigate the adverse effects of emergency measures.
 

### 2023-04-20 _targets.R
Use new profile argument
Yay https://github.com/ropensci/tarchetypes/issues/139 !

@@ -141,24 +141,10 @@ list(
   tar_target(h1_plot_data, build_h1_plot_data(m_h1, year_week_lookup)),
   tar_target(h2_plot_data, build_h2_plot_data(m_h2)),
   
-  ## Analysis notebook ----
-  tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE),
-  
-  # Temporary target here to switch profiles. Eventually this can be removed
-  # once quarto::quarto_render() adds support for profiles
-  # (https://github.com/quarto-dev/quarto-r/issues/95) and
-  # tarchetypes::tar_quarto() gets expanded to use that support
-  # (https://github.com/ropensci/tarchetypes/issues/139)
-  tar_target(manuscript_manuscripty, {
-    manuscript_nice  # Force a dependency
-    
-    # Render manuscript-y output with the ms profile
-    withr::with_envvar(new = c("QUARTO_PROFILE" = "ms"), {
-      quarto::quarto_render("manuscript", quiet = FALSE)
-      return(unlist(tarchetypes::tar_quarto_files("manuscript")))
-    })
-  }, format = "file"),
-  
+  ## Manuscript and analysis notebook ----
+  tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE, profile = "nice"),
+  tar_quarto(manuscript_manuscripty, path = "manuscript", quiet = FALSE, profile = "ms"),
+
   tar_quarto(website, path = ".", quiet = FALSE),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
   tar_target(deploy, {

### 2023-04-20 manuscript/manuscript.qmd
Add keywords

@@
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
+keywords:
+  - human rights
+  - treaty derogations
+  - COVID-19
+  - emergency policies
 bibliography: bib.bib

### 2023-04-20 manuscript/manuscript.qmd
Add Greene citation

@@
 
-While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency (Greene 2020, 36).
+While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency [@Greene:2020, 36].
 These include rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
@@
 As such, many states used the pandemic an as excuse to centralize authority and implemented measures that did not meet international legal standards.
-Many states imposed in discriminatory and abusive policy enforcement--they imposed restrictions on travel or movement by minority or opposition groups.
+Many states imposed in discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
 For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
@@
 Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.
-We focused specifically on the impact of derogations--an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
+We focused specifically on the impact of derogations—an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
 Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties without concern about the inability to adapt to times of emergency.

### 2023-04-20 manuscript/manuscript.qmd
Remove callout note + extra empty lines

@@
 
-\begingroup\small\sffamily
-
-::: {.callout-note}
-## Scope of this paper
-
-This paper is not a traditional research article with statistical tests of novel theories.
-Instead it is a research note aimed at examining unexplored relationships between derogations, human rights, and civil society.
-We look forward to feedback and ideas to make it a strong research note!
-
-:::
-
-\endgroup
-
 What is the impact of escape clauses or emergency provisions in international law on human rights?
@@
 
-These findings suggest that states may not be discerning the significance of derogable v/s.
+These findings suggest that states may not be discerning the significance of derogable vs.
 non-derogable rights status during emergencies.
@@
 
+AUDREY: Talk about the map of countries that derogated. The figure will go here-ish, I think
+
 ## Derogations from the ICCPR
@@
 In this research note, we look at states' derogations from the International Covenant on Civil and Political Rights (ICCPR) during the pandemic.
-ICCPR is a foundational human rights treaty with broad human rights coverage.
+The ICCPR is a foundational human rights treaty with broad human rights coverage.
 Exceptional circumstances brought on by the global pandemic have led to numerous, extensive derogations to the ICCPR removing state obligations to important rights areas of assembly, movement, and others.
@@
 
-AUDREY - write expectations about what how derogating and non-derogating states differed in their likelihood of imposing emergency policies - canceling public events, gathering restrictions, closing public transit, movement and international travel
+AUDREY: write expectations about what how derogating and non-derogating states differed in their likelihood of imposing emergency policies - canceling public events, gathering restrictions, closing public transit, movement and international travel
 
@@
 
-The pandemic also led to the violation of non-derogable rights, or rights that cannot be suspendended in any circumstances, in many states.
+The pandemic also led to the violation of non-derogable rights, or rights that cannot be suspended in any circumstances, in many states.
 For instance, Greek officials intercepted and turned back boats filled with asylum-seekers and also summarily returned asylum-seekers at the land border with Turkey.

### 2023-04-20 manuscript/manuscript.qmd
Temporarily remove appendix for word count purposes

@@
 
-# Appendix {.appendix}
-
-### H~1~: Derogations and human rights (ordered logistic regression)
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
-&\ \textbf{Models for distribution parameters} \\
-\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
-&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
-&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
-&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
-&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
-&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
-&\ \textbf{Priors} \\
-\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
-\alpha_k \sim&\ \mathcal{N}(0, 1)
-\end{aligned}
-$$
-
-### H~1~: Derogations and human rights (logistic regression)
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
-&\ \textbf{Distribution parameters} \\
-\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
-&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
-&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
-&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
-&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
-&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
-&\ \textbf{Priors} \\
-\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
-\end{aligned}
-$$
-
-### H~2~: Derogations, civil society, and human rights (ordered logistic regression)
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
-&\ \textbf{Models for distribution parameters} \\
-\phi_{it_j} =&\ (\beta_0 + b_{0_j})\ + \\
-&\ \beta_1 \text{Derogation in effect}_{it} + \beta_2 \text{Civil society repression}_{it} \\
-&\ \beta_3\ {(\text{Derogation} \times \text{Civil society repression})}_{i}\ + \\
-&\ \beta_4\ \text{New cases}_{it}\ + \beta_5\ \text{Cumulative cases}_{it}\ + \\
-&\ \beta_6\ \text{New deaths}_{it}\ + \beta_7\ \text{Cumulative deaths}_{it}\ + \\
-&\ \beta_8\ \text{Past ICCPR derogation}_{it}\ + \beta_9\ \text{Past ICCPR action}_{it}\ + \\
-&\ \beta_{10}\ \text{Rule of law index}_{it}\ + \beta_{11}\ \text{Civil liberties index}_{it}\ + \\
-&\ \beta_{12}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
-&\ \textbf{Priors} \\
-\beta_{0 \dots 12} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
-\alpha_k \sim&\ \mathcal{N}(0, 1)
-\end{aligned}
-$$
-
-### H~2~: Derogations, civil society, and human rights (logistic regression)
-
-$$
-\begin{aligned}
-&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
-\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
-&\ \textbf{Distribution parameters} \\
-\pi_{it_j} =&\ (\beta_0 + b_{0_j})\ + \\
-&\ \beta_1 \text{Derogation in effect}_{it} + \beta_2 \text{Civil society repression}_{it} \\
-&\ \beta_3\ {(\text{Derogation} \times \text{Civil society repression})}_{i}\ + \\
-&\ \beta_4\ \text{New cases}_{it}\ + \beta_5\ \text{Cumulative cases}_{it}\ + \\
-&\ \beta_6\ \text{New deaths}_{it}\ + \beta_7\ \text{Cumulative deaths}_{it}\ + \\
-&\ \beta_8\ \text{Past ICCPR derogation}_{it}\ + \beta_9\ \text{Past ICCPR action}_{it}\ + \\
-&\ \beta_{10}\ \text{Rule of law index}_{it}\ + \beta_{11}\ \text{Civil liberties index}_{it}\ + \\
-&\ \beta_{12}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
-&\ \textbf{Priors} \\
-\beta_{0 \dots 12} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
-\end{aligned}
-$$
-
-\newpage
-
-```{r tbl-results-full-prelim}
-#| tbl-cap: Results from models showing relationship between derogations and emergency policies
-#| include: true
-#| echo: false
-notes <- c(
-  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
-  "95% credible intervals (highest density posterior interval, or HDPI) in brackets.",
-  "Total \\(R^2\\) considers the variance of both population and group effects;",
-  "marginal \\(R^2\\) only takes population effects into account."
-)
-
-if (fmt_out == "latex") {
-  notes <- notes %>% 
-    str_replace_all(r"[\\]", r"[\\\\]") %>% 
-    str_replace_all("%", r"(\\\\%)")
-} else {
-  notes <- paste(notes, collapse = " ")
-}
-
-if (fmt_out == "latex") { 
-  modelsummary(models_tbl_prelim,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "kableExtra",
-               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
-               escape = FALSE) %>% 
-    kable_styling(htmltable_class = "table-sm light-border",
-                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
-    column_spec(1:6, width = "0.16\\\\textwidth")
-} else {
-  modelsummary(models_tbl_prelim,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "gt",
-               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2)) %>% 
-    tab_footnote(footnote = notes) %>% 
-    tab_style(style = cell_text(weight = "bold"),
-              locations = cells_column_labels())
-}
-```
-
-\newpage
-
-```{r tbl-results-full-h1}
-#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~1~)
-#| include: true
-#| echo: false
-if (fmt_out == "latex") {
-  modelsummary(models_tbl_h1,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "kableExtra",
-               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
-               escape = FALSE) %>% 
-    kable_styling(htmltable_class = "table-sm light-border",
-                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
-    column_spec(1:5, width = "0.18\\\\textwidth")
-} else {
-  modelsummary(models_tbl_h1,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "gt",
-               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
-               escape = FALSE) %>% 
-    tab_footnote(footnote = notes) %>% 
-    tab_style(style = cell_text(weight = "bold"),
-              locations = cells_column_labels())
-}
-```
-
-\newpage
-
-```{r tbl-results-full-h2}
-#| tbl-cap: Models showing relationship between derogations, civil society, and human rights (H~2~)
-#| include: true
-#| echo: false
-if (fmt_out == "latex") {
-  modelsummary(models_tbl_h2,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "kableExtra",
-               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
-               escape = FALSE) %>% 
-    kable_styling(htmltable_class = "table-sm light-border",
-                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
-    column_spec(1:5, width = "0.18\\\\textwidth")
-} else {
-  modelsummary(models_tbl_h2,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "gt",
-               fmt =  list(estimate = 2, conf.low = 2, conf.high = 2),
-               escape = FALSE) %>% 
-    tab_footnote(footnote = notes) %>% 
-    tab_style(style = cell_text(weight = "bold"),
-              locations = cells_column_labels())
-}
-```
-
-\clearpage
-
 # References

### 2023-04-20 manuscript/manuscript.qmd
Add note about Oxford strictness

@@
 Our results support the role that flexibility mechanisms such as derogations play in international law.
-date: March 8, 2023
+date: April 20, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
+> **Hypothesis 1~a~**: If states blah, then blah.
+>
+> **Hypothesis 2~b~**: If states blah, then blah.
+>
+
 ## Derogations and human rights
@@
 
-We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited seriously, and therefore more likely to impose measures that violated these criteria.
+We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to impose measures that violated these criteria.
 If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.
 
-> **Hypothesis 1~a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
+> **Hypothesis 2~a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
 >
-> **Hypothesis 1~b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
+> **Hypothesis 2~b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
 >
-> **Hypothesis 1~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
+> **Hypothesis 2~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
 
@@
   "ICCPR derogation",                          "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
-  "Civil society repression",                  "≈ −3 to 3; higher values indicate less repression",                 "Annual",    "Varieties of Democracy",
   "New and cumulative COVID deaths and cases", "Counts",                                                            "Weekly",    "World Health Organization",
@@
   kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA)) %>% 
-  pack_rows(index = c(" " = 2, "Primary explanatory variables" = 2, "Secondary explanatory variables" = 5),
+  pack_rows(index = c(" " = 2, "Primary explanatory variable" = 1, "Secondary explanatory variables" = 5),
             indent = FALSE) %>% 
@@
 
-We examine the relationship between formal treaty derogations and different human rights and policy outcomes in three stages (see @tbl-vars).
-First, to explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
-Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.
-As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in a given week.
+We examine the relationship between formal treaty derogations and different policy and human rights outcomes in two stages (see @tbl-vars).
+First, to test our first set of hypotheses and explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
+Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in a given week.
+
+[^severity]: Importantly, the Oxford Government Response Tracker measures the *stringency* or strictness of policy enforcement, not the *quality* of policies, so the values do not necessarily reflect how well countries implemented these emergency measures.
 
@@
 | General question                           | Outcomes                                                                                                                                  | Main explanatory variable  | Other explanatory variables                                                                                                                                              |
-|----------------|------------------|----------------|----------------------|
-| Derogations and emergency measures                  | Cancelling public events, Gathering restrictions, Close public transit, Restrictions on internal movement, Limits on international travel | ICCPR derogation in effect                  | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions                                                                     |
-| Derogations and human rights (H~1~)                 | Discriminatory policy, Violations of non-derogable rights, No time limit measures, Abusive enforcement                                    | ICCPR derogation in effect                  | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index, core civil society index |
-| Derogations, human rights, and civil society (H~2~) | Discriminatory policy, Violations of non-derogable rights, Limiting media, Abusive enforcement                                            | ICCPR derogation × civil society repression | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index                           |
+|-----------------|-----------------|-----------------|---------------------|
+| Derogations and emergency measures (H~1~)  | Cancelling public events, Gathering restrictions, Close public transit, Restrictions on internal movement, Limits on international travel | ICCPR derogation in effect | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions                                                                     |
+| Derogations and human rights (H~2~)        | Discriminatory policy, Violations of non-derogable rights, No time limit measures, Abusive enforcement                                    | ICCPR derogation in effect | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index, core civil society index |
 
@@
 
-Next, to test our first hypothesis, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
+Next, to test our second set of hypotheses, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
 Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations.
 We do not collapse these variables into binary indicators and instead work with them as ordered categories.
-Finally, we test our second hypothesis by exploring the more complex relationship treaty derogations, civil society regulations, and four policy outcomes from the PanDem project.
-We look at the same four policies as the first hypothesis with one exception: since civil society is often concerned with advocacy and information provision during public emergencies, we replace whether emergency measures have a time limit with an ordinal measure of whether limits were placed on media freedom as part of a state's pandemic response.
 Importantly, since these PanDem variables are all measured quarterly, while our main data is recorded weekly, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response each week.
@@
 
-### Key explanatory variables
+### Key explanatory variable
 
-We use three different primary independent variables to explain changes in policy and human rights outcomes.
-For our preliminary models and in models related to our first hypothesis, we rely on a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
+For both sets of hypotheses, we explain changes in policy and human rights outcomes using a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
 We collected and coded this data from real time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023].
@@
 
-For our second hypothesis, we use a measure of civil society repression from the Varieties of Democracy (V-Dem) project [@vdem-v12], derived from a latent Bayesian item response theory model of expert opinions, with values ranging from roughly −3 to 3 (where higher values represent less repression).
-This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti-civil society environment.
+### Secondary explanatory variables
 
-### Confounding variables
-
-Finally, we include several other variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
+We include several additionally variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
 To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
@@
 In all model specifications, we include new and cumulative COVID-19 cases and deaths as controls.
-In our first hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society.
-In our second hypothesis, we omit the core civil society index and instead include the interaction of civil society repression and derogations.
+In our second hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society.
 
@@
 
-Before exploring whether COVID-induced derogations are associated with specific human rights abuses, it is useful to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
+Before exploring whether COVID-induced derogations are associated with specific human rights abuses, we first explore our first set of hypotheses to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
 @fig-preds-mfx-prelim-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
@@
 
-While these results do not speak directly to our main hypotheses, they do reveal important insights into the motivations of derogating states.
+These results reveal important insights into the motivations of derogating states.
 Countries that undertook emergency safety measures seem to have taken the time to follow international law and uphold their treaty obligations.
@@
 Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
-We thus find strong support only for H~1c~—that derogating states are less likely to impose measures with no time limits.
-We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~1b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~1c~).
+We thus find strong support only for H~2c~—that derogating states are less likely to impose measures with no time limits.
+We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~2b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
 Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
-
-## Derogations, civil society, and human rights
-
-Civil society repression, on the other hand, adds an additional wrinkle to the story.
-Countries with severe level of repression (i.e.
-negative repression scores) that do not derogate have a 25% predicted probability of minor to major discriminatory policies, while similarly repressive countries that do derogate have nearly a 0% predicted probability of discriminatory policies (see @fig-preds-mfx-h2).
-This represents a substantial 5–10 percentage point difference between derogating and non-derogating countries with repressive civil societies.
-The derogation process could thus serve as a strong signal that emergency measures are indeed planned to be limited and non-discriminatory even in countries where civil society is under threat.
-
-However, there are no other substantial differences in predicted probabilities across derogation status for another other outcomes.
-Derogating and non-derogating states are generally equally likely to violate non-derogable rights (H~2a~), limit media (H~2d~), and engage in abusive enforcement (H~2c~) regardless of their level of civil society repression.
-We thus only find support for H~2b~—if states are repressive toward civil society organizations, they are more likely to use discriminatory policies, while derogations strongly negate this trend.
-This may be because advocacy-focused civil society organizations in particular tend to deal with on-the-ground monitoring, advocacy, and shaming, and is more likely to call out governments for unequal and unjust implementation of their emergency policies.
-
-```{r fig-preds-mfx-h2, fig.width=13, fig.height=6, out.width="100%"}
-#| include: true
-#| echo: false
-#| layout-ncol: 1
-#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
-#| fig-subcap:
-#|   - "Predicted probabilities"
-#|   - "Contrasts between predicted probabilities"
-
-h2_plots <- h2_plot_data %>% 
-  # Predicted probabilities
-  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
-    if (.family == "cumulative") {
-      .data %>% 
-        ggplot(aes(
-          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
-          color = .category, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = "Civil society repression", y = "Predicted probability", 
-             color = "Severity", fill = "Severity",
-             title = .nice_name) +
-        facet_wrap(vars(derogation_ineffect)) +
-        theme_pandem()
-    } else {
-      .data %>% 
-        ggplot(aes(
-          x = v2csreprss, y = .epred, ymin = .lower, ymax = .upper,
-          color = derogation_ineffect, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_y_continuous(labels = label_percent()) +
-        labs(x = "Civil society repression", y = "Predicted probability", 
-             color = "Derogation in effect", fill = "Derogation in effect",
-             title = .nice_name) +
-        theme_pandem()
-    }
-  })) %>% 
-  # Marginal effects / contrasts
-  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
-    if (.family == "cumulative") {
-      .data %>% 
-        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
-        ggplot(aes(
-          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
-          color = group, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3) +
-        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)]), guide = "none") +
-        labs(x = NULL,
-             y = "Percentage point difference\nin probability of outcome level",
-             title = .nice_name) +
-        facet_wrap(vars(group), ncol = 2) +
-        theme_pandem()
-    } else {
-      .data %>% 
-        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
-        ggplot(aes(
-          x = v2csreprss, y = draw, ymin = .lower, ymax = .upper,
-          group = forcats::fct_rev(ordered(.width))
-        )) +
-        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
-        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        labs(x = NULL,
-             y = "Percentage point difference\nin probability of outcome",
-             title = .nice_name) +
-        theme_pandem()
-    }
-  }))
-
-wrap_plots(h2_plots$pred_plot)
-wrap_plots(h2_plots$mfx_plot)
-```
 
@@
 
-The Covid-19 pandemic raised numerous concerns that the cris would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
+The Covid-19 pandemic raised numerous concerns that the crisis would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
 Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.

### 2023-04-20 analysis/analysis_policies.qmd
Rename; go to 5946ab5 for previous history
Going to 5946ab5 and seeing the history of /analysis/analysis_prelim.qmd and /analysis/analysis_h1.qmd will show the whole past history

--- 2023-03-06 analysis/analysis_prelim.qmd
@@
 \begin{aligned}
-\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\
+\text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\[0.75em]
 \operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{country} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
@@
 &\ \beta_6\ \text{Week number}_{i} \\
-\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\
-\\
+\alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\[0.75em]
 \bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
@@
 ```{r}
-prelim_trace <- function(model) {
-  model %>% 
-    gather_draws(b_Intercept, b_derogation_ineffect, b_year_week_num, sd_country_name__Intercept) %>% 
-    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
-    geom_line(linewidth = 0.1) +
-    scale_color_viridis_d(option = "rocket", end = 0.85) +
-    labs(color = "Chain") +
-    facet_wrap(vars(.variable), scales = "free_y") +
-    theme_pandem()
-}
-
-prelim_trank <- function(model) {
-  model %>% 
-    gather_draws(b_Intercept, b_derogation_ineffect, b_year_week_num, sd_country_name__Intercept) %>% 
-    group_by(.variable) %>% 
-    mutate(draw_rank = rank(.value)) %>% 
-    ggplot(aes(x = draw_rank, color = factor(.chain))) +
-    stat_bin(geom = "step", binwidth = 200, position = position_identity(), boundary = 0) +
-    scale_color_viridis_d(option = "rocket", end = 0.85) +
-    labs(color = "Chain") +
-    facet_wrap(vars(.variable), scales = "free_y") +
-    theme_pandem()
-}
-
-prelim_pp <- function(model) {
-  pp_check(model, ndraws = 100, type = "bars")
-}
+params_to_show <- c("b_Intercept", "b_derogation_ineffect", 
+                    "b_year_week_num", "sd_country_name__Intercept")
 ```
@@
 ```{r}
-prelim_trace(m_prelim_derog$model[[1]])
+plot_trace(m_prelim_derog$model[[1]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_prelim_derog$model[[1]])
+plot_trank(m_prelim_derog$model[[1]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_prelim_derog$model[[1]])
+plot_pp(m_prelim_derog$model[[1]])
 ```
@@
 ```{r}
-prelim_trace(m_prelim_derog$model[[2]])
+plot_trace(m_prelim_derog$model[[2]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_prelim_derog$model[[2]])
+plot_trank(m_prelim_derog$model[[2]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_prelim_derog$model[[2]])
+plot_pp(m_prelim_derog$model[[2]])
 ```
@@
 ```{r}
-prelim_trace(m_prelim_derog$model[[3]])
+plot_trace(m_prelim_derog$model[[3]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_prelim_derog$model[[3]])
+plot_trank(m_prelim_derog$model[[3]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_prelim_derog$model[[3]])
+plot_pp(m_prelim_derog$model[[3]])
 ```
@@
 ```{r}
-prelim_trace(m_prelim_derog$model[[4]])
+plot_trace(m_prelim_derog$model[[4]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_prelim_derog$model[[4]])
+plot_trank(m_prelim_derog$model[[4]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_prelim_derog$model[[4]])
+plot_pp(m_prelim_derog$model[[4]])
 ```
@@
 ```{r}
-prelim_trace(m_prelim_derog$model[[5]])
+plot_trace(m_prelim_derog$model[[5]], params_to_show)
 ```
@@
 ```{r}
-prelim_trank(m_prelim_derog$model[[5]])
+plot_trank(m_prelim_derog$model[[5]], params_to_show)
 ```
@@
 ```{r}
-prelim_pp(m_prelim_derog$model[[5]])
+plot_pp(m_prelim_derog$model[[5]])
 ```

### 2023-04-20 manuscript/manuscript.qmd
Audrey additions

@@
 
-AUDREY: Talk about the map of countries that derogated. The figure will go here-ish, I think
+Between January 2020 and June 2021, 23 states issued a total of 114 ICCPR derogations.
+[ INSERT MAP ABOUT HERE ] Of derogations that detailed specific obligation suspension lengths, the average length was 41 days.
+Descriptively, we see states across multiple regions derogated with notable concentration in South America, which has a robust history of engaging with human rights law.
+We also note that it is not necessarily the 'usual suspects' of strong democracies derogating.
+Notably, Nordic states, which typically are very active in human rights law engagement did not issue any derogations during this period.
 
@@
 During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21).
+As our descriptive data shows, not all states issued derogations.
+Derogating states were in the minority of states during the pandemic.
 
-AUDREY: write expectations about what how derogating and non-derogating states differed in their likelihood of imposing emergency policies - canceling public events, gathering restrictions, closing public transit, movement and international travel
+ We expect that states that derogate during COVID-19 with the intention to impose emergency policies will follow through.
+Our expectation comes from the specific, time-limited nature of derogations and the limited audience they speak to.
+Critics of international law may expect states ratify treaties with the expectation of reputational and other benefits.
+However, derogations have no such public audience and are not signaling an increased commitment to human rights.
+There is no clear or expected connection with foreign aid, trade, or other theorized 'rewards' for larger legal actions.
+Instead, derogations are a very specific legal action that states voluntarily enter into after commitment has fully taken place.
+Given this, we expect that derogating states took the legal action seriously and generally followed through with the derogations they submitted.
 
-> **Hypothesis 1~a~**: If states blah, then blah.
+> **Hypothesis 1~a~**: If states issued COVID-19 derogations, then we expect that they will be more likely to impose relevant emergency polices such as canceling public events, restricting gatherings, closing public transit, and restricting domestic and international movement.
 >
-> **Hypothesis 2~b~**: If states blah, then blah.
+> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose related emergency policies.  
 >

### 2023-04-20 manuscript/manuscript.qmd
Consistent capitalization of COVID

@@
 Did derogating and non-derogating states implement emergency measures differently, and if so, how?
-Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
+Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
 On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
@@
 
-Using data from the Varieties of Democracy PanDem dataset, the Oxford Covid-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation among derogating and non-derogating states on the implementation of emergency measures.
+Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation among derogating and non-derogating states on the implementation of emergency measures.
 We then analyze the effect of derogating (or not) on human rights practices during the pandemic.
@@
 
-# The impact of derogations during Covid-19
+# The impact of derogations during COVID-19
 
@@
 While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, human rights, and state actions during the pandemic.
-We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
+We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
 We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021.
@@
   ~`Outcome variables`,                        ~Description,                                                        ~Frequency,  ~Source,
-  "Emergency public health measures",          "Binary indicator",                                                  "Weekly",    "Oxford Covid-19 Government Response Tracker",
+  "Emergency public health measures",          "Binary indicator",                                                  "Weekly",    "Oxford COVID-19 Government Response Tracker",
   "Human rights and policy outcomes",          "Ordered categories for the severity of abuses",                     "Quarterly", "Varieties of Democracy PanDem",
@@
 We examine the relationship between formal treaty derogations and different policy and human rights outcomes in two stages (see @tbl-vars).
-First, to test our first set of hypotheses and explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford Covid-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
+First, to test our first set of hypotheses and explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
 Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in a given week.
@@
 
-The Covid-19 pandemic raised numerous concerns that the crisis would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
+The COVID-19 pandemic raised numerous concerns that the crisis would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
 Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.

### 2023-04-20 analysis/analysis_human-rights.qmd
Remove H2, upgrade prelim models to H1

--- 2023-03-08 analysis/analysis_h1.qmd
@@
 ---
-title: "H<sub>1</sub>: Derogations and human rights"
+title: "H<sub>2</sub>: Derogations and human rights"
 format:
@@
 # Models
-tar_load(m_h1)
-tar_load(h1_plot_data)
-tar_load(models_tbl_h1)
+tar_load(m_human_rights)
+tar_load(human_rights_plot_data)
+tar_load(models_tbl_human_rights)
 
@@
 
-model_h1_prior_only <- brm(
+model_human_rights_prior_only <- brm(
   bf(pandem_nolimit ~ derogation_ineffect + 
@@
 )
-model_h1_prior_only
+model_human_rights_prior_only
 ```
@@
 ```{r}
-datagrid(model = model_h1_prior_only,
+datagrid(model = model_human_rights_prior_only,
          year_week_num = 1:69,
          derogation_ineffect = 0:1) %>% 
-  add_epred_draws(model_h1_prior_only, ndraws = 100) %>% 
+  add_epred_draws(model_human_rights_prior_only, ndraws = 100) %>% 
   left_join(year_week_lookup, by = "year_week_num") %>% 
@@
 
-```{r tbl-h1, cache=TRUE, warning=FALSE}
+```{r tbl-human-rights, cache=TRUE, warning=FALSE}
 notes <- c(
@@
 
-modelsummary(models_tbl_h1,
+modelsummary(models_tbl_human_rights,
              estimate = "{estimate}",
@@
 
-### `r m_h1$nice[[1]]`
+### `r m_human_rights$nice[[1]]`
 
 ```{r}
-plot_trace(m_h1$model[[1]], params_to_show)
+plot_trace(m_human_rights$model[[1]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_h1$model[[1]], params_to_show)
+plot_trank(m_human_rights$model[[1]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_h1$model[[1]])
+plot_pp(m_human_rights$model[[1]])
 ```
 
-### `r m_h1$nice[[2]]`
+### `r m_human_rights$nice[[2]]`
 
 ```{r}
-plot_trace(m_h1$model[[2]], params_to_show)
+plot_trace(m_human_rights$model[[2]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_h1$model[[2]], params_to_show)
+plot_trank(m_human_rights$model[[2]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_h1$model[[2]])
+plot_pp(m_human_rights$model[[2]])
 ```
 
-### `r m_h1$nice[[3]]`
+### `r m_human_rights$nice[[3]]`
 
 ```{r}
-plot_trace(m_h1$model[[3]], params_to_show)
+plot_trace(m_human_rights$model[[3]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_h1$model[[3]], params_to_show)
+plot_trank(m_human_rights$model[[3]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_h1$model[[3]])
+plot_pp(m_human_rights$model[[3]])
 ```
 
-### `r m_h1$nice[[4]]`
+### `r m_human_rights$nice[[4]]`
 
 ```{r}
-plot_trace(m_h1$model[[4]], params_to_show)
+plot_trace(m_human_rights$model[[4]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_h1$model[[4]], params_to_show)
+plot_trank(m_human_rights$model[[4]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_h1$model[[4]])
+plot_pp(m_human_rights$model[[4]])
 ```
@@
 ```{r fig.width=13, fig.height=6, out.width="100%"}
-h1_plots <- h1_plot_data %>% 
+human_rights_plots <- human_rights_plot_data %>% 
   # Predicted probabilities
@@
 
-wrap_plots(h1_plots$pred_plot)
+wrap_plots(human_rights_plots$pred_plot)
 ```
@@
 ```{r fig.width=13, fig.height=6, out.width="100%"}
-wrap_plots(h1_plots$mfx_plot)
+wrap_plots(human_rights_plots$mfx_plot)
 ```

### 2023-04-20 analysis/analysis_policies.qmd
Remove H2, upgrade prelim models to H1

@@
 ---
-title: "Derogations and emergency policies"
+title: "H<sub>1</sub>: Derogations and emergency policies"
 format:
@@
 # Models
-tar_load(m_prelim_derog)
-tar_load(prelim_plot_data)
-tar_load(models_tbl_prelim)
+tar_load(m_policies)
+tar_load(policies_plot_data)
+tar_load(models_tbl_policies)
 
@@
 
-model_prelim_prior_only <- brm(
+model_policies_prior_only <- brm(
   bf(c7_internal_movement_bin ~ derogation_ineffect + 
@@
 )
-model_prelim_prior_only
+model_policies_prior_only
 ```
@@
 ```{r}
-datagrid(model = model_prelim_prior_only,
+datagrid(model = model_policies_prior_only,
          year_week_num = 1:69,
          derogation_ineffect = 0:1) %>% 
-  add_epred_draws(model_prelim_prior_only, ndraws = 100) %>% 
+  add_epred_draws(model_policies_prior_only, ndraws = 100) %>% 
   left_join(year_week_lookup, by = "year_week_num") %>% 
@@
 
-```{r tbl-prelim, cache=TRUE}
+```{r tbl-policies, cache=TRUE}
 notes <- c(
@@
 
-modelsummary(models_tbl_prelim,
+modelsummary(models_tbl_policies,
              estimate = "{estimate}",
@@
 
-### `r m_prelim_derog$nice[[1]]`
+### `r m_policies$nice[[1]]`
 
 ```{r}
-plot_trace(m_prelim_derog$model[[1]], params_to_show)
+plot_trace(m_policies$model[[1]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_prelim_derog$model[[1]], params_to_show)
+plot_trank(m_policies$model[[1]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_prelim_derog$model[[1]])
+plot_pp(m_policies$model[[1]])
 ```
 
-### `r m_prelim_derog$nice[[2]]`
+### `r m_policies$nice[[2]]`
 
 ```{r}
-plot_trace(m_prelim_derog$model[[2]], params_to_show)
+plot_trace(m_policies$model[[2]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_prelim_derog$model[[2]], params_to_show)
+plot_trank(m_policies$model[[2]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_prelim_derog$model[[2]])
+plot_pp(m_policies$model[[2]])
 ```
 
-### `r m_prelim_derog$nice[[3]]`
+### `r m_policies$nice[[3]]`
 
 ```{r}
-plot_trace(m_prelim_derog$model[[3]], params_to_show)
+plot_trace(m_policies$model[[3]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_prelim_derog$model[[3]], params_to_show)
+plot_trank(m_policies$model[[3]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_prelim_derog$model[[3]])
+plot_pp(m_policies$model[[3]])
 ```
 
-### `r m_prelim_derog$nice[[4]]`
+### `r m_policies$nice[[4]]`
 
 ```{r}
-plot_trace(m_prelim_derog$model[[4]], params_to_show)
+plot_trace(m_policies$model[[4]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_prelim_derog$model[[4]], params_to_show)
+plot_trank(m_policies$model[[4]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_prelim_derog$model[[4]])
+plot_pp(m_policies$model[[4]])
 ```
 
-### `r m_prelim_derog$nice[[5]]`
+### `r m_policies$nice[[5]]`
 
 ```{r}
-plot_trace(m_prelim_derog$model[[5]], params_to_show)
+plot_trace(m_policies$model[[5]], params_to_show)
 ```
@@
 ```{r}
-plot_trank(m_prelim_derog$model[[5]], params_to_show)
+plot_trank(m_policies$model[[5]], params_to_show)
 ```
@@
 ```{r}
-plot_pp(m_prelim_derog$model[[5]])
+plot_pp(m_policies$model[[5]])
 ```
@@
 ```{r fig.width=12, fig.height=6, out.width="100%"}
-prelim_plots <- prelim_plot_data %>% 
+policies_plots <- policies_plot_data %>% 
   mutate(pred_plot = map2(pred_data, nice, ~{
@@
 
-wrap_plots(prelim_plots$pred_plot)
+wrap_plots(policies_plots$pred_plot)
 ```
@@
 ```{r fig.width=12, fig.height=6, out.width="100%"}
-wrap_plots(prelim_plots$mfx_plot)
+wrap_plots(policies_plots$mfx_plot)
 ```

### 2023-04-20 analysis/hypotheses.qmd
Remove H2, upgrade prelim models to H1

--- 2023-02-15 analysis/hypotheses.qmd
@@
 
-## H~1~: Formal derogations and pandemic-related human rights violations
+## H~1~: Formal derogations and pandemic-related emergency policies
 
-- **H~1a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
-- **H~1b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights
-- **H~1c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
+- **H~1a~**: If states issued COVID-19 derogations, then we expect that they will be more likely to impose relevant emergency polices such as canceling public events, restricting gatherings, closing public transit, and restricting domestic and international movement.
+- **H~1b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose related emergency policies.
 
-## H~2~: Civil society repression and pandemic-related human rights violations
+## H~2~: Formal derogations and pandemic-related human rights violations
 
-- **H~2a~**: If states are repressive towards civil society organizations, then they are more likely to violate non-derogable rights.
-- **H~2b~**: If states are repressive towards civil society organizations, then they are more likely to use discriminatory policy measures during the pandemic.
-- **H~2c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.
-- **H~2d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom.
+- **H~2a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
+- **H~2b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
+- **H~2c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.

### 2023-04-20 analysis/model-details.qmd
Remove H2, upgrade prelim models to H1

@@
 # Models
-tar_load(c(m_prelim_derog, m_h1, m_h2))
+tar_load(c(m_policies, m_human_rights))
 ```
@@
 ```{r calculate-model-times}
-model_times <- bind_rows(list("**Preliminary models: Derogations and emergency policies**" = m_prelim_derog, 
-                              "**H<sub>1</sub>: Derogations and human rights**" = m_h1, 
-                              "**H<sub>2</sub>: Derogations, civil society, and human rights**" = m_h2), 
+model_times <- bind_rows(list("**H<sub>1</sub>: Derogations and emergency policies**" = m_policies, 
+                              "**H<sub>2</sub>: Derogations and human rights**" = m_human_rights), 
                          .id = "hypothesis") %>%  

### 2023-04-20 manuscript/manuscript.qmd
Remove H2, upgrade prelim models to H1

@@
 
-tar_load(c(prelim_plot_data, h1_plot_data, h2_plot_data))
-tar_load(c(models_tbl_prelim, models_tbl_h1, models_tbl_h2))
+tar_load(c(policies_plot_data, human_rights_plot_data))
+tar_load(c(models_tbl_policies, models_tbl_human_rights))
 
@@
 
-## Derogations under COVID-19
+## Derogations during COVID-19
 
@@
 > **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose related emergency policies.
->
 
@@
 Before exploring whether COVID-induced derogations are associated with specific human rights abuses, we first explore our first set of hypotheses to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
-@fig-preds-mfx-prelim-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
-In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-prelim-2 for the percentage-point differences between the predictions for the two groups of states).
+@fig-preds-mfx-policies-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
+In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states).
 
@@
 
-```{r fig-preds-mfx-prelim, fig.width=12, fig.height=6, out.width="100%", cache=FALSE}
+```{r fig-preds-mfx-policies, fig.width=12, fig.height=6, out.width="100%", cache=FALSE}
 #| include: true
@@
 
-prelim_plots <- prelim_plot_data %>% 
+policies_plots <- policies_plot_data %>% 
   mutate(pred_plot = map2(pred_data, nice, ~{
@@
 
-wrap_plots(prelim_plots$pred_plot)
-wrap_plots(prelim_plots$mfx_plot)
+wrap_plots(policies_plots$pred_plot)
+wrap_plots(policies_plots$mfx_plot)
 ```
@@
 
-The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-h1).
+The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights).
 There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement.
@@
 
-```{r fig-preds-mfx-h1, fig.width=13, fig.height=6, out.width="100%"}
+```{r fig-preds-mfx-human-rights, fig.width=13, fig.height=6, out.width="100%"}
 #| include: true
@@
 
-h1_plots <- h1_plot_data %>% 
+human_rights_plots <- human_rights_plot_data %>% 
   # Predicted probabilities
@@
 
-wrap_plots(h1_plots$pred_plot)
-wrap_plots(h1_plots$mfx_plot)
+wrap_plots(human_rights_plots$pred_plot)
+wrap_plots(human_rights_plots$mfx_plot)
 ```

### 2023-04-20 R/funs_models.R

@@ -28,7 +28,7 @@ build_modelsummary <- function(model_df) {
 #   "elpd.se",   "ELPD (SE)", 1,    FALSE
 # )
 # 
-# modelsummary(models_prelim,
+# modelsummary(models_policies,
 #              estimate = "{estimate}",
 #              statistic = "conf.int",
 #              gof_map = gm,
@@ -43,8 +43,6 @@ gof_map <- tribble(
 
 coef_map <- c(
   "b_derogation_ineffect" = "Derogation in effect",
-  "b_v2csreprss" = "Civil society repression",
-  "b_derogation_ineffect:v2csreprss" = "Derogation × Civil society repression",
   "b_new_cases_z" = "New cases (standardized)",
   "b_cumulative_cases_z" = "Cumulative cases (standardized)",
   "b_new_deaths_z" = "New deaths (standardized)",
@@ -65,7 +63,7 @@ coef_map <- c(
 
 
 # Model definitions
-f_prelim_derog <- function(panel) {
+f_policies <- function(panel) {
   BAYES_SEED <- 1757  # From random.org
   
   panel <- panel %>% 
@@ -83,7 +81,7 @@ f_prelim_derog <- function(panel) {
                     prior(student_t(1, 0, 3), class = b),
                     prior(cauchy(0, 1), class = sd, lb = 0))
   
-  prelim_derog_model <- function(y, data) {
+  policies_model <- function(y, data) {
     form <- glue::glue(y, " ~ derogation_ineffect + new_cases_z + cumulative_cases_z + ",
                        "new_deaths_z + cumulative_deaths_z + ",
                        "prior_iccpr_derogations + prior_iccpr_other_action + ",
@@ -105,7 +103,7 @@ f_prelim_derog <- function(panel) {
     )
   }
   
-  preliminary_models <- tribble(
+  policies_models <- tribble(
     ~nice, ~y, ~never,
     "Cancel Public Events", "c3_cancel_events_bin", "c3_cancel_events_never",
     "Gathering Restrictions", "c4_gatherings_bin", "c4_gatherings_never",
@@ -114,13 +112,13 @@ f_prelim_derog <- function(panel) {
     "International Travel", "c8_intl_travel_bin", "c8_intl_travel_never"
   ) %>% 
     mutate(data = map(never, ~never_filter(.))) %>% 
-    mutate(model = map2(y, data, ~prelim_derog_model(.x, .y)))
[…]

### 2023-04-20 R/funs_plots.R

@@ -53,7 +53,7 @@ plot_pp <- function(model) {
 
 library(marginaleffects)
 
-build_prelim_plot_data <- function(model_df, year_week_lookup) {
+build_policies_plot_data <- function(model_df, year_week_lookup) {
   model_df_preds_mfx <- model_df %>% 
     mutate(pred_data = map(model, ~{
       datagrid(model = .x,
@@ -83,7 +83,7 @@ build_prelim_plot_data <- function(model_df, year_week_lookup) {
 }
 
 
-build_h1_plot_data <- function(model_df, year_week_lookup) {
+build_human_rights_plot_data <- function(model_df, year_week_lookup) {
   model_df_preds_mfx <- model_df %>% 
     mutate(pred_data = map2(model, family, ~{
       if (.y == "cumulative") {
@@ -131,58 +131,6 @@ build_h1_plot_data <- function(model_df, year_week_lookup) {
           tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
       }
       
-      mfx
-    }))
-  
-  return(model_df_preds_mfx)
-}
-
-
-build_h2_plot_data <- function(model_df) {
-  model_df_preds_mfx <- model_df %>% 
-    mutate(pred_data = pmap(list(model, family), \(.model, .family) {
-      if (.family == "cumulative") {
-        df <- datagrid(model = .model,
-                       v2csreprss = seq(-2.5, 3, by = 0.1),
-                       derogation_ineffect = 0:1) %>% 
-          tidybayes::add_epred_draws(.model) %>% 
-          mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                              levels = 0:1,
-                                              labels = c("No derogation", "Derogation in effect"),
-                                              ordered = TRUE)) %>% 
-          group_by(v2csreprss, .category, derogation_ineffect) %>% 
-          tidybayes::median_qi(.epred, .width = c(0.5, 0.8, 0.95))
-      } else {
-        df <- datagrid(model = .model,
-                       v2csreprss = seq(-2.5, 3, by = 0.1),
-                       derogation_ineffect = 0:1) %>% 
-          tidybayes::add_epred_draws(.model) %>% 
-          mutate(derogation_ineffect = factor(derogation_ineffect, 
-                                              levels = 0:1,
-                                              labels = c("No", "Yes"),
[…]

### 2023-04-20 _targets.R

@@ -121,26 +121,23 @@ list(
              format = "file"),
   
   ## Models ----
-  tar_target(m_prelim_derog, f_prelim_derog(weekly_panel)),
-  tar_target(m_h1, f_h1(weekly_panel)),
-  tar_target(m_h2, f_h2(weekly_panel)),
+  tar_target(m_policies, f_policies(weekly_panel)),
+  tar_target(m_human_rights, f_human_rights(weekly_panel)),
 
   ## Model tables ----
   # Build tables here because they take a while
   tar_target(modelsummary_functions, lst(coef_map, gof_map)),
-  tar_target(models_tbl_prelim, build_modelsummary(m_prelim_derog)),
-  tar_target(models_tbl_h1, build_modelsummary(m_h1)),
-  tar_target(models_tbl_h2, build_modelsummary(m_h2)),
+  tar_target(models_tbl_policies, build_modelsummary(m_policies)),
+  tar_target(models_tbl_human_rights, build_modelsummary(m_human_rights)),
 
   ## Graphics ----
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
   tar_target(diagnostic_functions, lst(plot_trace, plot_trank, plot_pp)),
   
   ## Plots ----
-  tar_target(prelim_plot_data, build_prelim_plot_data(m_prelim_derog, year_week_lookup)),
-  tar_target(h1_plot_data, build_h1_plot_data(m_h1, year_week_lookup)),
-  tar_target(h2_plot_data, build_h2_plot_data(m_h2)),
-  
+  tar_target(policies_plot_data, build_policies_plot_data(m_policies, year_week_lookup)),
+  tar_target(human_rights_plot_data, build_human_rights_plot_data(m_human_rights, year_week_lookup)),
+
   ## Manuscript and analysis notebook ----
   tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE, profile = "nice"),
   tar_quarto(manuscript_manuscripty, path = "manuscript", quiet = FALSE, profile = "ms"),

### 2023-04-20 R/funs_data-cleaning.R
Remove Civicus stuff

@@ -268,60 +268,6 @@ load_world_map <- function(path) {
   return(world_map)
 }
 
-# Civicus Monitor
-# We downloaded the standalone embeddable widget
-# (https://monitor.civicus.org/widgets/world/) as an HTML file with
-# `wget https://monitor.civicus.org/widgets/world/` and saved it as index_2021-03-19.html
-#
-# We then extracted the COUNTRIES_DATA variable embedded in a <script> tag
-# (xpath = /html/body/script[5]), which is JSON-ish, but not quite. jsonlite
-# can't parse it for whatever reason, but some online JSON formatter and
-# validator could, so we ran it through that and saved the resulting clean file
-load_clean_civicus <- function(path) {
-  civicus_raw <- read_json(path) %>% as_tibble() %>% slice(1)
-  
-  civicus_lookup <- tribble(
-    ~value, ~category,
-    1, "Closed",
-    2, "Repressed",
-    3, "Obstructed",
-    4, "Narrowed",
-    5, "Open"
-  ) %>%
-    mutate(category = fct_inorder(category, ordered = TRUE))
-  
-  civicus_clean <- civicus_raw %>%
-    pivot_longer(everything(), names_to = "name", values_to = "value") %>%
-    mutate(value = map_chr(value, ~.)) %>%
-    mutate(value = parse_number(value, na = c("", "NA", "None"))) %>%
-    mutate(country_name = countrycode(name, "iso3c", "country.name",
-                                      custom_match = c("KOSOVO" = "XKK",
-                                                       "SVT" = "VCT")),
-           iso3c = countrycode(country_name, "country.name", "iso3c",
-                               custom_match = c("XKK" = "Kosovo",
-                                                "VCT" = "Saint Vincent and the Grenadines"))) %>%
-    left_join(civicus_lookup, by = "value") %>%
-    select(-name, -value, -country_name)
-  
-  return(civicus_clean)
-}
-
-create_civicus_map_data <- function(civicus, map) {
-  map_with_civicus <- map %>%
-    # Fix some Natural Earth ISO weirdness
-    mutate(ISO_A3 = ifelse(ISO_A3 == "-99", as.character(ISO_A3_EH), as.character(ISO_A3))) %>%
-    mutate(ISO_A3 = case_when(
-      .$ISO_A3 == "GRL" ~ "DNK",
-      .$NAME == "Norway" ~ "NOR",
-      .$NAME == "Kosovo" ~ "XKK",
-      TRUE ~ ISO_A3
[…]

### 2023-04-20 _targets.R

@@ -52,14 +52,9 @@ list(
              here_rel("data", "raw_data", "ne_110m_admin_0_countries",
                       "ne_110m_admin_0_countries.shp"),
              format = "file"),
-  tar_target(civicus_raw_file,
-             here_rel("data", "raw_data", "Civicus", "civicus_2021-03-19.json"),
-             format = "file"),
   
   ## Process and clean data ----
   tar_target(world_map, load_world_map(naturalearth_raw_file)),
-  tar_target(civicus_clean, load_clean_civicus(civicus_raw_file)),
-  tar_target(civicus_map_data, create_civicus_map_data(civicus_clean, world_map)),
   
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
   tar_target(iccpr_action_clean, clean_iccpr_action(iccpr_treaty_action_file)),

### 2023-04-20 R/funs_graphics.R
Use linewidth instead of size

@@ -13,10 +13,10 @@ theme_pandem <- function(base_size = 11, base_family = "Noto Sans", prior = FALS
                                        family = "Noto Sans", face = "plain"),
           plot.caption = element_text(margin = margin(t = 10), size = rel(0.6),
                                       family = "Noto Sans", face = "plain"),
-          panel.border = element_rect(color = "grey50", fill = NA, size = 0.15),
+          panel.border = element_rect(color = "grey50", fill = NA, linewidth = 0.15),
           panel.spacing = unit(1, "lines"),
           panel.grid.minor = element_blank(),
-          panel.grid.major = element_line(size = 0.25, colour = "grey90"),
+          panel.grid.major = element_line(linewidth = 0.25, colour = "grey90"),
           axis.line = element_blank(),
           axis.ticks = element_blank(),
           axis.title = element_text(size = rel(0.8),

### 2023-04-20 analysis/data-details.qmd
Add map of derogations

@@
 library(targets)
+library(sf)
 library(DT)
@@
 
-# Load data 
-withr::with_dir(here::here(), {
-  # Final data
-  tar_load(daily_panel)
-})
+# Load targets stuff
+tar_config_set(store = here::here('_targets'),
+               script = here::here('_targets.R'))
+
+invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
+tar_load(c(daily_panel, derogation_map_data))
 ```
@@
   datatable()
+```
+
+And here's a map of all the countries that derogated between January 2020 and June 2021:
+
+```{r derogations-map, fig.width=6.5, fig.height=3.5}
+ggplot() +
+  geom_sf(data = derogation_map_data, aes(fill = derogations_1plus),
+          linewidth = 0.1, color = "white") +
+  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
+  scale_fill_binned(low = colorspace::lighten(clrs[1], amount = 0.8), 
+                    high = clrs[1], na.value = "grey85",
+                    breaks = c(1, 4, 8, 12, 16),
+                    name = "Derogations filed: ",
+                    guide = guide_colorsteps(barwidth = 7, barheight = 0.4,
+                                             title.vjust = 1)) +
+  theme_pandem() +
+  theme(panel.border = element_blank(),
+        legend.title = element_text(face = "bold"),
+        legend.justification = "center")
 ```

### 2023-04-20 R/funs_data-cleaning.R

@@ -268,6 +268,30 @@ load_world_map <- function(path) {
   return(world_map)
 }
 
+make_derogation_count <- function(weekly_panel) {
+  new_derogations <- weekly_panel %>% 
+    group_by(iso3) %>% 
+    summarize(derogations = sum(iccpr_derogation_filed))
+  
+  return(new_derogations)
+}
+
+make_derogation_map_data <- function(derogation_count, map) {
+  map_with_derogations <- map %>%
+    # Fix some Natural Earth ISO weirdness
+    mutate(ISO_A3 = ifelse(ISO_A3 == "-99", as.character(ISO_A3_EH), as.character(ISO_A3))) %>%
+    mutate(ISO_A3 = case_when(
+      .$ISO_A3 == "GRL" ~ "DNK",
+      .$NAME == "Norway" ~ "NOR",
+      .$NAME == "Kosovo" ~ "XKK",
+      TRUE ~ ISO_A3
+    )) %>%
+    left_join(derogation_count, by = join_by(ISO_A3 == iso3)) %>% 
+    mutate(derogations_1plus = ifelse(derogations == 0, NA, derogations))
+  
+  return(map_with_derogations)
+}
+
 # When using a file-based target, {targets} requires that the function that
 # saves the file returns a path to the file. write_csv() and write_dta() both
 # invisibly return the data frame being written, and saveRDS() returns NULL, so

### 2023-04-20 _targets.R

@@ -54,8 +54,6 @@ list(
              format = "file"),
   
   ## Process and clean data ----
-  tar_target(world_map, load_world_map(naturalearth_raw_file)),
-  
   tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
   tar_target(iccpr_action_clean, clean_iccpr_action(iccpr_treaty_action_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
@@ -73,6 +71,11 @@ list(
   tar_target(weekly_panel, make_weekly_data(daily_panel)),
   
   tar_target(year_week_lookup, make_year_week_lookup(weekly_panel)),
+  
+  tar_target(world_map, load_world_map(naturalearth_raw_file)),
+  tar_target(derogation_count, make_derogation_count(weekly_panel)),
+  tar_target(derogation_map_data, make_derogation_map_data(derogation_count, world_map)),
+
   
   ## Save data ----
   ### Daily ----

### 2023-04-20 analysis/analysis_policies.qmd
Add V-Dem covariates to new H1

@@
        prior_iccpr_derogations + prior_iccpr_other_action + 
+       v2x_rule + v2x_civlib + v2xcs_ccsi +
        year_week_num + (1 | country_name)), 

### 2023-04-20 R/funs_models.R

@@ -49,7 +49,6 @@ coef_map <- c(
   "b_cumulative_deaths_z" = "Cumulative deaths (standardized)",
   "b_prior_iccpr_derogationsTRUE" = "Past ICCPR derogation",
   "b_prior_iccpr_other_actionTRUE" = "Past ICCPR action",
-  "b_year_week_num" = "Year-week",
   "b_v2x_rule" = "Rule of law",
   "b_v2x_civlib" = "Civil liberties",
   "b_v2xcs_ccsi" = "Core civil society index",
@@ -57,6 +56,7 @@ coef_map <- c(
   "b_Intercept[1]" = "Cut 1",
   "b_Intercept[2]" = "Cut 2",
   "b_Intercept[3]" = "Cut 3",
+  "b_year_week_num" = "Year-week",
   "sd_country_name__Intercept" = "Country random effects σ",
   "sd_who_region__Intercept" = "Region random effects σ"
 )
@@ -85,6 +85,7 @@ f_policies <- function(panel) {
     form <- glue::glue(y, " ~ derogation_ineffect + new_cases_z + cumulative_cases_z + ",
                        "new_deaths_z + cumulative_deaths_z + ",
                        "prior_iccpr_derogations + prior_iccpr_other_action + ",
+                       "v2x_rule + v2x_civlib + v2xcs_ccsi + ",
                        "year_week_num + (1 | country_name)") %>% 
       as.formula()
     

### 2023-04-21 R/funs_plots.R
Use nice theme for pp plots

@@ -25,10 +25,9 @@ plot_trank <- function(model, params) {
 }
 
 plot_pp <- function(model) {
-  bayesplot::pp_check(model, ndraws = 100, type = "bars")
+  bayesplot::pp_check(model, ndraws = 100, type = "bars") +
+    theme_pandem()
 }
-
-
 
 
 # Storing ggplot objects as rds files is BAD 

### 2023-04-21 analysis/analysis_policies.qmd
Update model definition

@@
 \text{COVID policy}_{i, t} \sim&\ \operatorname{Bernoulli}(\pi_{i, t}) \\[0.75em]
-\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{country} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
+\operatorname{logit}(\pi_{i, t}) =&\ \alpha_{\text{region} [i]} + \beta_1\ \text{Derogation in effect}_{i}\ + \\
 &\ \beta_2\ \text{New cases}_{i}\ + \beta_3\ \text{Cumulative cases}_{i}\ + \\
 &\ \beta_4\ \text{New deaths}_{i}\ + \beta_5\ \text{Cumulative deaths}_{i}\ + \\
-&\ \beta_6\ \text{Week number}_{i} \\
+&\ \beta_6\ \text{Past ICCPR derogation}_{i}\ + \beta_7\ \text{Past ICCPR action}_{i}\ + \\
+&\ \beta_8\ \text{Rule of law index}_{i}\ + \beta_9\ \text{Civil liberties index}_{i}\ + \\
+&\ \beta_{10}\ \text{Core civil society index}_{i}\ + \beta_{11}\ \text{Week number}_{i} \\
 \alpha_j \sim&\ \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \text{, for } j \text{ in } 1 \dots J \\[0.75em]
 \bar{\alpha} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
-\beta_{1 \dots 6} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
+\beta_{1 \dots 11} \sim&\ \operatorname{Student t}(\nu = 1, \mu = 0, \sigma = 3) \\
 \sigma_a \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0

### 2023-04-21 manuscript/manuscript.qmd
Add map and list of countries

@@
 library(scales)
+library(sf)
 library(lubridate)
@@
 tar_load(c(models_tbl_policies, models_tbl_human_rights))
+tar_load(c(derogation_map_data, derogation_count))
 
@@
 invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
+```
+
+```{r make-var-lists-and-tbls}
+derog_detail <- derogation_count %>% 
+  filter(derogations > 0) %>% 
+  summarize(countries = n(),
+            total = sum(derogations)) %>% 
+  as.list()
+
+map_data_nice <- derogation_count %>% 
+  filter(derogations > 0) %>% 
+  arrange(desc(derogations)) %>% 
+  mutate(Country = countrycode::countrycode(iso3, "iso3c", "country.name")) %>% 
+  select(Country, Derogations = derogations)
 ```
@@
 
-Between January 2020 and June 2021, 23 states issued a total of 114 ICCPR derogations.
-[ INSERT MAP ABOUT HERE ] Of derogations that detailed specific obligation suspension lengths, the average length was 41 days.
+```{r fig-derogation-map, fig.width=6.5, fig.height=3.5}
+#| include: true
+#| echo: false
+#| fig-cap: "Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures"
+ggplot() +
+  geom_sf(data = derogation_map_data, aes(fill = derogations_1plus),
+          linewidth = 0.1, color = "white") +
+  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
+  scale_fill_binned(low = colorspace::lighten(clrs[1], amount = 0.8), 
+                    high = clrs[1], na.value = "grey85",
+                    breaks = c(1, 5, 10, 15, 20),
+                    limits = c(1, 20),
+                    name = "Derogations filed: ",
+                    guide = guide_colorsteps(barwidth = 7, barheight = 0.4,
+                                             title.vjust = 1)) +
+  theme_pandem() +
+  theme(panel.border = element_blank(),
+        legend.title = element_text(face = "bold"),
+        legend.text = element_text(size = rel(0.6)),
+        legend.justification = "center")
+```
+
+```{r tbl-derogation-countries}
+#| tbl-cap: Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures 
+#| include: true
+#| echo: false
+bind_cols(dplyr::slice(map_data_nice, 1:12),
+          add_row(dplyr::slice(map_data_nice, 13:n()))) %>% 
+  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]",
+      col.names = rep(c("Country", "Derogations"), 2)) %>% 
+  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))
+```
+
+Between January 2020 and June 2021, `r derog_detail$countries` states issued a total of `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
+Of derogations that detailed specific obligation suspension lengths, the average length was 41 days.
 Descriptively, we see states across multiple regions derogated with notable concentration in South America, which has a robust history of engaging with human rights law.
-We also note that it is not necessarily the 'usual suspects' of strong democracies derogating.
-Notably, Nordic states, which typically are very active in human rights law engagement did not issue any derogations during this period.
+We also note that it is not necessarily the "usual suspects" of strong democracies that derogate.
+Notably, Nordic states—which typically are very active in human rights law engagement—did not issue any derogations during this period.
 
@@
 
-In all model specifications, we include new and cumulative COVID-19 cases and deaths as controls.
-In our second hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society.
+In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society.
 

### 2023-04-21 R/funs_data-cleaning.R

@@ -268,8 +268,8 @@ load_world_map <- function(path) {
   return(world_map)
 }
 
-make_derogation_count <- function(weekly_panel) {
-  new_derogations <- weekly_panel %>% 
+make_derogation_count <- function(data) {
+  new_derogations <- data %>% 
     group_by(iso3) %>% 
     summarize(derogations = sum(iccpr_derogation_filed))
   

### 2023-04-21 _targets.R

@@ -73,7 +73,7 @@ list(
   tar_target(year_week_lookup, make_year_week_lookup(weekly_panel)),
   
   tar_target(world_map, load_world_map(naturalearth_raw_file)),
-  tar_target(derogation_count, make_derogation_count(weekly_panel)),
+  tar_target(derogation_count, make_derogation_count(iccpr_who_clean)),
   tar_target(derogation_map_data, make_derogation_map_data(derogation_count, world_map)),
 
   

### 2023-05-04 manuscript/manuscript.qmd
Audrey edits

@@
 abstract: |
+  This research note asks whether states issuing pandemic-era human rights treaty derogations followed through with the emergency provisions or used them to increase discriminatory policies during a time of crisis.
 In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
-Did states actually end up complying with their pandemic-era derogations?
-Did derogating and non-derogating states implement emergency measures differently, and if so, how?
-Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
+Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states that derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
 On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
 Our results support the role that flexibility mechanisms such as derogations play in international law.
-date: April 20, 2023
+The findings contribute to earlier studies of emergency provisions in international law by applying the COVID-19 context.
+date: May 1, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
-What is the impact of escape clauses or emergency provisions in international law on human rights?
-Combating a deadly virus like SARS-CoV-2 required extraordinary public health measures.
-In an effort to fight the COVID-19 pandemic, governments across the world imposed restrictions on the freedom of movement, association, assembly and privacy.
-While countries were ostensibly free to instate these policies within their own borders, the imposition of these restrictions raised important questions about states' commitments to international human rights law.
-Some human rights treaties recognize states' needs for emergency measures during public health concerns and contain escape clauses that establish safeguards to prevent the permanent erosion of these rights during emergencies.
-The temporary suspension of state compliance with its international treaty obligations occurs through a legal action known as derogation.
+What is the impact of international legal emergency provisions on human rights?
+Combating a deadly virus like COVID-19 required extraordinary public health measures.
+In an effort to fight the COVID-19 pandemic, governments across the world introduced restrictions on the freedom of movement, association, and assembly.
+While countries were legally able to introduce the policies within their borders, the restrictions raised important questions about states’ commitments to international human rights law.
+Some human rights treaties recognize states’ ability to use emergency measures during times of crisis and contain "escape clauses" that safeguard the permanent erosion of human rights during and following emergencies.
+The temporary suspension of state compliance from international treaty obligations occurs through a legal action called derogation.
 However, the potential abuse of emergency measures during derogations raises concerns about their impact on human rights.
 Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice  [@Stasavage2020; @LuhrmannRooney2021].
+Did states use pandemic era derogations how they were intended or were they used as a "pandemic pass" to justify violating human rights?
 
-Given recent medical advancements and dwindling concerns about the pandemic, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023, lifting measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
-With this move, we can assess important descriptive and analytical questions raised regarding states' respect for international law and human rights during this emergency period.
-What did derogations mean for compliance with international law?
-Did derogating and non-derogating states differ in terms of how they implemented emergency measures during COVID-19?
-Were derogating states more likely to exploit emergency provisions and to abuse human rights during COVID-19?
+Given recent medical advancements such as distribution of vaccines and the declining concern about COVID-19 severity, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023.
+Doing so lifts measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
+As such, it is a worthwhile time to assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency period.
+This research note specifically examines what derogating meant for compliance with international law during the COVID-19 pandemic.
+In it, we examine whether states lived up to their expected derogations and investigate the broader concern about whether states used derogations to exploit emergency provisions and to abuse human rights during COVID-19.
 These questions are important for the study of international law, compliance, and human rights during times of crises.
 
-Typically, when international relations scholars write about human rights treaty compliance, they write of positive human rights recognition.
-Compliance is an adherence to the expectations and written standards within the treaties, often described as a "state of conformity or identity between an actor’s behavior and a specified rule" [@RaustialaSlaughter:2002, 539].
-However, there are treaty qualifications that are legal addendums to commitment that alter or even limit compliance expectations.
-When states comply with these legal actions, they are in fact, limiting their obligations to human rights standards.
-In this research note, we focus on one such legal addendum—the treaty derogation.
-If and when states adhere to the written expectations of treaty derogations, they will be suspending human rights recognition during a set time of domestic emergency and crisis.
-If states are complying with the derogation expectations, we should in fact expect a reduction in compliance with the overall human rights treaty during that time.
-We explore how the potential decline in compliance may actually support human rights during the COVID-19 period since emergency mitigation strategies also require certain rules and standards to prevent the abuse of rights.
+When states comply with derogations, they actually limit their obligations to human rights treaties overall.
+If states comply with the derogation conditions, we should in fact expect a reduction in compliance with the overall human rights treaty during that time.
+We explore how the potential decline in compliance may actually be a good indication that states are using treaty design as intended.
 
@@
 @Simmons2009; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations.
-Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of the types of states that are likely to comply with these escape clauses.
+Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding adherence to derogations.
 The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally.
@@
 @Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
-We find that overall, derogating states during the pandemic do not increase abuse and violations.
-This finding speaks to international legal compliance scholarship and supports the idea that states are being sincere about their intentions and not, byinlarge, using actions like derogation to cover abusive behavior.
+We find that overall, derogating states during the pandemic did not increase abuse and violations.
+This finding speaks to international legal compliance scholarship and supports the idea that states are being sincere about their intentions and not, generally, using actions like derogation to cover abusive behavior.
 
@@
 
-Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility in responding to events.
+Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility to respond to events.
 Times of crises can include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
@@
 
-While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power.” [@burchill2005, 97] An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency [@Greene:2020, 36].
-These include rights such as right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
+While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97].
+An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency [@Greene:2020, 36].
+These include rights such as the right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
 Research finds results that support this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
@@
 non-derogable rights status during emergencies.
-This also raises the question of the extent to which this distinction is a valuable one during the pandemic.
+This also raises the question of the extent to which this distinction is valuable during the pandemic.
 Most derogations to the ICCPR have been for internal conflict or natural disaster.
@@
 
-A first question of interest before we examine the potential impact of derogations is the nature of states that are more likely to submit them in the first place.
-Little research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the United Nations (UN) rather than the effects of these derogations.
+A first question of interest is the nature of states that are more likely to submit them in the first place.
+Little research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the UN rather than the effects of these derogations.
 This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss].
 The expectation is that states with strong democratic institutions need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
-Additionally, democratic states understand international law and international legal mechanisms to be more legitimate and recognize their importance more so than other states.
-Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important.
+Additionally, democratic states understand international law and international legal mechanisms to be more legitimate and recognize the importance of them more so than other states.
+Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important to engage with.
+Scholars were split on expectations about COVID-era derogations.
+@Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
+However, @richardson2020 see derogations as a less clear path for states during COVID-19, describing the actions, their conditions, and human rights as a "tangled morass" (p.
+106).
 
@@
 Exceptional circumstances brought on by the global pandemic have led to numerous, extensive derogations to the ICCPR removing state obligations to important rights areas of assembly, movement, and others.
-The ICCPR was adopted by the UN General Assembly on December 19, 1966 and entered into force on March 23, 1976.
+The ICCPR was adopted by the UN General Assembly on December 19, 1966, and entered into force on March 23, 1976.
 It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights.
@@
 As of 2023, there are 173 states party to the ICCPR.
-This means that most, but not all of the UN’s 193 members had the ability to submit derogations during COVID-19.
+This means that most, but not all the UN’s 193 members had the ability to submit derogations during COVID-19.
 Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.
@@
 Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination.
-However, not every country formally derogated—a number of countries implemented emergency measures without formally derogating from ICCPR.
+However, not every country formally derogated—a number of countries implemented emergency measures without formally derogating from the ICCPR.
 Over 100 countries issued emergency declarations while fewer than 30 submitted COVID-19 related derogations [@icnl-covid].
@@
 As such, many states used the pandemic an as excuse to centralize authority and implemented measures that did not meet international legal standards.
-Many states imposed in discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
+Many states imposed discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
 For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
@@
 The pandemic also led to the violation of non-derogable rights, or rights that cannot be suspended in any circumstances, in many states.
-For instance, Greek officials intercepted and turned back boats filled with asylum-seekers and also summarily returned asylum-seekers at the land border with Turkey.
-In doing so, they both violated the non-derogable principle of non-refoulement, as well as Article 7 of ICCPR is the person sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
+For instance, Greek officials intercepted and turned back boats filled with asylum-seekers and summarily returned asylum-seekers at the land border with Turkey.
+In doing so, they both violated the non-derogable principle of non-refoulement.
+Additionally, Article 7 of ICCPR prohibits people being sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
 Many other states changed rules of confinement affecting detainees, which also potentially violated non-derogable rights.
@@
 
-We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions and submit derogations were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to impose measures that violated these criteria.
+We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to impose measures that violated these criteria.
 If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.
@@
 
-While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to more broadly examine the relationship between treaty derogations, human rights, and state actions during the pandemic.
+While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to examine the relationship more broadly between treaty derogations, human rights, and state actions during the pandemic.
 We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
 We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021.
-While the Oxford Government Response Tracker provides daily data on a variety of government actions, submitting formal treaty responses to the ICCPR is less likely to be as immediate of an action as implementing domestic emergency actions like stay at home orders.
+While the Oxford Government Response Tracker provides daily data on a variety of government actions, submitting formal treaty responses to the ICCPR is less likely to be as immediate of an action as implementing domestic emergency actions like stay-at-home orders.
 Accordingly, we collapse this daily data into weekly totals, resulting in a panel with `r info$n_rows` country-week observations.
@@
 First, to test our first set of hypotheses and explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
-Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in a given week.
+Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.
 
@@
 
-We include several additionally variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
+We include several additional variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
 To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
@@
 Accordingly, we model our outcome variables with multilevel or hierarchical models, with region-specific random offsets for each intercept and a fixed global year trend.
-When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, use use ordered logistic regression models.
-We take a Bayesian approach to better model the uncertainty inherent in measuring the effect of derogations on these broad policies.
+When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, we use ordered logistic regression models.
+We take a Bayesian approach to better model and explore the uncertainty inherent in measuring the effect of derogations on these broad policies.
 
@@
 Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in the appendix).
-Instead we rely on probability-scale predictions and marginal effects whenever possible.
+Instead, we rely on probability-scale predictions and marginal effects whenever possible.
 To do so, we calculate predicted probabilities holding all explanatory variables at their typical values (median or mode).
@@
 
-Before exploring whether COVID-induced derogations are associated with specific human rights abuses, we first explore our first set of hypotheses to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
+We begin by exploring our first set of hypotheses to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
 @fig-preds-mfx-policies-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
@@
 
-For instance, all countries were highly likely to cancel public events in March and April 2020, with greater than a 95% chance.
+For instance, all countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance.
 Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021.
@@
 Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations.
-Further, as previous research shows, during high levels of stress and fear, decisionmaking is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
+Further, as previous research shows, during high levels of stress and fear, decision making is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
 As such, these differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations and claim that emergency measures will be implemented proportional to the threat at hand.
@@
 We thus find strong support only for H~2c~—that derogating states are less likely to impose measures with no time limits.
-We find weak support for our hypothesis that derogating states will be less likely to violated non-derogable rights (H~2b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
+We find weak support for our hypothesis that derogating states will be less likely to violate non-derogable rights (H~2b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
 Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
@@
 International legal scholarship has been divided over expectations of derogations during the pandemic.
-@Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
-However, @richardson2020 see derogations as a less clear path for states during COVID-19, describing the actions, their conditions, and human rights as a "tangled morass" (p.
-106).
 
-With the caveat that our results are by no means causal—we have not attempted to use any formal strategy to identify an unconfounded causal effect, and our findings are merely descriptive—we find that we find that there is good news for compliance and derogations.
-Derogators did not systematically use this legal "escape clause" opportunistically—derogators used the process correctly.
+With the caveat that our results are not causal—we have not attempted to use any formal strategy to identify an unconfounded causal effect, and our findings are therefore descriptive—we find that there is good news for compliance and derogations.
+Derogators did not systematically use this legal "escape clause" opportunistically—derogators used the process as intended.
 Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
-Though a majority of states did not submit COVID-19 related derogations, the ones that did used derogations as intended by international law.
+Though most states did not submit COVID-19 related derogations, the ones that did used derogations as intended by international law.
 Once the derogation ended, states readjusted to typical compliance once more.
@@
 We suggest that future research look to different incentives facing state officials during crisis.
-There may be key differences in enforcement between local level bureaucrats and higher level elites.
+There may be key differences in enforcement between local level bureaucrats and higher-level elites.
 While higher level diplomats submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.

### 2023-05-04 manuscript/manuscript.qmd
Suparna edits

@@
 short-title: "Pandemic Pass?"
-subtitle: "Research note"
+subtitle: "(Research note)"
 author:
@@
 abstract: |
-  This research note asks whether states issuing pandemic-era human rights treaty derogations followed through with the emergency provisions or used them to increase discriminatory policies during a time of crisis.
+  This research note asks whether states issuing pandemic-era human rights treaty derogations implemented emergency provisions as intended or used them to abuse human rights during a time of crisis.
 In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
-Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states that derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
+Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states that derogated from their international human rights obligations imposed emergency measures that were temporary and did not violate non-derogable rights.
 On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
-Our results support the role that flexibility mechanisms such as derogations play in international law.
-The findings contribute to earlier studies of emergency provisions in international law by applying the COVID-19 context.
-date: May 1, 2023
+Our results support the role that flexibility mechanisms such as derogations play in international law and show that states are being sincere about their intentions and not, generally, using these mechanisms to cover abusive behavior.
+date: May 2, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 In an effort to fight the COVID-19 pandemic, governments across the world introduced restrictions on the freedom of movement, association, and assembly.
-While countries were legally able to introduce the policies within their borders, the restrictions raised important questions about states’ commitments to international human rights law.
-Some human rights treaties recognize states’ ability to use emergency measures during times of crisis and contain "escape clauses" that safeguard the permanent erosion of human rights during and following emergencies.
+While countries were legally able to introduce these policies within their borders, the restrictions raised important questions about states’ commitments to international human rights law.
+Some human rights treaties recognize states’ ability to use emergency measures during times of crises, and contain "escape clauses" that safeguard the permanent erosion of human rights during and following emergencies.
 The temporary suspension of state compliance from international treaty obligations occurs through a legal action called derogation.
@@
 
-Given recent medical advancements such as distribution of vaccines and the declining concern about COVID-19 severity, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023.
+Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023.
 Doing so lifts measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
 As such, it is a worthwhile time to assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency period.
-This research note specifically examines what derogating meant for compliance with international law during the COVID-19 pandemic.
-In it, we examine whether states lived up to their expected derogations and investigate the broader concern about whether states used derogations to exploit emergency provisions and to abuse human rights during COVID-19.
+This research note examines what derogating meant for compliance with international law during the COVID-19 pandemic.
+Specifically, we examine whether states actually derogated as per international legal principles and investigate the broader concern about whether states used derogations to exploit emergency provisions and to abuse human rights during COVID-19.
 These questions are important for the study of international law, compliance, and human rights during times of crises.
-
-When states comply with derogations, they actually limit their obligations to human rights treaties overall.
-If states comply with the derogation conditions, we should in fact expect a reduction in compliance with the overall human rights treaty during that time.
-We explore how the potential decline in compliance may actually be a good indication that states are using treaty design as intended.
 
@@
 We then analyze the effect of derogating (or not) on human rights practices during the pandemic.
-We find that states that derogated and specified time limits to the emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights.
-On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely to impose discriminatory policies, violate non-derogable rights (the right to life, freedom from torture, and freedom from slavery), and impose emergency measures without time limits.
+We find that states that derogated and specified time limits to these emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights (the right to life, freedom from torture, and freedom from slavery).
+On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely to impose discriminatory policies, violate non-derogable rights, and impose emergency measures without time limits.
 
@@
 @Simmons2009; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations.
-Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding adherence to derogations.
-The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally.
+Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of adherence to derogations.
+When states comply with derogations, they are actually limiting their obligations to the human rights treaty overall.
+As such, we should in fact expect a reduction in compliance with the overall human rights treaty during that time.
+We explore how the potential decline in compliance may actually be a good indication that states are using treaty design as intended.
+
+The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster, typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally.
 In making this comparison, this research note makes an important descriptive contribution showing how the implementation of policy measures differed across derogating and non-derogating states during the pandemic.
@@
 While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
-@Lebret:2020; @helfer2021] as part of a broader trend of targeting groups during times of crises [@Grasse2021].
+@Lebret:2020; @helfer2021], as part of a broader trend of targeting groups during times of crises [@Grasse2021].
 We find that overall, derogating states during the pandemic did not increase abuse and violations.
@@
 
-While legal critical assessments of derogations find importance in their use [@richardson2020], others argue that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97].
+Many legal critical assessments of derogations find importance in their use [@richardson2020], with some arguing that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97].
 An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency [@Greene:2020, 36].
 These include rights such as the right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
-Research finds results that support this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
+Research supports this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
 Even states that officially derogate from their international legal obligations are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].
@@
 
-Between January 2020 and June 2021, `r derog_detail$countries` states issued a total of `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
+We find that between January 2020 and June 2021, `r derog_detail$countries` states issued a total of `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
 Of derogations that detailed specific obligation suspension lengths, the average length was 41 days.
-Descriptively, we see states across multiple regions derogated with notable concentration in South America, which has a robust history of engaging with human rights law.
+Descriptively, we see states across multiple regions derogate, with notable concentration in South America, which has a robust history of engaging with human rights law.
 We also note that it is not necessarily the "usual suspects" of strong democracies that derogate.
@@
 It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights.
-The ICCPR is one of only a few select treaties that requires states to report on the suspension or derogation of certain civil and political rights only under specific situations of "public emergency which threatens the life of the nation" [@iccpr].
+The ICCPR is one of only a few select human rights treaties that require states to report on the suspension or derogation of certain civil and political rights only under specific situations of "public emergency which threatens the life of the nation" [@iccpr].
 The other treaties with this requirement include the European Convention Protection of Human Rights and Fundamental Freedoms and the American Convention on Human Rights.
 However, the ICCPR is the only one with global, rather than regional, membership and participation to track during the pandemic.
-As of 2023, there are 173 states party to the ICCPR.
+As of 2023, there are 173 states that are party to the ICCPR.
 This means that most, but not all the UN’s 193 members had the ability to submit derogations during COVID-19.
@@
 
-To legally derogate, government must first officially declare a state of emergency (Article 4).
+To legally derogate, governments must first officially declare a state of emergency (Article 4).
 Derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory.
@@
 The international legal community viewed these rights as so fundamental and important that no type of crisis could justify their suspension.
-Other types of rights are seen as able to be suspended during crises.
-Several ICCPR Articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
+However, other rights can be suspended during crises.
+Several ICCPR articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
 
@@
 Our expectation comes from the specific, time-limited nature of derogations and the limited audience they speak to.
-Critics of international law may expect states ratify treaties with the expectation of reputational and other benefits.
+Critics of international law argue that states ratify treaties with the expectation of reputational and other benefits.
 However, derogations have no such public audience and are not signaling an increased commitment to human rights.
@@
 >
-> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose related emergency policies.
+> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose relevant emergency policies.
 
@@
 If states determine that emergency measures are justified without regard for international law, the processes for respecting rights during crises and restoring rights at the end of the emergency may not be followed.
-As such, many states used the pandemic an as excuse to centralize authority and implemented measures that did not meet international legal standards.
-Many states imposed discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
+As such, many states used the pandemic as an excuse to centralize authority and implemented measures that did not meet international legal standards.
+Many government engaged in discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
 For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
@@
 
-The pandemic also led to the violation of non-derogable rights, or rights that cannot be suspended in any circumstances, in many states.
+The pandemic also led to the violation of non-derogable rights in many states.
 For instance, Greek officials intercepted and turned back boats filled with asylum-seekers and summarily returned asylum-seekers at the land border with Turkey.
-In doing so, they both violated the non-derogable principle of non-refoulement.
-Additionally, Article 7 of ICCPR prohibits people being sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
+In doing so, they violated the non-derogable principle of non-refoulement—Article 7 of ICCPR prohibits people being sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020].
 Many other states changed rules of confinement affecting detainees, which also potentially violated non-derogable rights.
@@
 
-While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we examine a global database to examine the relationship more broadly between treaty derogations, human rights, and state actions during the pandemic.
+While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we use global data to examine the relationship more broadly between treaty derogations, human rights, and state actions during the pandemic.
 We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
@@
 To examine this first step, we code two binary variables: (1) past ICCPR derogation behavior before 2020 and (2) past ICCPR post-commitment action behavior before 2020.
-This allows us to empirically test for the first step of engagement before analyzing any impact COVID-19 era derogations had on different rights practices.
+This allows us to empirically test for the first step of engagement before analyzing any impact that COVID-19-era derogations had on different rights practices.
 There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019.
@@
 While elites or diplomats are the ones submitting derogations and committing to complying with international law, they are not the actors implementing emergency measures.
-During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially when it relates to human rights.
+During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially as it relates to human rights.
 Lower-level officials such as those in the police, state security forces and regional bureaucrats are often the ones enforcing emergency measures.
@@
 Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.
-We focused specifically on the impact of derogations—an international legal means to communicate to both the UN and global community that certain rights will be suspended during times of crisis.
+We focused specifically on the impact of derogations—an international legal means to communicate to both the UN and the global community that certain rights will be suspended during times of crisis.
 Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties without concern about the inability to adapt to times of emergency.
@@
 Once the derogation ended, states readjusted to typical compliance once more.
-States, therefore complied with the intent of derogations and the purpose of the treaty.
+States therefore complied with the intent of derogations and the purpose of the treaty.
 This is an important finding for international legal behavior and compliance studies.
@@
 
-We suggest that future research look to different incentives facing state officials during crisis.
+We suggest that future research examine different incentives facing state officials during crisis.
 There may be key differences in enforcement between local level bureaucrats and higher-level elites.
-While higher level diplomats submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.
+While elites submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.
 non-derogable rights.

### 2023-05-04 analysis/analysis_human-rights.qmd
remove ordinal bc we're bayesians

@@
 library(kableExtra)
-library(ordinal)
 library(brms)

### 2023-05-04 _targets_packages.R
renvification

@@ -1,0 +1,14 @@
+# Generated by targets::tar_renv(): do not edit by hand
+library(bs4Dash)
+library(clustermq)
+library(future)
+library(gt)
+library(markdown)
+library(pingr)
+library(qs)
+library(rstudioapi)
+library(shiny)
+library(shinybusy)
+library(shinyWidgets)
+library(tidyverse)
+library(visNetwork)

### 2023-05-04 _dependencies.R
Add some invisible packages

@@ -1,0 +1,4 @@
+# Things that need to be installed but didn't get installed automatically as 
+# dependencies for whatever reason
+library(quarto)
+library(collapse)

### 2023-05-05 analysis/environment.qmd
Add computing environment

---
title: "Computing environment"
---

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
```

## Platform

Here are the details of the R environment we used when running this analysis:

```{r get-session-info}
#| class-output: text
# Injecting Stan info into the platform info adapted from TJ Mahr:
# https://github.com/tjmahr/tjmahr.github.io/commit/b8be089
session_info <- sessioninfo::session_info()
session_info$platform$`stan (cmdstanr)` <- cmdstanr::cmdstan_version()

session_info$platform
```

## Packages

We use the [{renv} package](https://rstudio.github.io/renv/) to create a stable version-specific library of packages. The official list of packages is included in [`renv.lock`](https://github.com/andrewheiss/mountainous-mackerel/blob/main/renv.lock) and running this from the root of this project should install everything:

```r
renv::restore()
```

Here's the full list of packages in a less JSON-y format:

```{r get-package-info}
#| class-output: text
packages_raw <- jsonlite::read_json(here::here("renv.lock"))$Packages %>% 
  enframe()

packages_in_project <- sessioninfo::package_info(packages_raw$name) %>% 
  as_tibble()

packages_in_project %>%
  select(package, ondiskversion, date, source) %>% 
  print(n = Inf)
```

## Makevars

Because {brms} generates Stan code that is then compiled and run as binary executable programs, knowing about the compilation environment is helpful. Here are all the things that influence compiling (different paths, flags, external libraries, etc.) that are included in `~/.R/Makevars`:

```{r}
#| class-output: sh
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))
```

### 2023-05-05 analysis/targets.qmd
Minor formatting change
Add computing environment

@@
 ---
-title: "Targets workflow"
+title: "Targets pipeline"
 format:
@@
 
-We use [the magical **`targets`** package](https://docs.ropensci.org/targets/) to run our analysis and keep track of all dependencies automatically.
+We use [the magical **targets** package](https://docs.ropensci.org/targets/) to run our analysis and keep track of all dependencies automatically.
 

### 2023-05-05 analysis/data-final.qmd
Add targets as dependencies

@@
 ---
+
+```{r include=FALSE}
+# Add these final data files as targets dependencies
+library(targets)
+tar_load(c(data_daily_stata, data_daily_csv, data_daily_rds,
+           data_weekly_stata, data_weekly_csv, data_weekly_rds))
+tar_load(c(data_daily_stata_website, data_daily_csv_website, data_daily_rds_website,
+           data_weekly_stata_website, data_weekly_csv_website, data_weekly_rds_website))
+```
 

### 2023-05-05 analysis/data-details.qmd
Use {gt} instead of {DT}

--- 2023-02-15 analysis/data-details.qmd
@@
 library(targets)
-library(DT)
+library(sf)
+library(gt)
 library(kableExtra)
@@
 
+clrs_light <- MetBrewer::met.brewer("Tiepolo") %>% colorspace::lighten(0.8)
+
 # Generated via random.org
@@
 ```{r show-iccpr}
+#| column: page-right
 daily_panel %>%
@@
          derogation_start, derogation_ineffect, derogation_end) %>% 
-  datatable()
+  gt() %>% 
+  cols_width(country_name ~ px(250),
+             iso3 ~ px(60)) %>% 
+  data_color(columns = contains("derogation"),
+             fn = scales::col_factor(c(clrs_light[1], clrs_light[8]), 
+                                     domain = c(FALSE, TRUE))) %>% 
+  opt_interactive(use_compact_mode = TRUE, use_highlight = TRUE)
+```
+
+And here's a map of all the countries that derogated between January 2020 and June 2021:
+
+```{r derogations-map, fig.width=6.5, fig.height=3.5}
+ggplot() +
+  geom_sf(data = derogation_map_data, aes(fill = derogations_1plus),
+          linewidth = 0.1, color = "white") +
+  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
+  scale_fill_binned(low = colorspace::lighten(clrs[1], amount = 0.8), 
+                    high = clrs[1], na.value = "grey85",
+                    breaks = c(1, 5, 10, 15, 20),
+                    limits = c(1, 20),
+                    name = "Derogations filed: ",
+                    guide = guide_colorsteps(barwidth = 7, barheight = 0.4,
+                                             title.vjust = 1)) +
+  theme_pandem() +
+  theme(panel.border = element_blank(),
+        legend.title = element_text(face = "bold"),
+        legend.justification = "center")
 ```
@@
 ```{r show-oxford}
+#| column: page-right
 daily_panel %>%
@@
          c8_intl_travel, e1_income_support, e2_debt_relief) %>% 
-  datatable()
+  gt() %>% 
+  cols_width(country_name ~ px(250),
+             iso3 ~ px(60),
+             day ~ px(150)) %>% 
+  opt_interactive(use_compact_mode = TRUE, use_highlight = TRUE)
 ```

### 2023-05-05 README.qmd
Update and quartoify and targetify README

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

# Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

<!-- badges: start -->
<!-- badges: end -->

[Suparna Chaudhry](https://www.suparnachaudhry.com/) • Lewis and Clark College  
[Audrey Comstock](https://audreylcomstock.weebly.com/) • Arizona State University  
[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

---

## Abstract

This research note asks whether states issuing pandemic-era human rights treaty derogations implemented emergency provisions as intended or used them to abuse human rights during a time of crisis. In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations. Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states that derogated from their international human rights obligations imposed emergency measures that were temporary and did not violate non-derogable rights. On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights. Our results support the role that flexibility mechanisms such as derogations play in international law and show that states are being sincere about their intentions and not, generally, using these mechanisms to cover abusive behavior.

---

## How to download and replicate

You can either [download the compendium as a ZIP file](/archive/main.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

We use the [**renv** package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [**targets** package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis. ([See this for a short helpful walkthrough of **targets**.](https://books.ropensci.org/targets/walkthrough.html)).

To reproduce the findings and re-run the analysis, do the following:

1. Download and install these fonts (if you’re using Windows, make sure you right click on the font files and choose “Install for all users” when installing these fonts):
    - [Noto Sans](https://fonts.google.com/specimen/Noto+Sans)
    - [Linux Libertine O](https://www.cufonfonts.com/font/linux-libertine-o) (also [here](https://sourceforge.net/projects/linuxlibertine/))
    - [Libertinus Math](https://github.com/alerque/libertinus)
2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library
    - If you’re using Windows, [install RTools too](https://cran.r-project.org/bin/windows/Rtools/) and add it to your PATH so that you can install packages from source if needed
3. Open `mountainous-mackerel.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
4. In the terminal, run `quarto install tinytex` to ensure that you have a working LaTeX installation.
5. If it's not installed already, R *should* try to install the **renv** package when you open the RStudio Project for the first time. If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
6. Run `renv::restore()` in the R console to install all the required packages for this project.
7. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.

Running `targets::tar_make()` will create several helpful outputs:

1. All project data in `data/`
2. An analysis notebook website in `_site/`
3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/` and `manuscript/output-ms`

## `r emoji_glue(":snow_capped_mountain::fish:")`: Note on "mountainous mackerel" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [**codename**](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.4.0

codename(seed = "covid and ngos", type = "ubuntu")
#> [1] "mountainous mackerel"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

### 2023-05-05 _targets.R

@@ -147,5 +147,8 @@ list(
     website
     # Run the deploy script
     if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
-  })
+  }),
+  
+  ## Render the README ----
+  tar_quarto(readme, here_rel("README.qmd"))
 )

### 2023-05-05 R/funs_plots.R
Add p-direction to plot data

@@ -75,7 +75,10 @@ build_policies_plot_data <- function(model_df, year_week_lookup) {
         posteriordraws() %>% 
         left_join(year_week_lookup, by = "year_week_num") %>% 
         group_by(year_week_day) %>% 
-        tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
+        reframe(post_medians = tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95)),
+                p_gt_0 = sum(draw > 0) / n()) %>% 
+        unnest(post_medians) %>% 
+        rename(draw = y, .lower = ymin, .upper = ymax)
     }))
   
   return(model_df_preds_mfx)
@@ -123,11 +126,17 @@ build_human_rights_plot_data <- function(model_df, year_week_lookup) {
         mfx <- mfx %>% 
           mutate(group = factor(group, levels = levels(.model$data[[.y]]), ordered = TRUE)) %>% 
           group_by(year_week_day, group) %>% 
-          tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
+          reframe(post_medians = tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95)),
+                  p_gt_0 = sum(draw > 0) / n()) %>% 
+          unnest(post_medians) %>% 
+          rename(draw = y, .lower = ymin, .upper = ymax)
       } else {
         mfx <- mfx %>% 
           group_by(year_week_day) %>% 
-          tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95))
+          reframe(post_medians = tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95)),
+                  p_gt_0 = sum(draw > 0) / n()) %>% 
+          unnest(post_medians) %>% 
+          rename(draw = y, .lower = ymin, .upper = ymax)
       }
       
       mfx

### 2023-05-05 manuscript/manuscript.qmd
woohoo dynamic in-text numbers

@@
   select(Country, Derogations = derogations)
+
+fmt_p_inline <- function(x, direction) {
+  x <- round(x, 2)
+
+  if (direction == "gt") {
+    out <- glue::glue(r"[$p(\Delta > 0) = {x}$]")
+  } else {
+    out <- glue::glue(r"[$p(\Delta < 0) = {x}$]")
+  }
+
+  return(out)
+}
+
+policies_pred <- policies_plot_data %>% 
+  unnest(pred_data) %>% 
+  filter(.width == 0.95) %>% 
+  select(nice, year_week_day, derogation_ineffect, .epred, .lower, .upper) %>% 
+  group_by(nice, derogation_ineffect) %>% 
+  mutate(rank = dense_rank(.epred)) %>% 
+  filter(rank == 1 | rank == max(rank)) %>% 
+  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
+  group_by(rank, derogation_ineffect) %>% 
+  mutate(outcome = janitor::make_clean_names(nice)) %>%
+  rename(draw = .epred, min = .lower, max = .upper) %>% 
+  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0)))) %>% 
+  split(~outcome + derogation_ineffect + rank)
+
+policies_mfx <- policies_plot_data %>% 
+  unnest(mfx_data) %>% 
+  filter(.width == 0.95) %>% 
+  select(nice, year_week_day, draw, .lower, .upper, p_gt_0, .width) %>% 
+  group_by(nice) %>% 
+  mutate(rank = dense_rank(draw)) %>% 
+  filter(rank == 1 | rank == max(rank)) %>% 
+  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
+  group_by(rank) %>% 
+  mutate(outcome = janitor::make_clean_names(nice)) %>%
+  rename(min = .lower, max = .upper) %>% 
+  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0))),
+         p_lt_0 = p_gt_0 - 1) %>% 
+  mutate(p_gt = fmt_p_inline(p_gt_0, "gt"),
+         p_lt = fmt_p_inline(p_lt_0, "lt")) %>% 
+  split(~outcome + rank)
 ```
@@
 Instead, we rely on probability-scale predictions and marginal effects whenever possible.
-To do so, we calculate predicted probabilities holding all explanatory variables at their typical values (median or mode).
+To do so, we calculate conditional predicted probabilities holding all explanatory variables at their typical values (median or mode), and setting all region-specific random offsets ($b_{0_j}$) to 0, meaning that the predicted values refer to the effect of derogations in a typical region.
 For logistic models, we estimate the probability of having a specific policy or violating a specific human right; for ordered models, we estimate the probability of each possible level of severity.
@@
 Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability.
-The difference in predicted probabilities between the two types of countries ranged from 3 to 8 percentage points ($p(\Delta > 0) = 1$).
+The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice` to `r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`).
 The differences in policy implementation are starker for other types of emergency measures.
-Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between 80% and 90%, representing a 15–18 percentage point difference ($p(\Delta > 0) = 1$).
-Derogating states were also far more likely to close public transit, beginning with a roughly 80% probability in March 2020 and ending in June 2021 with a 60% probability.
-Non-derogating states followed a similar downward trend over time, but 23–25 percentage points lower than their derogating counterparts ($p(\Delta > 0) = 1$).
-Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to roughly 70% by June 2021.
-Non-derogating states, on the other hand, began the pandemic with a predicted 75% probability of limiting movement, dropping to only 25% by June 2021, representing a 20–40 percentage point difference ($p(\Delta > 0) = 1$).
+Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between `r policies_pred$gathering_restrictions.No.min$draw_nice`% and `r policies_pred$gathering_restrictions.No.max$draw_nice`%, representing a `r policies_mfx$gathering_restrictions.min$draw_nice`–`r policies_mfx$gathering_restrictions.max$draw_nice` percentage point difference (`r policies_mfx$gathering_restrictions.max$p_gt`).
+Derogating states were also far more likely to close public transit, beginning with a `r policies_pred$close_public_transit.Yes.max$draw_nice`% probability in March 2020 and ending in June 2021 with a `r policies_pred$close_public_transit.Yes.min$draw_nice`% probability.
+Non-derogating states followed a similar downward trend over time, but `r policies_mfx$close_public_transit.min$draw_nice`–`r policies_mfx$close_public_transit.max$draw_nice` percentage points lower than their derogating counterparts (`r policies_mfx$close_public_transit.max$p_gt`).
+Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to `r policies_pred$movement.Yes.min$draw_nice`% by June 2021.
+Non-derogating states, on the other hand, began the pandemic with a predicted `r policies_pred$movement.No.max$draw_nice`% probability of limiting movement, dropping to only `r policies_pred$movement.No.min$draw_nice`% by June 2021, representing a `r policies_mfx$movement.min$draw_nice`–`r policies_mfx$movement.max$draw_nice` percentage point difference (`r policies_mfx$movement.max$p_gt`).
 The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel.

### 2023-05-05 manuscript/manuscript.qmd
Add more about priors
Create table of summarized predicted probabilities as target

@@
 library(colorspace)
-library(modelsummary)
 library(kableExtra)
@@
 tar_load(c(policies_plot_data, human_rights_plot_data))
-tar_load(c(models_tbl_policies, models_tbl_human_rights))
 tar_load(c(derogation_map_data, derogation_count))
@@
 
-fmt_p_inline <- function(x, direction) {
-  x <- round(x, 2)
-
-  if (direction == "gt") {
-    out <- glue::glue(r"[$p(\Delta > 0) = {x}$]")
-  } else {
-    out <- glue::glue(r"[$p(\Delta < 0) = {x}$]")
-  }
-
-  return(out)
-}
-
-policies_pred <- policies_plot_data %>% 
-  unnest(pred_data) %>% 
-  filter(.width == 0.95) %>% 
-  select(nice, year_week_day, derogation_ineffect, .epred, .lower, .upper) %>% 
-  group_by(nice, derogation_ineffect) %>% 
-  mutate(rank = dense_rank(.epred)) %>% 
-  filter(rank == 1 | rank == max(rank)) %>% 
-  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
-  group_by(rank, derogation_ineffect) %>% 
-  mutate(outcome = janitor::make_clean_names(nice)) %>%
-  rename(draw = .epred, min = .lower, max = .upper) %>% 
-  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0)))) %>% 
+policies_pred <- policies_table_data$policies_pred_tbl %>% 
   split(~outcome + derogation_ineffect + rank)
 
-policies_mfx <- policies_plot_data %>% 
-  unnest(mfx_data) %>% 
-  filter(.width == 0.95) %>% 
-  select(nice, year_week_day, draw, .lower, .upper, p_gt_0, .width) %>% 
-  group_by(nice) %>% 
-  mutate(rank = dense_rank(draw)) %>% 
-  filter(rank == 1 | rank == max(rank)) %>% 
-  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
-  group_by(rank) %>% 
-  mutate(outcome = janitor::make_clean_names(nice)) %>%
-  rename(min = .lower, max = .upper) %>% 
-  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0))),
-         p_lt_0 = p_gt_0 - 1) %>% 
-  mutate(p_gt = fmt_p_inline(p_gt_0, "gt"),
-         p_lt = fmt_p_inline(p_lt_0, "lt")) %>% 
+policies_mfx <- policies_table_data$policies_mfx_tbl %>% 
   split(~outcome + rank)
@@
 In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society.
+We use Student *t* priors clustered around 0 with a scale parameter of 3, since (1) these model coefficients are measured on the logged odds scale and rarely result in large numbers, and (2) the Student *t* distribution has heavier tails than a standard normal distribution.
 
@@
 Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability.
-The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice` to `r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`).
+The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice`–`r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`) between March 2020 and June 2021.
 The differences in policy implementation are starker for other types of emergency measures.

### 2023-05-05 R/funs_plots.R

@@ -29,6 +29,18 @@ plot_pp <- function(model) {
     theme_pandem()
 }
 
+fmt_p_inline <- function(x, direction) {
+  x <- round(x, 2)
+  
+  if (direction == "gt") {
+    out <- glue::glue(r"[$p(\Delta > 0) = {x}$]")
+  } else {
+    out <- glue::glue(r"[$p(\Delta < 0) = {x}$]")
+  }
+  
+  return(out)
+}
+
 
 # Storing ggplot objects as rds files is BAD 
 #   (https://github.com/hadley/ggplot2-book/issues/344)
@@ -143,4 +155,39 @@ build_human_rights_plot_data <- function(model_df, year_week_lookup) {
     }))
   
   return(model_df_preds_mfx)
+}
+
+build_policies_table_data <- function(model_df) {
+  policies_pred_tbl <- model_df %>% 
+    unnest(pred_data) %>% 
+    filter(.width == 0.95) %>% 
+    select(nice, year_week_day, derogation_ineffect, .epred, .lower, .upper) %>% 
+    group_by(nice, derogation_ineffect) %>% 
+    mutate(rank = dense_rank(.epred)) %>% 
+    filter(rank == 1 | rank == max(rank)) %>% 
+    mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
+    group_by(rank, derogation_ineffect) %>% 
+    mutate(outcome = janitor::make_clean_names(nice)) %>%
+    rename(draw = .epred, min = .lower, max = .upper) %>% 
+    mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0)))) %>% 
+    ungroup()
+
+  policies_mfx_tbl <- model_df %>% 
+    unnest(mfx_data) %>% 
+    filter(.width == 0.95) %>% 
+    select(nice, year_week_day, draw, .lower, .upper, p_gt_0, .width) %>% 
+    group_by(nice) %>% 
+    mutate(rank = dense_rank(draw)) %>% 
+    filter(rank == 1 | rank == max(rank)) %>% 
+    mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
+    group_by(rank) %>% 
+    mutate(outcome = janitor::make_clean_names(nice)) %>%
+    rename(min = .lower, max = .upper) %>% 
+    mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0))),
[…]

### 2023-05-05 _targets.R

@@ -132,8 +132,9 @@ list(
   tar_target(graphic_functions, lst(theme_pandem, set_annotation_fonts, clrs)),
   tar_target(diagnostic_functions, lst(plot_trace, plot_trank, plot_pp)),
   
-  ## Plots ----
+  ## Plots and tables ----
   tar_target(policies_plot_data, build_policies_plot_data(m_policies, year_week_lookup)),
+  tar_target(policies_table_data, build_policies_table_data(policies_plot_data)),
   tar_target(human_rights_plot_data, build_human_rights_plot_data(m_human_rights, year_week_lookup)),
 
   ## Manuscript and analysis notebook ----

### 2023-05-05 manuscript/appendix.qmd
Add appendix

---
title: 'Online appendix for "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"'
short-title: 'Online appendix for "Pandemic Pass?"'
subtitle: "(Research note)"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  orcid: 0000-0003-3506-7590
  url: https://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Audrey L. Comstock
  affiliation: Arizona State University
  orcid: 0000-0003-1772-5350
  url: https://audreylcomstock.weebly.com/
  email: audrey.comstock@asu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  orcid: 0000-0002-3948-3914
  url: https://www.andrewheiss.com
  email: aheiss@gsu.edu
date: May 2, 2023
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
bibliography: bib.bib
link-citations: true
crossref: 
  tbl-title: Appendix Table
  fig-title: Appendix Figure
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
                      warning = FALSE, message = FALSE)

knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE)

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r libraries-data}
library(tidyverse)
library(modelsummary)
library(scales)
library(kableExtra)
library(gt)
library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Generated via random.org
set.seed(8511)

tar_load(c(policies_table_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))

# modelsummary functions
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))

add_breaks <- function(x, fmt) {
  if (fmt == "latex") {
    kableExtra::linebreak(x, align = "c")
  } else (
    str_replace(x, "\\n", "<br>")
  )
}
```

# Modeling approach

We use Stan `r cmdstanr::cmdstan_version()` [@stan] through R `r paste0(version$major, ".", version$minor)` [@r-project] and {brms} `r packageVersion("brms")` [@brms-jss] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/MTR6X>. Our general modeling approach can be summarized as follows:

## Model definitions

### Logistic regression

$$
\begin{aligned}
&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
\end{aligned}
$$

### Ordered logistic regression

$$
\begin{aligned}
&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
&\ \textbf{Models for distribution parameters} \\
\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
\alpha_k \sim&\ \mathcal{N}(0, 1)
\end{aligned}
$$

# H~1~: Derogations and emergency policies

## Predicted probabilities and group contrasts

```{r tbl-policies-summary}
#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries
#| include: true
#| echo: false
tbl_part1 <- policies_table_data$policies_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, time, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num")

tbl_part2 <- policies_table_data$policies_mfx_tbl %>% 
  mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, year_week_day, diff_type, num, p_gt_0) %>% 
  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_week_day")) %>% 
  mutate(derogation_ineffect = "No")

tbl_part1 %>% 
  left_join(tbl_part2, by = join_by(nice, derogation_ineffect)) %>% 
  mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
  select(derogation_ineffect, `March 2020`, `June 2021`,
         `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
         `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`) %>% 
  kbl(col.names = c("Derogation", "March 2020", "June 2021",
                    "Week", "Difference", "p > 0", "Week", "Difference", "p > 0"),
      align = "lcccccccc",
      booktabs = TRUE, linesep = "",
      escape = FALSE) %>% 
  kable_styling(font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
  column_spec(3, border_right = T) %>% 
  pack_rows(index = table(fct_inorder(tbl_part1$nice))) %>% 
  add_header_above(c(" " = 1, "Predicted probabilities" = 2, 
                     "Smallest difference" = 3, "Largest difference" = 3))
```

## Complete model results

```{r tbl-results-full-policies}
#| tbl-cap: Results from models showing relationship between derogations and emergency policies
#| include: true
#| echo: false
notes <- c(
  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
  "95% credible intervals (highest density posterior interval, or HDPI) in brackets.",
  "Total \\(R^2\\) considers the variance of both population and group effects;",
  "marginal \\(R^2\\) only takes population effects into account."
)

if (fmt_out == "latex") {
  notes <- notes %>% 
    str_replace_all(r"[\\]", r"[\\\\]") %>% 
    str_replace_all("%", r"(\\\\%)")
} else {
  notes <- paste(notes, collapse = " ")
}

if (fmt_out == "latex") { 
  modelsummary(models_tbl_policies,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "kableExtra",
               fmt = fmt_significant(2),
               escape = FALSE) %>% 
    kable_styling(htmltable_class = "table-sm light-border",
                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
    column_spec(1:6, width = "0.16\\\\textwidth")
} else {
  modelsummary(models_tbl_policies,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "gt",
               fmt = fmt_significant(2)) %>% 
    tab_footnote(footnote = notes) %>% 
    tab_style(style = cell_text(weight = "bold"),
              locations = cells_column_labels())
}
```

# H~2~: Derogations and human rights

## Predicted probabilities and group contrasts

## Complete model results

```{r tbl-results-full-human-rights}
#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~1~)
#| include: true
#| echo: false
if (fmt_out == "latex") {
  modelsummary(models_tbl_human_rights,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "kableExtra",
               fmt = fmt_significant(2),
               escape = FALSE) %>% 
    kable_styling(htmltable_class = "table-sm light-border",
                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
    column_spec(1:5, width = "0.18\\\\textwidth")
} else {
  modelsummary(models_tbl_human_rights,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "gt",
               fmt = fmt_significant(2),
               escape = FALSE) %>% 
    tab_footnote(footnote = notes) %>% 
    tab_style(style = cell_text(weight = "bold"),
              locations = cells_column_labels())
}
```

# References

### 2023-05-05 manuscript/manuscript.qmd
Add appendix

@@
 
-Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in the appendix).
+Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in Appendix Tables 2 and 4).
 Instead, we rely on probability-scale predictions and marginal effects whenever possible.
@@
 @fig-preds-mfx-policies-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
-In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states).
+In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
 

### 2023-05-05 analysis/environment.qmd
Add quarto version

@@
 session_info <- sessioninfo::session_info()
+session_info$platform$`quarto` <- as.character(quarto::quarto_version())
 session_info$platform$`stan (cmdstanr)` <- cmdstanr::cmdstan_version()

### 2023-05-05 manuscript/appendix.qmd
Add actual OSF DOI

@@
 
-Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/MTR6X>.
+Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/AQVNK>.
 Our general modeling approach can be summarized as follows:

### 2023-05-05 manuscript/appendix.qmd
Anonymize link

@@
 
-Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/AQVNK>.
+Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW>.
 Our general modeling approach can be summarized as follows:

### 2023-05-05 manuscript/appendix.qmd
Add human rights models summaries + tables + inline stats

@@
 
-tar_load(c(policies_table_data))
+tar_load(c(policies_table_data, hr_table_data))
 tar_load(c(models_tbl_policies, models_tbl_human_rights))
@@
 ```{r tbl-policies-summary}
-#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries
+#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for emergency policy models
 #| include: true
@@
 
+```{r tbl-hr-summary}
+#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models
+#| include: true
+#| echo: false
+tbl_part1 <- hr_table_data$hr_pred_tbl %>% 
+  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
+  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
+  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
+  select(nice, time, group = category, derogation_ineffect, num) %>% 
+  pivot_wider(names_from = "time", values_from = "num") %>% 
+  mutate(group = recode(group, "NA" = "—"))
+
+tbl_part2 <- hr_table_data$hr_mfx_tbl %>% 
+  mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
+  mutate(group = recode(group, "Logit" = "—")) %>% 
+  mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
+  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
+  select(nice, year_week_day, group, diff_type, num, p_gt_0) %>% 
+  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_week_day")) %>% 
+  mutate(derogation_ineffect = "No")
+
+tbl_part1 %>% 
+  left_join(tbl_part2, by = join_by(nice, derogation_ineffect, group)) %>% 
+  mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
+  select(derogation_ineffect, group, `March 2020`, `June 2021`,
+         `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
+         `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`) %>% 
+  kbl(col.names = c("Derogation", "Level", "March 2020", "June 2021",
+                    "Week", "Difference", "p > 0", "Week", "Difference", "p > 0"),
+      align = "lcccccccc",
+      booktabs = TRUE, linesep = "",
+      escape = FALSE) %>% 
+  kable_styling(font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
+  column_spec(4, border_right = T) %>% 
+  pack_rows(index = table(fct_inorder(tbl_part1$nice))) %>%
+  add_header_above(c(" " = 2, "Predicted probabilities" = 2, 
+                     "Smallest difference" = 3, "Largest difference" = 3))
+```
+
 ## Complete model results

### 2023-05-05 manuscript/manuscript.qmd
oops load this object
Add human rights models summaries + tables + inline stats

@@
 tar_load(c(policies_plot_data, human_rights_plot_data))
+tar_load(c(policies_table_data, hr_table_data))
 tar_load(c(derogation_map_data, derogation_count))
@@
 
+hr_pred <- hr_table_data$hr_pred_tbl %>% 
+  split(~outcome + derogation_ineffect + category + rank, drop = TRUE)
+
 policies_mfx <- policies_table_data$policies_mfx_tbl %>% 
   split(~outcome + rank)
+
+hr_mfx <- hr_table_data$hr_mfx_tbl %>% 
+  split(~outcome + group + rank, drop = TRUE)
 ```
@@
 There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement.
-Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an 85% predicted probability of no discriminatory policy in March 2020 rising to nearly 100% by June 2021.
+Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an `r hr_pred$discriminatory_policy.No.None.min$draw_nice`% predicted probability of no discriminatory policy in March 2020 rising to `r hr_pred$discriminatory_policy.No.None.max$draw_nice`% by June 2021.
 The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.
@@
 In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status.
-For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability).
+For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a `r hr_pred$no_time_limit_measures.No.None.max$draw_nice`% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with `r hr_pred$no_time_limit_measures.No.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.No.Moderate.max$draw_nice`% probability).
 The middle ground of only minor issues with time limits is exceptionally unlikely.
-Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues.
+Countries that formally derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only an `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues.
 The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
-Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)
+Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights (`r hr_mfx$non_derogable_rights.Logit.max$p_lt`)
 

### 2023-05-05 R/funs_plots.R

@@ -148,7 +148,8 @@ build_human_rights_plot_data <- function(model_df, year_week_lookup) {
           reframe(post_medians = tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95)),
                   p_gt_0 = sum(draw > 0) / n()) %>% 
           unnest(post_medians) %>% 
-          rename(draw = y, .lower = ymin, .upper = ymax)
+          rename(draw = y, .lower = ymin, .upper = ymax) %>% 
+          mutate(group = "Logit")
       }
       
       mfx
@@ -190,4 +191,51 @@ build_policies_table_data <- function(model_df) {
     ungroup()
   
   return(lst(policies_pred_tbl, policies_mfx_tbl))
+}
+
+build_hr_table_data <- function(model_df) {
+  hr_pred_tbl <- model_df %>% 
+    mutate(pred_data = map(pred_data, ~{
+      .x %>% 
+        mutate(derogation_ineffect = as.character(derogation_ineffect),
+               derogation_ineffect = recode(derogation_ineffect,
+                                            "No derogation" = "No",
+                                            "Derogation in effect" = "Yes"))
+    })) %>% 
+    unnest(pred_data) %>% 
+    filter(.width == 0.95) %>% 
+    select(nice, year_week_day, derogation_ineffect, .category, .epred, .lower, .upper) %>% 
+    group_by(nice, derogation_ineffect, .category) %>% 
+    mutate(rank = dense_rank(.epred)) %>% 
+    filter(rank == 1 | rank == max(rank)) %>% 
+    mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
+    group_by(rank, derogation_ineffect, .category) %>% 
+    mutate(outcome = janitor::make_clean_names(nice)) %>%
+    rename(draw = .epred, min = .lower, max = .upper) %>% 
+    mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0)))) %>% 
+    ungroup() %>% 
+    mutate(category = ifelse(is.na(.category), "NA", as.character(.category)))
+  
+  hr_mfx_tbl <- model_df %>% 
+    mutate(mfx_data = map(mfx_data, ~{
+      .x %>% 
+        mutate(group = as.character(group))
+    })) %>% 
+    unnest(mfx_data) %>% 
[…]

### 2023-05-05 _targets.R

@@ -136,6 +136,7 @@ list(
   tar_target(policies_plot_data, build_policies_plot_data(m_policies, year_week_lookup)),
   tar_target(policies_table_data, build_policies_table_data(policies_plot_data)),
   tar_target(human_rights_plot_data, build_human_rights_plot_data(m_human_rights, year_week_lookup)),
+  tar_target(hr_table_data, build_hr_table_data(human_rights_plot_data)),
 
   ## Manuscript and analysis notebook ----
   tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE, profile = "nice"),

### 2023-05-05 analysis/appendix.qmd
Add appendix to website

---
title: Appendix
toc: false
page-layout: full
---

```{r include=FALSE}
# Add the manuscript target as a dependency to the overall website
targets::tar_load(c(manuscript_nice, manuscript_manuscripty))
```

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/appendix.html" target="appendix">
  <i class="fa-solid fa-file-lines"></i> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/appendix.html" target="_blank">
  <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/appendix.pdf" target="appendix">
  <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
  </a>

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/appendix.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>

  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/appendix.pdf" target="appendix">
  <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
  </a>

  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/appendix.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>
  </a>
</p>

<iframe src="/manuscript/output/appendix.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="appendix"></iframe>
```

### 2023-05-05 analysis/manuscript.qmd
Add appendix to website

@@
 
-  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/manuscript-ms.pdf" target="manuscript">
+  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/manuscript.pdf" target="manuscript">
   <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
@@
 
-  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript-ms.pdf">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript.pdf">
   <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>

### 2023-05-05 manuscript/appendix.qmd
Add appendix to website

@@
   fig-title: Appendix Figure
+# Quarto struggles with adding the support CSS/JS files to this when rendering 
+# all the .qmd files in the manuscript folder, so we embed them in this file
+format:
+  html:
+    embed-resources: true
 ---

### 2023-05-05 manuscript/manuscript.qmd
Mention WHO

@@
 
-Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023.
-Doing so lifts measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well.
-As such, it is a worthwhile time to assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency period.
+Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, pandemic-related policies are less focused on rapid emergency responses.
+On May 5, 2023, the World Health Organization declared an end to the COVID-19 global health emergency, and many countries have made similar moves—the Biden administration, for instance, announced the end the United States' emergency measures on May 11, 2023.
+As such, now is a worthwhile time to assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency period.
 This research note examines what derogating meant for compliance with international law during the COVID-19 pandemic.

### 2023-05-06 manuscript/manuscript.qmd
Add a bunch of citations

@@
 First, while there is a significant literature on compliance with international law, especially human rights treaties [e.g.
-@Simmons2009; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations.
+@Simmons2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations.
 Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of adherence to derogations.
@@
 
-Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Simmons2009; @Merry:2009].
+Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @GoodliffeHawkins:2006; @Simmons2009; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
 While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
@@
 Our expectation comes from the specific, time-limited nature of derogations and the limited audience they speak to.
-Critics of international law argue that states ratify treaties with the expectation of reputational and other benefits.
+Critics of international law argue that states ratify treaties with the expectation of reputational and other benefits [@Hathaway:2007; @Hafner-BurtonLeVeckVictor:2017].
 However, derogations have no such public audience and are not signaling an increased commitment to human rights.

### 2023-05-06 manuscript/manuscript.qmd
Better math
Equation labels for crossref + no more mathrlap

@@
 
-\newpage
-
 ## Explanatory variables
@@
 Here we include two simplified model specifications to help demonstrate the intuition behind these multilevel TSCS models.
-For binary outcomes, we use a logistic model with country-specific offsets:
+For binary outcomes, we use a logistic model with region-specific offsets (see @eq-logit), while for ordered outcomes, we use ordered logistic regression with region-specific offsets (see @eq-ordered-logit).
 
@@
 \begin{aligned}
-&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
+&\ \textbf{Binary outcome $i$ across week $t$ within country $j$} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
@@
 &\ \textbf{Priors} \\
-\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-region variability,}
+\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Global averages and effects} \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Between-region variability}
 \end{aligned}
-$$
+$$ {#eq-logit}
 
 :::
-
-\noindent while for ordered outcomes, we use ordered logistic regression with country-specific offsets:
 
@@
 \begin{aligned}
-&\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within each country $j$}} \\
+&\ \textbf{Outcome level $i$ across week $t$ within country $j$} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
@@
 &\ \textbf{Priors} \\
-\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-region variability} \\
+\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Global averages and effects} \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Between-region variability} \\
 \alpha_k \sim&\ \mathcal{N}(0, 1) & \text{Boundaries between thresholds}
 \end{aligned}
-$$
+$$ {#eq-ordered-logit}
 

### 2023-05-06 manuscript/manuscript.qmd
Connect back to H1

@@
 In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
+This provides strong support for both components of our first hypothesis—states that derogated were substantially more likely to impose relevant emergency policies aimed at curtailing the spread of the virus (H~1a~), while states that did not issue formal derogations were less likely to engage in similar policies (H~1b~).
 

### 2023-05-06 manuscript/manuscript.qmd
Add cover page metadata

@@
 - name: Andrew Heiss
-  affiliation: Georgia State University
+  affiliations: 
+    - id: gsu
+      name: Georgia State University
+      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
+      address: "55 Park Place NE, #464"
+      city: Atlanta
+      region: GA
+      country: USA
+      postal-code: 30303
   orcid: 0000-0002-3948-3914
@@
   email: aheiss@gsu.edu
+  attributes:
+    corresponding: true
+correspondence-prefix: "Correspondence concerning this article should be addressed to"
 abstract: |
@@
 Our results support the role that flexibility mechanisms such as derogations play in international law and show that states are being sincere about their intentions and not, generally, using these mechanisms to cover abusive behavior.
+thanks: |
+  We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
 date: May 2, 2023

### 2023-05-06 _targets.R
Make separate profiles (and targets) for the manuscript and appendix
Quarto + targets does something weird with output_dir and individual files, so this is the best workaround. See https://github.com/ropensci/tarchetypes/discussions/107

@@ -140,7 +140,9 @@ list(
 
   ## Manuscript and analysis notebook ----
   tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE, profile = "nice"),
+  tar_quarto(appendix_nice, path = "manuscript", quiet = FALSE, profile = "appendix-nice"),
   tar_quarto(manuscript_manuscripty, path = "manuscript", quiet = FALSE, profile = "ms"),
+  tar_quarto(appendix_manuscripty, path = "manuscript", quiet = FALSE, profile = "appendix-ms"),
 
   tar_quarto(website, path = ".", quiet = FALSE),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),

### 2023-05-06 manuscript/appendix.qmd
Update date

--- 2023-05-05 manuscript/appendix.qmd
@@
   email: aheiss@gsu.edu
-date: May 2, 2023
+date: May 6, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2023-05-06 manuscript/manuscript.qmd
Update date

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: May 2, 2023
+date: May 6, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2023-05-09 analysis/appendix.qmd
Link appendix targets to here

@@
 # Add the manuscript target as a dependency to the overall website
-targets::tar_load(c(manuscript_nice, manuscript_manuscripty))
+targets::tar_load(c(appendix_nice, appendix_manuscripty))
 ```

### 2023-05-10 README.qmd
Pull abstract and title from manuscript.qmd

@@
 library(emoji)
+
+manuscript_target <- targets::tar_read(manuscript_nice)
+
+yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
+title <- yaml_metadata$title
+abstract <- yaml_metadata$abstract
 ```
 
-# Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19
+# `r title`
 
@@
 
-This research note asks whether states issuing pandemic-era human rights treaty derogations implemented emergency provisions as intended or used them to abuse human rights during a time of crisis.
-In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
-Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states that derogated from their international human rights obligations imposed emergency measures that were temporary and did not violate non-derogable rights.
-On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
-Our results support the role that flexibility mechanisms such as derogations play in international law and show that states are being sincere about their intentions and not, generally, using these mechanisms to cover abusive behavior.
+`r abstract`
 

### 2023-05-10 index.qmd
Pull abstract and title from manuscript.qmd

@@
     code-tools: false
+    toc: false
 sidebar: false
@@
 
-In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations.
-Did states actually end up complying with their pandemic-era derogations?
-Did derogating and non-derogating states implement emergency measures differently, and if so, how?
-Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory.
-On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights.
-Our results support the role that flexibility mechanisms such as derogations play in international law.
+```{r echo=FALSE}
+yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
+title <- yaml_metadata$title
+abstract <- yaml_metadata$abstract
+```
+
+### Abstract
+
+`r abstract`

### 2023-06-03 presentations/2022-11-04_peace-science/index.qmd
Better handling of slide resources

--- 2023-03-11 analysis/presentations/2022-11-04_peace-science/index.qmd
@@
 toc: false
-resources: slides
 abstract: |

### 2023-08-14 manuscript/manuscript.qmd
Fix table formatting
Suparna and Audrey edits

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: May 6, 2023
+date: August 14, 2023
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
-What is the impact of international legal emergency provisions on human rights?
-Combating a deadly virus like COVID-19 required extraordinary public health measures.
-In an effort to fight the COVID-19 pandemic, governments across the world introduced restrictions on the freedom of movement, association, and assembly.
-While countries were legally able to introduce these policies within their borders, the restrictions raised important questions about states’ commitments to international human rights law.
-Some human rights treaties recognize states’ ability to use emergency measures during times of crises, and contain "escape clauses" that safeguard the permanent erosion of human rights during and following emergencies.
-The temporary suspension of state compliance from international treaty obligations occurs through a legal action called derogation.
+Did states misuse international emergency provisions during the COVID-19 crisis to justify human rights abuse or did they follow international law?
+Responding to the pandemic required extraordinary public health measures.
+Many governments restricted freedom of movement, association, and assembly, raising questions about states’ commitments to international human rights law.
+Some human rights treaties recognize states’ ability to use emergency measures during crises, and contain “escape clauses” that safeguard the permanent erosion of human rights during and following emergencies.
+The temporary suspension of state compliance from international treaties occurs through a legal action called derogation.
 However, the potential abuse of emergency measures during derogations raises concerns about their impact on human rights.
 Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].
-Did states use pandemic era derogations how they were intended or were they used as a "pandemic pass" to justify violating human rights?
+Did states use pandemic-era derogations how they were intended or were they used as a “pandemic pass” to justify violating human rights?
 
-Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, pandemic-related policies are less focused on rapid emergency responses.
-On May 5, 2023, the World Health Organization declared an end to the COVID-19 global health emergency, and many countries have made similar moves—the Biden administration, for instance, announced the end the United States' emergency measures on May 11, 2023.
-As such, now is a worthwhile time to assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency period.
-This research note examines what derogating meant for compliance with international law during the COVID-19 pandemic.
-Specifically, we examine whether states actually derogated as per international legal principles and investigate the broader concern about whether states used derogations to exploit emergency provisions and to abuse human rights during COVID-19.
-These questions are important for the study of international law, compliance, and human rights during times of crises.
+This research note explores state derogations and human rights behavior during the pandemic.
+Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, on May 5, 2023, the World Health Organization declared an end to the global health emergency.
+Many countries followed suit—the Biden administration, for instance, announced the end of United States’ emergency measures on May 11, 2023.
+We now assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency.
+This research note examines what derogating meant for compliance with international law during the pandemic.
 
-Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation among derogating and non-derogating states on the implementation of emergency measures.
-We then analyze the effect of derogating (or not) on human rights practices during the pandemic.
+Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker, and United Nations treaty derogation data, we ask: What is the impact of international legal emergency provisions on human rights?
+How did derogating and non-derogating states vary in their implementation of emergency measures and human rights practices?
 We find that states that derogated and specified time limits to these emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights (the right to life, freedom from torture, and freedom from slavery).
-On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely to impose discriminatory policies, violate non-derogable rights, and impose emergency measures without time limits.
+States that did not derogate from their international legal obligations and still imposed emergency measures were more likely to implement discriminatory policies, violate non-derogable rights, and have emergency measures without time limits.
 
 This research note makes two contributions.
-First, while there is a significant literature on compliance with international law, especially human rights treaties [e.g.
-@Simmons2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations.
-Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of adherence to derogations.
-When states comply with derogations, they are actually limiting their obligations to the human rights treaty overall.
-As such, we should in fact expect a reduction in compliance with the overall human rights treaty during that time.
-We explore how the potential decline in compliance may actually be a good indication that states are using treaty design as intended.
+First, it expands our understanding of compliance with flexibility mechanisms and derogations.
+While there is a significant literature on compliance with international law, especially human rights treaties [e.g.
+@Simmons2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2005], we know little about states’ compliance with flexibility mechanisms such as derogations.
+We explore how the potential decline in overall treaty compliance may actually be a good indication that states are using treaty design, and derogation option, as intended.
 
@@
 
-Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @GoodliffeHawkins:2006; @Simmons2009; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
-While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
-@Lebret:2020; @helfer2021], as part of a broader trend of targeting groups during times of crises [@Grasse2021].
-We find that overall, derogating states during the pandemic did not increase abuse and violations.
-This finding speaks to international legal compliance scholarship and supports the idea that states are being sincere about their intentions and not, generally, using actions like derogation to cover abusive behavior.
+Second, this study contributes to the study of the relationship between international law and strategic repression by examining derogation behavior during a potentially tempting time for states to misuse them to increase repression and power during a time of global crisis.
+Previous derogation studies examine their use broadly, not during a defined global crisis.
+There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @Simmons2009; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
+While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
+@Lebret:2020; @helfer2021].
+Our findings are informative for international law, human rights, and repression.
+We find that derogating states during the pandemic did not increase abuse and violations.
+This supports the idea that states are being sincere about their intentions to comply with international law and not, generally, using actions like derogation to cover abusive behavior.
+This is an important finding for treaty design, including flexibility mechanisms, and the relationship between exploitation of international law to cover abuse.
 
-The rest of the paper proceeds as follows.
-First, we explore the purpose of derogations in international law, and then specifically during the pandemic.
-In doing so, we focus on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
-Following this, we introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic.
+Below, we first explore the purpose of derogations in international law, and then specifically during the pandemic.
+We focus on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
+We then introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state’s human rights practices during the pandemic.
 We conclude with the implications of our research note for the study of international law during times of crises.
@@
 
-Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility to respond to events.
+Derogations in international law authorize states to temporarily suspend their treaty commitments, providing them with flexibility to respond to crises [@HBHelferFariss].
 Times of crises can include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
-Derogations afford states the opportunity to adjust to these changing conditions [@HBHelferFariss].
-When states derogate from their international legal obligations, the derogation provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
+When states derogate from their international legal obligations, it provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
 This information allows these actors—at least in principle—to challenge measures that are excessive, vague, or outlast the intended time frame of their implementation [@helfer2021].
 
-Many legal critical assessments of derogations find importance in their use [@richardson2020], with some arguing that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97].
-An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency [@Greene:2020, 36].
-These include rights such as the right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies.
-Research supports this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
-Even states that officially derogate from their international legal obligations are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].
+Many legal assessments of derogations find importance in their use [@richardson2020], with some arguing that, “When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97].
+An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a *de jure* emergency [@Greene:2020, 36].
+Research supports this notion—there is no meaningful variation in states’ respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
+Even states that officially derogate from international treaties are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].
 
-These findings suggest that states may not be discerning the significance of derogable vs.
-non-derogable rights status during emergencies.
-This also raises the question of the extent to which this distinction is valuable during the pandemic.
-Most derogations to the ICCPR have been for internal conflict or natural disaster.
-Is a public health emergency different from other states of emergency?
+These findings suggest that states may not discern the difference between derogable vs.
+non-derogable rights during emergencies.
 Overall, there is a general expectation that following derogations, we may observe some decline in respect for human rights.
-However, previous studies have not explored the relationship between derogations and public health emergencies.
+However, previous studies have not explored the relationship between derogations and public health emergencies, as most derogations to the ICCPR have been for conflict or natural disasters.
 
@@
 
-A first question of interest is the nature of states that are more likely to submit them in the first place.
-Little research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the UN rather than the effects of these derogations.
-This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss].
-The expectation is that states with strong democratic institutions need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
-Additionally, democratic states understand international law and international legal mechanisms to be more legitimate and recognize the importance of them more so than other states.
-Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important to engage with.
+What kinds of states are more likely to submit derogations in the first place?
+Existing research focuses on the determinants of submitting derogations to the UN rather than the effects of these derogations.
+It finds that democracies are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not issue derogations [@HBHelferFariss].
+The former is because democracies typically need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
 Scholars were split on expectations about COVID-era derogations.
-@Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
-However, @richardson2020 see derogations as a less clear path for states during COVID-19, describing the actions, their conditions, and human rights as a "tangled morass" (p.
+@Lebret:2020 argued that derogations were the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
+However, @richardson2020 saw derogations as a less clear path for states during COVID-19, describing the actions and human rights as a “tangled morass” (p.
 106).
@@
 
-We find that between January 2020 and June 2021, `r derog_detail$countries` states issued a total of `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
-Of derogations that detailed specific obligation suspension lengths, the average length was 41 days.
-Descriptively, we see states across multiple regions derogate, with notable concentration in South America, which has a robust history of engaging with human rights law.
-We also note that it is not necessarily the "usual suspects" of strong democracies that derogate.
+Between January 2020 and June 2021, `r derog_detail$countries` states issued `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
+The average length of a derogation was 41 days.
+States across multiple regions derogated, with notable concentration in South America, which has a robust history of engaging with human rights law.
 Notably, Nordic states—which typically are very active in human rights law engagement—did not issue any derogations during this period.
@@
 
-In this research note, we look at states' derogations from the International Covenant on Civil and Political Rights (ICCPR) during the pandemic.
 The ICCPR is a foundational human rights treaty with broad human rights coverage.
-Exceptional circumstances brought on by the global pandemic have led to numerous, extensive derogations to the ICCPR removing state obligations to important rights areas of assembly, movement, and others.
-The ICCPR was adopted by the UN General Assembly on December 19, 1966, and entered into force on March 23, 1976.
-It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights.
-The ICCPR is one of only a few select human rights treaties that require states to report on the suspension or derogation of certain civil and political rights only under specific situations of "public emergency which threatens the life of the nation" [@iccpr].
+It is among a few human rights treaties that require states to report on the suspension or derogation of civil and political rights only under specific situations of “public emergency which threatens the life of the nation” [@iccpr].
 The other treaties with this requirement include the European Convention Protection of Human Rights and Fundamental Freedoms and the American Convention on Human Rights.
-However, the ICCPR is the only one with global, rather than regional, membership and participation to track during the pandemic.
-As of 2023, there are 173 states that are party to the ICCPR.
-This means that most, but not all the UN’s 193 members had the ability to submit derogations during COVID-19.
-Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.
+However, the ICCPR is the only one with global membership.
+As of 2023, 173 states are party to the ICCPR.
+This means that most of the UN's 193 members had the ability to submit derogations.
 
-To legally derogate, governments must first officially declare a state of emergency (Article 4).
-Derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory.
-However, it is worth noting that states cannot derogate from the entirety of the ICCPR.
-Seven rights enshrined in the ICCPR are non-derogable: the right to life, the prohibition of torture or cruel, inhumane punishment, the prohibition of slavery, the right to recognition as a person before the law, and the protection of freedom of thought, conscience and religion.
-The international legal community viewed these rights as so fundamental and important that no type of crisis could justify their suspension.
-However, other rights can be suspended during crises.
-Several ICCPR articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
+To derogate, governments must first officially declare a state of emergency (Article 4).
+Derogations must fulfill three criteria: they should be temporary, proportional to the threat and non-discriminatory.
+Seven rights cannot be suspended or are non-derogable: the right to life, the prohibition of torture or cruel, inhumane punishment, the prohibition of slavery, the right to recognition as a person before the law, and the protection of freedom of thought, conscience and religion.
+However, several other ICCPR articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
 
-Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination.
-However, not every country formally derogated—a number of countries implemented emergency measures without formally derogating from the ICCPR.
-Over 100 countries issued emergency declarations while fewer than 30 submitted COVID-19 related derogations [@icnl-covid].
-These COVID-19 emergency declarations may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020].
-Below, we first theorize how derogations led to variation in the implementation of emergency measures during the pandemic, followed by their impact on human rights.
+Recognizing that public health exigencies would require states to curb rights, the UN Human Rights Office of the High Commissioner (OHCHR) stated that emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination.
+However, not every country derogated—a number of countries implemented emergency measures without formally derogating from the ICCPR.
+Over 100 countries issued emergency declarations while fewer than 30 submitted derogations [@icnl-covid].
+Below, we theorize how derogations led to variation in the implementation of emergency measures during the pandemic, and their impact on human rights.
 
@@
 
-During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21).
-As our descriptive data shows, not all states issued derogations.
-Derogating states were in the minority of states during the pandemic.
-
-We expect that states that derogate during COVID-19 with the intention to impose emergency policies will follow through.
-Our expectation comes from the specific, time-limited nature of derogations and the limited audience they speak to.
-Critics of international law argue that states ratify treaties with the expectation of reputational and other benefits [@Hathaway:2007; @Hafner-BurtonLeVeckVictor:2017].
+We expect that states that derogated with the intention to impose emergency policies will follow through.
+Many argue that states ratify treaties with the expectation of reputational and other benefits [@Hathaway:2007; @Hafner-BurtonLeVeckVictor:2017].
 However, derogations have no such public audience and are not signaling an increased commitment to human rights.
-There is no clear or expected connection with foreign aid, trade, or other theorized 'rewards' for larger legal actions.
+There is no clear or expected connection with foreign aid, trade, or other theorized ‘rewards’ for larger legal actions.
 Instead, derogations are a very specific legal action that states voluntarily enter into after commitment has fully taken place.
@@
 
-> **Hypothesis 1~a~**: If states issued COVID-19 derogations, then we expect that they will be more likely to impose relevant emergency polices such as canceling public events, restricting gatherings, closing public transit, and restricting domestic and international movement.
+> **Hypothesis 1~a~**: If states issued COVID-19 derogations, then they will be more likely to impose emergency policies that cancel public events, restrict gatherings, close public transit, and restrict movement.
 >
-> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose relevant emergency policies.
+> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, then they will be less likely to impose relevant emergency policies.
 
@@
 If states determine that emergency measures are justified without regard for international law, the processes for respecting rights during crises and restoring rights at the end of the emergency may not be followed.
-As such, many states used the pandemic as an excuse to centralize authority and implemented measures that did not meet international legal standards.
-Many government engaged in discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
+Many states used the pandemic as an excuse to centralize authority.
+Governments engaged in discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups.
 For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
-In Bulgaria, checkpoint controls went into effect in neighborhoods largely composed of the Roma community [@bulgaria2020].
-In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country's new citizenship law seen as discriminatory towards Muslims [@bbc2020].
+In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country’s new citizenship law seen as discriminatory towards Muslims [@bbc2020].
 
@@
 For instance, in March 2020, a French government order extended the duration of pre-trial detention, raising the issue of prison overcrowding and leading to inhumane and degrading treatment [@Lebret:2020].
-As such, this was not just a violation of a non-derogable right, but also raised questions regarding the proportionality of emergency measures.
+Besides violating a non-derogable right, this also raised questions regarding the proportionality of emergency measures.
 
-We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to impose measures that violated these criteria.
-If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.
+We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to violate these criteria.
 
@@
 
-While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we use global data to examine the relationship more broadly between treaty derogations, human rights, and state actions during the pandemic.
-We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
+We use global data to examine the relationship between treaty derogations, human rights, and state actions during the pandemic—specifically, we use data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
 We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021.
@@
 
-We examine the relationship between formal treaty derogations and different policy and human rights outcomes in two stages (see @tbl-vars).
-First, to test our first set of hypotheses and explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
-Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.
+We examine the relationship between derogations and policy and human rights outcomes in two stages (see @tbl-vars).
+First, to test whether derogations were associated with emergency state responses, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
+Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] We collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.
 
@@
 \begingroup\footnotesize
+
+\renewcommand*{\arraystretch}{1.75}
+
+::: {.content-visible when-profile="ms"}
+\singlespacing
+:::
 
@@
 
-Next, to test our second set of hypotheses, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
+Next, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
 Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations.
@@
 
-For both sets of hypotheses, we explain changes in policy and human rights outcomes using a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
+We use a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
 We collected and coded this data from real time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023].
@@
 
-We include several additional variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
-To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
+Several additional variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
+To account for the severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
 
-Since formally derogating from treaties can serve as a signal that countries take their obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR post-commitment actions, such as RUDs, prior to the pandemic.
+Since derogations signal that countries take their legal obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR post-commitment actions, such as RUDs, prior to the pandemic.
 Some states new to derogations introduced them during the COVID-19 pandemic [@Comstock:2023].
@@
 To examine this first step, we code two binary variables: (1) past ICCPR derogation behavior before 2020 and (2) past ICCPR post-commitment action behavior before 2020.
-This allows us to empirically test for the first step of engagement before analyzing any impact that COVID-19-era derogations had on different rights practices.
+This allows us to test for the first step of engagement before analyzing any impact that COVID-19-era derogations had on different rights practices.
 There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019.
@@
 \begin{aligned}
-&\ \textbf{Binary outcome $i$ across week $t$ within country $j$} \\
+&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within country $j$}} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
@@
 \begin{aligned}
-&\ \textbf{Outcome level $i$ across week $t$ within country $j$} \\
+&\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within country $j$}} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
@@
 
-In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society.
-We use Student *t* priors clustered around 0 with a scale parameter of 3, since (1) these model coefficients are measured on the logged odds scale and rarely result in large numbers, and (2) the Student *t* distribution has heavier tails than a standard normal distribution.
-
 Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in Appendix Tables 2 and 4).
@@
 
-We begin by exploring our first set of hypotheses to examine how derogating and non-derogating states approached the implementation of their emergency public health measures.
+We first examine how derogating and non-derogating states implemented emergency public health measures.
 @fig-preds-mfx-policies-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
-In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
-This provides strong support for both components of our first hypothesis—states that derogated were substantially more likely to impose relevant emergency policies aimed at curtailing the spread of the virus (H~1a~), while states that did not issue formal derogations were less likely to engage in similar policies (H~1b~).
+Countries with a derogation in effect are more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
+This provides strong support for both components of our first hypothesis—states that derogated were substantially more likely to impose relevant emergency policies aimed at curtailing the spread of the virus (H~1a~), while states that did not issue derogations were less likely to engage in similar policies (H~1b~).
 
-For instance, all countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance.
+All countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance.
 Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021.
-Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability.
+Countries that derogated, however, regularly canceled with a nearly 100% probability.
 The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice`–`r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`) between March 2020 and June 2021.
@@
 Non-derogating states followed a similar downward trend over time, but `r policies_mfx$close_public_transit.min$draw_nice`–`r policies_mfx$close_public_transit.max$draw_nice` percentage points lower than their derogating counterparts (`r policies_mfx$close_public_transit.max$p_gt`).
-Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to `r policies_pred$movement.Yes.min$draw_nice`% by June 2021.
-Non-derogating states, on the other hand, began the pandemic with a predicted `r policies_pred$movement.No.max$draw_nice`% probability of limiting movement, dropping to only `r policies_pred$movement.No.min$draw_nice`% by June 2021, representing a `r policies_mfx$movement.min$draw_nice`–`r policies_mfx$movement.max$draw_nice` percentage point difference (`r policies_mfx$movement.max$p_gt`).
-The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel.
+Derogating states were more likely to restrict movement, with a nearly 100% probability, dropping to `r policies_pred$movement.Yes.min$draw_nice`% by June 2021.
+Non-derogating states began the pandemic with a predicted `r policies_pred$movement.No.max$draw_nice`% probability of limiting movement, dropping to only `r policies_pred$movement.No.min$draw_nice`% by June 2021, representing a `r policies_mfx$movement.min$draw_nice`–`r policies_mfx$movement.max$draw_nice` percentage point difference (`r policies_mfx$movement.max$p_gt`).
+The only policy where there was no sizable difference between the two types of states is restrictions on international travel.
 Both derogating and non-derogating states had a nearly 100% probability of limiting international movement across all 15 months.
@@
 
-These results reveal important insights into the motivations of derogating states.
-Countries that undertook emergency safety measures seem to have taken the time to follow international law and uphold their treaty obligations.
-International law thus appears to not just be an empty commitment—states that took the time to formally derogate were more likely to keep these emergency measures in place for longer and take their derogations seriously.
+These results show that countries implementing emergency measures seem to have taken the time to follow international law and uphold their treaty obligations.
+International law appears to not just be an empty commitment—states that derogated were more likely to keep these emergency measures in place for longer and take their derogations seriously.
 
@@
 
-The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights).
+Derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights).
 There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement.
@@
 
-The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems.
-While elites or diplomats are the ones submitting derogations and committing to complying with international law, they are not the actors implementing emergency measures.
-During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially as it relates to human rights.
-Lower-level officials such as those in the police, state security forces and regional bureaucrats are often the ones enforcing emergency measures.
+The mechanisms behind this reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems.
+While elites submit derogations and commit to complying with international law, they are not the actors who implement emergency measures.
+There may be immense variation in states' abilities to enforce these policies as intended, especially as it relates to human rights.
+Lower-level officials such as those in the police, state security forces and regional bureaucrats often enforce emergency measures.
 Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations.
-Further, as previous research shows, during high levels of stress and fear, decision making is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
-As such, these differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations and claim that emergency measures will be implemented proportional to the threat at hand.
+Further, during high levels of stress and fear, decision-making is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016].
+These differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations.
 
 In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status.
-For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a `r hr_pred$no_time_limit_measures.No.None.max$draw_nice`% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with `r hr_pred$no_time_limit_measures.No.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.No.Moderate.max$draw_nice`% probability).
+For non-derogating countries, the two most likely predicted outcomes are to have no issues with time limits (with a `r hr_pred$no_time_limit_measures.No.None.max$draw_nice`% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with `r hr_pred$no_time_limit_measures.No.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.No.Moderate.max$draw_nice`% probability).
 The middle ground of only minor issues with time limits is exceptionally unlikely.
-Countries that formally derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only an `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues.
+Countries that derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only an `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues.
 The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
@@
 
-Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies.
+Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abusive enforcement of those policies.
 We thus find strong support only for H~2c~—that derogating states are less likely to impose measures with no time limits.
 We find weak support for our hypothesis that derogating states will be less likely to violate non-derogable rights (H~2b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
-Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
+Derogations might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
 
@@
 
-The COVID-19 pandemic raised numerous concerns that the crisis would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide.
-Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations.
+The COVID-19 pandemic raised concerns about human rights violations worldwide.
+This research note examines state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations and state repression.
 We focused specifically on the impact of derogations—an international legal means to communicate to both the UN and the global community that certain rights will be suspended during times of crisis.
-Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties without concern about the inability to adapt to times of emergency.
-International legal scholarship has been divided over expectations of derogations during the pandemic.
-
 With the caveat that our results are not causal—we have not attempted to use any formal strategy to identify an unconfounded causal effect, and our findings are therefore descriptive—we find that there is good news for compliance and derogations.
-Derogators did not systematically use this legal "escape clause" opportunistically—derogators used the process as intended.
-Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
-Though most states did not submit COVID-19 related derogations, the ones that did used derogations as intended by international law.
+Derogating states did not use this legal “escape clause” opportunistically.
+Most derogations included time limitations, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
+Though most states did not submit COVID-19 related derogations, the ones that did, used derogations as intended by international law.
 Once the derogation ended, states readjusted to typical compliance once more.
@@
 
-We suggest that future research examine different incentives facing state officials during crisis.
-There may be key differences in enforcement between local level bureaucrats and higher-level elites.
+We suggest that future research examines different incentives facing state officials during a crisis.
+There may be key differences in enforcement between local bureaucrats and higher-level elites.
 While elites submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.
@@
 
-Overall, we find that states used derogations as they were intended during the COVID-19 pandemic and no large-scale misuse of the legal action took place.
+Overall, we find that states used derogations as they were intended during the COVID-19 pandemic and no large-scale misuse of the legal action took place to repress human rights.
 It can be an important signal for states to submit derogations, both to international and domestic communities, about intentions to mitigate the spread of disease through the temporary suspension of human rights.

### 2024-04-17 _targets.R
Set targets store and script

@@ -12,6 +12,14 @@ options(mc.cores = 4,
 
 set.seed(8588)  # From random.org
 
+# This hardcodes the absolute path in _targets.yaml, so to make this more
+# portable, we rewrite it every time this pipeline is run (and we don't track
+# _targets.yaml with git)
+tar_config_set(
+  store = here::here("_targets"),
+  script = here::here("_targets.R")
+)
+
 tar_option_set(
   packages = c("tidyverse"),
   format = "qs",

### 2024-04-17 deploy.sh
Change remote server details

@@ -1,5 +1,5 @@
-REMOTE_HOST="cloud"
-REMOTE_DIR="~/sites/stats/public_html/mountainous-mackerel"
+REMOTE_HOST="ath-cloud"
+REMOTE_DIR="~/sites/stats.andrewheiss.com/public_html/mountainous-mackerel"
 REMOTE_DEST=$REMOTE_HOST:$REMOTE_DIR
 
 echo "Uploading new changes to remote server..."

### 2024-04-17 manuscript/manuscript.qmd
Use {tinytable}!

@@
 library(colorspace)
+library(tinytable)
 library(kableExtra)
-library(gt)
 library(targets)
@@
 #| echo: false
+#| message: false
+
 bind_cols(dplyr::slice(map_data_nice, 1:12),
           add_row(dplyr::slice(map_data_nice, 13:n()))) %>%
-  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]",
-      col.names = rep(c("Country", "Derogations"), 2)) %>% 
-  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))
+  set_names(c("Country", "Derogations", "Country", "Derogations")) %>% 
+  tt(width = c(0.35, 0.15, 0.35, 0.15) * 0.9) %>% 
+  format_tt() %>% 
+  style_tt(j = 1:4, align = "lrlr") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm table-striped",
+    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}"
+  )
 ```

### 2024-04-18 analysis/appendix.qmd
Stop using profiles; use pandoc extension modifiers!
Use pandoc extension modifiers to make different versions of the PDFs instead of profiles (https://github.com/quarto-dev/quarto-cli/issues/7371)

@@
 # Add the manuscript target as a dependency to the overall website
-targets::tar_load(c(appendix_nice, appendix_manuscripty))
+targets::tar_load(c(manuscript))
 ```
@@
 
-  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/appendix.pdf" target="appendix">
+  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output/appendix+ms.pdf" target="appendix">
   <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
@@
 
-  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/appendix.pdf">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/appendix+ms.pdf">
   <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>

### 2024-04-18 analysis/paper.qmd
Rename manuscript page to "paper"
Stop using profiles; use pandoc extension modifiers!
Use pandoc extension modifiers to make different versions of the PDFs instead of profiles (https://github.com/quarto-dev/quarto-cli/issues/7371)

--- 2023-05-05 analysis/manuscript.qmd
@@
 ---
-title: Manuscript
+title: Paper
 toc: false
@@
 # Add the manuscript target as a dependency to the overall website
-targets::tar_load(c(manuscript_nice, manuscript_manuscripty))
+targets::tar_load(c(manuscript))
 ```
@@
 
-  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/manuscript.pdf" target="manuscript">
+  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output/manuscript+ms.pdf" target="manuscript">
   <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
@@
 
-  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript.pdf">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript+ms.pdf">
   <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>

### 2024-04-18 _targets.R

@@ -147,10 +147,7 @@ list(
   tar_target(hr_table_data, build_hr_table_data(human_rights_plot_data)),
 
   ## Manuscript and analysis notebook ----
-  tar_quarto(manuscript_nice, path = "manuscript", quiet = FALSE, profile = "nice"),
-  tar_quarto(appendix_nice, path = "manuscript", quiet = FALSE, profile = "appendix-nice"),
-  tar_quarto(manuscript_manuscripty, path = "manuscript", quiet = FALSE, profile = "ms"),
-  tar_quarto(appendix_manuscripty, path = "manuscript", quiet = FALSE, profile = "appendix-ms"),
+  tar_quarto(manuscript, path = "manuscript", quiet = FALSE),
 
   tar_quarto(website, path = ".", quiet = FALSE),
   tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),

### 2024-04-18 manuscript/manuscript.qmd
Use tinytable for all manuscript tables

@@
 tribble(
-  ~`Outcome variables`,                        ~Description,                                                        ~Frequency,  ~Source,
+  ~`Variable`,                                 ~Description,                                                        ~Frequency,  ~Source,
   "Emergency public health measures",          "Binary indicator",                                                  "Weekly",    "Oxford COVID-19 Government Response Tracker",
@@
 ) %>% 
-  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]") %>% 
-  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA)) %>% 
-  pack_rows(index = c(" " = 2, "Primary explanatory variable" = 1, "Secondary explanatory variables" = 5),
-            indent = FALSE) %>% 
-  column_spec(c(1, 2, 4), width = "0.25\\\\textwidth") %>% 
-  column_spec(3, width = "0.1\\\\textwidth")
+  tt(width = c(0.3, 0.3, 0.1, 0.3)) %>% 
+  group_tt(i = list(
+    "Outcome variables" = 1,
+    "Primary explanatory variable" = 3,
+    "Secondary explanatory variables" = 4
+  ), indent = 0) %>% 
+  style_tt(j = 1:4, align = "l") %>% 
+  style_tt(i = c(1, 4, 6), bold = TRUE, background = "#e6e6e6", align = "c") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm table-striped",
+    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}, row{4,6}={belowsep+=2em}"
+  )
 ```
@@
 
-\begingroup\footnotesize
-
-\renewcommand*{\arraystretch}{1.75}
-
-::: {.content-visible when-profile="ms"}
-\singlespacing
-:::
-
-| General question                           | Outcomes                                                                                                                                  | Main explanatory variable  | Other explanatory variables                                                                                                                                              |
-|-----------------|-----------------|-----------------|---------------------|
-| Derogations and emergency measures (H~1~)  | Cancelling public events, Gathering restrictions, Close public transit, Restrictions on internal movement, Limits on international travel | ICCPR derogation in effect | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions                                                                     |
-| Derogations and human rights (H~2~)        | Discriminatory policy, Violations of non-derogable rights, No time limit measures, Abusive enforcement                                    | ICCPR derogation in effect | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index, core civil society index |
-
-: Overview of outcomes and explanatory variables included in models {#tbl-vars}
-
-\endgroup
+```{r tbl-vars}
+#| tbl-cap: Overview of outcomes and explanatory variables included in models
+#| include: true
+#| echo: false
+tibble(
+  `General question` = c(
+    "Derogations and emergency measures (H~1~)",
+    "Derogations and human rights (H~2~)"
+  ),
+  Outcomes = c(
+    "Cancelling public events; Gathering restrictions; Close public transit; Restrictions on internal movement; Limits on international travel",
+    "Discriminatory policy; Violations of non-derogable rights; No time limit measures; Abusive enforcement"
+  ),
+  `Main explanatory variable` = c(
+    "ICCPR derogation in effect", 
+    "ICCPR derogation in effect"
+  ),
+  `Other explanatory variables` = c(
+    "New and cumulative COVID cases; New and cumulative COVID deaths; Prior ICCPR derogations and actions",
+    "New and cumulative COVID cases; New and cumulative COVID deaths; Prior ICCPR derogations and actions; Rule of law index; Civil liberties index; Core civil society index"
+  )
+) %>% 
+  tt(width = c(0.2, 0.3, 0.2, 0.3)) %>% 
+  style_tt(j = 1:4, align = "l") %>% 
+  format_tt(j = 1:2, markdown = TRUE) %>% 
+  style_tt(
+    bootstrap_class = "table table-sm table-striped",
+    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}, row{1}={valign=b}"
+  )
+```
 

### 2024-04-18 manuscript/manuscript.qmd
Update date + minor tweaks

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: August 14, 2023
+date: April 17, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
+::: {.content-visible when-format="html"}
+# Introduction {.hide}
+:::
+
 Did states misuse international emergency provisions during the COVID-19 crisis to justify human rights abuse or did they follow international law?
@@
 
-```{r fig-derogation-map, fig.width=6.5, fig.height=3.5}
+```{r fig-derogation-map, fig.width=6.5, fig.height=3.5, out.width="100%"}
 #| include: true
@@
 
-```{r fig-preds-mfx-policies, fig.width=12, fig.height=6, out.width="100%", cache=FALSE}
+```{r fig-preds-mfx-policies, fig.width=12, fig.height=6, out.width="100%"}
 #| include: true

### 2024-04-19 manuscript/appendix.qmd
Phew, complex tinytable stuff throughout the appendix!

@@
   email: aheiss@gsu.edu
-date: May 6, 2023
+date: April 18, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 crossref: 
-  tbl-title: Appendix Table
-  fig-title: Appendix Figure
+  tbl-title: "Table&nbsp;A"
+  fig-title: "Figure&nbsp;A"
 ---
@@
 library(gt)
+library(tinytable)
 library(targets)
@@
 add_breaks <- function(x, fmt) {
+  sapply(x, \(text) {
+    if (is.na(text)) {
+      return(NA)
+    }
+
     if (fmt == "latex") {
-    kableExtra::linebreak(x, align = "c")
-  } else (
-    str_replace(x, "\\n", "<br>")
-  )
+      text <- str_replace_all(text, c("\\[" = "{[}", "\\]" = "{]}"))
+      if (str_detect(text, "\\n")) {
+        str_c("{", str_replace(text, "\\n", r"{ \\\\ }"), "}")
+      } else {
+        text
+      }
+    } else {
+      str_replace(text, "\\n", "<br>")
+    }
+  }) %>% unname()
 }
@@
 
+::: {.content-visible when-format="html"}
 # H~1~: Derogations and emergency policies
@@
 ## Predicted probabilities and group contrasts
+:::
+
+\blandscape
 
@@
 #| echo: false
+
 tbl_part1 <- policies_table_data$policies_pred_tbl %>% 
@@
   mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
-  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
+  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
   select(nice, time, derogation_ineffect, num) %>% 
@@
   mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
-  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
+  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
   select(nice, year_week_day, diff_type, num, p_gt_0) %>% 
@@
 
+tbl_groups <- tbl_part1 %>%
+  select(nice) %>% 
+  mutate(row = 1:n()) %>%
+  distinct(nice, .keep_all = TRUE) %>% 
+  mutate(row_cumulative = row + row_number() - 1)
+
 tbl_part1 %>%
@@
   mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
-  select(derogation_ineffect, `March 2020`, `June 2021`,
+  select(
+    derogation_ineffect, `March 2020`, `June 2021`,
     `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
-         `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`) %>% 
-  kbl(col.names = c("Derogation", "March 2020", "June 2021",
-                    "Week", "Difference", "p > 0", "Week", "Difference", "p > 0"),
-      align = "lcccccccc",
-      booktabs = TRUE, linesep = "",
-      escape = FALSE) %>% 
-  kable_styling(font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-  column_spec(3, border_right = T) %>% 
-  pack_rows(index = table(fct_inorder(tbl_part1$nice))) %>% 
-  add_header_above(c(" " = 1, "Predicted probabilities" = 2, 
-                     "Smallest difference" = 3, "Largest difference" = 3))
+    `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`
+  ) %>%
+  set_names(
+    c(
+      "Derogation", "March 2020", "June 2021",
+      "Week", "$\\Delta$", "p > 0", "Week", "$\\Delta$", "p > 0"
+    )
+  ) %>%
+  tt(width = c(0.7, 1, 1, 1, 0.8, 0.5, 1, 0.8, 0.5) |> {\(x) x / sum(x)}()) %>%
+  format_tt() %>%
+  style_tt(j = 1:9, align = "lcccccccc") %>%
+  style_tt(i = 1:20, j = 3, line = "r", line_width = 0.1, line_color = "#000000") %>%
+  group_tt(
+    i = pull(tbl_groups, row, name = nice) %>% as.list(),
+    j = list(
+      "Predicted probabilities" = 2:3, 
+      "Smallest difference" = 4:6, 
+      "Largest difference" = 7:9),
+    indent = 0) %>% 
+  style_tt(
+    i = tbl_groups$row_cumulative, 
+    bold = TRUE, background = "#e6e6e6", align = "l"
+  ) %>% 
+  style_tt(
+    bootstrap_class = "table table-sm",
+    tabularray_inner = "rowsep=1pt, cells={font=\\scriptsize}, row{2}={valign=b}"
+  )
 ```
 
+\elandscape
+
+::: {.content-visible when-format="html"}
 ## Complete model results
+:::
 
 ```{r tbl-results-full-policies}
-#| tbl-cap: Results from models showing relationship between derogations and emergency policies
+#| tbl-cap: Complete results from models showing relationship between derogations and emergency policies (H~1~)
 #| include: true
 #| echo: false
-notes <- c(
-  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
+
+notes <- paste(
+  "Note: Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
   "95% credible intervals (highest density posterior interval, or HDPI) in brackets.",
@@
 if (fmt_out == "latex") {
-  notes <- notes %>% 
-    str_replace_all(r"[\\]", r"[\\\\]") %>% 
-    str_replace_all("%", r"(\\\\%)")
-} else {
-  notes <- paste(notes, collapse = " ")
+  notes <- str_replace_all(notes, "%", r"(\\%)")
 }
 
-if (fmt_out == "latex") { 
+inner <- "
+row{odd[3-27]}={abovesep=0pt, belowsep=6pt},
+row{even[2-26]}={rowsep=0pt},
+cells={font=\\scriptsize}, row{1}={valign=b}
+"
+
+css <- "
+.coef-table-1 tbody tr:not(:nth-child(n+26)):nth-child(odd) td {
+    padding-bottom: 0;
+}
+.coef-table-1 tbody tr:not(:nth-child(n+27)):nth-child(even) td {
+    padding-top: 0;
+}
+"
+
 modelsummary(models_tbl_policies,
@@
   gof_map = gof_map,
-               output = "kableExtra",
+  output = "tinytable",
   fmt = fmt_significant(2),
-               escape = FALSE) %>% 
-    kable_styling(htmltable_class = "table-sm light-border",
-                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
-    column_spec(1:6, width = "0.16\\\\textwidth")
-} else {
-  modelsummary(models_tbl_policies,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "gt",
-               fmt = fmt_significant(2)) %>% 
-    tab_footnote(footnote = notes) %>% 
-    tab_style(style = cell_text(weight = "bold"),
-              locations = cells_column_labels())
-}
+  notes = notes,
+  width = c(0.2, rep(0.16, 5)),  
+) %>% 
+  style_tt(i = seq(1, 25, 2), j = 1, rowspan = 2, alignv = "t") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm coef-table-1",
+    bootstrap_css_rule = css,
+    tabularray_inner = inner
+  )
 ```
 
+::: {.content-visible when-format="html"}
 # H~2~: Derogations and human rights
@@
 ## Predicted probabilities and group contrasts
+:::
 
-```{r tbl-hr-summary}
-#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models
+\blandscape
+
+```{r tb-hr-summary}
+#| tbl-cap: "Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models"
 #| include: true
 #| echo: false
+
 tbl_part1 <- hr_table_data$hr_pred_tbl %>% 
@@
   mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
-  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
+  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
   select(nice, time, group = category, derogation_ineffect, num) %>% 
@@
   mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
-  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
+  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
   select(nice, year_week_day, group, diff_type, num, p_gt_0) %>% 
@@
   mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
-  select(derogation_ineffect, group, `March 2020`, `June 2021`,
+  select(
+    derogation_ineffect, group, `March 2020`, `June 2021`,
     `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
-         `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`) %>% 
-  kbl(col.names = c("Derogation", "Level", "March 2020", "June 2021",
-                    "Week", "Difference", "p > 0", "Week", "Difference", "p > 0"),
-      align = "lcccccccc",
-      booktabs = TRUE, linesep = "",
-      escape = FALSE) %>% 
-  kable_styling(font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-  column_spec(4, border_right = T) %>% 
-  pack_rows(index = table(fct_inorder(tbl_part1$nice))) %>%
-  add_header_above(c(" " = 2, "Predicted probabilities" = 2, 
-                     "Smallest difference" = 3, "Largest difference" = 3))
+    `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`
+  ) %>%
+  set_names(
+    c(
+      "Derogation", "Level", "March 2020", "June 2021",
+      "Week", "$\\Delta$", "p > 0", "Week", "$\\Delta$", "p > 0"
+    )
+  ) %>% 
+  tt(
+    width = c(0.7, 0.7, 1, 1, 1, 0.8, 0.5, 1, 0.8, 0.5) |> {\(x) x / sum(x)}(),
+    caption = "Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models \\label{tbl-hr-summary}") %>%
+  format_tt() %>%
+  style_tt(j = 1:10, align = "lccccccccc") %>%
+  style_tt(i = 1:30, j = 4, line = "r", line_width = 0.1, line_color = "#000000") %>%
+  group_tt(
+    i = pull(tbl_groups, row, name = nice) %>% as.list(),
+    j = list(
+      "Predicted probabilities" = 3:4, 
+      "Smallest difference" = 5:7, 
+      "Largest difference" = 8:10),
+    indent = 0) %>% 
+  style_tt(
+    i = tbl_groups$row_cumulative, 
+    bold = TRUE, background = "#e6e6e6", align = "l"
+  ) %>% 
+  style_tt(
+    bootstrap_class = "table table-sm",
+    tabularray_outer = "theme=long",
+    tabularray_inner = "rowsep=1pt, cells={font=\\scriptsize}, row{2}={valign=b}"
+  ) %>% 
+  theme_tt("multipage", rowhead = 2)
 ```
 
+\elandscape
+
+::: {.content-visible when-format="html"}
 ## Complete model results
+:::
 
 ```{r tbl-results-full-human-rights}
-#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~1~)
+#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~2~)
 #| include: true
 #| echo: false
-if (fmt_out == "latex") {
+
+inner <- "
+row{odd[3-33]}={abovesep=0pt, belowsep=2pt},
+row{even[2-34]}={abovesep=0pt, belowsep=0pt},
+cells={font=\\scriptsize}, row{1}={valign=b}
+"
+
+css <- "
+.coef-table-2 tbody tr:not(:nth-child(n+32)):nth-child(odd) td {
+    padding-bottom: 0;
+}
+.coef-table-2 tbody tr:not(:nth-child(n+33)):nth-child(even) td {
+    padding-top: 0;
+}
+"
+
 modelsummary(models_tbl_human_rights,
@@
   gof_map = gof_map,
-               output = "kableExtra",
+  output = "tinytable",
   fmt = fmt_significant(2),
-               escape = FALSE) %>% 
-    kable_styling(htmltable_class = "table-sm light-border",
-                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
-    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
-    column_spec(1:5, width = "0.18\\\\textwidth")
-} else {
-  modelsummary(models_tbl_human_rights,
-               estimate = "{estimate}",
-               statistic = "[{conf.low}, {conf.high}]",
-               coef_map = coef_map,
-               gof_map = gof_map,
-               output = "gt",
-               fmt = fmt_significant(2),
-               escape = FALSE) %>% 
-    tab_footnote(footnote = notes) %>% 
-    tab_style(style = cell_text(weight = "bold"),
-              locations = cells_column_labels())
-}
+  notes = notes,
+  width = rep(0.2, 5),  
+) %>% 
+  style_tt(i = seq(1, 25, 2), j = 1, rowspan = 2, alignv = "t") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm coef-table-2",
+    bootstrap_css_rule = css,
+    tabularray_inner = inner
+  )
 ```
 
+\clearpage
+
 # References

### 2024-04-30 manuscript/appendix.qmd
Mega revisions for R2

@@
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
-bibliography: bib.bib
+bibliography: references.json
+csl: csl/chicago-author-date.csl
 link-citations: true
@@
 
+```{r tikz-stuff, include=FALSE}
+# Necessary for using dvisvgm on macOS
+# See https://www.andrewheiss.com/blog/2021/08/27/tikz-knitr-html-svg-fun/
+Sys.setenv(LIBGS = "/usr/local/share/ghostscript/9.53.3/lib/libgs.dylib.9.53")
+
+font_opts <- list(
+  extra.preamble = c(
+    "\\usepackage{libertine}",
+    "\\usepackage{libertinust1math}"
+  ),
+  # dvisvgm.opts = "--font-format=woff"
+  dvisvgm.opts = "--no-fonts"
+)
+```
+
 ```{r libraries-data}
 library(tidyverse)
+library(patchwork)
 library(modelsummary)
@@
 
+who_regions_raw <- tar_read(iccpr_who_clean) %>% 
+  distinct(country_name, iso3, who_region, who_region_long)
+
+tar_load(year_quarter_lookup)
 tar_load(c(policies_table_data, hr_table_data))
@@
 
-# modelsummary functions
+# Plotting and modelsummary functions
+invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
 invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
@@
 
+# WHO regions
+
+```{r tbl-who-regions}
+#| tbl-cap: WHO regions
+#| include: true
+#| echo: false
+who_regions <- who_regions_raw %>%
+  mutate(
+    region_title = glue::glue("{who_region}: {who_region_long}")
+  ) %>% 
+  group_by(region_title) %>% 
+  summarize(
+    countries = str_flatten(country_name, collapse = ", ", last = ", and ")
+  ) %>% 
+  pivot_longer(cols = everything()) %>% 
+  select(-name) %>% 
+  as.data.frame() %>%  # tibbles don't like NULL column names
+  set_names(NULL)
+
+who_regions %>% 
+  tt() %>% 
+  style_tt(i = seq(1, nrow(who_regions), by = 2), bold = TRUE, background = "#e6e6e6") %>% 
+  style_tt(i = seq(2, nrow(who_regions), by = 2), indent = 2)
+```
+
 # Modeling approach
 
-We use Stan `r cmdstanr::cmdstan_version()` [@stan] through R `r paste0(version$major, ".", version$minor)` [@r-project] and {brms} `r packageVersion("brms")` [@brms-jss] to estimate our models.
+We use Stan `r cmdstanr::cmdstan_version()` [@stan] through R `r paste0(version$major, ".", version$minor)` [@rproject-4-3-3] and {brms} `r packageVersion("brms")` [@Burkner:2017] to estimate our models.
 We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup.
@@
 Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW>.
-Our general modeling approach can be summarized as follows:
+
+## Priors
+
+We follow the suggestion of @GelmanJakulinPittau:2008 and use weakly informative priors for our logistic and ordered logistic regression models.
+For consistency with prior specification, and for computation efficiency, we mean-center all nonbinary variables so that parameter estimates represent changes from the mean.
+We use two general priors (see @fig-priors):
+
+- For all $\beta$ terms, we use a Student t distribution with a mean of 0 and a standard deviation of 3.
+This keeps most parameter estimates around −5 to 5, with thicker tails that allow for some possibility of extreme values.
+- For $\sigma$ terms related to the variance or standard deviation of parameter distributions, which must be positive, we use a half Cauchy distribution, centered at 0 with a $\gamma$ of 1
+
+These priors give more weight to realistic areas of parameter values and downweight values in unrealistic spaces.
+For instance, since logit-scale coefficient values greater than 4 or 5 are highly unlikely, our Student t prior puts more weight on smaller values.
+Additionally, weakly informative priors allow reasonable and considerable uncertainty in possible parameter estimates.
+
+```{r}
+#| label: fig-priors
+#| fig-cap: "Density plots of prior distributions for model parameters"
+#| fig-width: 6.5
+#| fig-height: 2.5
+#| out-width: "100%"
+#| include: true
+#| echo: false
+
+p1 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
+                fill = clrs[2]) +
+  xlim(c(-20, 20)) +
+  labs(x = "βs") +
+  facet_wrap(vars("β: Student t(ν = 1, µ = 0, σ = 3)")) +
+  theme_pandem(prior = TRUE)
+
+p2 <- ggplot() +
+  stat_function(geom = "area", 
+                fun = ~dcauchy(., 0, 1), 
+                fill = clrs[4]) +
+  xlim(c(0, 10)) +
+  labs(x = "σ") +
+  facet_wrap(vars("σ: Cauchy(x = 0, γ = 1), lower bound = 0")) +
+  theme_pandem(prior = TRUE)
+
+plot_priors <- p1 | p2
+plot_priors
+# ggsave("submissions/plot_priors.svg", plot_priors, width = 6.5, height = 2.5)
+```
 
@@
 
-### Logistic regression
+### H~1~: Logistic regression
+
+{{< include tikz/h1.md >}}
 
@@
 
-### Ordered logistic regression
+The actual R code for these models is included in the replication code at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW> This is a simplified representation of the {brms} [@Burkner:2017] model code:
+
+```r
+# H1: Logistic regression for binary outcomes
+brm(
+  bf(outcome ~ derogation_ineffect + 
+      new_cases_z + cumulative_cases_z + new_deaths_z + cumulative_deaths_z +
+      prior_iccpr_derogations + prior_iccpr_other_action + 
+      v2x_rule + v2x_civlib + v2xcs_ccsi + 
+      year_week_num + (1 | country_name)),
+  family = bernoulli(),
+  prior = c(
+    prior(student_t(1, 0, 3), class = Intercept),
+    prior(student_t(1, 0, 3), class = b),
+    prior(cauchy(0, 1), class = sd, lb = 0)),
+  ...
+)
+```
+
+### H~2~: Ordered logistic regression
+
+{{< include tikz/h2.md >}}
 
@@
 \begin{aligned}
-&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
+&\ \mathrlap{\textbf{Model of outcome level $i$ across quarter $t$ within each region $j$}} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
@@
 &\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
-&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
+&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Quarter number}_{it} \\
 b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
@@
 $$
+
+The actual R code for these models is included in the replication code at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW> This is a simplified representation of the {brms} [@Burkner:2017] model code:
+
+```r
+# H2: Ordinal logistic regression for ordered outcomes
+brm(
+  bf(outcome ~ derogation_ineffect + 
+      new_cases_z + cumulative_cases_z + new_deaths_z + cumulative_deaths_z +
+      prior_iccpr_derogations + prior_iccpr_other_action + 
+      v2x_rule + v2x_civlib + v2xcs_ccsi + 
+      year_quarter_num + (1 | who_region)),
+  family = cumulative(),
+  prior = c(
+    prior(student_t(1, 0, 3), class = Intercept),
+    prior(student_t(1, 0, 3), class = b),
+    prior(cauchy(0, 1), class = sd, lb = 0)),
+  ...
+)
+```
 
@@
 ```{r tbl-results-full-policies}
-#| tbl-cap: Complete results from models showing relationship between derogations and emergency policies (H~1~)
+#| tbl-cap: "Complete results from models showing relationship between derogations and emergency policies (H~1~)"
 #| include: true
@@
   mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
-  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
+  mutate(time = ifelse(year_quarter_day == "2020-01-01", "2020-Q1", "2021-Q2")) %>% 
   mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
@@
 tbl_part2 <- hr_table_data$hr_mfx_tbl %>% 
+  left_join(year_quarter_lookup, by = join_by(year_quarter_day)) %>% 
   mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
@@
   mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
-  select(nice, year_week_day, group, diff_type, num, p_gt_0) %>% 
-  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_week_day")) %>% 
+  select(nice, year_quarter, group, diff_type, num, p_gt_0) %>% 
+  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_quarter")) %>% 
   mutate(derogation_ineffect = "No")
@@
   select(
-    derogation_ineffect, group, `March 2020`, `June 2021`,
-    `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
-    `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`
+    derogation_ineffect, group, `2020-Q1`, `2021-Q2`,
+    `year_quarter_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
+    `year_quarter_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`
   ) %>%
@@
     c(
-      "Derogation", "Level", "March 2020", "June 2021",
+      "Derogation", "Level", "2020-Q2", "2021-Q2",
       "Week", "$\\Delta$", "p > 0", "Week", "$\\Delta$", "p > 0"

### 2024-04-30 manuscript/manuscript.qmd
Mega revisions for R2

@@
   - emergency policies
-bibliography: bib.bib
+bibliography: references.json
+csl: csl/chicago-author-date.csl
 link-citations: true
@@
   mutate(year_week_day = ymd(year_week_day))
+
+quarterly_panel <- tar_read(quarterly_panel) %>% 
+  mutate(year_quarter_day = ymd(year_quarter_day))
 
@@
 While there is a significant literature on compliance with international law, especially human rights treaties [e.g.
-@Simmons2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2005], we know little about states’ compliance with flexibility mechanisms such as derogations.
+@Simmons:2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock:2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2005], we know little about states’ compliance with flexibility mechanisms such as derogations.
 We explore how the potential decline in overall treaty compliance may actually be a good indication that states are using treaty design, and derogation option, as intended.
@@
 Previous derogation studies examine their use broadly, not during a defined global crisis.
-There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @Simmons2009; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
+There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @Simmons:2009; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
 While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
@@
 Existing research focuses on the determinants of submitting derogations to the UN rather than the effects of these derogations.
-It finds that democracies are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not issue derogations [@HBHelferFariss].
+It finds that democracies are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons:2009; @comstock2019] and that most states facing an emergency do not issue derogations [@HBHelferFariss].
 The former is because democracies typically need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
@@
   n_weeks = length(unique(weekly_panel$year_week)),
+  n_quarters = length(unique(quarterly_panel$year_quarter)),
   n_countries = length(unique(weekly_panel$iso3)),
+  n_regions = length(unique(weekly_panel$who_region)),
+  n_subregions = length(unique(quarterly_panel$who_subregion)),
   n_rows = label_comma()(nrow(weekly_panel)),
@@
 
-We use global data to examine the relationship between treaty derogations, human rights, and state actions during the pandemic—specifically, we use data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023].
+We use global data to examine the relationship between treaty derogations, human rights, and state actions during the pandemic—specifically, we use data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WorldHealthOrganization:2023].
 We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021.
@@
 We do not collapse these variables into binary indicators and instead work with them as ordered categories.
-Importantly, since these PanDem variables are all measured quarterly, while our main data is recorded weekly, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response each week.
 
@@
 
-Our data is structured as a balanced time series cross-sectional (TSCS) panel that has a natural hierarchical structure, with `r info$n_weeks` weekly observations nested within `r info$n_countries` countries.
-Accordingly, we model our outcome variables with multilevel or hierarchical models, with region-specific random offsets for each intercept and a fixed global year trend.
-When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, we use ordered logistic regression models.
-We take a Bayesian approach to better model and explore the uncertainty inherent in measuring the effect of derogations on these broad policies.
+Since our data is structured as a balanced time series cross-sectional (TSCS) panel, we model our outcomes with multilevel or hierarchical models, with country- or region-specific random offsets for each intercept and a fixed time trend.
+For our emergency measures outcomes (H~1~), there are `r info$n_weeks` weekly observations nested within `r info$n_countries` countries.
+For our human rights outcomes (H~2~), there are `r info$n_quarters` quarterly observations.
+These observations are also measured at a country level, but due to the slow moving nature of these PanDem human rights variables, there is not enough variation within countries to fit a model with country-level effects.
+Since geography plays an important role in pandemic, emergency policy, and human rights responses, we instead nest country-quarter observations within `r info$n_regions` WHO regions (see Appendix Table 1).
+Importantly, several of our variables move at different rates—some are measured weekly, others quarterly, others annually, and others (like country name) are time-invariant.
+One advantage of using multilevel models is that the model structure can flexibly handle these different levels of variation [@SingerWillett:2003], incorporating both time-varying and time-invariant predictors.
+For instance, the slow-moving annual V-Dem rule of law index effectively captures the overall level of a country's respect for rule of law across each week- or quarter-based observations, similar to a country-level fixed effect.
 
-We include complete formal specifications of all our models, likelihoods, and priors in the appendix.
+When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working with ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, we use ordered logistic regression models.
+Because the the panel data represents a complete or apparent sample of all countries [@BerkWesternWeiss:1995], we use a Bayesian approach to model the probabilities of the various outcomes and better estimate the uncertainty in the model parameters.
+We use weakly informative priors [@GelmanJakulinPittau:2008] for all model coefficients to keep estimates within realistic ranges and allow for reasonable uncertainty in parameter estimates.
+
+We include complete formal specifications and visualizations of all our models, likelihoods, and priors in the appendix.
 Here we include two simplified model specifications to help demonstrate the intuition behind these multilevel TSCS models.
-For binary outcomes, we use a logistic model with region-specific offsets (see @eq-logit), while for ordered outcomes, we use ordered logistic regression with region-specific offsets (see @eq-ordered-logit).
+For binary outcomes in H~1~, we use a logistic model with country-specific offsets and a weekly time trend (see @eq-logit), while for ordered outcomes in H~2~, we use ordered logistic regression with region-specific offsets and a quarterly time trend (see @eq-ordered-logit).[^fixed-note]
+
+[^fixed-note]: In these formulas, we refer to all the $\beta$ coefficients as "population averages and coefficients." The nomenclature for these terms is inconsistent across disciplines, though, and they are also often referred to as "global effects" or "fixed effects".
+Broadly speaking, these $\beta$ terms represent the effects of each variable on the outcome across all regions or countries and do not incorporate region- or country-specific differences.
 
@@
 \begin{aligned}
-&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within country $j$}} \\
+&\ \mathrlap{\mathbf{H_1} \textbf{: Binary outcome $i$ across week $t$ within country $j$}} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
@@
 &\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
-b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random region offsets} \\[0.75em]
+b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random country offsets} \\[0.75em]
 &\ \textbf{Priors} \\
-\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Global averages and effects} \\
-\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Between-region variability}
+\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Population averages and coefficients} \\
+\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Between-country variability}
 \end{aligned}
@@
 \begin{aligned}
-&\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within country $j$}} \\
+&\ \mathrlap{\mathbf{H_2} \textbf{: Outcome level $i$ across quarter $t$ within region $j$}} \\
 \text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
@@
 \phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + & \text{Cumulative probability of outcome} \\
-&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
+&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Quarter number}_{it} \\
 b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random region offsets} \\[0.75em]
 &\ \textbf{Priors} \\
-\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Global averages and effects} \\
+\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Population averages and effects} \\
 \sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Between-region variability} \\
@@
 
-Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in Appendix Tables 2 and 4).
+Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete results in Appendix Tables 3 and 5).
 Instead, we rely on probability-scale predictions and marginal effects whenever possible.
@@
 For logistic models, we estimate the probability of having a specific policy or violating a specific human right; for ordered models, we estimate the probability of each possible level of severity.
-To determine the effect of derogations, we calculate the contrast in predicted probabilities when derogation is set to true vs.
+To determine the average differences associated with derogations, we calculate the contrast in predicted probabilities when derogation is set to true vs.
 when it is set to false (e.g., in ordered logistic models we compare the probability of major discrimination under derogation with the probability of major discrimination without derogation).
@@
 Derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights).
-There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement.
-Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an `r hr_pred$discriminatory_policy.No.None.min$draw_nice`% predicted probability of no discriminatory policy in March 2020 rising to `r hr_pred$discriminatory_policy.No.None.max$draw_nice`% by June 2021.
+There is no measurable difference in the probability of violating non-derogable rights across derogating- and non-derogating-countries across the entire sample.
+However, in first 15 months of the pandemic, derogating countries were 4–5 percentage points more likely (`r hr_mfx$discriminatory_policy.None.max$p_gt`) to have no discriminatory policies than those that did not derogate, though the difference disappears by June 2021.
 The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.
-Moreover, countries that derogate were substantially less likely to engage in no abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic.
-Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.
+Moreover, countries that derogate were slightly less likely to not engage in abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic.
+Minor and moderate levels of excessive enforcement are each a few percentage points more likely for derogating countries than for their non-derogating counterparts, though with substantial uncertainty.
 
@@
       .data %>% 
+        mutate(year_quarter = case_when(
+          year_quarter_day == "2020-01-01" ~ "2020-Q1",
+          TRUE ~ year_quarter
+        )) %>% 
+        mutate(year_quarter = str_replace(year_quarter, "-", "\n")) %>% 
         ggplot(aes(
-          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
+          x = year_quarter, y = .epred, ymin = .lower, ymax = .upper,
           color = .category, fill = after_scale(color),
@@
         )) +
-        geom_lineribbon(alpha = 0.3) +
+        geom_line(aes(group = .category), linewidth = 0.25) +
+        geom_pointinterval(fatten_point = 1.5) +
         scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
         scale_y_continuous(labels = label_percent()) +
@@
       .data %>% 
+        mutate(year_quarter = case_when(
+          year_quarter_day == "2020-01-01" ~ "2020-Q1",
+          TRUE ~ year_quarter
+        )) %>% 
+        mutate(year_quarter = str_replace(year_quarter, "-", "\n")) %>% 
         ggplot(aes(
-          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
-          color = derogation_ineffect, fill = after_scale(color),
-          group = forcats::fct_rev(ordered(.width))
+          x = year_quarter, y = .epred, ymin = .lower, ymax = .upper,
+          color = derogation_ineffect, fill = after_scale(color)
         )) +
-        geom_lineribbon(alpha = 0.3) +
+        geom_line(aes(group = derogation_ineffect), linewidth = 0.25) +
+        geom_pointinterval(position = position_dodge(width = 0.5), fatten_point = 1.5) +
         scale_color_manual(values = c(clrs[1], clrs[8])) +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
         scale_y_continuous(labels = label_percent()) +
@@
       .data %>% 
+        mutate(year_quarter = case_when(
+          year_quarter_day == "2020-01-01" ~ "2020-Q1",
+          TRUE ~ year_quarter
+        )) %>% 
+        mutate(year_quarter = str_replace(year_quarter, "-", "\n")) %>% 
         mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
         ggplot(aes(
-          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+          x = year_quarter, y = draw, ymin = .lower, ymax = .upper,
           color = group, fill = after_scale(color),
@@
         )) +
-        geom_lineribbon(alpha = 0.3) +
+        geom_line(aes(group = group), linewidth = 0.25) +
+        geom_pointinterval(fatten_point = 1.5) +
         geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
         scale_color_manual(values = darken(clrs[c(2, 4, 6, 7)], 0.5, space = "combined"), guide = "none") +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
         labs(x = NULL,
@@
       .data %>% 
+        mutate(year_quarter = case_when(
+          year_quarter_day == "2020-01-01" ~ "2020-Q1",
+          TRUE ~ year_quarter
+        )) %>% 
+        mutate(year_quarter = str_replace(year_quarter, "-", "\n")) %>% 
         mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
         ggplot(aes(
-          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
+          x = year_quarter, y = draw, ymin = .lower, ymax = .upper,
           group = forcats::fct_rev(ordered(.width))
         )) +
-        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
+        geom_line(color = clrs[5], linewidth = 0.25) +
+        geom_pointinterval(color = clrs[5], fatten_point = 1.5) +
         geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
-        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
         labs(x = NULL,
@@
 
-The mechanisms behind this reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems.
+The mechanisms behind this reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way—may be attributable to principal-agent problems.
 While elites submit derogations and commit to complying with international law, they are not the actors who implement emergency measures.
@@
 The middle ground of only minor issues with time limits is exceptionally unlikely.
-Countries that derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only an `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues.
+Countries that derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only a `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues.
 The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
-Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights (`r hr_mfx$non_derogable_rights.Logit.max$p_lt`)
 
@@
 We thus find strong support only for H~2c~—that derogating states are less likely to impose measures with no time limits.
-We find weak support for our hypothesis that derogating states will be less likely to violate non-derogable rights (H~2b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
+We find no support for our hypothesis that derogating states will be less likely to violate non-derogable rights (H~2b~) and weak support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
 Derogations might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.

### 2024-04-30 R/funs_data-cleaning.R

@@ -5,7 +5,7 @@ library(countrycode)
 library(jsonlite)
 suppressPackageStartupMessages(library(sf))
 
-clean_iccpr_who <- function(path) {
+clean_iccpr_who <- function(path, subregions_path) {
   who_regions <- tribble(
     ~who_region, ~who_region_long,
     "AFRO", "Regional Office for Africa",
@@ -15,6 +15,48 @@ clean_iccpr_who <- function(path) {
     "EMRO", "Regional Office for the Eastern Mediterranean",
     "WPRO", "Regional Office for the Western Pacific"
   )
+  
+  who_subregions_manual <- tribble(
+    ~member_states, ~region_abbr, ~stratum, ~subregion_imputed,
+    "American Samoa", "WPR", "B", TRUE,
+    "Anguilla", "AMR", "B", TRUE,
+    "Aruba", "AMR", "B", TRUE,
+    "Bermuda", "AMR", "B", TRUE,
+    "British Virgin Islands", "AMR", "B", TRUE,
+    "Cayman Islands", "AMR", "B", TRUE,
+    "Curaçao", "AMR", "B", TRUE,
+    "French Guiana", "AMR", "B", TRUE,
+    "French Polynesia", "WPR", "B", TRUE,
+    "Kosovo", "EUR", "B", TRUE,
+    "Liechtenstein", "EUR", "A", TRUE,
+    "Palestinian Territories", "EMR", "B", TRUE,
+    "St. Barthélemy", "AMR", "B", TRUE,
+    "St. Helena", "AFR", "E", TRUE,
+    "St. Martin (French part)", "AMR", "B", TRUE,
+    "Sint Maarten", "AMR", "B", TRUE,
+    "Turks & Caicos Islands", "AMR", "B", TRUE
+  )
+  
+  who_subregions <- read_excel(subregions_path) %>% 
+    set_names(c("subregions", "member_states")) %>% 
+    separate(subregions, into = c("region_abbr", "stratum"), sep = " ") %>% 
+    mutate(member_states = str_split(member_states, "; ")) %>% 
+    unnest(member_states) %>% 
+    bind_rows(who_subregions_manual) %>% 
+    mutate(
+      who_region = paste0(region_abbr, "O"),
+      who_subregion = paste0(region_abbr, stratum)
+    ) %>% 
+    replace_na(list(subregion_imputed = FALSE)) %>% 
+    mutate(
+      country_name = countrycode(
[…]

### 2024-04-30 R/funs_models.R

@@ -139,15 +139,14 @@ f_human_rights <- function(panel) {
   # (1 | country_name) is included, the model predicts a 100% chance of 
   # pandem_discrim == 1 and a 0% chance of pandem_discrim == 0, which is
   # annoying (and it takes 45 minutes to run, ugh)
-  
-  # So instead we use region random effects instead? That only takes 10 minutes.
-  # Or no random effects at all?
+  #
+  # So instead we use region random effects
   human_rights_model <- function(y, family, prior) {
     form <- glue::glue(y, " ~ derogation_ineffect + new_cases_z + cumulative_cases_z + ",
                        "new_deaths_z + cumulative_deaths_z + ", 
                        "prior_iccpr_derogations + prior_iccpr_other_action + ",
                        "v2x_rule + v2x_civlib + v2xcs_ccsi + ",
-                       "year_week_num + (1 | who_region)") %>% 
+                       "year_quarter_num + (1 | who_region)") %>% 
       as.formula()
     
     # Use rlang::inject() to evaluate the formula object before running the model

### 2024-04-30 R/funs_plots.R

@@ -84,7 +84,7 @@ build_policies_plot_data <- function(model_df, year_week_lookup) {
         comparisons(newdata = datagrid(year_week_num = 1:69),
                     variables = "derogation_ineffect",
                     type = "response")  %>%  
-        posteriordraws() %>% 
+        posterior_draws() %>% 
         left_join(year_week_lookup, by = "year_week_num") %>% 
         group_by(year_week_day) %>% 
         reframe(post_medians = tidybayes::median_qi(draw, .width = c(0.5, 0.8, 0.95)),
@@ -97,54 +97,54 @@ build_policies_plot_data <- function(model_df, year_week_lookup) {
 }
 
 
-build_human_rights_plot_data <- function(model_df, year_week_lookup) {
+build_human_rights_plot_data <- function(model_df, year_quarter_lookup) {
   model_df_preds_mfx <- model_df %>% 
     mutate(pred_data = map2(model, family, ~{
       if (.y == "cumulative") {
         df <- datagrid(model = .x,
-                       year_week_num = 1:69,
+                       year_quarter_num = 1:6,
                        derogation_ineffect = 0:1) %>% 
           tidybayes::add_epred_draws(.x) %>% 
-          left_join(year_week_lookup, by = "year_week_num") %>% 
+          left_join(year_quarter_lookup, by = "year_quarter_num") %>% 
           mutate(derogation_ineffect = factor(derogation_ineffect, 
                                               levels = 0:1,
                                               labels = c("No derogation", "Derogation in effect"),
                                               ordered = TRUE)) %>% 
-          group_by(year_week_day, .category, derogation_ineffect) %>% 
+          group_by(year_quarter_day, year_quarter, .category, derogation_ineffect) %>% 
           tidybayes::median_qi(.epred, .width = c(0.5, 0.8, 0.95))
       } else {
         df <- datagrid(model = .x,
-                       year_week_num = 1:69,
+                       year_quarter_num = 1:6,
                        derogation_ineffect = 0:1) %>% 
           tidybayes::add_epred_draws(.x) %>% 
-          left_join(year_week_lookup, by = "year_week_num") %>% 
+          left_join(year_quarter_lookup, by = "year_quarter_num") %>% 
           mutate(derogation_ineffect = factor(derogation_ineffect, 
[…]

### 2024-04-30 _targets.R

@@ -43,6 +43,11 @@ list(
   tar_target(iccpr_who_raw_file,
              here_rel("data", "raw_data", "ICCPR Derogation and WHO case data 1 3 2020 to 6 30 2021.xlsx"),
              format = "file"),
+  # https://doi.org/10.1371/journal.pmed.1001923
+  # https://plos.figshare.com/articles/dataset/World_Health_Organization_WHO_Member_States_by_subregion_/2579569
+  tar_target(who_subregion_raw_file,
+             here_rel("data", "raw_data", "Havelaar et al Table 2.xls"),
+             format = "file"),
   tar_target(iccpr_treaty_action_file,
              here_rel("data", "raw_data", "ICCPR Treaty Action Variables.xlsx"),
              format = "file"),
@@ -62,7 +67,7 @@ list(
              format = "file"),
   
   ## Process and clean data ----
-  tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file)),
+  tar_target(iccpr_who_clean, clean_iccpr_who(iccpr_who_raw_file, who_subregion_raw_file)),
   tar_target(iccpr_action_clean, clean_iccpr_action(iccpr_treaty_action_file)),
   tar_target(oxford_clean, clean_oxford(oxford_raw_file)),
   tar_target(pandem_clean, clean_pandem(pandem_raw_file)),
@@ -77,8 +82,10 @@ list(
                              oxford_clean, pandem_clean, vdem_clean)),
   
   tar_target(weekly_panel, make_weekly_data(daily_panel)),
+  tar_target(quarterly_panel, make_quarterly_data(daily_panel)),
   
   tar_target(year_week_lookup, make_year_week_lookup(weekly_panel)),
+  tar_target(year_quarter_lookup, make_year_quarter_lookup(quarterly_panel)),
   
   tar_target(world_map, load_world_map(naturalearth_raw_file)),
   tar_target(derogation_count, make_derogation_count(iccpr_who_clean)),
@@ -128,7 +135,7 @@ list(
   
   ## Models ----
   tar_target(m_policies, f_policies(weekly_panel)),
-  tar_target(m_human_rights, f_human_rights(weekly_panel)),
+  tar_target(m_human_rights, f_human_rights(quarterly_panel)),
 
   ## Model tables ----
   # Build tables here because they take a while
@@ -143,7 +150,7 @@ list(
   ## Plots and tables ----
[…]

### 2024-04-30 manuscript/appendix.qmd
Use tikz PDF in PDFs

@@
 
-font_opts <- list(
+# Conditional tikz output types; use PDF if this is LaTeX, otherwise use SVG with outlined fonts
+if (fmt_out == "latex") {
+  knitr::opts_template$set(
+    tikz_settings = list(
+      fig.ext = "pdf", 
+      fig.align = "center", 
+      extra.preamble = c(
+        "\\usepackage{libertine}",
+        "\\usepackage{libertinust1math}"
+      )
+    )
+  )
+} else {
+  knitr::opts_template$set(
+    tikz_settings = list(
+      fig.ext = "svg", 
+      fig.align = "center", 
       extra.preamble = c(
@@
       ),
-  # dvisvgm.opts = "--font-format=woff"
-  dvisvgm.opts = "--no-fonts"
+      engine.opts = list(dvisvgm.opts = "--no-fonts")
     )
+  )
+}
 ```
@@
 plot_priors
-# ggsave("submissions/plot_priors.svg", plot_priors, width = 6.5, height = 2.5)
+# ggsave("submissions/figs/plot_priors.svg", plot_priors, width = 6.5, height = 2.5)
 ```

### 2024-04-30 manuscript/appendix.qmd
Update date

@@
   email: aheiss@gsu.edu
-date: April 18, 2024
+date: April 30, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2024-04-30 manuscript/manuscript.qmd
Update date

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: April 17, 2024
+date: April 30, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2024-05-02 manuscript/appendix.qmd
Address another reviewer concern

@@
 
-# WHO regions
-
-```{r tbl-who-regions}
-#| tbl-cap: WHO regions
-#| include: true
-#| echo: false
-who_regions <- who_regions_raw %>%
-  mutate(
-    region_title = glue::glue("{who_region}: {who_region_long}")
-  ) %>% 
-  group_by(region_title) %>% 
-  summarize(
-    countries = str_flatten(country_name, collapse = ", ", last = ", and ")
-  ) %>% 
-  pivot_longer(cols = everything()) %>% 
-  select(-name) %>% 
-  as.data.frame() %>%  # tibbles don't like NULL column names
-  set_names(NULL)
-
-who_regions %>% 
-  tt() %>% 
-  style_tt(i = seq(1, nrow(who_regions), by = 2), bold = TRUE, background = "#e6e6e6") %>% 
-  style_tt(i = seq(2, nrow(who_regions), by = 2), indent = 2)
+```{=latex}
+\listoftables
+\listoffigures
 ```
@@
   bf(outcome ~ derogation_ineffect + 
-      new_cases_z + cumulative_cases_z + new_deaths_z + cumulative_deaths_z +
+      new_cases_z + cumulative_cases_z + 
+      new_deaths_z + cumulative_deaths_z +
       prior_iccpr_derogations + prior_iccpr_other_action + 
@@
 
+```{r tbl-who-regions}
+#| tbl-cap: WHO regions
+#| include: true
+#| echo: false
+who_regions <- who_regions_raw %>%
+  mutate(
+    region_title = glue::glue("{who_region}: {who_region_long}")
+  ) %>% 
+  group_by(region_title) %>% 
+  summarize(
+    countries = str_flatten(country_name, collapse = ", ", last = ", and ")
+  ) %>% 
+  pivot_longer(cols = everything()) %>% 
+  select(-name) %>% 
+  as.data.frame() %>%  # tibbles don't like NULL column names
+  set_names(NULL)
+
+who_regions %>% 
+  tt(width = 1) %>% 
+  format_tt(escape = TRUE) %>% 
+  style_tt(i = seq(1, nrow(who_regions), by = 2), bold = TRUE, background = "#e6e6e6") %>% 
+  style_tt(i = seq(2, nrow(who_regions), by = 2), indent = 2, align = "l") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm",
+    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}"
+  )
+```
+
+\clearpage
+
 # References

### 2024-05-02 manuscript/manuscript.qmd
Address another reviewer concern

@@
 First, to test whether derogations were associated with emergency state responses, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
-Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] We collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.
+Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing stringency of emergency measures.[^severity] To allow for comparisons across outcomes, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.[^collapse]
 
 [^severity]: Importantly, the Oxford Government Response Tracker measures the *stringency* or strictness of policy enforcement, not the *quality* of policies, so the values do not necessarily reflect how well countries implemented these emergency measures.
+
+[^collapse]: Since values are not comparable across outcomes (e.g.
+some outcomes have five levels, some have three, and categories do not represent similar concepts across the outcomes), dichotomizing the outcome allows us to better compare changes in the probability of emergency measures across all outcomes.
 

### 2024-05-16 manuscript/appendix.qmd
Add Audrey edits

@@
   email: aheiss@gsu.edu
-date: April 30, 2024
+date: May 16, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 options(digits = 3, width = 120,
-        dplyr.summarise.inform = FALSE,
-        knitr.kable.NA = "")
+        dplyr.summarise.inform = FALSE)
 ```
@@
 library(scales)
-library(kableExtra)
-library(gt)
 library(tinytable)
@@
 
+```{r tbl-achr-echr-iccpr}
+#| tbl-cap: States that derogated from regional human rights conventions
+#| include: true
+#| echo: false
+
+achr_echr_iccpr <- tribble(
+                 ~ACHR, ~Also,             ~ECHR, ~Also1,
+           "Argentina", "Yes",         "Albania",   "No",
+             "Bolivia",  "No",         "Armenia",  "Yes",
+               "Chile", "Yes",      "Azerbaijan",  "Yes",
+            "Colombia", "Yes",         "Estonia",  "Yes",
+  "Dominican Republic", "Yes",         "Georgia",  "Yes",
+             "Ecuador", "Yes",          "Latvia",  "Yes",
+         "El Salvador", "Yes",         "Moldova",  "Yes",
+           "Guatemala", "Yes", "North Macedonia",   "No",
+            "Honduras",  "No",         "Romania",  "Yes",
+             "Jamaica",  "No",      "San Marino",  "Yes",
+              "Panama",  "No",          "Serbia",   "No",
+            "Paraguay", "Yes",                NA,     NA,
+                "Peru", "Yes",                NA,     NA,
+            "Suriname",  "No",                NA,     NA,
+           "Venezuela",  "No",                NA,     NA
+  )
+
+achr_echr_iccpr %>% 
+  as.data.frame() %>% 
+  set_names(
+    c(
+      "States that derogated from the ACHR",
+      "Also submitted ICCPR derogation",
+      "States that derogated from the ECHR",
+      "Also submitted ICCPR derogation"
+    )
+  ) %>% 
+  tt(
+    width = c(0.3, 0.2, 0.3, 0.2),
+    notes = "ACHR = American Convention on Human Rights; ECHR = European Convention on Human Rights"
+  ) %>% 
+  format_tt(replace = list(" " = NA)) %>% 
+  style_tt(j = 1:4, align = "lclc") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm",
+    tabularray_inner = "cells={font=\\scriptsize}, row{1}={valign=b}"
+  )
+```
+
+\clearpage
+
+```{r tbl-results-determinants}
+#| tbl-cap: Results from Authors (2024) showing the determinants of derogation
+#| include: true
+#| echo: false
+
+det_models <- readRDS(here::here("data", "raw_data",  "determinants_models.rds"))
+
+modelsummary(
+  det_models$models, 
+  exponentiate = TRUE,
+  estimate = "{estimate} ({std.error}){stars}",
+  statistic = NULL,
+  coef_map = det_models$coef_map,
+  gof_map = det_models$gof_map,
+  output = "tinytable",
+  fmt = fmt_significant(2),
+  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
+  notes = list(
+    "Dependent variable = derogation submission",
+    "Logistic regression models; odds ratios (standard error)",
+    "* p < 0.1, ** p < 0.05, *** p < 0.01"
+  ),
+  width = c(0.5, 0.25, 0.25)
+) %>% 
+  style_tt(
+    bootstrap_class = "table table-sm",
+    tabularray_inner = "cells={font=\\scriptsize}"
+  )
+```
+
+\clearpage
+
 # References

### 2024-05-16 manuscript/manuscript.qmd
Add Audrey edits

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: April 30, 2024
+date: May 16, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 It is among a few human rights treaties that require states to report on the suspension or derogation of civil and political rights only under specific situations of “public emergency which threatens the life of the nation” [@iccpr].
-The other treaties with this requirement include the European Convention Protection of Human Rights and Fundamental Freedoms and the American Convention on Human Rights.
-However, the ICCPR is the only one with global membership.
-As of 2023, 173 states are party to the ICCPR.
+The other treaties with this requirement include the European Convention on Human Rights (ECHR) and the American Convention on Human Rights (ACHR).
+Most of the states that derogated from the regional human rights conventions also derogated from the ICCPR, demonstrating that regional derogation activity did not preclude or replace ICCPR derogation activity.[^regional] The ICCPR is the only treaty with global membership---as of 2023, 173 states are party to the ICCPR.
 This means that most of the UN's 193 members had the ability to submit derogations.
+As such, we focus on ICCPR derogation activity in this research.[^future]
+
+[^regional]: See Table A5 for a list of countries that derogated from both the ACHR/ECHR and the ICCPR.
+
+[^future]: This project specifically examines what happens after derogations are filed.
+In a separate project by Authors (2024) we explore the determinants of derogations and find that (1) pandemic-era backsliding and (2) the severity of the pandemic as measured by national-level deaths are significant and positive indicators of pandemic-related derogation submission (see Table A6 in the appendix).
 
@@
 However, several other ICCPR articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).
+
+States follow different pathways to the actual submission of ICCPR derogations based on national level norms and policies.
+The UN does not specify the set pathway for national derogation approval.
+Most states followed a pathway of executive level instruction and authorization.
+For example, Argentina's derogations were described, upon submission to the UN as following an emergency decree issued by its executive branch [@un-derog-arg].
+Guatemala directly referenced the will of the President of the state in initiating the derogation policy [@un-derog-gtm].
+Some states such as the Dominican Republic, specified the role of the legislature in advancing the derogation provision [@un-derog-dom].
+Other states, such as Azerbaijan, referenced how the executive and legislature were involved in approving [@un-derog-aze].
+In reading through the written derogations, eighteen of the issuing states indicated that the executive or head of state directed the derogation, and five indicated the national legislature's role in approving it.
+All written derogations referenced a national declaration of emergency.
+
+Almost all states that issued derogations specified time limits, as the derogation format requires.
+Fifteen of the derogating states issued derogations that extended or renewed prior ones.
+The frequency of renewal signifies that states took the times specified within the derogations seriously and updated the treaty action when their national government expected the times of crises to continue.
+We understand the time specification to be a high level of precision about obligation, applying the dimensions of legalization [@AbbottKeohaneMoravcsik:2000] which contributes to a higher expectation of compliance, or adherence with the treaty.
 
@@
 However, derogations have no such public audience and are not signaling an increased commitment to human rights.
-There is no clear or expected connection with foreign aid, trade, or other theorized ‘rewards’ for larger legal actions.
+There is no clear or expected connection with foreign aid, trade, or other theorized 'rewards' for larger legal actions.
 Instead, derogations are a very specific legal action that states voluntarily enter into after commitment has fully taken place.
@@
 Of the derogations that indicated specific time limits, the average derogation duration was 41 days.
+Fifteen of the derogating states issued renewals/extensions of derogations that had time limits.
 We mark the start and end date of each derogation so that each country-week observation indicates whether a state has derogated.

### 2024-05-16 README.qmd
Stop using profiles; use pandoc extension modifiers!
Use pandoc extension modifiers to make different versions of the PDFs instead of profiles (https://github.com/quarto-dev/quarto-cli/issues/7371)
Use new manuscript target name

@@
 
-manuscript_target <- targets::tar_read(manuscript_nice)
+manuscript_target <- targets::tar_read(manuscript)
 
@@
 
-We use the [**renv** package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [**targets** package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis.
-([See this for a short helpful walkthrough of **targets**.](https://books.ropensci.org/targets/walkthrough.html)).
+We use the [{renv} package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [{targets} package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis.
+([See this for a short helpful walkthrough of {targets}.](https://books.ropensci.org/targets/walkthrough.html)).
 
@@
 5.
-If it's not installed already, R *should* try to install the **renv** package when you open the RStudio Project for the first time.
+If it's not installed already, R *should* try to install the {renv} package when you open the RStudio Project for the first time.
 If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
@@
 2. An analysis notebook website in `_site/`
-3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/` and `manuscript/output-ms`
+3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/`
 
@@
 
-Because project titles change all the time with revisions, rewriting, and peer review, we used [**codename**](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.
+Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.
 

### 2024-07-10 manuscript/appendix.qmd
Update date

@@
   email: aheiss@gsu.edu
-date: May 16, 2024
+date: July 10, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2024-07-10 manuscript/manuscript.qmd
Update date

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: May 16, 2024
+date: July 10, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2024-07-13 manuscript/appendix.qmd
Specify column
https://github.com/vincentarelbundock/tinytable/issues/306

@@
   tt(width = 1) %>% 
-  format_tt(escape = TRUE) %>% 
+  format_tt(j = 1, escape = TRUE) %>% 
   style_tt(i = seq(1, nrow(who_regions), by = 2), bold = TRUE, background = "#e6e6e6") %>% 

### 2024-07-20 manuscript/manuscript.qmd
Suparna edits

@@
 Responding to the pandemic required extraordinary public health measures.
-Many governments restricted freedom of movement, association, and assembly, raising questions about states’ commitments to international human rights law.
-Some human rights treaties recognize states’ ability to use emergency measures during crises, and contain “escape clauses” that safeguard the permanent erosion of human rights during and following emergencies.
+Many governments restricted freedom of movement, association, and assembly, raising questions about states' commitments to international human rights law.
+Some human rights treaties recognize states' need for emergency measures during crises, and contain “escape clauses” that safeguard the permanent erosion of rights during emergencies.
 The temporary suspension of state compliance from international treaties occurs through a legal action called derogation.
-However, the potential abuse of emergency measures during derogations raises concerns about their impact on human rights.
+However, the potential abuse of measures during derogations raises concerns about their impact on human rights.
 Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].
@@
 
-This research note explores state derogations and human rights behavior during the pandemic.
-Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, on May 5, 2023, the World Health Organization declared an end to the global health emergency.
-Many countries followed suit—the Biden administration, for instance, announced the end of United States’ emergency measures on May 11, 2023.
-We now assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency.
+Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, we assess important descriptive and analytical questions about states' respect for international law and human rights during this emergency.
 This research note examines what derogating meant for compliance with international law during the pandemic.
+Looking at the use of derogations during a public health emergency is especially useful as the countries that derogated from the main international human rights treaty—the International Covenant on Civil and Political Rights (ICCPR)—do not appear to coincide with much of the conventional wisdom about which states are mostly likely to abide by international human rights treaty commitments.
+While countries with strong democracies and presence of international NGOs [@Neumayer:2005], or high state capacity and bureaucratic efficacy [@Cole:2015], or strong domestic and judicial institutions [@Hathaway:2002; @Simmons:2009; @PowellStaton:2009] are typically more likely to comply with human rights treaties, @tbl-derogation-countries shows that this was not exactly the set of countries that utilized derogations during the pandemic.
+Moreover, even these derogation patterns do not align with previous research that finds that democracies are most likely to derogate [@HBHelferFariss].
+These descriptive statistics highlight the uniqueness in the conditions under which states comply with their international human rights treaty commitments and when they decide to derogate from them, especially in public health emergencies.
+This question is thus important for the study of international law, compliance, and human rights during times of crises.
+
+```{r tbl-derogation-countries}
+#| tbl-cap: Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures 
+#| include: true
+#| echo: false
+#| message: false
+
+bind_cols(dplyr::slice(map_data_nice, 1:12),
+          add_row(dplyr::slice(map_data_nice, 13:n()))) %>%
+  set_names(c("Country", "Derogations", "Country", "Derogations")) %>% 
+  tt(width = c(0.35, 0.15, 0.35, 0.15) * 0.9) %>% 
+  format_tt() %>% 
+  style_tt(j = 1:4, align = "lrlr") %>% 
+  style_tt(
+    bootstrap_class = "table table-sm table-striped",
+    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}"
+  )
+```
 
@@
 How did derogating and non-derogating states vary in their implementation of emergency measures and human rights practices?
-We find that states that derogated and specified time limits to these emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights (the right to life, freedom from torture, and freedom from slavery).
-States that did not derogate from their international legal obligations and still imposed emergency measures were more likely to implement discriminatory policies, violate non-derogable rights, and have emergency measures without time limits.
+We find that states that derogated generally stuck to the time limits of emergency measures and adhered to required standards by avoiding discriminatory policies and not violating non-derogable rights (the right to life, freedom from torture, and freedom from slavery).
+States that did not derogate and still imposed emergency measures were more likely to implement discriminatory policies, violate non-derogable rights, and have emergency measures without time limits.
 
-This research note makes two contributions.
+This research note makes three contributions.
 First, it expands our understanding of compliance with flexibility mechanisms and derogations.
 While there is a significant literature on compliance with international law, especially human rights treaties [e.g.
-@Simmons:2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock:2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2005], we know little about states’ compliance with flexibility mechanisms such as derogations.
-We explore how the potential decline in overall treaty compliance may actually be a good indication that states are using treaty design, and derogation option, as intended.
-
-The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster, typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally.
-In making this comparison, this research note makes an important descriptive contribution showing how the implementation of policy measures differed across derogating and non-derogating states during the pandemic.
+@Simmons:2009; @Cole:2012; @NielsenSimmons:2015; @ZvobgoSandholtzMulesky:2020; @Comstock:2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2005], we know much less about states' compliance with flexibility mechanisms such as derogations.
+However, compliance with derogations may not necessarily be the same as compliance with treaty obligations.
+Treaty compliance is often motivated by international-level factors or concerns such as reciprocity, retaliation, signaling, and reputations [@SimmonsHopkins:2005; @Hathaway:2007; @Hafner-BurtonLeVeckVictor:2017].
+However, countries derogate primarily for domestic political or institutional reasons [@HBHelferFariss], and international-level factors were relatively unimportant during the pandemic where preventing the domestic transmission of a virus and restricting citizens' rights to this end were governments' main motivations to derogate or not.
+As such, compliance with derogations is distinct from compliance with treaties and the former is relevant to our understanding of the latter—without derogations, states may repress more rights during crises, including non-derogable rights.
+Analyzing the behavior of derogating states and if it systematically differs from those that do not has important implications for the design of international treaties and whether derogations actually provide the safety valves that governments need during crises.
 
 Second, this study contributes to the study of the relationship between international law and strategic repression by examining derogation behavior during a potentially tempting time for states to misuse them to increase repression and power during a time of global crisis.
-Previous derogation studies examine their use broadly, not during a defined global crisis.
-There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @Simmons:2009; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
+There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @SimmonsHopkins:2005; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
 While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
 @Lebret:2020; @helfer2021].
-Our findings are informative for international law, human rights, and repression.
-We find that derogating states during the pandemic did not increase abuse and violations.
+We find that overall these concerns are overstated, and derogating states during the pandemic did not increase abuse and violations.
 This supports the idea that states are being sincere about their intentions to comply with international law and not, generally, using actions like derogation to cover abusive behavior.
-This is an important finding for treaty design, including flexibility mechanisms, and the relationship between exploitation of international law to cover abuse.
+This is an important finding for the relationship between treaty design exploitation of international law to cover abuse.
+
+Finally, this research note looks at an understudied area in which derogations happen—public health emergencies.
+Derogations during the pandemic do not appear to hold the same patterns as found in previous literature according to which stable democracies and countries where domestic courts can exercise strong oversight of the executive are most likely to derogate.
+In addition, democracies are the most serial derogators in response to grave external threats such as terrorism [@HBHelferFariss].
+However, as @tbl-derogation-countries shows, this was not the case during the pandemic.
+The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster, typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally.
+Thus, this research note makes an important descriptive contribution showing how the implementation of policy measures differed across derogating and non-derogating states during the pandemic.
 
@@
 We focus on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
-We then introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state’s human rights practices during the pandemic.
+We then introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic.
 We conclude with the implications of our research note for the study of international law during times of crises.
@@
 Derogations in international law authorize states to temporarily suspend their treaty commitments, providing them with flexibility to respond to crises [@HBHelferFariss].
-Times of crises can include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
+These crises can include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies.
 When states derogate from their international legal obligations, it provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions.
@@
 Many legal assessments of derogations find importance in their use [@richardson2020], with some arguing that, “When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97].
-An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a *de jure* emergency [@Greene:2020, 36].
-Research supports this notion—there is no meaningful variation in states’ respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
-Even states that officially derogate from international treaties are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].
+An abuse of state power can also potentially lead to the violation of non-derogable rights [@Greene:2020, 36].
+Research supports this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013].
+Even states that derogate from international treaties are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].
 
@@
 Existing research focuses on the determinants of submitting derogations to the UN rather than the effects of these derogations.
-It finds that democracies are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons:2009; @comstock2019] and that most states facing an emergency do not issue derogations [@HBHelferFariss].
+Democracies are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons:2009; @comstock2019]; most states facing an emergency do not issue derogations [@HBHelferFariss].
 The former is because democracies typically need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies.
@@
 
-```{r tbl-derogation-countries}
-#| tbl-cap: Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures 
-#| include: true
-#| echo: false
-#| message: false
-
-bind_cols(dplyr::slice(map_data_nice, 1:12),
-          add_row(dplyr::slice(map_data_nice, 13:n()))) %>%
-  set_names(c("Country", "Derogations", "Country", "Derogations")) %>% 
-  tt(width = c(0.35, 0.15, 0.35, 0.15) * 0.9) %>% 
-  format_tt() %>% 
-  style_tt(j = 1:4, align = "lrlr") %>% 
-  style_tt(
-    bootstrap_class = "table table-sm table-striped",
-    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}"
-  )
-```
-
-Between January 2020 and June 2021, `r derog_detail$countries` states issued `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
+Between January 2020 and June 2021, we find that `r derog_detail$countries` states issued `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries).
 The average length of a derogation was 41 days.
 States across multiple regions derogated, with notable concentration in South America, which has a robust history of engaging with human rights law.
-Notably, Nordic states—which typically are very active in human rights law engagement—did not issue any derogations during this period.
+Notably, Nordic states—which typically are very active in human rights law engagement—did not issue any derogations.
 
@@
 
-The ICCPR is a foundational human rights treaty with broad human rights coverage.
-It is among a few human rights treaties that require states to report on the suspension or derogation of civil and political rights only under specific situations of “public emergency which threatens the life of the nation” [@iccpr].
+The ICCPR is a foundational human rights treaty and among the few treaties that requires states to report on the suspension or derogation of civil and political rights only under specific situations of “public emergency which threatens the life of the nation” [@iccpr].
 The other treaties with this requirement include the European Convention on Human Rights (ECHR) and the American Convention on Human Rights (ACHR).
 Most of the states that derogated from the regional human rights conventions also derogated from the ICCPR, demonstrating that regional derogation activity did not preclude or replace ICCPR derogation activity.[^regional] The ICCPR is the only treaty with global membership---as of 2023, 173 states are party to the ICCPR.
-This means that most of the UN's 193 members had the ability to submit derogations.
-As such, we focus on ICCPR derogation activity in this research.[^future]
+Thus, most of the UN's 193 members had the ability to submit derogations and we focus on ICCPR derogation activity in this research.[^future]
 
@@
 
-To derogate, governments must first officially declare a state of emergency (Article 4).
+To derogate, governments must first declare a state of emergency (Article 4).
 Derogations must fulfill three criteria: they should be temporary, proportional to the threat and non-discriminatory.
@@
 
-States follow different pathways to the actual submission of ICCPR derogations based on national level norms and policies.
-The UN does not specify the set pathway for national derogation approval.
+States follow different pathways to the actual submission of ICCPR derogations based on national level norms and policies; the UN does not specify a uniform procedure.
 Most states followed a pathway of executive level instruction and authorization.
-For example, Argentina's derogations were described, upon submission to the UN as following an emergency decree issued by its executive branch [@un-derog-arg].
+For example, Argentina's derogations were described, upon submission to the UN, as following an emergency decree issued by its executive branch [@un-derog-arg].
 Guatemala directly referenced the will of the President of the state in initiating the derogation policy [@un-derog-gtm].
 Some states such as the Dominican Republic, specified the role of the legislature in advancing the derogation provision [@un-derog-dom].
-Other states, such as Azerbaijan, referenced how the executive and legislature were involved in approving [@un-derog-aze].
+Others, such as Azerbaijan, referenced how the executive and legislature were involved in approving [@un-derog-aze].
 In reading through the written derogations, eighteen of the issuing states indicated that the executive or head of state directed the derogation, and five indicated the national legislature's role in approving it.
@@
 
-Almost all states that issued derogations specified time limits, as the derogation format requires.
-Fifteen of the derogating states issued derogations that extended or renewed prior ones.
-The frequency of renewal signifies that states took the times specified within the derogations seriously and updated the treaty action when their national government expected the times of crises to continue.
+Almost all states that issued derogations specified time limits.
+Fifteen states extended or renewed prior derogations.
+The frequency of renewal signifies that states took the time specified within the derogation seriously and updated the treaty action when their national government expected the times of crises to continue.
 We understand the time specification to be a high level of precision about obligation, applying the dimensions of legalization [@AbbottKeohaneMoravcsik:2000] which contributes to a higher expectation of compliance, or adherence with the treaty.
@@
 
-> **Hypothesis 1~a~**: If states issued COVID-19 derogations, then they will be more likely to impose emergency policies that cancel public events, restrict gatherings, close public transit, and restrict movement.
+> **Hypothesis 1~a~**: If states derogate from the ICCPR, then they will be more likely to impose emergency policies that cancel public events, restrict gatherings, close public transit, and restrict movement.
 >
-> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, then they will be less likely to impose relevant emergency policies.
+> **Hypothesis 1~b~**: If states did not derogate from the ICCPR, then they will be less likely to impose relevant emergency policies.
 
@@
 For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021].
-In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country’s new citizenship law seen as discriminatory towards Muslims [@bbc2020].
+In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country's new citizenship law seen as discriminatory towards Muslims [@bbc2020].
 
@@
 For instance, in March 2020, a French government order extended the duration of pre-trial detention, raising the issue of prison overcrowding and leading to inhumane and degrading treatment [@Lebret:2020].
-Besides violating a non-derogable right, this also raised questions regarding the proportionality of emergency measures.
 
 We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to violate these criteria.
+However, we posit that derogations may be associated with fewer discriminatory policies due to two potential mechanisms.
+First, derogating states may be more sincere about their treaty commitments, which may have motivated them to derogate in the first place.
+Second, states that derogate also invite increased domestic and international monitoring, leading to more rights-respecting behavior.
+The UN Human Rights Committee (UNHRC) reviews derogations, may challenge them, and also examines complaints filed by individuals.[^individuals] It can also lead to involvement of domestic stakeholders in the process; their monitoring can lead to increased incentive for the government to make sure they are not violating the terms of the derogations.
+For instance, after Armenia derogated from the ICCPR and the ECHR, civil society and media representatives provided feedback on emergency measures affecting them—a move that was welcomed both by the OSCE Media Freedom Representative and the Armenian media [@OSCE:2020].
+Further, for regional treaties not examined in this paper, such as the ECHR, individuals can also seek recourse at the European Court of Human Rights.
+Thus, there could be sincere as well as strategic reasons derogating states may behave differently from non-derogating states.
 
-> **Hypothesis 2~a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
+[^individuals]: Challenges to derogations by individuals and as part of the review of state reports is quite common [@Nowak:2005].
+
+> **Hypothesis 2~a~**: If states derogate from the ICCPR, then they are less likely to use discriminatory policy measures during the pandemic.
 >
-> **Hypothesis 2~b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
+> **Hypothesis 2~b~**: If states derogate from the ICCPR, then they are less likely to violate non-derogable rights.
 >
-> **Hypothesis 2~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.
+> **Hypothesis 2~c~**: If states derogate from the ICCPR, then they are less likely to impose measures with no time limits.
 
@@
 
-Next, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
+Next, we explore the relationship between formal treaty derogations and four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether emergency measures were discriminatory, (2) whether countries derogated from non-derogable rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
 Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations.
@@
 
-We use a binary indicator of whether a state was under ICCPR derogation conditions during a given week.
-We collected and coded this data from real time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023].
+We use a binary indicator of whether a state derogated from the ICCPR during a given week.
+We coded this data from real-time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023].
 During the 15 months covered in this project, there were 115 formal derogations submitted mentioning 8 specific ICCPR articles; 15 derogations made no reference to any specific article.
@@
 
-Several additional variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions.
+Several additional variables that are associated with and potentially confound both the decision to derogate and to enact specific policies.
 To account for the severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
@@
 When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working with ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, we use ordered logistic regression models.
-Because the the panel data represents a complete or apparent sample of all countries [@BerkWesternWeiss:1995], we use a Bayesian approach to model the probabilities of the various outcomes and better estimate the uncertainty in the model parameters.
+Because the panel data represents a complete or apparent sample of all countries [@BerkWesternWeiss:1995], we use a Bayesian approach to model the probabilities of the various outcomes and better estimate the uncertainty in the model parameters.
 We use weakly informative priors [@GelmanJakulinPittau:2008] for all model coefficients to keep estimates within realistic ranges and allow for reasonable uncertainty in parameter estimates.
@@
 
-[^fixed-note]: In these formulas, we refer to all the $\beta$ coefficients as "population averages and coefficients." The nomenclature for these terms is inconsistent across disciplines, though, and they are also often referred to as "global effects" or "fixed effects".
-Broadly speaking, these $\beta$ terms represent the effects of each variable on the outcome across all regions or countries and do not incorporate region- or country-specific differences.
+[^fixed-note]: In these formulas, we refer to all the coefficients as "population averages and coefficients." The nomenclature for these terms is inconsistent across disciplines, though, and they are also often referred to as "global effects" or "fixed effects".
+Broadly speaking, these terms represent the effects of each variable on the outcome across all regions or countries and do not incorporate region- or country-specific differences.
 
@@
 @fig-preds-mfx-policies-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic.
-Countries with a derogation in effect are more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
+Derogating countries are more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
 This provides strong support for both components of our first hypothesis—states that derogated were substantially more likely to impose relevant emergency policies aimed at curtailing the spread of the virus (H~1a~), while states that did not issue derogations were less likely to engage in similar policies (H~1b~).
@@
 All countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance.
-Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021.
+Countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021.
 Countries that derogated, however, regularly canceled with a nearly 100% probability.
@@
 
-These results show that countries implementing emergency measures seem to have taken the time to follow international law and uphold their treaty obligations.
-International law appears to not just be an empty commitment—states that derogated were more likely to keep these emergency measures in place for longer and take their derogations seriously.
+These results show that countries implementing emergency measures took the time to uphold their treaty obligations.
+International law was not just an empty promise—statse that derogated were more likely to keep these emergency measures in place for longer and take their derogations seriously.
 
@@
 The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.
-Moreover, countries that derogate were slightly less likely to not engage in abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic.
-Minor and moderate levels of excessive enforcement are each a few percentage points more likely for derogating countries than for their non-derogating counterparts, though with substantial uncertainty.
+Thus, contrary to our expectation, derogating states also violate non-derogable rights and implement some discriminatory policies.
+These findings can perhaps be attributed to increased domestic and international monitoring—since derogating states open themselves up to such monitoring and review by the UNHRC, we may be more likely to observe discrimination and non-derogable rights violations than we would in non-derogating states.
 
@@
 
-The mechanisms behind this reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way—may be attributable to principal-agent problems.
+The mechanisms behind this reversal—where countries that derogate end up enforcing emergency policies in a more abusive way—may be attributable to principal-agent problems.
 While elites submit derogations and commit to complying with international law, they are not the actors who implement emergency measures.
 There may be immense variation in states' abilities to enforce these policies as intended, especially as it relates to human rights.
-Lower-level officials such as those in the police, state security forces and regional bureaucrats often enforce emergency measures.
+Lower-level officials such as those in the police, state security forces, and regional bureaucrats often enforce emergency measures.
 Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations.
@@
 Countries that derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only a `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues.
-The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
+Countries that care about their treaty obligations and derogate appear to actually follow the time limits they establish.
 
-Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abusive enforcement of those policies.
+Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights.
+It is possible that if states already intended to implement limited emergency measures that respect the spirit of the treaty, then it is relatively low cost to submit a derogation.
+However, compliance with derogations may not necessarily be low cost—as part of UN Human Rights Committee reports, even individuals can file complaints.
+Further, for regional treaties not examined in this paper, such as the ECHR, individuals can also seek recourse to the European Court of Human Rights.
+Thus, sincere behavior and strategic behavior to avoid misusing derogations may not be observationally equivalent in all cases.
+It is also worth noting, derogations do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abusive enforcement of those policies.
 We thus find strong support only for H~2c~—that derogating states are less likely to impose measures with no time limits.
 We find no support for our hypothesis that derogating states will be less likely to violate non-derogable rights (H~2b~) and weak support for our hypothesis that derogations protect against discriminatory policies (H~2c~).
-Derogations might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.
+Derogations might protect against severe violations of human rights, but they do not appear to influence the entire implementation of emergency measures.
 
@@
 The COVID-19 pandemic raised concerns about human rights violations worldwide.
-This research note examines state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations and state repression.
+This research note examines state engagement with flexibility mechanisms in international human rights treaties, particularly the ICCPR, during the pandemic and any subsequent impact on human rights violations.
 We focused specifically on the impact of derogations—an international legal means to communicate to both the UN and the global community that certain rights will be suspended during times of crisis.
@@
 Most derogations included time limitations, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate.
-Though most states did not submit COVID-19 related derogations, the ones that did, used derogations as intended by international law.
+Though most states did not submit COVID-19-related derogations, the ones that did, used derogations as intended by international law.
 Once the derogation ended, states readjusted to typical compliance once more.
-States therefore complied with the intent of derogations and the purpose of the treaty.
+States therefore complied with the intent of derogations.
 This is an important finding for international legal behavior and compliance studies.
@@
 non-derogable rights.
-This distinction may explain some variation in policy and discrimination practices during the pandemic.
-We also recommend that future research analyze the impact of other factors that can mitigate the adverse impacts of emergency measures.
-In the absence of adequate review from the international community and treaty monitoring bodies during crises, certain domestic-level factors, such as a robust civil society and media, may be necessary to mitigate the adverse effects of emergency measures.
+This distinction may explain some variation in policy and discrimination practices during the pandemic and points to the need to examine not just the structure of flexibility mechanisms, but also their use in practice.
+We also recommend that future research analyze the impact of other factors, such as a robust civil society and media, that can mitigate the adverse impacts of emergency measures.
 
-Overall, we find that states used derogations as they were intended during the COVID-19 pandemic and no large-scale misuse of the legal action took place to repress human rights.
+Overall, we find that states used derogations as they were intended during the pandemic and no large-scale misuse of the legal action took place to repress human rights.
 It can be an important signal for states to submit derogations, both to international and domestic communities, about intentions to mitigate the spread of disease through the temporary suspension of human rights.

### 2024-07-20 manuscript/appendix.qmd
yay nicer format output names

@@
   email: aheiss@gsu.edu
-date: July 10, 2024
+date: July 20, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2024-07-20 manuscript/manuscript.qmd
yay nicer format output names

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: July 10, 2024
+date: July 20, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"

### 2024-07-20 analysis/data-details.qmd
unbritishize spelling

@@
 
-- **Cancel public events** (`c3_cancel_events`): Indicator of whether the country cancelled public events.
-  - **0**: no measures; **1**: recommend cancelling; **2**: require cancelling; **Blank**: no data
+- **Cancel public events** (`c3_cancel_events`): Indicator of whether the country canceled public events.
+  - **0**: no measures; **1**: recommend canceling; **2**: require canceling; **Blank**: no data
 

### 2024-07-20 manuscript/manuscript.qmd
Audrey edits

@@
 
-Did states misuse international emergency provisions during the COVID-19 crisis to justify human rights abuse or did they follow international law?
+Did states misuse international emergency provisions during the COVID-19 crisis to justify human rights abuse?
 Responding to the pandemic required extraordinary public health measures.
 Many governments restricted freedom of movement, association, and assembly, raising questions about states' commitments to international human rights law.
-Some human rights treaties recognize states' need for emergency measures during crises, and contain “escape clauses” that safeguard the permanent erosion of rights during emergencies.
-The temporary suspension of state compliance from international treaties occurs through a legal action called derogation.
-However, the potential abuse of measures during derogations raises concerns about their impact on human rights.
+Some treaties recognize states' need for emergency measures and contain “escape clauses” that safeguard the permanent erosion of rights during emergencies.
+The temporary suspension of compliance from international treaties occurs through a legal action called derogation.
+However, potential abuse during derogations raises concerns about their impact on human rights.
 Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice [@Stasavage2020; @LuhrmannRooney2021].
@@
 
-Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, we assess important descriptive and analytical questions about states' respect for international law and human rights during this emergency.
-This research note examines what derogating meant for compliance with international law during the pandemic.
+Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, this research note examines what derogating meant for compliance with international law during the pandemic.
 Looking at the use of derogations during a public health emergency is especially useful as the countries that derogated from the main international human rights treaty—the International Covenant on Civil and Political Rights (ICCPR)—do not appear to coincide with much of the conventional wisdom about which states are mostly likely to abide by international human rights treaty commitments.
@@
 Moreover, even these derogation patterns do not align with previous research that finds that democracies are most likely to derogate [@HBHelferFariss].
-These descriptive statistics highlight the uniqueness in the conditions under which states comply with their international human rights treaty commitments and when they decide to derogate from them, especially in public health emergencies.
+Descriptive statistics highlight the uniqueness in the conditions under which states comply with their international human rights treaty commitments and when they decide to derogate from them, especially in public health emergencies.
 This question is thus important for the study of international law, compliance, and human rights during times of crises.
@@
 
-Second, this study contributes to the study of the relationship between international law and strategic repression by examining derogation behavior during a potentially tempting time for states to misuse them to increase repression and power during a time of global crisis.
-There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Neumayer:2005; @SimmonsHopkins:2005; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
-While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
+Second, this research contributes to the study of the relationship between international law and strategic repression by examining derogation behavior during a potentially tempting time for states to misuse them through repression during global crisis.
+There is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights [@Neumayer:2005; @SimmonsHopkins:2005; @Merry:2009; @Hillebrecht:2014; @ConradRitter:2019].
+While derogations are intended to ensure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g.
 @Lebret:2020; @helfer2021].
-We find that overall these concerns are overstated, and derogating states during the pandemic did not increase abuse and violations.
-This supports the idea that states are being sincere about their intentions to comply with international law and not, generally, using actions like derogation to cover abusive behavior.
-This is an important finding for the relationship between treaty design exploitation of international law to cover abuse.
+We find that overall these concerns are overstated.
+Derogating states during the pandemic did not increase abuse and violations.
+This supports the idea that states are generally sincere about intentions to comply with international law and not, generally, using actions like derogation to cover abuse.
+This is an important finding for the relationship between treaty design and exploitation of flexibility mechanisms to conceal violations.
 
 Finally, this research note looks at an understudied area in which derogations happen—public health emergencies.
-Derogations during the pandemic do not appear to hold the same patterns as found in previous literature according to which stable democracies and countries where domestic courts can exercise strong oversight of the executive are most likely to derogate.
-In addition, democracies are the most serial derogators in response to grave external threats such as terrorism [@HBHelferFariss].
+Derogations during the pandemic do not appear to hold the same patterns as found in previous literature.
+Prior research finds that stable democracies and countries where domestic courts can exercise strong oversight of the executive are most likely to derogate and that derogations are made in response to grave external threats such as terrorism [@HBHelferFariss].
 However, as @tbl-derogation-countries shows, this was not the case during the pandemic.
@@
 Below, we first explore the purpose of derogations in international law, and then specifically during the pandemic.
-We focus on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime.
+We focus on derogations to the ICCPR—arguably the most significant and extensive treaty in the international human rights regime.
 We then introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic.
@@
 What kinds of states are more likely to submit derogations in the first place?
-Existing research focuses on the determinants of submitting derogations to the UN rather than the effects of these derogations.
+Existing research focuses on the determinants of submitting derogations to the UN rather than the effects of derogations.
 Democracies are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons:2009; @comstock2019]; most states facing an emergency do not issue derogations [@HBHelferFariss].
@@
 Scholars were split on expectations about COVID-era derogations.
-@Lebret:2020 argued that derogations were the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic.
+@Lebret:2020 argued that derogations were the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the pandemic.
 However, @richardson2020 saw derogations as a less clear path for states during COVID-19, describing the actions and human rights as a “tangled morass” (p.
@@
 
-## Derogations from the ICCPR
+## ICCPR derogations
 
-The ICCPR is a foundational human rights treaty and among the few treaties that requires states to report on the suspension or derogation of civil and political rights only under specific situations of “public emergency which threatens the life of the nation” [@iccpr].
+The ICCPR is a foundational human rights treaty and among the few treaties that requires states to report on the suspension, or derogation of, civil and political rights which is allowable only under specific situations of “public emergency which threatens the life of the nation” [@iccpr].
 The other treaties with this requirement include the European Convention on Human Rights (ECHR) and the American Convention on Human Rights (ACHR).
-Most of the states that derogated from the regional human rights conventions also derogated from the ICCPR, demonstrating that regional derogation activity did not preclude or replace ICCPR derogation activity.[^regional] The ICCPR is the only treaty with global membership---as of 2023, 173 states are party to the ICCPR.
+Most states that derogated from the regional human rights conventions also derogated from the ICCPR, demonstrating that regional derogation activity did not preclude or replace ICCPR derogation activity.[^regional] The ICCPR is the only treaty with global membership---as of 2023, 173 states are party to the ICCPR.
 Thus, most of the UN's 193 members had the ability to submit derogations and we focus on ICCPR derogation activity in this research.[^future]
@@
 
-States follow different pathways to the actual submission of ICCPR derogations based on national level norms and policies; the UN does not specify a uniform procedure.
-Most states followed a pathway of executive level instruction and authorization.
+States follow different pathways to the actual submission of ICCPR derogations based on national-level norms and policies; the UN does not specify a uniform procedure.
+Most states followed a pathway of executive-level instruction and authorization.
 For example, Argentina's derogations were described, upon submission to the UN, as following an emergency decree issued by its executive branch [@un-derog-arg].
@@
 Some states such as the Dominican Republic, specified the role of the legislature in advancing the derogation provision [@un-derog-dom].
-Others, such as Azerbaijan, referenced how the executive and legislature were involved in approving [@un-derog-aze].
+Others, such as Azerbaijan, highlighted how executive and legislative approval were involved in derogating [@un-derog-aze].
 In reading through the written derogations, eighteen of the issuing states indicated that the executive or head of state directed the derogation, and five indicated the national legislature's role in approving it.
@@
 Instead, derogations are a very specific legal action that states voluntarily enter into after commitment has fully taken place.
-Given this, we expect that derogating states took the legal action seriously and generally followed through with the derogations they submitted.
+Given this, we expect that derogating states took the legal action seriously and generally followed through with submitted derogations.
 
@@
 
-If states determine that emergency measures are justified without regard for international law, the processes for respecting rights during crises and restoring rights at the end of the emergency may not be followed.
+If states determine emergency measures are justified without regard for international law, the processes for respecting rights during crises and restoring rights at the end of the emergency may not be followed.
 Many states used the pandemic as an excuse to centralize authority.
@@
 Many other states changed rules of confinement affecting detainees, which also potentially violated non-derogable rights.
-For instance, in March 2020, a French government order extended the duration of pre-trial detention, raising the issue of prison overcrowding and leading to inhumane and degrading treatment [@Lebret:2020].
+In March 2020, a French government order extended the duration of pre-trial detention, raising the issue of prison overcrowding and leading to inhumane and degrading treatment [@Lebret:2020].
 
@@
 Second, states that derogate also invite increased domestic and international monitoring, leading to more rights-respecting behavior.
-The UN Human Rights Committee (UNHRC) reviews derogations, may challenge them, and also examines complaints filed by individuals.[^individuals] It can also lead to involvement of domestic stakeholders in the process; their monitoring can lead to increased incentive for the government to make sure they are not violating the terms of the derogations.
+The UN Human Rights Committee (UNHRC) reviews derogations, may challenge them, and also examines complaints filed by individuals.[^individuals] It can also lead to involvement of domestic stakeholders in the process; monitoring can lead to increased incentive for the government to make sure they are not violating the terms of the derogations.
 For instance, after Armenia derogated from the ICCPR and the ECHR, civil society and media representatives provided feedback on emergency measures affecting them—a move that was welcomed both by the OSCE Media Freedom Representative and the Armenian media [@OSCE:2020].
@@
 
-We examine the relationship between derogations and policy and human rights outcomes in two stages (see @tbl-vars).
-First, to test whether derogations were associated with emergency state responses, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
+We examine the relationship between derogations, policy, and human rights outcomes in two stages (see @tbl-vars).
+First, to test whether derogations were associated with emergency state responses, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) canceling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel.
 Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing stringency of emergency measures.[^severity] To allow for comparisons across outcomes, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.[^collapse]
@@
   Outcomes = c(
-    "Cancelling public events; Gathering restrictions; Close public transit; Restrictions on internal movement; Limits on international travel",
+    "Canceling public events; Gathering restrictions; Close public transit; Restrictions on internal movement; Limits on international travel",
     "Discriminatory policy; Violations of non-derogable rights; No time limit measures; Abusive enforcement"
@@
 
-Next, we explore the relationship between formal treaty derogations and four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether emergency measures were discriminatory, (2) whether countries derogated from non-derogable rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
+Next, we explore the relationship between treaty derogations and four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether emergency measures were discriminatory, (2) whether countries derogated from non-derogable rights, (3) whether emergency measures failed to set a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures.
 Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations.
@@
 
-Several additional variables that are associated with and potentially confound both the decision to derogate and to enact specific policies.
+Several additional variables are associated with and potentially confound the decision to derogate and to enact specific policies.
 To account for the severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.
@@
 Derogating countries are more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states; see Appendix Table 1 for numeric values).
-This provides strong support for both components of our first hypothesis—states that derogated were substantially more likely to impose relevant emergency policies aimed at curtailing the spread of the virus (H~1a~), while states that did not issue derogations were less likely to engage in similar policies (H~1b~).
+This provides strong support for both components of our first hypothesis—derogating states were substantially more likely to impose relevant emergency policies aimed at curtailing the spread of the virus (H~1a~), while states that did not issue derogations were less likely to engage in similar policies (H~1b~).
 
@@
 These results show that countries implementing emergency measures took the time to uphold their treaty obligations.
-International law was not just an empty promise—statse that derogated were more likely to keep these emergency measures in place for longer and take their derogations seriously.
+International law was not just an empty promise—states that derogated were more likely to keep these emergency measures in place for longer and take their derogations seriously.
 
@@
 The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.
-Thus, contrary to our expectation, derogating states also violate non-derogable rights and implement some discriminatory policies.
+Thus, contrary to our expectations, derogating states also violate non-derogable rights and implement some discriminatory policies.
 These findings can perhaps be attributed to increased domestic and international monitoring—since derogating states open themselves up to such monitoring and review by the UNHRC, we may be more likely to observe discrimination and non-derogable rights violations than we would in non-derogating states.
@@
 It is possible that if states already intended to implement limited emergency measures that respect the spirit of the treaty, then it is relatively low cost to submit a derogation.
-However, compliance with derogations may not necessarily be low cost—as part of UN Human Rights Committee reports, even individuals can file complaints.
+However, compliance with derogations may not necessarily be low cost—as part of UNHRC reports, even individuals can file complaints.
 Further, for regional treaties not examined in this paper, such as the ECHR, individuals can also seek recourse to the European Court of Human Rights.
@@
 
-We suggest that future research examines different incentives facing state officials during a crisis.
-There may be key differences in enforcement between local bureaucrats and higher-level elites.
-While elites submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.
+We suggest that future research examines different incentives facing state officials during crises.
+There may be key differences in enforcement between local-level bureaucrats and national-level elites.
+While elites submit derogations and may be more cautious about international law, local-level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs.
 non-derogable rights.
 This distinction may explain some variation in policy and discrimination practices during the pandemic and points to the need to examine not just the structure of flexibility mechanisms, but also their use in practice.
-We also recommend that future research analyze the impact of other factors, such as a robust civil society and media, that can mitigate the adverse impacts of emergency measures.
+We also recommend future research analyzing the impact of other factors, such as a robust civil society and media, that can mitigate adverse impacts of emergency measures.
 

### 2024-07-23 manuscript/appendix.qmd
Final changes pre-submission

--- 2024-07-10 manuscript/appendix.qmd
@@
   email: aheiss@gsu.edu
-date: July 10, 2024
+date: July 22, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
+::: {.content-visible when-meta="list-of-figures-normal"}
 ```{=latex}
@@
 ```
+:::
+
+::: {.content-visible when-meta="list-of-figures-single"}
+
+```{=latex}
+\begin{singlespace}
+\listoftables
+\listoffigures
+\end{singlespace}
+```
+
+:::
+
+\clearpage
 
@@
 
-```{r tb-hr-summary}
-#| tbl-cap: "Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models"
+```{r table-hr-summary}
 #| include: true
@@
   tt(width = 1) %>% 
-  format_tt(escape = TRUE) %>% 
+  format_tt(j = 1, escape = TRUE) %>% 
   style_tt(i = seq(1, nrow(who_regions), by = 2), bold = TRUE, background = "#e6e6e6") %>% 
-  style_tt(i = seq(2, nrow(who_regions), by = 2), indent = 2, align = "l") %>% 
+  style_tt(i = seq(2, nrow(who_regions), by = 2), indent = 0, align = "l") %>% 
   style_tt(

### 2024-07-23 manuscript/manuscript.qmd
Final changes pre-submission

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: July 20, 2024
+date: July 22, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
-[^regional]: See Table A5 for a list of countries that derogated from both the ACHR/ECHR and the ICCPR.
+[^regional]: See Table A6 for a list of countries that derogated from both the ACHR/ECHR and the ICCPR.
 
 [^future]: This project specifically examines what happens after derogations are filed.
-In a separate project by Authors (2024) we explore the determinants of derogations and find that (1) pandemic-era backsliding and (2) the severity of the pandemic as measured by national-level deaths are significant and positive indicators of pandemic-related derogation submission (see Table A6 in the appendix).
+In a separate project by Authors (2024) we explore the determinants of derogations and find that (1) pandemic-era backsliding and (2) the severity of the pandemic as measured by national-level deaths are significant and positive indicators of pandemic-related derogation submission, and that a country's level of rule of law is positively but not significantly predictive of derogation (see Table A7 in the appendix).
 
@@
 These observations are also measured at a country level, but due to the slow moving nature of these PanDem human rights variables, there is not enough variation within countries to fit a model with country-level effects.
-Since geography plays an important role in pandemic, emergency policy, and human rights responses, we instead nest country-quarter observations within `r info$n_regions` WHO regions (see Appendix Table 1).
+Since geography plays an important role in pandemic, emergency policy, and human rights responses, we instead nest country-quarter observations within `r info$n_regions` WHO regions (see Table A5).
 Importantly, several of our variables move at different rates—some are measured weekly, others quarterly, others annually, and others (like country name) are time-invariant.

### 2024-07-23 _targets_config.R
Add targets config file

@@ -1,0 +1,4 @@
+targets::tar_config_set(
+  store = here::here("_targets"),
+  script = here::here("_targets.R")
+)

### 2024-07-23 _targets.R
Add raw data

@@ -40,27 +40,40 @@ lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
 # Pipeline ----------------------------------------------------------------
 list(
   ## Raw data files ----
+  # Hand-coded by us
   tar_target(iccpr_who_raw_file,
              here_rel("data", "raw_data", "ICCPR Derogation and WHO case data 1 3 2020 to 6 30 2021.xlsx"),
              format = "file"),
+  tar_target(iccpr_treaty_action_file,
+            here_rel("data", "raw_data", "ICCPR Treaty Action Variables.xlsx"),
+            format = "file"),
+
   # https://doi.org/10.1371/journal.pmed.1001923
   # https://plos.figshare.com/articles/dataset/World_Health_Organization_WHO_Member_States_by_subregion_/2579569
   tar_target(who_subregion_raw_file,
              here_rel("data", "raw_data", "Havelaar et al Table 2.xls"),
              format = "file"),
-  tar_target(iccpr_treaty_action_file,
-             here_rel("data", "raw_data", "ICCPR Treaty Action Variables.xlsx"),
-             format = "file"),
+
+  # https://www.bsg.ox.ac.uk/research/covid-19-government-response-tracker
   tar_target(oxford_raw_file,
              here_rel("data", "raw_data", "Oxford Covid Response Data 1 3 2020 to 6 20 2021.xlsx"),
              format = "file"),
+  
+  # PanDem v6.0
+  # https://github.com/vdeminstitute/pandem/tree/master/older%20versions/PanDem%20v6
   tar_target(pandem_raw_file,
              here_rel("data", "raw_data", "pandem", "datasets", "pandem_TS_v6.xlsx"),
              format = "file"),
+
+  # V-Dem v12
+  # https://www.v-dem.net/data/dataset-archive/
   tar_target(vdem_raw_file,
              here_rel("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v12",
                       "V-Dem-CY-Full+Others-v12.rds"),
              format = "file"),
+
+  # Natural Earth
+  # https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
   tar_target(naturalearth_raw_file,
              here_rel("data", "raw_data", "ne_110m_admin_0_countries",
                       "ne_110m_admin_0_countries.shp"),

### 2024-07-24 manuscript/appendix.qmd
Only set LIBGS on macOS

@@
 ```{r tikz-stuff, include=FALSE}
+if (Sys.info()["sysname"] == "Darwin") {
   # Necessary for using dvisvgm on macOS
@@
   Sys.setenv(LIBGS = "/usr/local/share/ghostscript/9.53.3/lib/libgs.dylib.9.53")
+}
 

### 2024-07-24 _targets.R
Make sure targets knows its working in a subdirectory for the manuscript

@@ -37,6 +37,11 @@ here_rel <- function(...) {fs::path_rel(here::here(...))}
 # Load R scripts with functions to use in the pipeline
 lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
 
+# Set some conditional flags
+should_deploy <- identical(Sys.getenv("UPLOAD_WEBSITES1"), "TRUE")
+is_docker <- identical(Sys.getenv("IS_DOCKER"), "TRUE")
+
+
 # Pipeline ----------------------------------------------------------------
 list(
   ## Raw data files ----
@@ -167,17 +172,18 @@ list(
   tar_target(hr_table_data, build_hr_table_data(human_rights_plot_data)),
 
   ## Manuscript and analysis notebook ----
-  tar_quarto(manuscript, path = "manuscript", quiet = FALSE),
+  tar_quarto(manuscript, path = "manuscript", working_directory = "manuscript", quiet = FALSE),
 
   tar_quarto(website, path = ".", quiet = FALSE),
-  tar_target(deploy_script, here_rel("deploy.sh"), format = "file"),
+
+  tar_target(deploy_script, here_rel("deploy.sh"), format = "file", cue = tar_cue_skip(!should_deploy)),
   tar_target(deploy, {
     # Force a dependency
     website
     # Run the deploy script
-    if (Sys.getenv("UPLOAD_WEBSITES") == "TRUE") processx::run(paste0("./", deploy_script))
-  }),
-  
+    if (should_deploy) processx::run(paste0("./", deploy_script))
+  }, cue = tar_cue_skip(!should_deploy)),
+
   ## Render the README ----
   tar_quarto(readme, here_rel("README.qmd"))
 )

### 2024-07-25 analysis/data-final.qmd
Add link to docker repo + rearrange stuff

--- 2022-07-24 analysis/data-final.qmd
@@
     code-tools: false
+    toc: false
 ---
 
-<table class="table">
-  <thead>
-    <tr>
-      <th scope="col">Program</th>
-      <th scope="col">File</th>
-    </tr>
-  </thead>
-  <tbody>
-    <tr>
-      <td>CSV file for any program</td>
-      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa table >}} &ensp;<span class="font-monospace">daily_panel.csv</span></a></td>
-    </tr>
-    <tr>
-      <td><span class="font-monospace">.rds</span> file for R<br><span class="small">(load with `df <- readRDS("daily_panel.rds")`)</small></td>
-      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa brands r-project >}} &ensp;<span class="font-monospace">daily_panel.rds</span></a></td>
-    </tr>
-    <tr>
-      <td><span class="font-monospace">.dta</span> file for Stata<br><span class="small">(load with `use "daily_panel.dta"`)</small></td>
-      <td><a class="btn btn-primary" target="_blank" href="#">{{< fa table-cells >}} &ensp;<span class="font-monospace">daily_panel.dta</span></a></td>
-    </tr>
-  </tbody>
-</table>
+```{r include=FALSE}
+# Add these final data files as targets dependencies
+library(targets)
+tar_load(c(data_weekly_stata, data_weekly_csv, data_weekly_rds,
+           data_quarterly_stata, data_quarterly_csv, data_quarterly_rds))
+tar_load(c(data_weekly_stata_website, data_weekly_csv_website, data_weekly_rds_website,
+           data_quarterly_stata_website, data_quarterly_csv_website, data_quarterly_rds_website))
+```
+
+The cleaned data is available in three formats:
+
+- CSV file for any program
+- `.rds` file for R <span class="small">(load with `df <- readRDS("weekly_panel.rds")`)</span>
+- `.dta` file for Stata <span class="small">(load with `use "weekly_panel.dta"`)</span>
+
+## Weekly data
+
+<a class="btn btn-primary btn-sm" target="_blank" href="data/weekly_panel.csv">{{< fa table >}} &ensp;<span class="font-monospace">weekly_panel.csv</span></a> &emsp; <a class="btn btn-success btn-sm" target="_blank" href="data/weekly_panel.rds">{{< fa brands r-project >}} &ensp;<span class="font-monospace">weekly_panel.rds</span></a> &emsp; <a class="btn btn-warning btn-sm" target="_blank" href="data/weekly_panel.dta">{{< fa table-cells >}} &ensp;<span class="font-monospace">weekly_panel.dta</span></a>
+
+## Quarterly data
+
+<a class="btn btn-primary btn-sm" target="_blank" href="data/quarterly_panel.csv">{{< fa table >}} &ensp;<span class="font-monospace">quarterly_panel.csv</span></a> &emsp; <a class="btn btn-success btn-sm" target="_blank" href="data/quarterly_panel.rds">{{< fa brands r-project >}} &ensp;<span class="font-monospace">quarterly_panel.rds</span></a> &emsp; <a class="btn btn-warning btn-sm" target="_blank" href="data/quarterly_panel.dta">{{< fa table-cells >}} &ensp;<span class="font-monospace">quarterly_panel.dta</span></a>

### 2024-07-25 analysis/replication.qmd
Add link to docker repo + rearrange stuff

---
title: "Replication Docker container"
format:
  html:
    code-tools: false
---

To maximize replicability, we wrote this manuscript using [Quarto](https://quarto.org/), which allowed us to mix computational figures, text, and tables with the actual prose of the manuscript. This means that there's no need to rely on comments within the code to identify the location of each appropriate result in the manuscript—all results are programmatically included when rendering the document.

We use the [{renv} package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of R packages, and we use the [{targets} package](https://docs.ropensci.org/targets/index.html) to manage all the file dependencies and run the analysis. {targets} is especially helpful with long-running objects like the main models, which take ≈40 minutes to run—as long as upstream dependencies don't change, the models only need to run once, and can be loaded from {targets}'s data store thereafter.

Because it can sometimes be difficult to set up and configure version-specific libraries of R packages and install specific versions of Stan, we provide two methods for replicating our analysis: (1) a Docker container built and orchestrated with Docker Compose, or (2) restoring a {renv} environment on your local computer.

We have [a separate GitHub repository with a Dockerfile and complete instructions for replicating the results](https://github.com/andrewheiss/mountainous-mackerel-docker) (and the manuscript and the companion results notebook—this very website, actually!)

::: {.d-flex .justify-content-center .align-items-center}

```{=html}
<!-- svg generated at https://gh-card.dev/ -->
<svg xmlns="http://www.w3.org/2000/svg" xmlns:xlink="http://www.w3.org/1999/xlink" width="442" height="109" version="1.2" baseProfile="tiny" data-reactroot=""><defs></defs><g fill="none" stroke="black" stroke-width="1" fill-rule="evenodd" stroke-linecap="square" stroke-linejoin="bevel"><g fill="#ffffff" fill-opacity="1" stroke="none" transform="matrix(1,0,0,1,0,0)"><rect x="0" y="0" width="440" height="109"></rect></g><rect x="0" y="0" width="441" height="108" stroke="#eaecef" stroke-width="2"></rect><g fill="#586069" fill-opacity="1" stroke="none" transform="matrix(1.25,0,0,1.25,17,21)"><path vector-effect="none" fill-rule="evenodd" d="M4,9 L3,9 L3,8 L4,8 L4,9 M4,6 L3,6 L3,7 L4,7 L4,6 M4,4 L3,4 L3,5 L4,5 L4,4 M4,2 L3,2 L3,3 L4,3 L4,2 M12,1 L12,13 C12,13.55 11.55,14 11,14 L6,14 L6,16 L4.5,14.5 L3,16 L3,14 L1,14 C0.45,14 0,13.55 0,13 L0,1 C0,0.45 0.45,0 1,0 L11,0 C11.55,0 12,0.45 12,1 M11,11 L1,11 L1,13 L3,13 L3,12 L6,12 L6,13 L11,13 L11,11 M11,1 L2,1 L2,10 L11,10 L11,1"></path></g><g fill="#0366d6" fill-opacity="1" stroke="#0366d6" stroke-opacity="1" stroke-width="1" stroke-linecap="square" stroke-linejoin="bevel" transform="matrix(1,0,0,1,0,0)"><a target="" href="https://github.com/andrewheiss/mountainous-mackerel-docker"><text fill="#0366d6" fill-opacity="1" stroke="none" xml:space="preserve" x="41" y="33" font-family="sans-serif" font-size="16" font-weight="630" font-style="normal">mountainous-mackerel-docker</text></a></g><g fill="#586069" fill-opacity="1" stroke="#586069" stroke-opacity="1" stroke-width="1" stroke-linecap="square" stroke-linejoin="bevel" transform="matrix(1,0,0,1,0,0)"><text fill="#586069" fill-opacity="1" stroke="none" xml:space="preserve" x="17" y="65" font-family="sans-serif" font-size="14" font-weight="400" font-style="normal">Docker container for replicating &quot;Pandemic Pass&quot; </text></g><g fill="#24292e" fill-opacity="1" stroke="#24292e" stroke-opacity="1" stroke-width="1" stroke-linecap="square" stroke-linejoin="bevel" transform="matrix(1,0,0,1,0,0)"><text fill="#24292e" fill-opacity="1" stroke="none" xml:space="preserve" x="33" y="91" font-family="sans-serif" font-size="12" font-weight="400" font-style="normal">Dockerfile</text></g><g fill="#000000" fill-opacity="1" stroke="none" transform="matrix(1,0,0,1,110,78)"></g><g fill="#586069" fill-opacity="1" stroke="#586069" stroke-opacity="1" stroke-width="1" stroke-linecap="square" stroke-linejoin="bevel" transform="matrix(1,0,0,1,0,0)"></g><g fill="#000000" fill-opacity="1" stroke="none" transform="matrix(1,0,0,1,173,78)"></g><g fill="#586069" fill-opacity="1" stroke="#586069" stroke-opacity="1" stroke-width="1" stroke-linecap="square" stroke-linejoin="bevel" transform="matrix(1,0,0,1,0,0)"></g><circle cx="23" cy="86" r="7" stroke="none" fill="#384d54"></circle></g></svg>
```

:::

### 2024-07-25 _targets.R
Update target definitions

@@ -38,7 +38,7 @@ here_rel <- function(...) {fs::path_rel(here::here(...))}
 lapply(list.files("R", full.names = TRUE, recursive = TRUE), source)
 
 # Set some conditional flags
-should_deploy <- identical(Sys.getenv("UPLOAD_WEBSITES1"), "TRUE")
+should_deploy <- identical(Sys.getenv("UPLOAD_WEBSITES"), "TRUE")
 is_docker <- identical(Sys.getenv("IS_DOCKER"), "TRUE")
 
 
@@ -111,26 +111,6 @@ list(
 
   
   ## Save data ----
-  ### Daily ----
-  tar_target(data_daily_stata, 
-             save_dta(daily_panel, here_rel("data", "derived_data", "daily_panel.dta")),
-             format = "file"),
-  tar_target(data_daily_stata_website, 
-             save_dta(daily_panel, here_rel("analysis", "data", "daily_panel.dta")),
-             format = "file"),
-  tar_target(data_daily_csv, 
-             save_csv(daily_panel, here_rel("data", "derived_data", "daily_panel.csv")),
-             format = "file"),
-  tar_target(data_daily_csv_website, 
-             save_csv(daily_panel, here_rel("analysis", "data", "daily_panel.csv")),
-             format = "file"),
-  tar_target(data_daily_rds, 
-             save_r(daily_panel, here_rel("data", "derived_data", "daily_panel.rds")),
-             format = "file"),
-  tar_target(data_daily_rds_website, 
-             save_r(daily_panel, here_rel("analysis", "data", "daily_panel.rds")),
-             format = "file"),
-  
   ### Weekly ----
   tar_target(data_weekly_stata, 
              save_dta(weekly_panel, here_rel("data", "derived_data", "weekly_panel.dta")),
@@ -149,6 +129,26 @@ list(
              format = "file"),
   tar_target(data_weekly_rds_website, 
              save_r(weekly_panel, here_rel("analysis", "data", "weekly_panel.rds")),
+             format = "file"),
+  
+  ### Quarterly ----
+  tar_target(data_quarterly_stata, 
+             save_dta(quarterly_panel, here_rel("data", "derived_data", "quarterly_panel.dta")),
+             format = "file"),
+  tar_target(data_quarterly_stata_website, 
+             save_dta(quarterly_panel, here_rel("analysis", "data", "quarterly_panel.dta")),
+             format = "file"),
+  tar_target(data_quarterly_csv, 
[…]

### 2024-07-25 README.qmd
Add link to Docker repo

@@
 
-To reproduce the findings and re-run the analysis, do the following:
+To replicate the findings and re-run the analysis, you can use a Docker container at [`mountainous-mackerel-docker`](https://github.com/andrewheiss/mountainous-mackerel-docker), which will create a complete computing envirionment for running everything.
+Full details and instructions [are available there](https://github.com/andrewheiss/mountainous-mackerel-docker).
+
+Alternatively, everything can be run on your local computer too. Do the following:
 

### 2024-07-25 manuscript/appendix.qmd
Final edits pre-II

@@
   email: aheiss@gsu.edu
-date: July 22, 2024
+date: July 25, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 
-Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW>.
+Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/7GVDA>.
 
@@
 ```{r tbl-results-determinants}
-#| tbl-cap: Results from Authors (2024) showing the determinants of derogation
+#| tbl-cap: Results from @ChaudhryComstockHeiss:2024 showing the determinants of derogation
 #| include: true

### 2024-07-25 manuscript/manuscript.qmd
Use references div
Final edits pre-II

@@
   We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
-date: July 22, 2024
+date: July 25, 2024
 code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
@@
 [^future]: This project specifically examines what happens after derogations are filed.
-In a separate project by Authors (2024) we explore the determinants of derogations and find that (1) pandemic-era backsliding and (2) the severity of the pandemic as measured by national-level deaths are significant and positive indicators of pandemic-related derogation submission, and that a country's level of rule of law is positively but not significantly predictive of derogation (see Table A7 in the appendix).
+In a separate project by @ChaudhryComstockHeiss:2024 we explore the determinants of derogations and find that (1) pandemic-era backsliding and (2) the severity of the pandemic as measured by national-level deaths are significant and positive indicators of pandemic-related derogation submission, and that a country's level of rule of law is positively but not significantly predictive of derogation (see Table A7 in the appendix).
 
@@
 # References
+
+::: {#refs}
+:::

### 2024-07-25 manuscript/manuscript.qmd
lol oops, typo

@@
 thanks: |
-  We would like to thank particpants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
+  We would like to thank participants of the 2022 Peace Science Society conference and the 2023 International Studies Association (ISA) convention for their comments and feedback.
 date: July 25, 2024

### 2024-07-25 README.qmd
Add DOI badges
Add preprint badge

@@
 <!-- badges: start -->
+[![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
 <!-- badges: end -->

### 2024-08-06 README.qmd
Add instructions for installing r-universe stuff

@@
     - [Libertinus Math](https://github.com/alerque/libertinus)
+
 2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
@@
     - If you’re using Windows, [install RTools too](https://cran.r-project.org/bin/windows/Rtools/) and add it to your PATH so that you can install packages from source if needed
+
 3. Open `mountainous-mackerel.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
+
 4. In the terminal, run `quarto install tinytex` to ensure that you have a working LaTeX installation.
+
 5.
@@
 If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
-6. Run `renv::restore()` in the R console to install all the required packages for this project.
-7. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.
+
+6.
+Run this to install {cmdstanr}.
+This is supposed to happen automatically as part of `renv::restore()` below, since {cmdstanr} is in the lockfile, but [due to an issue with {renv}](https://github.com/rstudio/renv/issues/1961) (fixed in the development version as of 2024-08-06), it doesn't install correctly because it is hosted at <https://stan-dev.r-universe.dev> instead of CRAN.
+So for now, until the next stable release of {renv}, it's easiest to install {cmdstanr} in a separate step.
+
+   ```r
+   install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", "https://packagemanager.posit.co/cran/latest"))
+   ```
+
+7. Run `renv::restore()` in the R console to install all the required packages for this project.
+
+8. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.
 

### 2024-09-22 manuscript/appendix.qmd
Use homebrew ghostscript

@@
   # See https://www.andrewheiss.com/blog/2021/08/27/tikz-knitr-html-svg-fun/
-  Sys.setenv(LIBGS = "/usr/local/share/ghostscript/9.53.3/lib/libgs.dylib.9.53")
+  Sys.setenv(LIBGS = "/opt/homebrew/opt/ghostscript/lib/libgs.dylib")
 }

### 2024-09-22 README.qmd
Add instructions for XQuartz

@@
 2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
-    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library
+    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library.
+Also [install Homebrew](https://brew.sh/) and run `brew install ghostscript`.
     - If you’re using Windows, [install RTools too](https://cran.r-project.org/bin/windows/Rtools/) and add it to your PATH so that you can install packages from source if needed

### 2024-10-08 manuscript/manuscript.qmd
Fix missing word

@@
 There is no measurable difference in the probability of violating non-derogable rights across derogating- and non-derogating-countries across the entire sample.
-However, in first 15 months of the pandemic, derogating countries were 4–5 percentage points more likely (`r hr_mfx$discriminatory_policy.None.max$p_gt`) to have no discriminatory policies than those that did not derogate, though the difference disappears by June 2021.
+However, in the first 15 months of the pandemic, derogating countries were 4–5 percentage points more likely (`r hr_mfx$discriminatory_policy.None.max$p_gt`) to have no discriminatory policies than those that did not derogate, though the difference disappears by June 2021.
 The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status.

### 2024-11-26 README.qmd
Add official DOI

@@
 
-<!-- badges: start -->
-[![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
-<!-- badges: end -->
-
 [Suparna Chaudhry](https://www.suparnachaudhry.com/) • Lewis and Clark College  
@@
 [Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University
+
+---
+
+<!-- badges: start -->
+[![Global Policy DOI](https://img.shields.io/badge/International%20Interactions%20DOI-10.1080%2F03050629.2024.2413965-brightgreen)](https://doi.org/10.1080/03050629.2024.2413965) [![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
+<!-- badges: end -->
+
+> Suparna Chaudhry, Audrey L.
+Comstock, and Andrew Heiss, “Pandemic Pass?
+Treaty Derogations and Human Rights Practices during COVID-19,” *International Interactions* 50, no.
+6 (2024): 1064–86, <https://doi.org/10.1080/03050629.2024.2413965>.
 

### 2024-11-26 README.qmd
Better badge for Zenodo

@@
 <!-- badges: start -->
-[![Global Policy DOI](https://img.shields.io/badge/International%20Interactions%20DOI-10.1080%2F03050629.2024.2413965-brightgreen)](https://doi.org/10.1080/03050629.2024.2413965) [![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
+[![International Interactions DOI](https://img.shields.io/badge/International%20Interactions-10.1080%2F03050629.2024.2413965-brightgreen)](https://doi.org/10.1080/03050629.2024.2413965)  
+[![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF%20project-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://img.shields.io/badge/Code-10.5281%2Fzenodo.12817616-blue)](https://doi.org/10.5281/zenodo.12817616)
 <!-- badges: end -->

### 2024-11-26 README.qmd
Better green in the badge

--- 2024-11-26 README.qmd
@@
 <!-- badges: start -->
-[![Global Policy DOI](https://img.shields.io/badge/International%20Interactions%20DOI-10.1080%2F03050629.2024.2413965-brightgreen)](https://doi.org/10.1080/03050629.2024.2413965) [![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
+[![International Interactions DOI](https://img.shields.io/badge/International%20Interactions-10.1080%2F03050629.2024.2413965-3D9970)](https://doi.org/10.1080/03050629.2024.2413965)  
+[![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF%20project-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://img.shields.io/badge/Code-10.5281%2Fzenodo.12817616-blue)](https://doi.org/10.5281/zenodo.12817616)
 <!-- badges: end -->
